---
title: "Pigment Analysis"
author: "Sebastian Di Geronimo"
date: "2024-01-29"
format: html
editor: source
fig_width: 5 
fig_height: 5 
---

# 1.0 ---- Summary of Document ----

+ setup 
  - load libraries and data

+ load-data
  - load data to be used in this script

: concentration-vs-time 
  - one pigment through time for each region, colored by season

: scatter-plot-pigment-tchlor-a 
  - scatter plot of pigment x to Tchlor-a per region
  
: hist-viol 
  - function to make plots for histogram or violin
  - hist and viol plots for all pigments and regions
  
: chla-boxplot
  - boxplot for each region of chlor-a for each season


# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
    # additional
    ggthemes, 
    ggtext,
    ggstatsplot,
    patchwork
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )
  
source(here("scripts", "misc_functions.R"))

# set size options of figures
# knitr::opts_chunk$set(fig.width = 19, fig.height = 9.7)

# set ggplot theme
theme_set(theme_calc(base_size = 12, base_family = "serif"))

# skip some plotting chunks
skip <- TRUE
```

```{r hurricane-dates}
hurricane_dates <-
  tribble(
    ~date, ~name,
    # "2015-08-30", "Erika", ##
    # "2016-10-12", "Nicole", ##
    # "2017-07-31", "Emily", ##
    "2017-09-10", "Irma",
    # "2017-10-29", "Philippe", ##
    "2018-09-03", "Gordon",
    # "2019-09-01", "Dorian", ##
    # "2020-05-25", "Bertha", ##
    "2020-08-02", "Isaias",
    # "2020-08-26", "Laura", ##
    "2020-09-11", "Sally",
    "2020-11-09", "Eta",
  ) %>%
  mutate(
    date = ymd(date),
    nearoff = "Nearshore"
  ) %T>%
  print()
```


## 2.2 Load Dataset

Merged pigment data is created in `01_load_data.Rmd`
- "combined_pig_dat.csv"
- includes: 
  - metadata 
  - pigment concentrations
  - size class calculations

```{r load-data}
# run snippet after saving file
# set location to save data and plots
dir_data_save <- here("data", "processed", "03_1_pigment")
dir_plt_save  <- here("data", "plots", "03_1_pigment")


# set season names and order
szn <- c("Winter", "Spring", "Summer", "Autumn")

# load data
pig_dat_og <-
  here("data", "processed", "load_data", "combined_pig_dat.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(season = fct(season, levels = szn)) %>%
  filter(
    .by = c(station, cruise_id),
    depth == min(depth) 
    & depth < 20
  ) %T>%
  print()
```



```{r}
pig_dat_og %>%
  filter(
    str_detect(name_of_water_body, "Keys")
    & !str_detect(station, "60|21.5|24|^3$|9.5|9B")
  ) %>%
    .$tot_chl_a %>%
      max()

pig_dat_og %>%
  filter(
    str_detect(name_of_water_body, "Keys")
    & !str_detect(station, "60|21.5|24|^3$|9.5|9B")
  ) %>%
  summarise(
    chl_avg  = mean(tot_chl_a, na.rm = TRUE),
    chl_med  = median(tot_chl_a, na.rm = TRUE),
    chl_sd   = sd(tot_chl_a, na.rm = TRUE),
    chl_sd2  = sd(tot_chl_a, na.rm = TRUE) * 2,
    chl_lwr  = mean(tot_chl_a, na.rm = TRUE) - sd(tot_chl_a, na.rm = TRUE),
    chl_upr  = mean(tot_chl_a, na.rm = TRUE) + sd(tot_chl_a, na.rm = TRUE),
    chl_25   = quantile(tot_chl_a, probs = 0.25, na.rm = TRUE),
    chl_75   = quantile(tot_chl_a, probs = 0.75, na.rm = TRUE),
    chl_5    = quantile(tot_chl_a, probs = 0.05, na.rm = TRUE),
    chl_95   = quantile(tot_chl_a, probs = 0.95, na.rm = TRUE),
    chl_ci95 = 1.96*(sd(tot_chl_a)/(sqrt(length(tot_chl_a))))
  )

pig_dat_og %>%
  filter(
    str_detect(name_of_water_body, "Keys")
    & !str_detect(station, "60|21.5|24|^3$|9.5|9B")
  ) %>%
  summarise(
    .by = season,
    chl_avg = mean(tot_chl_a, na.rm = TRUE),
    chl_med = median(tot_chl_a, na.rm = TRUE),
    chl_sd  = sd(tot_chl_a, na.rm = TRUE),
    chl_25  = quantile(tot_chl_a, probs = 0.25, na.rm = TRUE),
    chl_75  = quantile(tot_chl_a, probs = 0.75, na.rm = TRUE)
  )

pig_dat_og %>%
  filter(
    str_detect(name_of_water_body, "Keys")
    & !str_detect(station, "60|21.5|24|^3$|9.5|9B")
    &  tot_chl_a <= 1
  ) %>%
  summarise(
    .by = season,
    chl_avg = mean(tot_chl_a, na.rm = TRUE),
    chl_med = median(tot_chl_a, na.rm = TRUE),
    chl_sd  = sd(tot_chl_a, na.rm = TRUE),
    chl_25  = quantile(tot_chl_a, probs = 0.25, na.rm = TRUE),
    chl_75  = quantile(tot_chl_a, probs = 0.75, na.rm = TRUE)
  )

pig_dat_og %>%
  filter(
    str_detect(name_of_water_body, "Keys")
    & !str_detect(station, "60|21.5|24|^3$|9.5|9B")
  ) %>%
  summarise(
    .by = year,
    chl_avg = mean(tot_chl_a, na.rm = TRUE),
    chl_med = median(tot_chl_a, na.rm = TRUE),
    chl_sd = sd(tot_chl_a, na.rm = TRUE),
    chl_25 = quantile(tot_chl_a, probs = 0.25, na.rm = TRUE),
    chl_75 = quantile(tot_chl_a, probs = 0.75, na.rm = TRUE)
  )

pig_dat_og %>%
  filter(
    str_detect(name_of_water_body, "Keys")
    & !str_detect(station, "60|21.5|24|^3$|9.5|9B")
    &  tot_chl_a <= 1
  ) %>%
  summarise(
    .by = year,
    chl_avg = mean(tot_chl_a, na.rm = TRUE),
    chl_med = median(tot_chl_a, na.rm = TRUE),
    chl_sd = sd(tot_chl_a, na.rm = TRUE),
    chl_25 = quantile(tot_chl_a, probs = 0.25, na.rm = TRUE),
    chl_75 = quantile(tot_chl_a, probs = 0.75, na.rm = TRUE)
  )


pig_dat_og %>%
 filter(
    str_detect(name_of_water_body, "Keys")
    & !str_detect(station, "60|21.5|24|^3$|9.5|9B")
  ) %>%
  ggplot(aes(x = date_time_utc, y = tot_chl_a)) +
  geom_point(aes(color = station)) #+
  # geom_line(aes(color = station)) 

pig_dat_og %>%
  filter(
    str_detect(name_of_water_body, "Keys")
    & !str_detect(station, "60|21.5|24|^3$|9.5|9B")
    &  tot_chl_a <= 1
  ) %>%
  ggplot(aes(x = date_time_utc, y = tot_chl_a)) +
  # geom_point(aes(color = station)) #+
  geom_point(aes(color = season)) #+
  # geom_line(aes(color = station)) 

pig_dat_og %>%
  filter(
    str_detect(name_of_water_body, "Keys")
    & !str_detect(station, "60|21.5|24|^3$|9.5|9B")
    # &  tot_chl_a <= 1
  ) %>%
  ggplot(aes(x = tot_chl_a)) +
  # geom_point(aes(color = station)) #+
  geom_histogram(fill = "white", color = "black", binwidth = 0.05) +
  geom_vline(xintercept = c(0.1388, 0.614), color = "red") +
  scale_y_continuous(breaks = seq(5, 40, 5))
  # geom_line(aes(color = station)) 


```

## 2.2.1 Remove Outliers?

```{r rm-outlier}
# remove_out <- menu(c("yes", "no"), title = "Remove Outliers?")

remove_out <- 2
if (remove_out == 1) {
  message("Filtering out outliers!")
  pig_dat <-
    pig_dat_og %>%
    filter(
      tot_chl_a <= 1
    )
  
  file_suffix <- "_no_out"
} else {
  message("Keeping outliers!")
  pig_dat <- pig_dat_og
  file_suffix <- ""
}


pig_dat_keys <-
  pig_dat %>%
  # reframe(
  #   .by = c(cruise_id, date_time_utc, station, depth),
  #   pick(1:lat),
  #   across(
  #     tot_chl_a:last_col(),
  #     \(x) mean(x, na.rm = TRUE)
  #   )
  # ) %>%
  mutate(station = str_replace(station, "9B", "9.5")) %>%
  filter(
    str_detect(name_of_water_body, "Keys") &
      !str_detect(station, "60|21.5|24|^3$|9.5|9B")
  ) %T>%
  print() %>%
  # select(cruise_id, date_time_utc, station, depth, season, tot_chl_a) %>%
  distinct(cruise_id, date_time_utc, station, depth, .keep_all = TRUE)  %>%
  mutate(
    .after = station,
    nearoff =
      if_else(
        str_detect(station, "10|16|7"),
        "Nearshore",
        "Offshore"
      )
  )
```


# 3.0 ---- Plotting ----

## 3.1 Scatter: Concentration of Chlorophyll-a vs Time

Chlorophyll-a plotted as a time series separated by region

```{r concentration-vs-time, fig.height=15, fig.asp=0.5}
if (!skip) {
# ---- over all water bodies ---- # 
pig_dat %>%
  ggplot(
    aes(
      x = date_time_utc,
      y = tot_chl_a,
      color = season,
      shape = name_of_water_body
    )
  ) +
  geom_point(size = 2) +
  labs(
    x = "Year",
    y = "Concentration (mg m<sup>-3</sup>)",
    color = "Season",
    shape = "Region"
  ) +
  scale_y_continuous(expand = c(0, 0), na.value = 0) +
  scale_x_datetime(
    date_breaks = "3 month",
    date_labels = "%B\n%Y"
  ) +
  facet_wrap(~name_of_water_body,
    nrow = 3,
    scales = "free_y"
  ) +
  theme(axis.title.y = element_markdown(margin = margin(r = 30, l = 15)))

# ---- Florida Keys ---- #
pig_dat %>%
  filter(
    str_detect(name_of_water_body, "Keys")
    # str_detect(name_of_water_body, "Keys")
    
    ) %>%
  ggplot(
    aes(
      x = date_time_utc,
      y = tot_chl_a,
      color = season,
      shape = station
    )
  ) +
  geom_point(size = 2) +
  labs(
    x = "Year",
    y = "Concentration (mg m<sup>-3</sup>)",
    color = "Season"
  ) +
  scale_y_continuous(expand = c(0, 0), na.value = 0) +
  scale_x_datetime(
    date_breaks = "1 year",
    date_labels = "%Y"
  ) +
  facet_wrap(~name_of_water_body,
    nrow = 3,
    scales = "free_y"
  ) +
  theme(axis.title.y = element_markdown(margin = margin(r = 30, l = 15)))

# ---- Florida Keys Stations ---- # 
pig_dat %>%
  filter(str_detect(name_of_water_body, "Keys")) %>%
  ggplot(
    aes(
      x = date_time_utc,
      y = tot_chl_a,
      color = season
    )
  ) +
  geom_point(size = 2) +
  labs(
    x = "Year",
    y = "Concentration (mg m<sup>-3</sup>)",
    color = "Season"
  ) +
  scale_y_continuous(
      expand = c(0, 0), 
      na.value = 0, 
      breaks = c(seq(0, 1, 0.1), 1:4)
      ) +
  scale_x_datetime(
    date_breaks = "1 year",
    date_labels = "%Y"
  ) +
  facet_wrap(station ~ .,
    nrow = 3,
    scales = "free_y"
  ) +
  theme(axis.title.y = element_markdown(margin = margin(r = 30, l = 15))) +
  ylim(c(0, NA))

# ---- Florida Keys Box/Violin ---- #
pig_dat %>%
  filter(str_detect(name_of_water_body, "Keys")) %>%
  ggplot(
    aes(
      x = season,
      y = tot_chl_a,
      fill = season
      
    )
  ) +
  geom_boxplot(outliers = FALSE) +
  geom_violin(fill = NA) +
  geom_jitter(shape = 21, width = 0.2) +
  scale_y_continuous(breaks = c(seq(0, 1, 0.1), 1:4))
}
```

## 3.2 Scatter: Pigment Concentration vs Total Chlorophyll-a 

All pigments vs chlor-a by season and region

```{r scatter-plot-pigment-tchlor-a, fig.height=7, fig.width=10}
if (!skip) {
plt_pigment_vs_chlor <-
  pig_dat %>%
  pivot_longer(
    data      = .,
    cols      = c(tot_chl_b:tchl),
    names_to  = "pigment",
    values_to = "conc",
  ) %>%
  nest(.by = pigment) %>%
  mutate(
    plt_chla = map2(
      .x = pigment,
      .y = data,
      \(.x, .df) {
        ggplot(
          .df,
          aes(
            x = tot_chl_a,
            y = conc,
            fill = name_of_water_body,
            shape = name_of_water_body
          )
        ) +
          geom_abline(slope = 1) +
          geom_point(alpha = 0.50) +
          labs(
            title = str_replace_all(.x, "_", " "),
            x = "Total Chlorophyll-a",
            y = "Pigment Concentration",
            fill = "Region",
            shape = "Region"
          ) +
          scale_y_continuous(expand = c(0, 0), na.value = 0) +
          scale_x_continuous(expand = c(0, 0)) +
          scale_shape_manual(values = c(21, 22, 24)) +
          coord_cartesian(ylim = c(0, NA)) +
          facet_wrap(name_of_water_body ~ season, scales = "free") +
          theme_bw(base_family = "serif", base_size = 12)
      }
    ),
    plt_time = map2(
      .x = pigment,
      .y = data,
      \(.x, .df) {
        ggplot(
          .df,
          aes(
            x     = date_time_utc,
            y     = conc,
            color = season,
            shape = name_of_water_body
          )
        ) +
          geom_point(size = 2) +
          labs(
            title = str_replace_all(.x, "_", " "),
            x     = "Year",
            y     = "Concentration (mg m<sup>-3</sup>)",
            color = "Season",
            shape = "Region"
          ) +
          scale_y_continuous(expand = c(0, 0), na.value = 0) +
          scale_x_datetime(
            date_breaks = "6 month",
            date_labels = "%B\n%Y",
            limits = c(as_datetime("2015-11-01"), NA)
          ) +
          facet_wrap(
            ~name_of_water_body,
            nrow = 3,
            scales = "free_y"
          ) +
        theme_bw(base_family = "serif", base_size = 12) +
        theme(axis.title.y = element_markdown(margin = margin(r = 30, l = 15)))
      }
    ),
    plt_tod = map2(
      .x = pigment,
      .y = data,
      \(.x, .df) {
        ggplot(
          .df,
          aes(
            x     = date_time_utc,
            y     = conc,
            color = season,
            shape = name_of_water_body
          )
        ) +
          geom_point(size = 2) +
          labs(
            title = str_replace_all(.x, "_", " "),
            x     = "Year",
            y     = "Concentration (mg m<sup>-3</sup>)",
            color = "Season",
            shape = "Region"
          ) +
          scale_y_continuous(expand = c(0, 0), na.value = 0) +
          # scale_x_datetime(
          #   date_breaks = "6 month",
          #   date_labels = "%B\n%Y",
          #   limits = c(as_datetime("2015-11-01"), NA)
          # ) +
          facet_wrap(
            ~name_of_water_body,
            nrow = 3,
            scales = "free_y"
          ) +
        theme_bw(base_family = "serif", base_size = 12) +
        theme(axis.title.y = element_markdown(margin = margin(r = 30, l = 15)))
      }
    )
  )

plt_pigment_vs_chlor$plt_time
plt_pigment_vs_chlor$plt_chla[[1]]

# ---- save plot
# save
plt_pigment_vs_chlor %$%
  walk2(
    pigment,
    plt_chla,
    \(.x, .y) {
      save_gg(
        plt            = .y, 
        save_location  = here(dir_plt_save, "pigment_vs_chlor", str_remove(file_suffix, "^_")),
        save_name      = glue("{.x}_vs_chla"),
        verbose        = TRUE,
        time_stamp_fmt = NULL,
        device         = "jpeg",
        height         = 9.7,
        width          = 19,
        dpi            = 600,
        units          = "in",
        bg             = "white"
      )
    }
  )

# save
plt_pigment_vs_chlor %$%
  walk2(
    pigment,
    plt_time,
    \(.x, .y) {
      save_gg(
        plt            = .y,
        save_location  = here(dir_plt_save, "pigment_vs_time", str_remove(file_suffix, "^_")),
        save_name      = glue("{.x}_time"),
        verbose        = TRUE,
        time_stamp_fmt = NULL,
        device         = "jpeg",
        height         = 9.7,
        width          = 19,
        dpi            = 600,
        units          = "in",
        bg             = "white"
      )
    }
  )

}
```



## 3.3 Plot: Pigments vs Time of Day



```{r time-of-day}
if (!skip) {
pig_dat %>%
  pivot_longer(
    data      = .,
    cols      = c(tot_chl_b:tchl),
    names_to  = "pigment",
    values_to = "conc",
  ) %>%
  nest(.by = pigment) %>%
  # slice(1) %>%
  mutate(
    plt_tod = map2(
      .x = pigment,
      .y = data,
      \(.x, .df) {
        
        .df %>%
        filter(str_detect(name_of_water_body, "Florida Keys"))  %>%
        ggplot(
          
          aes(
            x     = time,
            y     = conc,
            color = as.factor(year),
            # shape = name_of_water_body
          )
        ) +
          geom_point(size = 2) +
          labs(
            title = str_replace_all(.x, "_", " "),
            x     = "Time of Day",
            y     = "Concentration (mg m<sup>-3</sup>)",
            color = "Season",
            shape = "Region"
          ) +
          scale_y_continuous(expand = c(0, 0), na.value = 0) +
          # scale_x_time(breaks = "1 hour") + 
          # scale_x_datetime(
          #   date_breaks = "6 month",
          #   date_labels = "%B\n%Y",
          #   limits = c(as_datetime("2015-11-01"), NA)
          # ) +
          facet_wrap(
            ~season,
            # nrow = 3,
            scales = "free_y"
          ) +
        theme_bw(base_family = "serif", base_size = 12) +
        theme(axis.title.y = element_markdown(margin = margin(r = 30, l = 15)))
      }
    ),
    plt_tod_stn = map2(
      .x = pigment,
      .y = data,
      \(.x, .df) {
        
        .df %>%
        filter(str_detect(name_of_water_body, "Florida Keys"))  %>%
        ggplot(
          
          aes(
            x     = time,
            y     = conc,
            color = as.factor(year),
            # shape = name_of_water_body
          )
        ) +
          geom_point(size = 2) +
          labs(
            title = str_replace_all(.x, "_", " "),
            x     = "Time of Day",
            y     = "Concentration (mg m<sup>-3</sup>)",
            color = "Season",
            shape = "Region"
          ) +
          scale_y_continuous(expand = c(0, 0), na.value = 0) +
          # scale_x_time(breaks = "1 hour") + 
          # scale_x_datetime(
          #   date_breaks = "6 month",
          #   date_labels = "%B\n%Y",
          #   limits = c(as_datetime("2015-11-01"), NA)
          # ) +
          facet_wrap(
            ~station,
            # nrow = 3,
            # scales = "free_y"
          ) +
        theme_bw(base_family = "serif", base_size = 12) +
        theme(axis.title.y = element_markdown(margin = margin(r = 30, l = 15)))
      }
    )
  ) %>%
  pull(plt_tod_stn)
    
}
```


## 3.4 Histogram and Violin: Each Pigment

Plot all pigments for each season as histogram and violin (density) plots.

The histograms are separated by year, season and region. 
The violin plots are separated by season and region

```{r hist-viol, fig.height=9.7, fig.width=30}
if (!skip) {
plt_pig_hist_viola <-
  pig_dat %>%
  pivot_longer(
    cols      = tot_chl_a:tchl_a_tpg,
    names_to  = "pigment",
    values_to = "conc"
  ) %>%
  select(season, pigment, conc, name_of_water_body, year) %>%
  nest(.by = pigment) %>%
  mutate(
    hist = map2(
      .x = data,
      .y = pigment,
      ~ .x %>%
        ggplot(aes(conc,
          fill = season
        )) +
        geom_histogram(
          position = "stack",
          bins = 40,
          na.rm = TRUE
        ) +
        labs(
          title = .y,
          x     = "Log<sub>\u0065</sub>([Pigment] + 1)",
          y     = "Frequency",
          fill  = "Season"
        ) +
        scale_x_continuous(
          trans = log1p_trans(),
          expand = c(0, 0)
        ) +
        scale_y_continuous(expand = c(0, 0)) +
        coord_cartesian(xlim = c(0, NA)) +
        facet_grid(
          name_of_water_body ~ as.character(year) + season,
          scales = "free_x"
        ) +
        theme(
          axis.title.x =
            element_markdown(margin = margin(t = 20, b = 5))
        )
    ),
    viol = map2(
      .x = data,
      .y = pigment,
      ~ .x %>%
        ggplot(aes(x = season, y = conc, color = season)) +
        geom_violin(
          trim = TRUE,
          na.rm = TRUE
        ) +
        geom_jitter(
          width = 0.1,
          size = 0.5,
          alpha = 0.5,
          color = "gray",
          na.rm = TRUE
        ) +
        labs(
          title = .y,
          y     = "Concentration",
          x     = "Season"
        ) +
        scale_y_continuous(expand = c(0, 0)) +
        coord_cartesian(ylim = c(0, NA)) +
        facet_wrap(~ name_of_water_body) +
        theme(legend.position = "none")
    )
  )

# show subset of plots
plt_pig_hist_viola %>%
    slice_sample(n = 5) %T>%
    walk(.x = .$hist,
         .f = ~ print(.x))  %>% 
    walk(.x = .$viol,
         .f = ~ print(.x)) 
}
```


# 4.0 ANOVA: Chlorophyll-a 


```{r data-filter}
# pig_dat_keys <-
#   pig_dat %>%
#   # reframe(
#   #   .by = c(cruise_id, date_time_utc, station, depth),
#   #   pick(1:lat),
#   #   across(
#   #     tot_chl_a:last_col(),
#   #     \(x) mean(x, na.rm = TRUE)
#   #   )
#   # ) %>%
#   mutate(station = str_replace(station, "9B", "9.5")) %>%
#   filter(
#     str_detect(name_of_water_body, "Keys") &
#       !str_detect(station, "60|21.5|24|^3$|9.5|9B")
#   ) %T>%
#   print() %>%
#   # select(cruise_id, date_time_utc, station, depth, season, tot_chl_a) %>%
#   distinct(cruise_id, date_time_utc, station, depth, .keep_all = TRUE)  %>%
#   mutate(
#     .after = station,
#     nearoff =
#       if_else(
#         str_detect(station, "10|16|7"),
#         "Nearshore",
#         "Offshore"
#       )
#   )

pig_dat_keys
pig_dat_keys %>%
    count(season)
pig_dat_keys %>%
    count(season, nearoff)

pig_dat_keys %>%
  datawizard::describe_distribution(select = tot_chl_a, quartiles = TRUE)
pig_dat_keys %>%
  group_by(season) %>%
  datawizard::describe_distribution(select = tot_chl_a, quartiles = TRUE)
pig_dat_keys %>%
  group_by(nearoff) %>%
  datawizard::describe_distribution(select = tot_chl_a, quartiles = TRUE)
pig_dat_keys %>%
  filter(str_detect(station, "10|12|16|18")) %>%
  group_by(nearoff) %>%
  datawizard::describe_distribution(select = tot_chl_a, quartiles = TRUE)
```


```{r}
shelf(emmeans, sjPlot)

chla_anova <-
  lmerTest::lmer(
    tot_chl_a ~ nearoff * season *  (1 | year),
    mutate(pig_dat_keys, year = factor(year))
  )

summary(chla_anova)

(chla_anova_means <- 
  emmeans(
    chla_anova,
    ~ nearoff 
    )
)


contrast(chla_anova_means, "pairwise")


tab_model(chla_anova)
```

```{r}
# shelf(emmeans, sjPlot)

chla_anova <-
  lmerTest::lmer(
    tot_chl_a ~ nearoff *  (1 | year),
    mutate(pig_dat_keys, year = factor(year)) %>%
    filter(season == "Autumn")
  )

summary(chla_anova)

(chla_anova_means <- 
  emmeans(
    chla_anova,
    ~ nearoff 
    )
)


contrast(chla_anova_means, "pairwise")


tab_model(chla_anova)


shelf(glmmTMB)

chla_glmm <- glmmTMB::glmmTMB(
    tot_chl_a ~ nearoff * (1 | year) * season,
    data = mutate(pig_dat_keys, year = factor(year))
    )

summary(chla_glmm)
tab_model(chla_glmm)
emmeans(
  chla_glmm,
  ~ nearoff
)
emmeans(
  chla_glmm,
  ~ season
)
emmeans(
  chla_glmm,
  ~ nearoff + season
)

chla_glmm <- glmmTMB::glmmTMB(
    tot_chl_a ~ nearoff * season,
    data = mutate(pig_dat_keys, year = factor(year))
    )

summary(chla_glmm)
tab_model(chla_glmm)
emmeans(
  chla_glmm,
  ~ nearoff
)
emmeans(
  chla_glmm,
  ~ season
)
emmeans(
  chla_glmm,
  ~ nearoff + season
)

glmmTMB::glmmTMB(
    tot_chl_a ~ nearoff * month ,
    data = mutate(pig_dat_keys, 
                  year = factor(year),
                  )
    ) %>%
    summary()

mutate(pig_dat_keys,  date_time_utc = as_date(date_time_utc))
```

## 4.1 Between Stations

```{r station-anova}
# ---- between stations ---- #
ggbetweenstats(
  data  = pig_dat_keys,
  x     = station,
  y     = tot_chl_a,
  title = "Station",
  xlab  = "Station",
  ylab  = "Concentration",
  type  = "nonparametric"
) %>%
  print()

# ---- parametric ---- #
pig_dat_keys %>%
  aov(
    data = .,
    tot_chl_a ~ station
  ) %>%
  summary()

# ---- nonparametric ---- #
pig_dat_keys %>%
  rstatix::kruskal_test(
    .,
    tot_chl_a ~ station
  ) 

```


## 4.2 Between Season


```{r season-anova}
ggbetweenstats(
  data  = pig_dat_keys,
  x     = season,
  y     = tot_chl_a,
  title = "Season",
  xlab  = "Season",
  ylab  = "Concentration",
  type  = "nonparametric"
) %>%
  print()

# ---- parametric ---- #
# pig_dat_keys %>%
#   aov(
#     data = .,
#     tot_chl_a ~ season
#   ) %>%
#   summary()
# 
# pig_dat_keys %>%
#   rstatix::tukey_hsd(
#     .,
#     tot_chl_a ~ season
#   )
# ---- nonparametric ---- #
pig_dat_keys %>%
  rstatix::kruskal_test(
    .,
    tot_chl_a ~ season
  ) 

# ---- pairwise Dunn's test ---- #
pig_dat_keys %>%
  rstatix::dunn_test(
    .,
    tot_chl_a ~ season
  ) %>%
  mutate(
      p.adj = round(p.adj, 3)
  )


```
```{r}
remove_out <- 1
if (remove_out == 1) {
  message("Filtering out outliers!")
  pig_dat_out <-
    pig_dat_og %>%
    filter(
      tot_chl_a <= 1
    )
  
  file_suffix <- "_no_out"
} else {
  message("Keeping outliers!")
  pig_dat <- pig_dat_og
  file_suffix <- ""
}

pig_dat_keys_out <-
  pig_dat_out %>%
  # reframe(
  #   .by = c(cruise_id, date_time_utc, station, depth),
  #   pick(1:lat),
  #   across(
  #     tot_chl_a:last_col(),
  #     \(x) mean(x, na.rm = TRUE)
  #   )
  # ) %>%
  distinct(cruise_id, date_time_utc, station, depth, .keep_all = TRUE)  %>%
  mutate(station = str_replace(station, "9B", "9.5")) %T>%
  print() %>%
  filter(
    str_detect(name_of_water_body, "Keys") &
      !str_detect(station, "60|21.5|24|^3$|9.5|9B")
  ) %>%
  mutate(
    .after = station,
    nearoff =
      if_else(
        str_detect(station, "10|16|7"),
        "Nearshore",
        "Offshore"
      )
  ) # %>% filter(tot_chl_a <= 1) 

pig_dat_keys_out
count(pig_dat_keys_out, season)
```


```{r}
ggbetweenstats(
  data  = pig_dat_keys_out,
  x     = season,
  y     = tot_chl_a,
  title = "Season",
  xlab  = "Season",
  ylab  = "Concentration",
  type  = "nonparametric"
) %>%
  print()

# ---- parametric ---- #
# pig_dat_keys_out %>%
#   aov(
#     data = .,
#     tot_chl_a ~ season
#   ) %>%
#   summary()
# 
# pig_dat_keys_out %>%
#   rstatix::tukey_hsd(
#     .,
#     tot_chl_a ~ season
#   )

# ---- nonparametric ---- #
pig_dat_keys_out %>%
  rstatix::kruskal_test(
    .,
    tot_chl_a ~ season
  ) 

# ---- pairwise Dunn's test ---- #
pig_dat_keys_out %>%
  rstatix::dunn_test(
    .,
    tot_chl_a ~ season
  ) %>%
  mutate(
      p.adj = round(p.adj, 3)
  )


```


## 4.3 Between Years

Both parametric and nonparametric are not significant

Fail to reject the null hypothesis
```{r year-anova}
# ---- between year ---- #
ggbetweenstats(
  data  = pig_dat_keys,
  x     = year,
  y     = tot_chl_a,
  title = "Year",
  xlab  = "Year",
  ylab  = "Concentration",
  type  = "nonparametric"
) %>%
  print()

# ---- parametric ---- #
pig_dat_keys %>%
  aov(
    data = .,
    tot_chl_a ~ year
  ) %>%
  summary()

# ---- nonparametric ---- #
pig_dat_keys %>%
  rstatix::kruskal_test(
    .,
    tot_chl_a ~ year
  ) 

# ---- pairwise Dunn's test ---- #
pig_dat_keys %>%
  rstatix::dunn_test(
    .,
    tot_chl_a ~ year
  )
```

## 4.4 Distance From Shore


```{r distance-anova}
pig_dat_keys_out %>%
  filter(str_detect(station, "10|12|16|18")) %>%
  ggbetweenstats(
    data  = .,
    x     = nearoff,
    y     = tot_chl_a,
    title = "Distance",
    xlab  = "Distance",
    ylab  = "Concentration",
    type  = "nonparametric"
  ) %>%
  print()

# ---- parametric ---- #
pig_dat_keys_out %>%
  filter(str_detect(station, "10|12|16|18")) %>%
  aov(
    data = .,
    tot_chl_a ~ nearoff
  ) %>%
  summary()

# ---- nonparametric ---- #
pig_dat_keys %>%
  filter(str_detect(station, "10|12|16|18")) %>%
    # group_by(season) %>%
    
  rstatix::kruskal_test(
    .,
    tot_chl_a ~ nearoff
  )

pig_dat_keys_out %>%
  filter(str_detect(station, "10|12|16|18")) %>%
    # group_by(season) %>%
    
  rstatix::kruskal_test(
    .,
    tot_chl_a ~ nearoff
  )

pig_dat_keys %>%
  filter(str_detect(station, "10|12|16|18")) %>%
  count(season, nearoff)

# ---- pairwise Dunn's test ---- #
pig_dat_keys %>%
  filter(str_detect(station, "10|12|16|18")) %>%
  rstatix::dunn_test(
    .,
    tot_chl_a ~ nearoff
  )

```

```{r distance-anova}
pig_dat_keys_out %>%
  filter(str_detect(station, "10|12|16|18")) %>%
ggbetweenstats(
  data  = ,
  x     = nearoff,
  y     = tot_chl_a,
  title = "Distance",
  xlab  = "Distance",
  ylab  = "Concentration",
  type  = "nonparametric"
) %>%
  print()

# ---- parametric ---- #
pig_dat_keys_out %>%
  filter(str_detect(station, "10|12|16|18")) %>%
  aov(
    data = .,
    tot_chl_a ~ nearoff
  ) %>%
  summary()

# ---- nonparametric ---- #
pig_dat_keys_out %>%
  filter(str_detect(station, "10|12|16|18")) %>%
  group_by(season) %>%
  rstatix::kruskal_test(
    .,
    tot_chl_a ~ nearoff
  )

# ---- pairwise Dunn's test ---- #
pig_dat_keys_out %>%
  filter(str_detect(station, "10|12|16|18")) %>%
  rstatix::dunn_test(
    .,
    tot_chl_a ~ nearoff
  )
```


```{r distance-anova}
pig_dat_keys %>%
  # filter(str_detect(station, "10|12|16|18")) %>%
ggbetweenstats(
  data  = ,
  x     = nearoff,
  y     = tot_chl_a,
  title = "Distance",
  xlab  = "Distance",
  ylab  = "Concentration",
  type  = "nonparametric"
) %>%
  print()

# ---- parametric ---- #
pig_dat_keys %>%
  # filter(str_detect(station, "10|12|16|18")) %>%
  aov(
    data = .,
    tot_chl_a ~ nearoff
  ) %>%
  summary()

# ---- nonparametric ---- #
pig_dat_keys %>%
  # filter(str_detect(station, "10|12|16|18")) %>%
  rstatix::kruskal_test(
    .,
    tot_chl_a ~ nearoff
  )

# ---- pairwise Dunn's test ---- #
pig_dat_keys %>%
  # filter(str_detect(station, "10|12|16|18")) %>%
  rstatix::dunn_test(
    .,
    tot_chl_a ~ nearoff
  )


```


## 4.5 Interactions

```{r plt-anova}
pig_dat_keys %>%
  filter(
    str_detect(station, "10|12|16|18")
    # & tot_chl_a < 2 
    ) %>%
  select(season, nearoff, station, tot_chl_a) %>%
  group_by(nearoff) %>%
  rstatix::get_summary_stats(probs = c(0.05, 0.25, 0.5, 0.75, 0.95), type = "quantile")

# ---- interactions ---- #
# season and year
pig_dat_keys %>%
  aov(
    data = .,
    tot_chl_a ~ year * season
  ) %>%
  summary()

# season and station
pig_dat_keys %>%
  aov(
    data = .,
    tot_chl_a ~ station * season
  ) %>%
  summary()

# year and station
pig_dat_keys %>%
  aov(
    data = .,
    tot_chl_a ~ station * year
  ) %>%
  summary()

# season, station and year
pig_dat_keys %>%
  aov(
    data = .,
    tot_chl_a ~ station * season * year
  ) %>%
  summary()

```


```{r random-values}
pig_dat_keys %>%
  summarise(
    .by    = year,
    mean   = mean(tot_chl_a),
    median = median(tot_chl_a),
    sd     = sd(tot_chl_a, na.rm = TRUE)
  ) %>%
  arrange(median)

pig_dat %>%
  filter(
    str_detect(name_of_water_body, "Keys")
    & str_detect(station, "24|3|7|9.5|9B")) %>%
  count(station)
```


Used to check differences in Total Chlorophyll-a between Seasons

```{r keys-season-histogram}
pig_dat_keys %>%
  filter(!str_detect(station, "^24$")  
         & depth < 6) %>%
  ggplot(aes(tot_chl_a)) +
  geom_histogram(binwidth = 0.02) +
  facet_wrap(~ season, ncol = 1) +
  scale_x_log10()

pig_dat_keys %>%
  filter(
    str_detect(station, "10|12|16|18")   
    # !str_detect(station, "^24$")  
    & depth < 6) %>%
  ggplot(aes(tot_chl_a)) +
  geom_histogram(binwidth = 0.02) +
  scale_x_log10() +
  labs(x = NULL, y = NULL) +
  facet_grid(nearoff ~ season) 
```


# 5.0 Boxplot: Total Chlorophyll-a vs Season

Box plots per region and season for chlor-a


## 5.1 Create Boxplots 

Removed `geom_jitter` in place of `ggforce::geom_sina` because plots a jitter
axis as a density plot

geom_jitter(
  shape = 21,
  color = "black",
  fill  = "black",
  width = 0.1, 
  size  = 0.1, 
  alpha = 1, 
  na.rm = TRUE
  ) +

```{r chla-boxplot, fig.width=10}
# theme_set(theme_calc(base_size = 12, base_family = "TT Arial"))
theme_set(theme_calc(base_size = 12, base_family = "sans"))
chl_a_box <-
  # pig_dat %>%
  pig_dat_keys %>%
  filter(!str_detect(station, "^24$")  
         & depth < 6) %>%
  nest(.by = name_of_water_body) %>%
  mutate(
    plt = map2(
      .x = data,
      .y = name_of_water_body,
      function(.x, .y) {
        
        if (str_detect(.y, "Florida Keys")) {
          ylims <- c(0, 0.8)
        }  else {
          ylims <- c(0, quantile(.x[["tot_chl_a"]], 0.99, na.rm = TRUE) + 0.5) 
        }
        
        # labels for x axis with number of samples
        szn_labels <- 
          .x %>%
          count(season) %>%
          # mutate(labels = glue("{season} (N = {n})"))
          mutate(labels = glue("{season} (<i>n</i> = {n})"))
        
        .x %>%
          ggplot(
            aes(
              x = season,
              y = tot_chl_a
            )
          ) +
          stat_boxplot(geom = "errorbar", linewidth = 0.2, width = 0.1) +
          geom_boxplot(
            # aes(fill = season),
            outlier.shape = NA, 
            linewidth     = 0.2,
            color = "black",
            na.rm = TRUE, 
            alpha = 1,
            ) +
          ggforce::geom_sina(
            data = ~ filter(.x, tot_chl_a < ylims[2]),
            shape = 16,
            color = "grey70",
            size  = 0.55,
            alpha = 0.5,
            na.rm = TRUE,
            seed  = 1234
          ) + 
            
          
          stat_summary(
            color = "#CC0000",
            alpha = 0.5,
            fun.data   = mean_sdl, # std dev * 2
            fun.args = list(mult = 1),
            # fun.data   = mean_sdl, # std dev * 2
            # fun.data   = mean_cl_normal, # lower and upper Gaussian confidence limits 
            # fun.data   = mean_cl_boot, # nonparametric bootstrap for confidence limits of population mean
            # geom  = "linerange",
            geom  = "errorbar",
            linewidth = 0.5,
            width = 0.1,
            # size  = 0.5,
            na.rm = TRUE
          ) +
          stat_summary(
            color = "#CC0000",
            fun   = mean,
            geom  = "point",
            shape = 20,
            size  = 2,
            na.rm = TRUE
          ) +
            
          geom_point(
              data = ~ filter(.x, tot_chl_a > ylims[2])  %>%
                       mutate(tot_chl_a = ylims[2] - 0.005 + 0.05),
              color = "orange",
              size = 0.25
          ) +
          # add top and right lines
          annotate(
            geom = 'segment',
            y    = c(Inf, -Inf),
            yend = c(Inf, Inf),
            x    = c(-Inf, Inf),
            xend = c(Inf, Inf)
          ) +
          labs(
            title = NULL,
            x     = NULL,
            y     = "Chlorophyll-a Concentration (mg m<sup>-3</sup>)"
          ) +
            
          guides(fill = "none") +
          scale_x_discrete(labels = szn_labels$labels) +
          scale_y_continuous(
            breaks = scales::breaks_pretty(n = 4),
            expand = expansion(add = c(0, 0.05)), 
            # limits = ylims,
            # oob = scales::squish
            # oob = scales::oob_keep
            ) +
          scale_fill_manual(
              values =  c("#88CCEE", "#44AA99", "#DDCC77","#CC6677")
              ) +
          coord_cartesian(y = ylims, clip = "on") +
          theme_classic(base_family = "sans") +
          theme(
            text = element_text(size = 10, face = "bold"),
            line = element_line(color = "black", linewidth = 0.2),
            rect = element_rect(color = "black", linewidth = 0.2),
            
            axis.title.y    = element_markdown(),
            axis.title.x    = element_markdown(),
            axis.text.x     = element_markdown(size = 10, color = "black"),
            axis.text.y     = element_text(color = "black"),
            axis.ticks.x    = element_blank(),
            plot.background = element_rect(color = "white"),
            
            legend.position   = c(0.8, 0.9),
            legend.text       = element_text(size = 5),
            legend.key.size   = unit(2.5, "mm"),
            legend.margin     = margin(0.5, 0.5, 0.5, 0.5),
            legend.background = element_rect(color = "black"),
            
            strip.text       = element_markdown(),
            strip.background = element_blank(),
            strip.placement  = "outside",
            
            panel.spacing      = unit(0, "lines"),
            panel.grid.major.y = element_blank(),
            plot.margin      = margin(t = 5)
           
          )
            
            
          # theme(
          #   text            = element_text(size = 6, face = "bold"),
          #   line            = element_line(color = "black", linewidth = 0.2),
          #   rect            = element_rect(color = "black", linewidth = 0.2),
          #   axis.title.y    = element_markdown(),
          #   axis.title.x    = element_markdown(),
          #   axis.text.x     = element_markdown(color = "black"),
          #   axis.text.y     = element_text(color = "black"),
          #   axis.ticks.x    = element_blank(),
          #   plot.background = element_rect(color = "white"),
          #   panel.border    = element_rect(color = "black"),
          #   panel.grid.minor.x = element_blank(),
          #   panel.grid.major.y = element_blank(),
          #   panel.grid.minor.y = element_blank(),
          # )
      }
    )
  )

chl_a_box %$% 
map2(
  .x = plt,
  .y = name_of_water_body,
  \(.x, .y) {
    .x +
    labs(title = .y)
    
    }
)

# pig_dat %>%
pig_dat_keys %>%
  filter(!str_detect(station, "^24$")  
         & depth < 6) %>%
  filter(
    str_detect(name_of_water_body, "Keys")  
    & tot_chl_a > 0.8
  ) %>%
  select(season, tot_chl_a) %>%
  arrange(season)

pig_dat_keys %>%
  summarise(
    .by = season,
    num = n(),
    mean = mean(tot_chl_a),
    sd = sd(tot_chl_a),
    se = sd(tot_chl_a)/(sqrt(n())),
    ci95 = sd(tot_chl_a)/(sqrt(n())) * 1.96,
    lwr = mean(tot_chl_a) - sd(tot_chl_a),
    hi = mean(tot_chl_a) + sd(tot_chl_a),
    med = median(tot_chl_a, na.rm = TRUE),
    ) %>%
  arrange(season)

pig_dat_keys %>%
  summarise(
    .by = c(season, nearoff),
    num = n(),
    mean = mean(tot_chl_a),
    sd = sd(tot_chl_a),
    se = sd(tot_chl_a)/(sqrt(n())),
    ci95 = sd(tot_chl_a)/(sqrt(n())) * 1.96,
    lwr = mean(tot_chl_a) - sd(tot_chl_a),
    hi = mean(tot_chl_a) + sd(tot_chl_a),
    med = median(tot_chl_a, na.rm = TRUE),
    ) %>%
  arrange(season)
```
## 5.2 Save: Chlorophyll-a Plots

```{r save-chla-boxplot}
chl_a_box  %>%
  filter(str_detect(name_of_water_body, "Keys")) %$%
  walk2(
    .x = name_of_water_body,
    .y = plt,
    overwrite = FALSE,
    # overwrite = TRUE,
    
    titles = "without", # "with", "both"
    \(.x, .y, overwrite, titles) {
      naming <-
        str_to_lower(.x) %>%
        str_replace_all("\\s", "_")

      # save with a title
      if (str_detect(titles, "with$|both")) {
        save_gg(
          .y +
          labs(title = .x),
          save_location  = here(dir_plt_save, "chla_box", str_remove(file_suffix, "^_")),
          save_name      = glue(naming, "_chla_title", file_suffix),
          verbose        = TRUE,
          overwrite      = overwrite,
          time_stamp_fmt = "%Y%m%d_%H%M%S",
          device         = "png",
          width          = 90,
          height         = 90 * 0.5625,
          dpi            = 600,
          units          = "mm",
          bg             = "white"
        )
      }
      # save without a title
      if (str_detect(titles, "without|both")) {
        save_gg(
          .y,
          save_location  = here(dir_plt_save, "chla_box", str_remove(file_suffix, "^_")),
          save_name      = glue(naming, "_chla", file_suffix),
          verbose        = TRUE,
          overwrite      = overwrite,
          time_stamp_fmt = "%Y%m%d_%H%M%S",
          device         = "png",
          width          = 90,
          height         = 90 * 0.5625,
          dpi            = 600,
          units          = "mm",
          bg             = "white"
        )
      }
    }
  )

```


## 5.3 Boxplot: Nearshore vs Offshore Total Chlorophyll-a

Figure 3

```{r old-chla-nearoff}
# chl_a_box <-
#   pig_dat %>%
#   filter(str_detect(station, "10|12|16|18")  
#          & depth < 6
#          & str_detect(name_of_water_body, "Keys")) %>%
#   mutate(
#     nearoff = 
#       if_else(
#         str_detect(station, "10|16"), 
#         "Nearshore", 
#         "Offshore"
#         )
#     ) %>%
#   nest(.by = name_of_water_body) %>%
#   mutate(
#     plt = map2(
#       .x = data,
#       .y = name_of_water_body,
#       function(.x, .y) {
#         
#         # ylims <- c(0, quantile(.x[["tot_chl_a"]], 0.99, na.rm = TRUE) + 0.5)
#         ylims <- c(0, 0.8)
# 
#         # labels for x axis with number of samples
#         szn_labels <- 
#           .x %>%
#           count(season) %>%
#           mutate(labels = glue("{season} (N = {n})"))
#         
#         plt <- 
#           .x %>%
#           ggplot(
#             aes(
#               x = season,
#               y = tot_chl_a,
#               fill = nearoff
#             )
#           ) +
#           geom_boxplot(
#             aes(fill = nearoff),
#             outlier.shape = NA, 
#             linewidth     = 0.5,
#             color = "black",
#             na.rm = TRUE, 
#             alpha = 1,
#             key_glyph = draw_key_rect
#             ) +
#           stat_summary(
#             color       = "#CC0000",
#             alpha       = 0.5,
#             fun.data    = mean_sdl, # std dev * 2
#             fun.args    = list(mult = 1),  
#             # fun.data    = mean_cl_normal, # lower and upper Gaussian confidence limits
#             # fun.data    = mean_cl_boot, # nonparametric bootstrap for confidence limits of population mean
#             geom        = "linerange",
#             linewidth   = 0.5,
#             size        = 0.5,
#             na.rm       = TRUE,
#             show.legend = FALSE,
#             position    = position_dodge2(0.75, preserve = "single"),
#           ) +
#           stat_summary(
#             color = "#CC0000",
#             fun   = mean,
#             geom  = "point",
#             shape = 20,
#             size  = 2,
#             na.rm = TRUE,
#             position = position_dodge2(0.75, preserve = "single"),
#             show.legend = FALSE,
#           ) +
#           labs(
#             title = NULL,
#             x     = NULL,
#             y     = "Chlorophyll-a Concentration (mg m<sup>-3</sup>)",
#             fill  = NULL
#           ) +
#             
#           # guides(fill = "none") +
#           scale_x_discrete(labels = szn_labels$labels) +
#           scale_y_continuous(
#             breaks = scales::breaks_pretty(n = 4),
#             expand = expansion(add = c(0, 0.009)), 
#             # limits = c(0, 1.5),
#             # oob = scales::squish
#             ) +
#           scale_fill_manual(values =  c("#88CCEE", "#882255")) +
#           coord_cartesian(y = ylims, clip = "on") +
# 
#           theme(
#             text              = element_text(size = 6),
#             line              = element_line(color = "black", linewidth = 0.2),
#             rect              = element_rect(color = "black", linewidth = 0.2),
#             axis.title.y      = element_markdown(),
#             axis.title.x      = element_markdown(),
#             axis.text.x       = element_text(color = "black"),
#             axis.text.y       = element_text(color = "black"),
#             axis.ticks.x      = element_blank(),
#             plot.background   = element_rect(color = "white"),
#             panel.border      = element_rect(color = "black"),
#             legend.position   = c(0.9, 0.9),
#             legend.text       = element_text(size = 5),
#             legend.key.size   = unit(2.5, "mm"),
#             legend.margin     = margin(0.5, 0.5, 0.5, 0.5),
#             legend.background = element_rect(color = "black")
#           )
#         
#         if (.env$remove_out == 2) {
#             plt <-
#               plt +
#             geom_point(
#             data  = tibble(
#               season    = c("Winter", "Winter"),
#               tot_chl_a = c(0.79, NA),
#               nearoff   = c("Nearshore", "Offshore")
#               ),
#             color = "#CC0000",
#             size  = 2,
#             position = position_dodge2(0.75, preserve = "single"),
#             show.legend = FALSE,
#             na.rm = TRUE
#           ) 
#         }
#         
#         return(plt)
#       }
#     )
#   )
# 
# chl_a_box$plt[[1]]

# average total chlorophyll-a for each season and distance from shore
pig_dat %>%
  filter(str_detect(station, "10|12|16|18")  
         & depth < 6
         & str_detect(name_of_water_body, "Keys")) %>%
  mutate(
    nearoff = 
      if_else(
        str_detect(station, "10|16"), 
        "Nearshore", 
        "Offshore"
        )
    ) %>%
  summarise(
    .by = c(season, nearoff),
    average = mean(tot_chl_a),
    sd = sd(tot_chl_a)
  ) %>%
  arrange(season, nearoff)

# # ---- save plot
# save_gg(
#   plt            = chl_a_box$plt[[1]], # edit if want specific plot
#   save_location  = here(dir_plt_save, "chla_nearoff_box", str_remove(file_suffix, "^_")),
#   save_name      = "chla_nearoff",
#   verbose        = TRUE,
#   # overwrite      = TRUE,
#   time_stamp_fmt = "%Y%m%d_%H%M%S",
#   device         = "png",
#   width          = 90,
#   height         = 90 * 0.5625,
#   dpi            = 600,
#   units          = "mm",
#   bg             = "white"
#   )

# shelf(camcorder)
# camcorder::gg_record(
#   dir    = here::here("data", "cam_test", "nearoff_chla"),
#   device         = "png",
#   height         = 90 * 0.5625,
#   width          = 90,
#   dpi            = 600,
#   units          = "mm",
#   bg             = "white"
# )

# shell.exec(here::here("data", "cam_test", "nearoff_chla"))
# gg_stop_recording()

# gg_resize_film(width = 90, height = 90 * 0.5625,  dpi = 600, units = "mm")

```


```{r season-labels}
# labels for x axis with number of samples
# szn_labels_nearoff <-
#   pig_dat_keys %>%
#   # filter(str_detect(station, "10|12|16|18")) %>%
#   count(season, nearoff) %>%
#   mutate(labels = glue("{nearoff} (N = {n})")) %T>%
#   print()

szn_labels <-
  pig_dat_keys %>%
  filter(str_detect(station, "10|12|16|18")) %>%
  count(season) %>%
  # mutate(labels = glue("{season} (N = {n})")) %T>%
  mutate(labels = glue("{season}\n(<i>n</i> = {n})")) %T>%
  print()

```

```{r stats-nearoff}
pig_dat_keys %>%
  filter(str_detect(station, "10|12|16|18")) %>%
  summarise(
    .by  = c(season, nearoff),
    num  = n(),
    mean = mean(tot_chl_a),
    sd   = sd(tot_chl_a),
    se   = sd(tot_chl_a)/(sqrt(n())),
    ci95 = sd(tot_chl_a)/(sqrt(n())) * 1.96,
    lwr  = mean(tot_chl_a) - sd(tot_chl_a),
    hi   = mean(tot_chl_a) + sd(tot_chl_a),
    med  = median(tot_chl_a, na.rm = TRUE),
    ) %>%
  arrange(season)



##%######################################################%##
#                                                          #
####               Near/Offshore Boxplot                ####
#                                                          #
##%######################################################%##
#' Near/Offshore Boxplot
#'
#' FUNCTION_DESCRIPTION
#'
#' @param .x Dataframe
#' @param .y Season Name
#' @param rm_out If outliers removed, 1 or 2
#'
#' @return Boxplot of Nearshore and Offshore
#' @examples
#' # ADD_EXAMPLES_HERE
#'
#' @author Sebastian Di Geronimo (October 28, 2024)
#'
boxplot_near_off <- function(.x, .y, rm_out) {
  # ylims <- c(0, quantile(.x[["tot_chl_a"]], 0.99, na.rm = TRUE) + 0.5)
  ylims <- c(0, 0.8)

  szn_labels_nearoff <-
    .x %>%
    count(nearoff) %>%
    # mutate(labels = glue("{nearoff}\n(N = {n})"))
    # mutate(labels = glue("{nearoff}<br>(<i>n</i> = {n})"))
    mutate(labels = glue("(<i>n</i> = {n})"))

  plt <-
    .x %>%
    ggplot(
      aes(
        x = nearoff,
        y = tot_chl_a,
        fill = nearoff
      )
    ) +
    stat_boxplot(geom = "errorbar", linewidth = 0.2, width = 0.1) +
    geom_boxplot(
      aes(fill = nearoff),
      outlier.shape = NA,
      linewidth = 0.2,
      color = "black",
      na.rm = TRUE,
      alpha = 1,
      key_glyph = draw_key_rect
    ) +
    stat_summary(
      color = "#CC0000",
      alpha = 0.5,
      fun.data = mean_sdl, # std dev * 2
      fun.args = list(mult = 1),
      # fun.data    = mean_cl_normal, # lower and upper Gaussian confidence limits
      # fun.data    = mean_cl_boot, # nonparametric bootstrap for confidence limits of population mean
      # geom        = "linerange",
      geom        = "errorbar",
      linewidth   = 0.5,
      width       = 0.1,
      # size        = 0.5,
      na.rm       = TRUE,
      show.legend = FALSE,
      position    = position_dodge2(0.75, preserve = "single"),
    ) +
    stat_summary(
      color = "#CC0000",
      fun = mean,
      geom = "point",
      shape = 20,
      size = 2,
      na.rm = TRUE,
      position = position_dodge2(0.75, preserve = "single"),
      show.legend = FALSE,
    ) +
      
    # add top line
    annotate(
        geom = 'segment',
        y    = Inf,
        yend = Inf,
        x    = -Inf,
        xend = Inf
    ) +
    labs(
      title = NULL,
      x     = NULL,
      y     = "Chlorophyll-a Concentration (mg m<sup>-3</sup>)",
      fill  = NULL
    ) +
    facet_wrap(~labels, strip.position = "bottom") +
    scale_x_discrete(labels = szn_labels_nearoff$labels) +
    scale_y_continuous(
      breaks = scales::breaks_pretty(n = 4),
      expand = expansion(add = c(0, 0.013)),
      # limits = c(0, 1.5),
      # oob = scales::squish
    ) +
    scale_fill_manual(values = c("#88CCEE", "#882255")) +
    coord_cartesian(y = ylims, clip = "on") +
    # theme_classic(base_family = "serif") +
    # theme_classic(base_family = "TT Arial") +
    theme_classic(base_family = "sans") +
    # theme_calc(base_family = "TT Arial") +
    theme(
      text = element_text(size = 10, face = "bold"),
      line = element_line(color = "black", linewidth = 0.2),
      rect = element_rect(color = "black", linewidth = 0.2),
      
      axis.title.y    = element_markdown(),
      axis.title.x    = element_markdown(),
      axis.text.x     = element_markdown(size = 6, color = "black"),
      axis.text.y     = element_text(color = "black"),
      axis.ticks.x    = element_blank(),
      plot.background = element_rect(color = "white"),
      
      legend.position   = c(0.8, 0.9),
      legend.text       = element_text(size = 5),
      legend.key.size   = unit(2.5, "mm"),
      legend.margin     = margin(0.5, 0.5, 0.5, 0.5),
      legend.background = element_rect(color = "black"),
      
      strip.text       = element_markdown(size = 10),
      strip.background = element_blank(),
      strip.placement  = "outside",
      
      panel.spacing      = unit(0, "lines"),
      
      panel.grid.major.y = element_blank()
      # panel.grid.major.y = element_line(color = "black")
    )

  # if (.env$remove_out == 2 &
  print(rm_out)
  if (rm_out == 2 &
    str_detect(.y, "Winter")
  ) {
    plt <-
      plt +
      geom_point(
        data = tibble(
          season    = c("Winter", "Winter"),
          tot_chl_a = c(0.79, NA),
          nearoff   = c("Nearshore", "Offshore")
        ),
        color = "#CC0000",
        size = 2,
        position = position_dodge2(0.75, preserve = "single"),
        show.legend = FALSE,
        na.rm = TRUE
      )
  }

  return(plt)
}
```

```{r chla-nearoff}
pig_keys_plt <-
  pig_dat_keys %>%
  select(-c(tot_chl_b:last_col())) %>%
  filter(str_detect(station, "10|12|16|18")) %>%
  left_join(szn_labels) %>%
  nest(.by = season) %>%
  arrange(season) %>%
  # slice(3) %>%
  mutate(
    plt = map2(
      .x = data,
      .y = season,
      boxplot_near_off,
      rm_out = remove_out
    )
  )

pull(pig_keys_plt, plt)[[1]]
```

```{r nearoff-total}
boxplot_total <- 
  pig_dat_keys %>%
  select(-c(tot_chl_b:last_col())) %>%
  filter(str_detect(station, "10|12|16|18")) %>%
  mutate(
    .by = nearoff,
    # label = glue("{nearoff}\n(N = {n()})")
    # label = glue("{nearoff}<br>(<i>n</i> = {n()})")
    label = glue("(<i>n</i> = {n()})")
      ) %>%
  # mutate(labels = glue("Total (N = {n()})")) %>%
  mutate(labels = glue("Total (<i>n</i> = {n()})")) %>%
  ggplot(
      aes(
        # x = nearoff,
        x = label,
        y = tot_chl_a,
        fill = nearoff
      )
    ) +
    stat_boxplot(geom = "errorbar", linewidth = 0.2, width = 0.1) +
    geom_boxplot(
      aes(fill = nearoff),
      outlier.shape = NA,
      linewidth = 0.2,
      color = "black",
      na.rm = TRUE,
      alpha = 1,
      key_glyph = draw_key_rect
    ) +
stat_summary(
      color = "#CC0000",
      alpha = 0.5,
      fun.data = mean_sdl, # std dev * 2
      fun.args = list(mult = 1),
      # fun.data    = mean_cl_normal, # lower and upper Gaussian confidence limits
      # fun.data    = mean_cl_boot, # nonparametric bootstrap for confidence limits of population mean
      # geom        = "linerange",
      geom = "errorbar",
      linewidth = 0.5,
      width = 0.1,
      size = 0.5,
      na.rm = TRUE,
      show.legend = FALSE,
      position = position_dodge2(0.75, preserve = "single"),
    ) +
    stat_summary(
      color = "#CC0000",
      fun = mean,
      geom = "point",
      shape = 20,
      size = 2,
      na.rm = TRUE,
      position = position_dodge2(0.75, preserve = "single"),
      show.legend = FALSE,
    ) +
    
    # add top and right lines
    annotate(
        geom = 'segment',
        y    = c(Inf, -Inf),
        yend = c(Inf, Inf),
        x    = c(-Inf, Inf),
        xend = c(Inf, Inf)
    ) +
    labs(
      title = NULL,
      x     = NULL,
      y     = "Chlorophyll-a Concentration (mg m<sup>-3</sup>)",
      fill  = NULL
    ) +
    facet_wrap(~labels, strip.position = "bottom") +
    # scale_x_discrete(labels = szn_labels_nearoff$labels) +
    # scale_x_discrete(labels =  label) +
    scale_y_continuous(
      breaks = scales::breaks_pretty(n = 4),
      expand = expansion(add = c(0, 0.013)),
      # limits = c(0, 1.5),
      # oob = scales::squish
    ) +
    scale_fill_manual(values = c("#88CCEE", "#882255")) +
    coord_cartesian(y = c(0, 0.8), clip = "on") +
    # theme_classic(base_family = "serif") +
    # theme_classic(base_family = "TT Arial") +
    theme_classic(base_family = "sans") +
    theme(
      text = element_text(size = 10, face = "bold"),
      line = element_line(color = "black", linewidth = 0.2),
      rect = element_rect(color = "black", linewidth = 0.2),
      
      axis.title.y    = element_markdown(),
      axis.title.x    = element_markdown(),
      axis.text.x     = element_markdown(size = 6, color = "black"),
      axis.text.y     = element_text(color = "black"),
      axis.ticks.x    = element_blank(),
      plot.background = element_rect(color = "white"),
      
      legend.position   = c(0.8, 0.9),
      legend.text       = element_text(size = 5),
      legend.key.size   = unit(2.5, "mm"),
      legend.margin     = margin(0.5, 0.5, 0.5, 0.5),
      legend.background = element_rect(color = "black"),
      
      strip.text       = element_markdown(size = 10),
      strip.background = element_blank(),
      strip.placement  = "outside",
      
      panel.spacing      = unit(0, "lines"),
      # panel.grid.major.y = element_line(color = "black")
      panel.grid.major.y = element_blank()
    )

boxplot_total
```


```{r save-nearoff-boxplot}
nearoff_plt <-
  (   
  pull(pig_keys_plt, plt)[[1]] + theme(legend.position = "none") |
      
  pull(pig_keys_plt, plt)[[2]] 
  + theme(legend.position = "inside", legend.position.inside = c(0.58, 0.9),
          legend.key.spacing.y = unit(0.5, "mm")) |
      
  pull(pig_keys_plt, plt)[[3]] + theme(legend.position = "none") |
      
  pull(pig_keys_plt, plt)[[4]] + theme(legend.position = "none") |
  # overall total
  boxplot_total + theme(legend.position = "none")
  ) +
 plot_layout(
    # ncol = 2,
    # guides = "keep",
    axes = "collect"
             )  &
  theme(
    panel.spacing    = unit(0, "lines"),
    plot.margin      = margin(t = 5),
    plot.background  = element_blank()
    
  )

nearoff_plt
 
# ---- save plot
save_gg(
  plt            = nearoff_plt, # edit if want specific plot
  save_location  = here(dir_plt_save, "chla_nearoff_box", str_remove(file_suffix, "^_")),
  save_name      = "chla_nearoff",
  verbose        = TRUE,
  # overwrite      = TRUE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "png",
  width          = 90,
  height         = 90 * 0.5625,
  dpi            = 600,
  units          = "mm",
  bg             = "white"
  ) 
```

5.3.3 Combine Seasonal Chlor-a and Near/Off

Doesn't work for some reason. Issue is that the seasonal chlor-a y axis ticks 
are separated from the grid panel


```{r combine-plts}
# does not include `total` box plot
nearoff_plt2 <-
  (   
  pull(pig_keys_plt, plt)[[1]] + theme(legend.position = "none") |
  pull(pig_keys_plt, plt)[[2]] 
  + theme(legend.position = "inside", legend.position.inside = c(0.58, 0.9),
          legend.key.spacing.y = unit(0.5, "mm")) |
  pull(pig_keys_plt, plt)[[3]] + theme(legend.position = "none") |
  pull(pig_keys_plt, plt)[[4]] + theme(legend.position = "none") +
      annotate(
        geom = 'segment',
        y    = c(Inf, -Inf),
        yend = c(Inf, Inf),
        x    = c(-Inf, Inf),
        xend = c(Inf, Inf)
    ) 
  ) +
 plot_layout(
    # ncol = 2,
    # guides = "keep",
    axes = "collect_y"
             )  &
  theme(
    panel.spacing    = unit(0, "lines"),
    plot.margin      = margin(t = 5),
    plot.background  = element_blank()
    
  )

combine_plt <-
  (chl_a_box$plt[[1]] / nearoff_plt2) +
  plot_layout(axis_titles = "collect")

combine_plt


save_gg(
  plt            = combine_plt, # edit if want specific plot
  save_location  = here(dir_plt_save, "chla_combine"),
  save_name      = "chla_combine",
  verbose        = TRUE,
  # overwrite      = TRUE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "png",
  width          = 180,
  height         = 180 * 0.5625,
  dpi            = 600,
  units          = "mm",
  bg             = "white"
  ) 
```
```{r cow-plot}
cowplot::plot_grid(
  NULL,
  chl_a_box$plt[[1]],
  NULL,
  nearoff_plt,
  nrow             = 4,
  label_fontfamily = "sans",
  label_size       = 10,
  labels           = c("", "a) Seasonal", "", "b) Distance from Shore"),
  vjust            = 0.25,
  hjust            = 0.0,
  label_x          = 0.0,
  rel_heights      = c(0.05, 1, 0.05, 1)
  ) %>%

  save_gg(
    plt            = ., # edit if want specific plot
    save_location  = here(dir_plt_save, "chla_combine"),
    save_name      = "chla_combine",
    verbose        = TRUE,
    # overwrite      = TRUE,
    time_stamp_fmt = "%Y%m%d_%H%M%S",
    device         = "png",
    width          = 170,
    height         = 170,
    dpi            = 600,
    units          = "mm",
    bg             = "white"
  )
```


## 5.4 Time Series Stations Tchla

```{r timeseries-chla}
# shelf(camcorder)
# camcorder::gg_record(
#   dir    = here::here("data", "cam_test", "ts_station"),
#   device         = "png",
#   height         = 170,
#   width          = 170,
#   dpi            = 600,
#   units          = "mm",
#   bg             = "white"
# )

# function used to set all date to same year
  same_year <- function(x, .year = 2000) {
    year(x) <- .year
    x
  }

dat_month_stk <-
  pig_dat_keys %>%
  select(-c(tot_chl_b:last_col())) %>%
  # filter(str_detect(nearoff, "Nearshore")) %>%
  # filter(str_detect(station, "10|12|16|18")) %>%
  # summarise(
  #   .by = c(name, dist, month_year),
  #   across(
  #     value,
  #     stat_summ
  #   )
  # )   %>%
  mutate(
    .after = 1,
    date = as_date(date_time_utc),
    # year2 = year(date),
    year2 = as.character(year),
    year_cat = paste(year2, nearoff),
    date2 = map(
      .x = date,
      .f = same_year
    ),
    nearoff = case_when(
        str_detect(nearoff, "Nearshore") ~ "a) Nearshore",
        str_detect(nearoff, "Offshore") ~ "b) Offshore"
    )
  ) %>%
  unnest(date2) %>%
  arrange(date) %T>%
  print() %>%
      
  # ---- start plot ---- #
  ggplot(
    aes(
      x     = date,
      y     = tot_chl_a,
      color = station,
      # shape = station
    )
  ) +
  geom_path(key_glyph = draw_key_rect) +
  geom_point(
    shape = 21, 
    fill = "white",
    show.legend = FALSE
    ) +
     
  #  ---- hurricane red lines ---- #
  geom_vline(
    xintercept = hurricane_dates$date, 
    color      = "red", 
    linetype   = "dashed"
    ) +
  # geom_hline(yintercept = 0.6, color = "gray", linetype = "dashed") +
  # ggrepel::geom_text_repel(
  
  #  ---- hurricane names ---- #
  # geom_text(
  geom_label(
    data = mutate(hurricane_dates, 
                  nearoff = "a) Nearshore"),
    aes(
      x     = date,
      y     = c(4, 4, 4, 3 , 3.7),
      label = name,
    ),
    angle = 90,
    vjust = -0.2,
    hjust = 1.0,
    size  = 4,
    inherit.aes = FALSE,
    fontface = "bold"
  ) +
  #  ---- end hurricane names ---- #

  labs(
    x = NULL,
    y = "Chlorophyll-a Concentration (mg m<sup>-3</sup>)",
    color = "Station",
    shape = "Station",
  ) +
  scale_x_date(
    date_labels = "%Y",
    date_breaks = "1 year"
  ) +
  scale_y_continuous(
    limits = c(-0.005, NA),
    expand = expansion(add = c(0, 0.1))
  ) +
  scale_color_manual(values = ggsci::pal_cosmic("hallmarks_light", alpha = 1)(10)[-c(1,8)]) +
  facet_wrap(
    ~nearoff,
    ncol   = 1,
    scales = "free_y"
  ) +
  theme_bw() +
  theme(
    text = element_text(face = "bold"),
    axis.text          = element_text(color = "black"),
    axis.title.y       = element_markdown(size = 10),
    axis.title.x       = element_markdown(),
    
    legend.position    = "inside",
    legend.position.inside = c(0.06, 0.85),
    legend.key.size    = unit(4, "mm"),
    legend.key.spacing.y = unit(0.5, "mm"), 
    legend.title       = element_text(size = 8, face = "bold"),
    legend.text        = element_text(size = 6),
    legend.background  = element_rect(
      color = "black",
      linewidth = 0.2
    ),
    
    # strip.background   = element_rect(fill = "white"),
    strip.background   = element_blank(),
    strip.text = element_text(face = "bold", hjust = 0),
    
    
    panel.grid.major.x = element_line(
      linetype  = "dashed",
      color     = "grey30",
      linewidth = 0.2
    ),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),

  )
dat_month_stk
# egg::tag_facet(dat_month_stk)

# shell.exec(here::here("data", "cam_test", "ts_station"))
# camcorder::gg_stop_recording()

# camcorder::gg_resize_film(width = 170, height = 170,  dpi = 600, units = "mm")
```



## 5.4 TEST Boxplot: Season and Year

Issues not being able to set the colors in seasons

```{r chla-year-season}
# shelf(camcorder)
# camcorder::gg_record(
#     dir    = here::here("data", "cam_test", "chla_box_season_year"),
#     device = "png",
#     width  = 90,
#     height = 90 * 0.5625,
#     dpi    = 600,
#     units  = "mm"
# )

# shell.exec(here::here("data", "cam_test", "chla_box_season_year"))
# gg_stop_recording()

n_fun <- function(x){
  return(data.frame(y = 0.95*30,
                    label = length(x)))
}


chl_a_box2 <- 
  pig_dat %>%
  filter(!str_detect(station, "^24$")
         & depth < 6
         ) %>%
  mutate(year_fct = fct_reorder(as.character(year), year, min)) %>%

  nest(.by = name_of_water_body) %>%
  mutate(
    plt = map2(
      .x = data,
      .y = name_of_water_body,
      function(.x, .y) {
        
        ylims <- c(0, quantile(.x[["tot_chl_a"]], 0.99, na.rm = TRUE) + 0.5)
        
        
        # labels for x axis with number of samples
        szn_labels <- 
          .x %>%
          count(season) %>%
          mutate(labels = glue("{season} (N = {n})"))
        
        .x %>%
          ggplot(
            aes(
              fill = year_fct,
              y = tot_chl_a,
              x = season
            )
          ) +
          geom_boxplot(
            # aes(fill = season),
            outlier.shape = NA, 
            linewidth     = 0.2,
            color = "black",
            na.rm = TRUE, 
            alpha = 1,
            position = position_dodge2(preserve = "single", padding = 0.2)
            ) +
          stat_summary(
              aes(group = year_fct, x = season, y = tot_chl_a),
            # aes(x = season, fill = year_fct),
            color = "#CC0000",
            fun   = mean,
            geom  = "point",
            shape = 20,
            size  = 1,
            na.rm = TRUE,
            position = position_dodge2(0.75, preserve = "single"),
            inherit.aes = FALSE
          ) +
          stat_summary(
            aes(group = year_fct, x = season, y = tot_chl_a),
            color = "black",
            fun.data = n_fun,
            geom  = "text",
            # shape = 20,
            size  = 1.5,
            na.rm = TRUE,
            position = position_dodge2(0.75, preserve = "single"),
            inherit.aes = FALSE, 
            hjust = 0.5,
            vjust = 1
          ) +
          labs(
            title = NULL,
            x     = NULL,
            y     = "Chlorophyll-a Concentration (mg m<sup>-3</sup>)",
            fill  = "Year",
            color = "Year"
          ) +
            
         guides(fill = guide_legend(direction = "horizontal", nrow = 1)) +
          scale_y_continuous(
            breaks = scales::breaks_pretty(n = 4),
            expand = expansion(add = c(0, 0.009)), 
            limits = ylims,
            oob = scales::squish) +
          scale_x_discrete(labels = szn_labels$labels, drop = FALSE) +
          scale_fill_manual(
              drop = FALSE,
              values =  c("#88CCEE", "#44AA99", "#117733","#CC6677", "#DDCC77", "#FFAABB")
              ) +
          scale_color_manual(
              drop = FALSE,
              values =  c("#88CCEE", "#44AA99", "#117733","#CC6677", "#DDCC77", "#FFAABB")
              ) +
          coord_cartesian(y = ylims, clip = "off") +

          theme(
            text            = element_text(size = 6),
            line            = element_line(color = "black", linewidth = 0.2),
            rect            = element_rect(color = "black", linewidth = 0.2),
            axis.title.y    = element_markdown(),
            axis.title.x    = element_markdown(),
            axis.text.x     = element_text(color = "black"),
            axis.text.y     = element_text(color = "black"),
            axis.ticks.x    = element_blank(),
            plot.background = element_rect(color = "white"),
            panel.border    = element_rect(color = "black"),
            legend.position = "bottom",
            legend.margin = margin(),
            legend.box.margin = margin(t = -10)
            # legend.direction = "horizontal",
            # legend.box = "horizontal"
          )
      }
    )
  )

chl_a_box2  %$% 
map2(
  .x = plt,
  .y = name_of_water_body,
  \(.x, .y) {
    .x +
    labs(title = .y)
    
    }
)

chl_a_box2 %>%
    slice(2) %>%
    pull(plt)

```



```{r chla-year-season}
# shelf(camcorder)
# camcorder::gg_record(
#     dir    = here::here("data", "cam_test", "chla_box_season_year"),
#     device = "png",
#     width  = 90,
#     height = 90 * 0.5625,
#     dpi    = 600,
#     units  = "mm"
# )

# shell.exec(here::here("data", "cam_test", "chla_box_season_year"))
# gg_stop_recording()


chl_a_box2 <- 
  pig_dat %>%
  filter(!str_detect(station, "^24$")) %>%
  mutate(year_fct = fct_reorder(as.character(year), year, min)) %>%

  nest(.by = name_of_water_body) %>%
  mutate(
    plt = map2(
      .x = data,
      .y = name_of_water_body,
      function(.x, .y) {
        
        ylims <- c(0, quantile(.x[["tot_chl_a"]], 0.99, na.rm = TRUE) + 0.5)
        
        # labels for x axis with number of samples
        yr_labels <- 
          .x %>%
          count(year) %>%
          mutate(labels = glue("{year} (N = {n})"))
        
        .x %>%
          ggplot(
            aes(
              x = year_fct,
              y = tot_chl_a,
              fill = season
            )
          ) +
          geom_boxplot(
            # aes(fill = season),
            outlier.shape = NA, 
            linewidth     = 0.2,
            color = "black",
            na.rm = TRUE, 
            alpha = 1,
            position = position_dodge2(preserve = "single", padding = 0.2)
            ) +
          # ggforce::geom_sina(
          #   aes(fill = year_fct, color = year_fct),
          #   shape = 22,
          #   # color = "black",
          #   # fill  = "black",
          #   # width = 0.1,
          #   size  = 0.1,
          #   alpha = 1,
          #   na.rm = TRUE,
          #   seed  = 1234,
          #   # maxwidth = 0.2, 
          #   scale = "width",
          #   position = position_dodge2(preserve = "single")

          # ) +
          stat_summary(
              aes(x = year_fct, group = season, y = tot_chl_a),
            # aes(x = season, fill = year_fct),
            color = "#CC0000",
            fun   = mean,
            geom  = "point",
            shape = 20,
            size  = 1.5,
            na.rm = TRUE,
            position = position_dodge2(0.75, preserve = "single"),
            inherit.aes = FALSE
          ) +
          stat_summary(
              aes(x = year_fct, group = season, y = tot_chl_a),
            # aes(x = season, fill = year_fct),
            color = "black",
            fun.data = n_fun,
            geom  = "text",
            # shape = 20,
            size  = 1.5,
            na.rm = TRUE,
            position = position_dodge2(0.75, preserve = "single"),
            inherit.aes = FALSE, 
            hjust = 0.5,
            vjust = 1
          ) +
          labs(
            title = NULL,
            x     = NULL,
            y     = "Chlorophyll-a Concentration (mg m<sup>-3</sup>)",
            fill  = "Season",
            color = "Season"
          ) +
            
          guides(fill = guide_legend(direction = "horizontal", nrow = 1)) +
          scale_x_discrete(labels = yr_labels$labels, drop = FALSE) +
          scale_y_continuous(
            breaks = scales::breaks_pretty(n = 4),
            expand = expansion(add = c(0, 0.009)), 
            limits = ylims,
            oob = scales::squish) +
          scale_fill_manual(
              drop = FALSE,
              values =  c("#88CCEE", "#44AA99", "#117733","#CC6677", "#DDCC77", "#FFAABB")
              ) +
          scale_color_manual(
              drop = FALSE,
              values =  c("#88CCEE", "#44AA99", "#117733","#CC6677", "#DDCC77", "#FFAABB")
              ) +
          coord_cartesian(y = ylims, clip = "off") +

          theme(
            text            = element_text(size = 6),
            line            = element_line(color = "black", linewidth = 0.2),
            rect            = element_rect(color = "black", linewidth = 0.2),
            axis.title.y    = element_markdown(),
            axis.title.x    = element_blank(),
            axis.text.x     = element_text(color = "black"),
            axis.text.y     = element_text(color = "black"),
            axis.ticks.x    = element_blank(),
            plot.background = element_rect(color = "white"),
            panel.border    = element_rect(color = "black"),
            legend.position = "bottom",
            legend.margin = margin(),
            legend.box.margin = margin(t = -10)
            # legend.direction = "horizontal",
            # legend.box = "horizontal"
          )
      }
    )
  )

chl_a_box2 %>%
    slice(2) %>%
    pull(plt)


chl_a_box2 %$% 
map2(
  .x = plt,
  .y = name_of_water_body,
  \(.x, .y) {
    .x +
    labs(title = .y)
    
    }
)

 pig_dat %>%
  filter(!str_detect(station, "^24$")) %>%
  mutate(year_fct = fct_reorder(as.character(year), year, min)) %>%
          count(name_of_water_body, year) %>%
          mutate(name_of_water_body, labels = glue("{year} (N = {n})"), .keep = "used")

```





