---
title: "Pigment Analysis"
author: "Sebastian Di Geronimo"
date: "2024-01-29"
format: html
editor: source
fig_width: 5 
fig_height: 5 
---

# 1.0 ---- Summary of Document ----

+ setup 
  - load libraries and data

+ load-data
  - load data to be used in this script

: concentration-vs-time 
  - one pigment through time for each region, colored by season

: scatter-plot-pigment-tchlor-a 
  - scatter plot of pigment x to Tchlor-a per region
  
: hist-viol 
  - function to make plots for histogram or violin
  - hist and viol plots for all pigments and regions
  
: chla-boxplot
  - boxplot for each region of chlor-a for each season


# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
    # additional
    ggthemes, ggtext,
    ggstatsplot
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )
  
source(here("scripts", "misc_functions.R"))

# set size options of figures
# knitr::opts_chunk$set(fig.width = 19, fig.height = 9.7)

# set ggplot theme
theme_set(theme_calc(base_size = 12, base_family = "serif"))
```

## 2.2 Load Dataset

Merged pigment data is created in `01_load_data.Rmd`
- "combined_pig_dat.csv"
- includes: 
  - metadata 
  - pigment concentrations
  - size class calculations

```{r load-data}
# run snippet after saving file
# set location to save data and plots
dir_data_save <- here("data", "processed", "03_1_pigment")
dir_plt_save  <- here("data", "plots", "03_1_pigment")


# set season names and order
szn <- c("Winter", "Spring", "Summer", "Autumn")

# load data
pig_dat <-
  here("data", "processed", "load_data", "combined_pig_dat.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(season = fct(season, levels = szn)) %T>%
  print()
```

# 3.0 ---- Plotting ----

## 3.1 Scatter: Concentration of Chlorophyll-a vs Time

Chlorophyll-a plotted as a time series separated by region

```{r concentration-vs-time, fig.height=15, fig.asp=0.5}
# ---- over all water bodies ---- # 
pig_dat %>%
  ggplot(
    aes(
      x = date_time_utc,
      y = tot_chl_a,
      color = season,
      shape = name_of_water_body
    )
  ) +
  geom_point(size = 2) +
  labs(
    x = "Year",
    y = "Concentration (mg m<sup>-3</sup>)",
    color = "Season",
    shape = "Region"
  ) +
  scale_y_continuous(expand = c(0, 0), na.value = 0) +
  scale_x_datetime(
    date_breaks = "3 month",
    date_labels = "%B\n%Y"
  ) +
  facet_wrap(~name_of_water_body,
    nrow = 3,
    scales = "free_y"
  ) +
  theme(axis.title.y = element_markdown(margin = margin(r = 30, l = 15)))

# ---- Florida Keys ---- #
pig_dat %>%
  filter(
    str_detect(name_of_water_body, "Keys")
    # str_detect(name_of_water_body, "Keys")
    
    ) %>%
  ggplot(
    aes(
      x = date_time_utc,
      y = tot_chl_a,
      color = season,
      shape = station
    )
  ) +
  geom_point(size = 2) +
  labs(
    x = "Year",
    y = "Concentration (mg m<sup>-3</sup>)",
    color = "Season"
  ) +
  scale_y_continuous(expand = c(0, 0), na.value = 0) +
  scale_x_datetime(
    date_breaks = "1 year",
    date_labels = "%Y"
  ) +
  facet_wrap(~name_of_water_body,
    nrow = 3,
    scales = "free_y"
  ) +
  theme(axis.title.y = element_markdown(margin = margin(r = 30, l = 15)))

# ---- Florida Keys Stations ---- # 
pig_dat %>%
  filter(str_detect(name_of_water_body, "Keys")) %>%
  ggplot(
    aes(
      x = date_time_utc,
      y = tot_chl_a,
      color = season
    )
  ) +
  geom_point(size = 2) +
  labs(
    x = "Year",
    y = "Concentration (mg m<sup>-3</sup>)",
    color = "Season"
  ) +
  scale_y_continuous(
      expand = c(0, 0), 
      na.value = 0, 
      breaks = c(seq(0, 1, 0.1), 1:4)
      ) +
  scale_x_datetime(
    date_breaks = "1 year",
    date_labels = "%Y"
  ) +
  facet_wrap(station ~ .,
    nrow = 3,
    scales = "free_y"
  ) +
  theme(axis.title.y = element_markdown(margin = margin(r = 30, l = 15))) +
  ylim(c(0, NA))

# ---- Florida Keys Box/Violin ---- #
pig_dat %>%
  filter(str_detect(name_of_water_body, "Keys")) %>%
  ggplot(
    aes(
      x = season,
      y = tot_chl_a,
      fill = season
      
    )
  ) +
  geom_boxplot(outliers = FALSE) +
  geom_violin(fill = NA) +
  geom_jitter(shape = 21, width = 0.2) +
  scale_y_continuous(breaks = c(seq(0, 1, 0.1), 1:4))
```

## 3.2 Scatter: Pigment Concentration vs Total Chlorophyll-a 

All pigments vs chlor-a by season and region

```{r scatter-plot-pigment-tchlor-a, fig.height=7, fig.width=10}
plt_pigment_vs_chlor <-
  pig_dat %>%
  pivot_longer(
    data      = .,
    cols      = c(tot_chl_b:tchl),
    names_to  = "pigment",
    values_to = "conc",
  ) %>%
  nest(.by = pigment) %>%
  mutate(
    plt_chla = map2(
      .x = pigment,
      .y = data,
      \(.x, .df) {
        ggplot(
          .df,
          aes(
            x = tot_chl_a,
            y = conc,
            fill = name_of_water_body,
            shape = name_of_water_body
          )
        ) +
          geom_abline(slope = 1) +
          geom_point(alpha = 0.50) +
          labs(
            title = str_replace_all(.x, "_", " "),
            x = "Total Chlorophyll-a",
            y = "Pigment Concentration",
            fill = "Region",
            shape = "Region"
          ) +
          scale_y_continuous(expand = c(0, 0), na.value = 0) +
          scale_x_continuous(expand = c(0, 0)) +
          scale_shape_manual(values = c(21, 22, 24)) +
          coord_cartesian(ylim = c(0, NA)) +
          facet_wrap(name_of_water_body ~ season, scales = "free") +
          theme_bw(base_family = "serif", base_size = 12)
      }
    ),
    plt_time = map2(
      .x = pigment,
      .y = data,
      \(.x, .df) {
        ggplot(
          .df,
          aes(
            x     = date_time_utc,
            y     = conc,
            color = season,
            shape = name_of_water_body
          )
        ) +
          geom_point(size = 2) +
          labs(
            title = str_replace_all(.x, "_", " "),
            x     = "Year",
            y     = "Concentration (mg m<sup>-3</sup>)",
            color = "Season",
            shape = "Region"
          ) +
          scale_y_continuous(expand = c(0, 0), na.value = 0) +
          scale_x_datetime(
            date_breaks = "6 month",
            date_labels = "%B\n%Y",
            limits = c(as_datetime("2015-11-01"), NA)
          ) +
          facet_wrap(
            ~name_of_water_body,
            nrow = 3,
            scales = "free_y"
          ) +
        theme_bw(base_family = "serif", base_size = 12) +
        theme(axis.title.y = element_markdown(margin = margin(r = 30, l = 15)))
      }
    ),
    plt_tod = map2(
      .x = pigment,
      .y = data,
      \(.x, .df) {
        ggplot(
          .df,
          aes(
            x     = date_time_utc,
            y     = conc,
            color = season,
            shape = name_of_water_body
          )
        ) +
          geom_point(size = 2) +
          labs(
            title = str_replace_all(.x, "_", " "),
            x     = "Year",
            y     = "Concentration (mg m<sup>-3</sup>)",
            color = "Season",
            shape = "Region"
          ) +
          scale_y_continuous(expand = c(0, 0), na.value = 0) +
          # scale_x_datetime(
          #   date_breaks = "6 month",
          #   date_labels = "%B\n%Y",
          #   limits = c(as_datetime("2015-11-01"), NA)
          # ) +
          facet_wrap(
            ~name_of_water_body,
            nrow = 3,
            scales = "free_y"
          ) +
        theme_bw(base_family = "serif", base_size = 12) +
        theme(axis.title.y = element_markdown(margin = margin(r = 30, l = 15)))
      }
    )
  )

plt_pigment_vs_chlor$plt_time
plt_pigment_vs_chlor$plt_chla[[1]]

# ---- save plot
# save
plt_pigment_vs_chlor %$%
  walk2(
    pigment,
    plt_chla,
    \(.x, .y) {
      save_gg(
        plt            = .y, 
        save_location  = here(dir_plt_save, "pigment_vs_chlor"),
        save_name      = glue("{.x}_vs_chla"),
        verbose        = TRUE,
        time_stamp_fmt = NULL,
        device         = "jpeg",
        height         = 9.7,
        width          = 19,
        dpi            = 600,
        units          = "in",
        bg             = "white"
      )
    }
  )

# save
plt_pigment_vs_chlor %$%
  walk2(
    pigment,
    plt_time,
    \(.x, .y) {
      save_gg(
        plt            = .y,
        save_location  = here(dir_plt_save, "pigment_vs_time"),
        save_name      = glue("{.x}_time"),
        verbose        = TRUE,
        time_stamp_fmt = NULL,
        device         = "jpeg",
        height         = 9.7,
        width          = 19,
        dpi            = 600,
        units          = "in",
        bg             = "white"
      )
    }
  )


```



## 3.3 Plot: Pigments vs Time of Day



```{r time-of-day}
 pig_dat %>%
  pivot_longer(
    data      = .,
    cols      = c(tot_chl_b:tchl),
    names_to  = "pigment",
    values_to = "conc",
  ) %>%
  nest(.by = pigment) %>%
  # slice(1) %>%
  mutate(
    plt_tod = map2(
      .x = pigment,
      .y = data,
      \(.x, .df) {
        
        .df %>%
        filter(str_detect(name_of_water_body, "Florida Keys"))  %>%
        ggplot(
          
          aes(
            x     = time,
            y     = conc,
            color = as.factor(year),
            # shape = name_of_water_body
          )
        ) +
          geom_point(size = 2) +
          labs(
            title = str_replace_all(.x, "_", " "),
            x     = "Time of Day",
            y     = "Concentration (mg m<sup>-3</sup>)",
            color = "Season",
            shape = "Region"
          ) +
          scale_y_continuous(expand = c(0, 0), na.value = 0) +
          # scale_x_time(breaks = "1 hour") + 
          # scale_x_datetime(
          #   date_breaks = "6 month",
          #   date_labels = "%B\n%Y",
          #   limits = c(as_datetime("2015-11-01"), NA)
          # ) +
          facet_wrap(
            ~season,
            # nrow = 3,
            scales = "free_y"
          ) +
        theme_bw(base_family = "serif", base_size = 12) +
        theme(axis.title.y = element_markdown(margin = margin(r = 30, l = 15)))
      }
    ),
    plt_tod_stn = map2(
      .x = pigment,
      .y = data,
      \(.x, .df) {
        
        .df %>%
        filter(str_detect(name_of_water_body, "Florida Keys"))  %>%
        ggplot(
          
          aes(
            x     = time,
            y     = conc,
            color = as.factor(year),
            # shape = name_of_water_body
          )
        ) +
          geom_point(size = 2) +
          labs(
            title = str_replace_all(.x, "_", " "),
            x     = "Time of Day",
            y     = "Concentration (mg m<sup>-3</sup>)",
            color = "Season",
            shape = "Region"
          ) +
          scale_y_continuous(expand = c(0, 0), na.value = 0) +
          # scale_x_time(breaks = "1 hour") + 
          # scale_x_datetime(
          #   date_breaks = "6 month",
          #   date_labels = "%B\n%Y",
          #   limits = c(as_datetime("2015-11-01"), NA)
          # ) +
          facet_wrap(
            ~station,
            # nrow = 3,
            # scales = "free_y"
          ) +
        theme_bw(base_family = "serif", base_size = 12) +
        theme(axis.title.y = element_markdown(margin = margin(r = 30, l = 15)))
      }
    )
  ) %>%
  pull(plt_tod_stn)
```


## 3.4 Histogram and Violin: Each Pigment

Plot all pigments for each season as histogram and violin (density) plots.

The histograms are separated by year, season and region. 
The violin plots are separated by season and region

```{r hist-viol, fig.height=9.7, fig.width=30}
plt_pig_hist_viola <-
  pig_dat %>%
  pivot_longer(
    cols      = tot_chl_a:tchl_a_tpg,
    names_to  = "pigment",
    values_to = "conc"
  ) %>%
  select(season, pigment, conc, name_of_water_body, year) %>%
  nest(.by = pigment) %>%
  mutate(
    hist = map2(
      .x = data,
      .y = pigment,
      ~ .x %>%
        ggplot(aes(conc,
          fill = season
        )) +
        geom_histogram(
          position = "stack",
          bins = 40,
          na.rm = TRUE
        ) +
        labs(
          title = .y,
          x     = "Log<sub>\u0065</sub>([Pigment] + 1)",
          y     = "Frequency",
          fill  = "Season"
        ) +
        scale_x_continuous(
          trans = log1p_trans(),
          expand = c(0, 0)
        ) +
        scale_y_continuous(expand = c(0, 0)) +
        coord_cartesian(xlim = c(0, NA)) +
        facet_grid(
          name_of_water_body ~ as.character(year) + season,
          scales = "free_x"
        ) +
        theme(
          axis.title.x =
            element_markdown(margin = margin(t = 20, b = 5))
        )
    ),
    viol = map2(
      .x = data,
      .y = pigment,
      ~ .x %>%
        ggplot(aes(x = season, y = conc, color = season)) +
        geom_violin(
          trim = TRUE,
          na.rm = TRUE
        ) +
        geom_jitter(
          width = 0.1,
          size = 0.5,
          alpha = 0.5,
          color = "gray",
          na.rm = TRUE
        ) +
        labs(
          title = .y,
          y     = "Concentration",
          x     = "Season"
        ) +
        scale_y_continuous(expand = c(0, 0)) +
        coord_cartesian(ylim = c(0, NA)) +
        facet_wrap(~ name_of_water_body) +
        theme(legend.position = "none")
    )
  )

# show subset of plots
plt_pig_hist_viola %>%
    slice_sample(n = 5) %T>%
    walk(.x = .$hist,
         .f = ~ print(.x))  %>% 
    walk(.x = .$viol,
         .f = ~ print(.x)) 
```


# 4.0 ANOVA: Chlorophyll-a 


```{r plt-anova}
# shelf(ggstatsplot)

n_fun <- function(x){
  return(data.frame(y = 0.95*30,
                    label = length(x)))
}

pig_dat_keys <- 
  pig_dat %>%
  filter(
    str_detect(name_of_water_body, "Keys")
    & !str_detect(station, "24|3|7|9.5|9B")) %>%
  mutate(
    labels = 
      if_else(
        str_detect(station, "10|16|7"), 
        "Nearshore", 
        "Offshore"
        )
    ) 


# between stations
ggbetweenstats(
  data  = pig_dat_keys,
  x     = station,
  y     = tot_chl_a,
  title = "Station",
  xlab  = "Station",
  ylab  = "Concentration",
  type  = "nonparametric"
) %>%
  print()

pig_dat_keys %>%
  aov(
    data = .,
    tot_chl_a ~ station
  ) %>%
  summary()

# between season
ggbetweenstats(
  data  = pig_dat_keys,
  x     = season,
  y     = tot_chl_a,
  title = "Season",
  xlab  = "Season",
  ylab  = "Concentration",
  type  = "nonparametric"
) %>%
  print()

pig_dat_keys %>%
  aov(
    data = .,
    tot_chl_a ~ season
  ) %>%
  summary()



# between year
ggbetweenstats(
  data  = pig_dat_keys,
  x     = year,
  y     = tot_chl_a,
  title = "Year",
  xlab  = "Year",
  ylab  = "Concentration",
  type  = "parametric"
) %>%
  print()

pig_dat_keys %>%
  aov(
    data = .,
    tot_chl_a ~ year
  ) %>%
  summary()



# between near/off
pig_dat_keys %>%
  filter(
    str_detect(station, "10|12|16|18")
    & tot_chl_a < 1.5
    ) %>%
ggbetweenstats(
  data  = ,
  x     = labels,
  y     = tot_chl_a,
  title = "Distance",
  xlab  = "Distance",
  ylab  = "Concentration",
  type  = "nonparametric"
) %>%
  print()

pig_dat_keys %>%
  aov(
    data = .,
    tot_chl_a ~ labels
  ) %>%
  summary()


pig_dat_keys %>%
  filter(str_detect(station, "10|12|16|18")) %>%
rstatix::dunn_test(
    .,
    tot_chl_a ~ labels
)

pig_dat_keys %>%
  filter(
    str_detect(station, "10|12|16|18")
    & tot_chl_a < 2 
    ) %>%
rstatix::kruskal_test(
    .,
    tot_chl_a ~ station
)

pig_dat_keys %>%
  filter(
    str_detect(station, "10|12|16|18")
    # & tot_chl_a < 2 
    ) %>%
  select(season, labels, station, tot_chl_a) %>%
  group_by(labels) %>%
  rstatix::get_summary_stats(probs = c(0.05, 0.25, 0.5, 0.75, 0.95), type = "quantile")

# ---- interactions ---- #
# season and year
pig_dat_keys %>%
  aov(
    data = .,
    tot_chl_a ~ year * season
  ) %>%
  summary()

# season and station
pig_dat_keys %>%
  aov(
    data = .,
    tot_chl_a ~ station * season
  ) %>%
  summary()

# season, station and year
pig_dat_keys %>%
  aov(
    data = .,
    tot_chl_a ~ station * season * year
  ) %>%
  summary()



```
```{r random-values}
pig_dat_keys %>%
  summarise(
    .by    = year,
    mean   = mean(tot_chl_a),
    median = median(tot_chl_a),
    sd     = sd(tot_chl_a, na.rm = TRUE)
  ) %>%
  arrange(median)

pig_dat %>%
  filter(
    str_detect(name_of_water_body, "Keys")
    & str_detect(station, "24|3|7|9.5|9B")) %>%
  count(station)
```

# 5.0 Boxplot: Total Chlorophyll-a vs Season

Box plots per region and season for chlor-a


## 5.1 Create Boxplots 

Removed `geom_jitter` in place of `ggforce::geom_sina` because plots a jitter
axis as a density plot

geom_jitter(
  shape = 21,
  color = "black",
  fill  = "black",
  width = 0.1, 
  size  = 0.1, 
  alpha = 1, 
  na.rm = TRUE
  ) +

```{r chla-boxplot, fig.width=3.54, fig.height=1.99}
chl_a_box <-
  pig_dat %>%
  filter(!str_detect(station, "^24$")  
         & depth < 6) %>%
  nest(.by = name_of_water_body) %>%
  mutate(
    plt = map2(
      .x = data,
      .y = name_of_water_body,
      function(.x, .y) {
        
        if (str_detect(.y, "Florida Keys")) {
          ylims <- c(0, 0.8)
        }  else {
          ylims <- c(0, quantile(.x[["tot_chl_a"]], 0.99, na.rm = TRUE) + 0.5) 
        }
        
        # labels for x axis with number of samples
        szn_labels <- 
          .x %>%
          count(season) %>%
          mutate(labels = glue("{season} (N = {n})"))
        
        .x %>%
          ggplot(
            aes(
              x = season,
              y = tot_chl_a
            )
          ) +
          geom_boxplot(
            aes(fill = season),
            outlier.shape = NA, 
            linewidth     = 0.2,
            color = "black",
            na.rm = TRUE, 
            alpha = 1,
            ) +
          ggforce::geom_sina(
            data = ~ filter(.x, tot_chl_a < ylims[2]),
            shape = 16,
            color = "grey70",
            size  = 0.55,
            alpha = 0.5,
            na.rm = TRUE,
            seed  = 1234
          ) + 
            
          geom_point(
              data = ~ filter(.x, tot_chl_a > ylims[2])  %>%
                       mutate(tot_chl_a = ylims[2] - 0.005 + 0.05),
              color = "orange",
              size = 0.25
          ) +
          stat_summary(
            color = "#CC0000",
            fun   = mean,
            geom  = "point",
            shape = 20,
            size  = 2,
            na.rm = TRUE
          ) +
          labs(
            title = NULL,
            x     = NULL,
            y     = "Chlorophyll-a Concentration (mg m<sup>-3</sup>)"
          ) +
            
          guides(fill = "none") +
          scale_x_discrete(labels = szn_labels$labels) +
          scale_y_continuous(
            breaks = scales::breaks_pretty(n = 4),
            expand = expansion(add = c(0, 0.05)), 
            # limits = ylims,
            # oob = scales::squish
            # oob = scales::oob_keep
            ) +
          scale_fill_manual(
              values =  c("#88CCEE", "#44AA99", "#DDCC77","#CC6677")
              ) +
          # coord_cartesian(y = ylims, clip = "off") +
          coord_cartesian(y = ylims, clip = "off") +
          theme(
            text            = element_text(size = 6),
            line            = element_line(color = "black", linewidth = 0.2),
            rect            = element_rect(color = "black", linewidth = 0.2),
            axis.title.y    = element_markdown(),
            axis.title.x    = element_markdown(),
            axis.text.x     = element_text(color = "black"),
            axis.text.y     = element_text(color = "black"),
            axis.ticks.x    = element_blank(),
            plot.background = element_rect(color = "white"),
            panel.border    = element_rect(color = "black")
          )
      }
    )
  )

chl_a_box %$% 
map2(
  .x = plt,
  .y = name_of_water_body,
  \(.x, .y) {
    .x +
    labs(title = .y)
    
    }
)

pig_dat %>%
  filter(!str_detect(station, "^24$")  
         & depth < 6) %>%
  filter(
    str_detect(name_of_water_body, "Keys")  
    & tot_chl_a > 0.8
  ) %>%
  select(season, tot_chl_a) %>%
  arrange(season)

```
## 5.2 Save: Chlorophyll-a Plots

```{r save-chla-boxplot}
chl_a_box %$%
  walk2(
    .x = name_of_water_body,
    .y = plt,
    # overwrite = FALSE,
    overwrite = TRUE,
    titles = "without", # "with", "both"
    \(.x, .y, overwrite, titles) {
      naming <-
        str_to_lower(.x) %>%
        str_replace_all("\\s", "_")

      # save with a title
      if (str_detect(titles, "with$|both")) {
        save_gg(
          .y +
          labs(title = .x),
          save_location  = here(dir_plt_save, "chla_box"),
          save_name      = glue(naming, "chla_title", .sep = "_"),
          verbose        = TRUE,
          overwrite      = overwrite,
          time_stamp_fmt = "%Y%m%d_%H%M%S",
          device         = "png",
          width          = 90,
          height         = 90 * 0.5625,
          dpi            = 600,
          units          = "mm",
          bg             = "white"
        )
      }
      # save without a title
      if (str_detect(titles, "without|both")) {
        save_gg(
          .y,
          save_location  = here(dir_plt_save, "chla_box"),
          save_name      = glue(naming, "chla", .sep = "_"),
          verbose        = TRUE,
          overwrite      = overwrite,
          time_stamp_fmt = "%Y%m%d_%H%M%S",
          device         = "png",
          width          = 90,
          height         = 90 * 0.5625,
          dpi            = 600,
          units          = "mm",
          bg             = "white"
        )
      }
    }
  )

```


## 5.3 Boxplot: Nearshore vs Offshore Total Chlorophyll-a

Figure 3

```{r chla-nearoff}
chl_a_box <-
  pig_dat %>%
  filter(str_detect(station, "10|12|16|18")  
         & depth < 6
         & str_detect(name_of_water_body, "Keys")) %>%
  mutate(
    nearoff = 
      if_else(
        str_detect(station, "10|16"), 
        "Nearshore", 
        "Offshore"
        )
    ) %>%
  nest(.by = name_of_water_body) %>%
  mutate(
    plt = map2(
      .x = data,
      .y = name_of_water_body,
      function(.x, .y) {
        
        # ylims <- c(0, quantile(.x[["tot_chl_a"]], 0.99, na.rm = TRUE) + 0.5)
        ylims <- c(0, 0.8)

        # labels for x axis with number of samples
        szn_labels <- 
          .x %>%
          count(season) %>%
          mutate(labels = glue("{season} (N = {n})"))
        
        .x %>%
          ggplot(
            aes(
              x = season,
              y = tot_chl_a,
              fill = nearoff
            )
          ) +
          geom_boxplot(
            aes(fill = nearoff),
            outlier.shape = NA, 
            linewidth     = 0.2,
            color = "black",
            na.rm = TRUE, 
            alpha = 1,
            key_glyph = draw_key_rect
            ) +
          geom_point(
            data  = tibble(
              season    = c("Winter", "Winter"), 
              tot_chl_a = c(0.79, NA), 
              nearoff   = c("Nearshore", "Offshore")
              ),
            color = "#CC0000",
            size  = 2,
            position = position_dodge2(0.75, preserve = "single"),
            show.legend = FALSE,
            na.rm = TRUE
          ) +
          stat_summary(
            color = "#CC0000",
            fun   = mean,
            geom  = "point",
            shape = 20,
            size  = 2,
            na.rm = TRUE,
            show.legend = FALSE,
            position = position_dodge2(0.75, preserve = "single"),
          ) +
          labs(
            title = NULL,
            x     = NULL,
            y     = "Chlorophyll-a Concentration (mg m<sup>-3</sup>)",
            fill  = NULL
          ) +
            
          # guides(fill = "none") +
          scale_x_discrete(labels = szn_labels$labels) +
          scale_y_continuous(
            breaks = scales::breaks_pretty(n = 4),
            expand = expansion(add = c(0, 0.009)), 
            # limits = c(0, 1.5),
            # oob = scales::squish
            ) +
          scale_fill_manual(values =  c("#88CCEE", "#882255")) +
          # coord_cartesian(y = c(0, 1.5), clip = "off") +
          coord_cartesian(y = ylims, clip = "off") +

          theme(
            text            = element_text(size = 6),
            line            = element_line(color = "black", linewidth = 0.2),
            rect            = element_rect(color = "black", linewidth = 0.2),
            axis.title.y    = element_markdown(),
            axis.title.x    = element_markdown(),
            axis.text.x     = element_text(color = "black"),
            axis.text.y     = element_text(color = "black"),
            axis.ticks.x    = element_blank(),
            plot.background = element_rect(color = "white"),
            panel.border    = element_rect(color = "black"),
            legend.position = c(0.9, 0.9),
            legend.text = element_text(size = 5),
            legend.key.size = unit(2.5, "mm"),
            legend.margin = margin(0.5, 0.5, 0.5, 0.5),
            legend.background = element_rect(color = "black")
          )
      }
    )
  )

chl_a_box$plt[[1]]

# average total chlorophyll-a for each season and distance from shore
pig_dat %>%
  filter(str_detect(station, "10|12|16|18")  
         & depth < 6
         & str_detect(name_of_water_body, "Keys")) %>%
  mutate(
    nearoff = 
      if_else(
        str_detect(station, "10|16"), 
        "Nearshore", 
        "Offshore"
        )
    ) %>%
  summarise(
    .by = c(season, nearoff),
    average = mean(tot_chl_a)
  ) %>%
  arrange(season, nearoff)

# ---- save plot
save_gg(
  plt            = chl_a_box$plt[[1]], # edit if want specific plot
  save_location  = here(dir_plt_save, "chla_nearoff_box"),
  save_name      = "chla_nearoff",
  verbose        = TRUE,
  overwrite      = FALSE,
  # overwrite      = TRUE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "png",
  width          = 90,
  height         = 90 * 0.5625,
  dpi            = 600,
  units          = "mm",
  bg             = "white"
  )
```


## 5.4 TEST Boxplot: Season and Year

Issues not being able to set the colors in seasons

```{r chla-year-season}
# shelf(camcorder)
# camcorder::gg_record(
#     dir    = here::here("data", "cam_test", "chla_box_season_year"),
#     device = "png",
#     width  = 90,
#     height = 90 * 0.5625,
#     dpi    = 600,
#     units  = "mm"
# )

# shell.exec(here::here("data", "cam_test", "chla_box_season_year"))
# gg_stop_recording()



chl_a_box2 <- 
  pig_dat %>%
  filter(!str_detect(station, "^24$")
         & depth < 6
         ) %>%
  mutate(year_fct = fct_reorder(as.character(year), year, min)) %>%

  nest(.by = name_of_water_body) %>%
  mutate(
    plt = map2(
      .x = data,
      .y = name_of_water_body,
      function(.x, .y) {
        
        ylims <- c(0, quantile(.x[["tot_chl_a"]], 0.99, na.rm = TRUE) + 0.5)
        
        
        # labels for x axis with number of samples
        szn_labels <- 
          .x %>%
          count(season) %>%
          mutate(labels = glue("{season} (N = {n})"))
        
        .x %>%
          ggplot(
            aes(
              fill = year_fct,
              y = tot_chl_a,
              x = season
            )
          ) +
          geom_boxplot(
            # aes(fill = season),
            outlier.shape = NA, 
            linewidth     = 0.2,
            color = "black",
            na.rm = TRUE, 
            alpha = 1,
            position = position_dodge2(preserve = "single", padding = 0.2)
            ) +
          stat_summary(
              aes(group = year_fct, x = season, y = tot_chl_a),
            # aes(x = season, fill = year_fct),
            color = "#CC0000",
            fun   = mean,
            geom  = "point",
            shape = 20,
            size  = 1,
            na.rm = TRUE,
            position = position_dodge2(0.75, preserve = "single"),
            inherit.aes = FALSE
          ) +
          stat_summary(
            aes(group = year_fct, x = season, y = tot_chl_a),
            color = "black",
            fun.data = n_fun,
            geom  = "text",
            # shape = 20,
            size  = 1.5,
            na.rm = TRUE,
            position = position_dodge2(0.75, preserve = "single"),
            inherit.aes = FALSE, 
            hjust = 0.5,
            vjust = 1
          ) +
          labs(
            title = NULL,
            x     = NULL,
            y     = "Chlorophyll-a Concentration (mg m<sup>-3</sup>)",
            fill  = "Year",
            color = "Year"
          ) +
            
         guides(fill = guide_legend(direction = "horizontal", nrow = 1)) +
          scale_y_continuous(
            breaks = scales::breaks_pretty(n = 4),
            expand = expansion(add = c(0, 0.009)), 
            limits = ylims,
            oob = scales::squish) +
          scale_x_discrete(labels = szn_labels$labels, drop = FALSE) +
          scale_fill_manual(
              drop = FALSE,
              values =  c("#88CCEE", "#44AA99", "#117733","#CC6677", "#DDCC77", "#FFAABB")
              ) +
          scale_color_manual(
              drop = FALSE,
              values =  c("#88CCEE", "#44AA99", "#117733","#CC6677", "#DDCC77", "#FFAABB")
              ) +
          coord_cartesian(y = ylims, clip = "off") +

          theme(
            text            = element_text(size = 6),
            line            = element_line(color = "black", linewidth = 0.2),
            rect            = element_rect(color = "black", linewidth = 0.2),
            axis.title.y    = element_markdown(),
            axis.title.x    = element_markdown(),
            axis.text.x     = element_text(color = "black"),
            axis.text.y     = element_text(color = "black"),
            axis.ticks.x    = element_blank(),
            plot.background = element_rect(color = "white"),
            panel.border    = element_rect(color = "black"),
            legend.position = "bottom",
            legend.margin = margin(),
            legend.box.margin = margin(t = -10)
            # legend.direction = "horizontal",
            # legend.box = "horizontal"
          )
      }
    )
  )

chl_a_box2  %$% 
map2(
  .x = plt,
  .y = name_of_water_body,
  \(.x, .y) {
    .x +
    labs(title = .y)
    
    }
)

chl_a_box2 %>%
    slice(2) %>%
    pull(plt)

```



```{r chla-year-season}
# shelf(camcorder)
# camcorder::gg_record(
#     dir    = here::here("data", "cam_test", "chla_box_season_year"),
#     device = "png",
#     width  = 90,
#     height = 90 * 0.5625,
#     dpi    = 600,
#     units  = "mm"
# )

# shell.exec(here::here("data", "cam_test", "chla_box_season_year"))
# gg_stop_recording()


chl_a_box2 <- 
  pig_dat %>%
  filter(!str_detect(station, "^24$")) %>%
  mutate(year_fct = fct_reorder(as.character(year), year, min)) %>%

  nest(.by = name_of_water_body) %>%
  mutate(
    plt = map2(
      .x = data,
      .y = name_of_water_body,
      function(.x, .y) {
        
        ylims <- c(0, quantile(.x[["tot_chl_a"]], 0.99, na.rm = TRUE) + 0.5)
        
        # labels for x axis with number of samples
        yr_labels <- 
          .x %>%
          count(year) %>%
          mutate(labels = glue("{year} (N = {n})"))
        
        .x %>%
          ggplot(
            aes(
              x = year_fct,
              y = tot_chl_a,
              fill = season
            )
          ) +
          geom_boxplot(
            # aes(fill = season),
            outlier.shape = NA, 
            linewidth     = 0.2,
            color = "black",
            na.rm = TRUE, 
            alpha = 1,
            position = position_dodge2(preserve = "single", padding = 0.2)
            ) +
          # ggforce::geom_sina(
          #   aes(fill = year_fct, color = year_fct),
          #   shape = 22,
          #   # color = "black",
          #   # fill  = "black",
          #   # width = 0.1,
          #   size  = 0.1,
          #   alpha = 1,
          #   na.rm = TRUE,
          #   seed  = 1234,
          #   # maxwidth = 0.2, 
          #   scale = "width",
          #   position = position_dodge2(preserve = "single")

          # ) +
          stat_summary(
              aes(x = year_fct, group = season, y = tot_chl_a),
            # aes(x = season, fill = year_fct),
            color = "#CC0000",
            fun   = mean,
            geom  = "point",
            shape = 20,
            size  = 1.5,
            na.rm = TRUE,
            position = position_dodge2(0.75, preserve = "single"),
            inherit.aes = FALSE
          ) +
          stat_summary(
              aes(x = year_fct, group = season, y = tot_chl_a),
            # aes(x = season, fill = year_fct),
            color = "black",
            fun.data = n_fun,
            geom  = "text",
            # shape = 20,
            size  = 1.5,
            na.rm = TRUE,
            position = position_dodge2(0.75, preserve = "single"),
            inherit.aes = FALSE, 
            hjust = 0.5,
            vjust = 1
          ) +
          labs(
            title = NULL,
            x     = NULL,
            y     = "Chlorophyll-a Concentration (mg m<sup>-3</sup>)",
            fill  = "Season",
            color = "Season"
          ) +
            
          guides(fill = guide_legend(direction = "horizontal", nrow = 1)) +
          scale_x_discrete(labels = yr_labels$labels, drop = FALSE) +
          scale_y_continuous(
            breaks = scales::breaks_pretty(n = 4),
            expand = expansion(add = c(0, 0.009)), 
            limits = ylims,
            oob = scales::squish) +
          scale_fill_manual(
              drop = FALSE,
              values =  c("#88CCEE", "#44AA99", "#117733","#CC6677", "#DDCC77", "#FFAABB")
              ) +
          scale_color_manual(
              drop = FALSE,
              values =  c("#88CCEE", "#44AA99", "#117733","#CC6677", "#DDCC77", "#FFAABB")
              ) +
          coord_cartesian(y = ylims, clip = "off") +

          theme(
            text            = element_text(size = 6),
            line            = element_line(color = "black", linewidth = 0.2),
            rect            = element_rect(color = "black", linewidth = 0.2),
            axis.title.y    = element_markdown(),
            axis.title.x    = element_blank(),
            axis.text.x     = element_text(color = "black"),
            axis.text.y     = element_text(color = "black"),
            axis.ticks.x    = element_blank(),
            plot.background = element_rect(color = "white"),
            panel.border    = element_rect(color = "black"),
            legend.position = "bottom",
            legend.margin = margin(),
            legend.box.margin = margin(t = -10)
            # legend.direction = "horizontal",
            # legend.box = "horizontal"
          )
      }
    )
  )

chl_a_box2 %>%
    slice(2) %>%
    pull(plt)


chl_a_box2 %$% 
map2(
  .x = plt,
  .y = name_of_water_body,
  \(.x, .y) {
    .x +
    labs(title = .y)
    
    }
)

 pig_dat %>%
  filter(!str_detect(station, "^24$")) %>%
  mutate(year_fct = fct_reorder(as.character(year), year, min)) %>%
          count(name_of_water_body, year) %>%
          mutate(name_of_water_body, labels = glue("{year} (N = {n})"), .keep = "used")

```





