---
title: "Size Fraction"
author: "Sebastian Di Geronimo"
date: "2024-01-29"
format: html
editor: source
---

# 1.0 ---- Summary of Document ----

+ setup 
  - load libraries and data

+ load-data
  - load data to be used in this script
  
: size-class-boxplots
  - boxplot for fraction of each size class, separated by season and region
  - boxplot for concentration of each size class, separated by season and region
  - boxplot for fraction of each size class
  
: size-frac-stack-bar
  - stacked bar charts separated by region with each season broken into average 
    for stations
  
: size-frac-table
  - table for seasonal average fraction for each size class   

# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
    # additional
    
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )
  
source(here("scripts", "misc_functions.R"))

# set size options of figures
knitr::opts_chunk$set(fig.width = 19, fig.height = 9.7)

# set ggplot theme
theme_set(theme_calc(base_size = 12, base_family = "serif"))
```

## 2.2 Load Dataset

Merged pigment data is created in `01_load_data.Rmd`
- "combined_pig_dat.csv"
- includes: 
  - metadata 
  - pigment concentrations
  - size class calculations

```{r load-data}
# set location to save data and plots
dir_data_save <- here("data", "processed", "03_2_size_fraction")
dir_plt_save  <- here("data", "plots", "03_2_size_fraction") 


# set season names and order
szn <- c("Winter", "Spring", "Summer", "Autumn")

# load data
pig_dat <-
  here("data", "processed", "load_data", "combined_pig_dat.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(season = fct(season, levels = szn)) %T>%
  print()
```


# 3.0 ---- Size Classes using Diagnostic Biomarker ----


## 3.1 Boxplots: Each Region and Season of Biomarkers for Each Size Classes
- micro (> 20um)
- nano (2 - 20um)
- pico (< 2um)

They are plotted by absolute value and percentages of the total


```{r size-class-boxplots}
shelf(patchwork)

size_frac_plt <-
  pig_dat %>%
  pivot_longer(
    cols      = f_micro:pico,
    names_to  = "size_frac",
    values_to = "conc"
  ) %>%
  select(name_of_water_body, season, size_frac, conc) %>%
  nest(.by = size_frac) %>%
  mutate(
    labels = case_when(
      str_detect(size_frac, "micro") ~ "Microplankton (>20 &mu;m)",
      str_detect(size_frac, "pico") ~ "Picoplankton (<2 &mu;m)",
      str_detect(size_frac, "nano") ~ "Nanoplankton (2 - 20 &mu;m)"
    ),
    labels = if_else(str_detect(size_frac, "f_"),
      str_c("Fraction of ", labels),
      labels
    ),
    plt = map2(
      .x = data,
      .y = labels,
      function(.x, .y) {
        if (str_detect(.y, "Frac")) {
          y_name <- "Percent (%)"
          y_lim <- c(0, 1)
        } else {
          y_name <- "Concentration (mg m<sup>-3</sup>)"
          y_lim <- c(0, quantile(.x[["conc"]], 0.99, na.rm = TRUE))
        }

        .x %>%
          ggplot(
            aes(
              x = season,
              y = conc,
              fill = name_of_water_body
            )
          ) +
          geom_boxplot(na.rm = TRUE, linewidth = 0.2) +
          {
            if (str_detect(y_name, "Percent")) {
              scale_y_continuous(
                labels = scales::label_percent(),
                expand = c(0, 0)
              )
            } else {
              scale_y_continuous(expand = c(0, 0))
            }
          } +
          labs(
            title = .y,
            x     = "Season",
            y     = y_name,
            fill  = "Location",
            color = "Location"
          ) +
          scale_fill_brewer(palette = "Dark2") +
          coord_cartesian(ylim = y_lim) +
          theme(
            plot.title   = element_markdown(),
            axis.text.x  = element_text(color = "black"),
            axis.text.y  = element_text(color = "black"),
            axis.title.y = element_markdown(
              margin = margin(
                r = 30,
                l = 15
              )
            ),
            axis.title.x = element_markdown(
              margin = margin(
                t = 20,
                b = 5
              )
            )
          )
      }
    )
  )

size_frac_plt$plt

size_frac_plt$plt[[1]] /
  size_frac_plt$plt[[2]] /
  size_frac_plt$plt[[3]]

# ---- save plot
# save size class fractions and absolute
size_frac_plt %$%
  walk2(
    size_frac,
    plt,
    \(.x, .y) {
      save_gg(
        plt            = .y,
        save_location  = here(dir_plt_save, "size_class"),
        save_name      = glue("size_class_{.x}"),
        verbose        = TRUE,
        time_stamp_fmt = NULL,
        device         = "jpeg",
        height         = 9.7,
        width          = 19,
        dpi            = 600,
        units          = "in",
        bg             = "white"
      )
    }
  )


# save stacked fractions
(size_frac_plt$plt[[1]] /
  size_frac_plt$plt[[2]] /
  size_frac_plt$plt[[3]]) %>%
  save_gg(
    plt            = .,
    save_location  = here(dir_plt_save, "size_class"),
    save_name      = glue("size_class_frac_stacked"),
    verbose        = TRUE,
    time_stamp_fmt = NULL,
    device         = "jpeg",
    height         = 9.7,
    width          = 19,
    dpi            = 600,
    units          = "in",
    bg             = "white"
  )

(size_frac_plt$plt[[4]] /
  size_frac_plt$plt[[5]] /
  size_frac_plt$plt[[6]]) %>%
  save_gg(
    plt            = .,
    save_location  = here(dir_plt_save, "size_class"),
    save_name      = glue("size_class_pct_stacked"),
    verbose        = TRUE,
    time_stamp_fmt = NULL,
    device         = "jpeg",
    height         = 9.7,
    width          = 19,
    dpi            = 600,
    units          = "in",
    bg             = "white"
  )

unshelf(patchwork)
```

## 3.2 Bar Graph: Size Classes for Each Station

Stacked bar charts for each station for each region as percentages.
Florida Keys - plotted West - East
Southwest and West Florida Shelf - plotted 


```{r size-frac-stack-bar}
plt_stack <-
  pig_dat %>%
  pivot_longer(
    cols      = f_micro:f_pico,
    names_to  = "size_frac",
    values_to = "frac"
  ) %>%
  nest(.by = name_of_water_body) %>%
  mutate(plt = map2(
    .x = data,
    .y = name_of_water_body,
    function(.x, .y) {
      factor_order <-
        ifelse(
          str_detect(.y, "(?i)southwest"),
          expression(fct_reorder(station, lat)),
          expression(fct_reorder(station, lon))
        )

      .x %>%
        mutate(
          water_name = .y,
          station = eval(factor_order)
        ) %>%
        ggplot(aes(
          x    = station,
          y    = frac,
          fill = size_frac
        )) +
        geom_bar(
          position = "fill",
          stat     = "identity",
          na.rm    = TRUE,
        ) +

        # labels
        labs(
          title = .y,
          x     = "Station",
          y     = "Percent (%)",
          fill  = "Phytoplankton Size Class"
        ) +
        scale_fill_brewer(
          palette = "Dark2",
          labels = c(
            "*f*<sub>micro</sub> (>20 &mu;m)",
            "*f*<sub>nano</sub> (2 - 20 &mu;m)",
            "*f*<sub>pico</sub> (<2 &mu;m)"
          )
        ) +
        scale_y_continuous(
          labels = label_percent(),
          expand = c(0, 0)
        ) +
        scale_x_discrete(guide = guide_axis(n.dodge = 1, angle = 90)) +
        facet_wrap(~season) +
        theme(
          legend.text = element_markdown(),
          panel.spacing.x = unit(25, "pt"),
          panel.border = element_rect(fill = NA)
        )
    }
  ))

plt_stack$plt

# save plots
plt_stack %$%
  walk2(
    .x = name_of_water_body,
    .y = plt,
    \(.x, .y) {
      .x %>%
        str_to_lower() %>%
        str_replace_all("\\s", "_") %>%
        glue(., "frac_bar", .sep = "_") %>%
        save_gg(
          .y,
          save_location  = dir_plt_save,
          save_name      = .,
          verbose        = TRUE,
          time_stamp_fmt = "%Y%m%d_%H%M%S",
          device         = "jpeg",
          height         = 9.7,
          width          = 19,
          dpi            = 600,
          units          = "in",
          bg             = "white"
        )
    }
  )


```

## 3.3 Table: Regional Size Classes Fractions for Each Season

Calculations for each size class and season:
- Average
- Standard deviation  

Table creation:
- use `gt`
- exported as `.html` and `.docx`


**NOTE**: 
when saving as word doc, the formatting **will not** adhere to what is expected 
and will need manual editing to use.

```{r size-frac-table}
frac_gt <-
  pig_dat %>%
  pivot_longer(
    cols      = f_micro:f_pico,
    names_to  = "size_frac",
    values_to = "frac"
  ) %>%
  summarise(
    .by      = c(name_of_water_body, season, size_frac),
    frac_avg = (mean(frac, na.rm = TRUE) * 100) %>% round(., 1),
    frac_sd  = (sd(frac, na.rm = TRUE) * 100) %>% round(., 1)
  ) %>%
  pivot_wider(
    names_from  = "size_frac",
    values_from = c(frac_avg, frac_sd)
  ) %>%
  arrange(name_of_water_body, season) %>%
  gt(
    rowname_col   = "season",
    groupname_col = "name_of_water_body"
  ) %>%
  cols_merge(
    contains("micro"),
    pattern = "{1} \u00B1 {2} %"
  ) %>%
  cols_merge(
    contains("nano"),
    pattern = "{1} \u00B1 {2} %"
  ) %>%
  cols_merge(
    contains("pico"),
    pattern = "{1} \u00B1 {2} %"
  ) %>%
  cols_label(
    season = "Season",
    contains("micro") ~ md("***f*<sub>micro</sub>** (>20 &mu;m)"),
    contains("nano") ~ md("***f*<sub>nano</sub>** (2 - 20 &mu;m)"),
    contains("pico") ~ md("***f*<sub>pico</sub>** (<2 &mu;m)")
  ) %>%
  tab_options(
    table_body.hlines.width      = 0,
    table_body.vlines.width      = 0,
    stub.border.width            = 0,
    row_group.as_column          = TRUE,
    row_group.padding.horizontal = 20,
    row_group.padding            = 200,
    stub_row_group.font.weight   = "bold",
    stub_row_group.border.width  = 0,
    row_group.border.top.width   = px(5)
  ) %>%
  cols_align("center") %>%
  tab_style(
    cell_text(
      v_align = "middle",
      align   = "center"
    ),
    locations = cells_row_groups()
  ) %>%
  opt_table_font("serif")

frac_gt

# save 2 tables as:
# - word docx (editable)
# - html (better formatted)
if (FALSE) {
  file_expr(
    loc            = dir_plt_save,
    file_base      = "all_fract_table",
    exts           = c("html", "docx"),
    time_stamp_fmt = "%Y%m%d_%H%M%S"
  ) %$%
    eval(file_expr) %>%
    walk(
      .,
      \(file_n) {
        gtsave(
          frac_gt,
          filename = file_n
        )
      }
    )
}
```