---
title: "Photosynthetic and Photoprotective Carotenoids"
author: "Sebastian Di Geronimo"
date: "2024-01-29"
format: html
editor: source
---

# 1.0 ---- Summary of Document ----

: photoprotective-index
  - calc ppc/tchlor-a
  - plot PI index against time and separate regions
  - boxplot of PI for seaon and separate regions with jitter axis
  
+ anova-pi-season-year
  - kruskal wallis and dunn test on:
    - PI per season
    - PI per year
  
: photoprotective-photosynthetic
  - plot PPC vs PSC and add linear regression for each season and region

# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
    # additional
    
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )
  
source(here("scripts", "misc_functions.R"))

# set size options of figures
knitr::opts_chunk$set(fig.width = 19, fig.height = 9.7)

# set ggplot theme
theme_set(theme_calc(base_size = 12, base_family = "serif"))
```

## 2.2 Load Dataset

Merged pigment data is created in `01_load_data.Rmd`
- "combined_pig_dat.csv"
- includes: 
  - metadata 
  - pigment concentrations
  - size class calculations

```{r load-data}
# set location to save data and plots
dir_data_save <- here("data", "processed", "03_3_ppc_psc")
dir_plt_save  <- here("data", "plots", "03_3_ppc_psc")

# set season names and order
szn <- c("Winter", "Spring", "Summer", "Autumn")

# load data
pig_dat <-
  here("data", "processed", "load_data", "combined_pig_dat.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(season = fct(season, levels = szn)) %T>%
  print()
```

# 4.0 ---- Photoprotective Index ----

## 4.1 PI Calculation

$$
\begin{align}
& PI~=~\frac{PPC}{\text{Tchl-a}} \\
& PPC~=~\text{photoproective carotenoids} \\
& \text{Tchl-a}~=~\text{Total chlorophyll-a (Chl-a + MV-Chl-a)}

\end{align}
$$
When PI < 1, means phyto are not being photoinhibited (?) because the total 
chlorophyll-a is higher, but if PI > 1, then photoinhibited


## 4.2 Scatter and Box Plots: PI:Chl-a Plots

- Scatter: Time vs PI 
- Box: Season vs PI and Season + Region vs PI

TODO: correlate temperature with PPC and maybe PI ratio?
```{r photoprotective-index}
shelf(ggbeeswarm)

pigm_pi <- mutate(pig_dat, pi = ppc/tot_chl_a)

plt_pi <- 
  ggplot(pigm_pi, 
    aes(x = date_time_utc, 
        y = pi, 
        color = season,
        shape = name_of_water_body)) +
  geom_point() +
  labs(color = "Season",
       shape = "Water Body",
       x = "Date",
       y = "Photoprotective Index \n\n(PPC/Total Chlorophyll-a)") +
# theme
    theme(
      axis.title.y     = element_markdown(),
      # legend.position  = c(0.6, 0.05),
      # legend.direction = "horizontal"
          )

plt_pi_all_water_body <- 
  plt_pi /
  (
    plt_pi + 
    facet_grid(name_of_water_body ~ .) +
    guides(color = "none",
           shape = "none")
   )


# with SeaBass calculation for PPC
plt_pi_box_base <- 
  ggplot(
    pigm_pi, 
    aes(
      y = pi,
      # fill = season
      )) +
    
    # geom_boxplot() +
    labs(
        x = "Season",
        y = "Photoprotection Index (PI)") +
    geom_hline(yintercept = 1, color = "gray", linetype = "dashed") +
    theme_classic(base_family = "serif", base_size = 12) +
    theme(legend.position = "none")

# top plot
pi_box_seasonal_water_body <- 
  (
   plt_pi_box_base +  
   geom_boxplot(aes( x = season, fill = season))
  ) /
  # bottom plot separated by water body
  (
    plt_pi_box_base + 
    geom_boxplot(
      aes(x    = name_of_water_body, 
          fill = name_of_water_body)) +
    geom_quasirandom(
      aes(x     = name_of_water_body, 
          fill  = name_of_water_body,
          color = as.factor(year))) +
    facet_grid( ~ season)
  )


# view plots
plt_pi_all_water_body
pi_box_seasonal_water_body


# save plots
# ratio vs time
save_gg(
  plt            = plt_pi_all_water_body,
  save_location  = dir_plt_save,
  save_name      = "pi_ppc_chl_all_water_body",
  verbose        = TRUE,
  overwrite      = FALSE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "jpeg",
  height         = 6,
  width          = 10,
  dpi            = 600,
  units          = "in",
  bg             = "white"
)

# boxplot ppc vs season
save_gg(
  plt            = pi_box_seasonal_water_body,
  save_location  = dir_plt_save,
  save_name      = "pi_box_ppc_chl_all_water_body",
  verbose        = TRUE,
  overwrite      = FALSE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "jpeg",
  height         = 6,
  width          = 10,
  dpi            = 600,
  units          = "in",
  bg             = "white"
)


```

## 4.3 ANOVA PI vs Season

- Kruskal-Wallis: 
    - non-parametric 
    - season and year
    
- Dunn's Test
    - pairwise test
    - bonferroni adjusted p-values
    
PI ~ Season
PI ~ Year

```{r anova-pi-season-year}
# function to run ANOVA on pi and arbitrary category
anova_func <- 
  list(
    kw = \(.df, category) { 
      formulas <- formula(paste("pi ~ ", category)) 
      kruskal_test(formulas, data = .df)
    },
    # Dunn's test - pairwise with adj
    dunn = \(.df, category) {
        formulas <- formula(paste("pi ~ ", category)) 
        dunn_test( 
          data = .df,
          formula = formulas,
          p.adjust.method = "bonferroni"
        ) %>%
          mutate(
            p = num(p, digits = 3),
            p.adj = num(p.adj, digits = 3)
          ) %>%
          arrange(desc(p.adj))
      }
)


# season factor
anova_func$kw(.df = pigm_pi, category = "season")
anova_func$dunn(.df = pigm_pi, category = "season")


# year factor
anova_func$kw(.df = pigm_pi, category = "year")
anova_func$dunn(.df = pigm_pi, category = "year")

# each region
pigm_anova <- 
  pigm_pi %>%
  nest(.by = name_of_water_body) %>%
  mutate(
    # by season
    season_kw =
      map2(
        data,
        "season",
        anova_func$kw
      ),
    season_dunn =
      map2(
        data,
        "season",
        anova_func$dunn
      ),
    
    # by year
    year_kw =
      map2(
        data,
        "year",
        anova_func$kw
      ),
    year_dunn =
      map2(
        data,
        "year",
        anova_func$dunn
      )
  )

# season
pigm_anova %>%
  unnest(
    c(season_kw, season_dunn),
    names_sep = "_"
  )

# year
pigm_anova %>%
  unnest(
    c(year_kw, year_dunn),
    names_sep = "_"
  )

```

## 4.4 Statsplot of ANOVA

Same as 5.3, but using `ggstatsplot` to show differences

- PI vs season
- PI vs year
- PI vs season + region
- PI vs year + region

```{r plt-anova}
shelf(ggstatsplot)

ggbetweenstats(
  data  = pigm_pi,
  x     = season, 
  y     = pi, 
  xlab = "Season",
  ylab = "PI",
  title = "Season: all regions",
  type  = "nonparametric"
  )

ggbetweenstats(
  data  = pigm_pi,
  x     = year, 
  y     = pi, 
  title = "Yearly: all regions",
  xlab = "Year",
  ylab = "PI",
  type  = "nonparametric"
  )


pigm_pi %>%
  nest(.by = name_of_water_body) %$%

  map2(
    name_of_water_body,
    data,
    (\(water_name, .df) {
      ggbetweenstats(
        data =  .df,
        x = season,
        y = pi,
        title = glue("Season: {water_name}"),
        xlab = "Season",
        ylab = "PI",
        type = "nonparametric"
      ) %>%
        print()

      ggbetweenstats(
        data = .df,
        x = year,
        y = pi,
        title = glue("Yearly: {water_name}"),
        xlab = "Year",
        ylab = "PI",
        type = "nonparametric"
      )
    })
  )

```


# 5.0 ---- PPC:PSC Ratio ----

## 5.1 Scatter Plots with Linear Regression

PPC vs PSC for region and season

Linear regression for year factor and overall regression for region and season

```{r photoprotective-photosynthetic, fig.height=15, fig.width=30}
# TODO: separate plots by region 
plt_ppc_psc <- 
  pigm_pi %>%
  filter(
    # !(str_detect(season, "Winter")
    # & str_detect(name_of_water_body, "Keys")
    # & ppc > 2)
      
    str_detect(name_of_water_body, "Keys") &
    ppc < 2
    ) %>%
  ggplot(
  aes(
    x = psc,
    y = ppc,
    # shape = season,
    # color = season
    # color = name_of_water_body
  )
) +
  geom_point(size = 4, alpha = 0.3, aes(color = as.factor(year))) +
  geom_abline(slope = 1, intercept = 0) +
  geom_smooth(
    aes(color = as.factor(year)),
    se        = FALSE,
    method    = "lm",
    formula   = y ~ x,
    linetype  = "dashed",
    linewidth = 1.5
  ) +
  geom_smooth(
    se        = FALSE,
    color     = "red",
    method    = "lm",
    formula   = y ~ x,
    linetype  = "dashed",
    linewidth = 2
  ) +
  labs(
    y = "PPC",
    x = "PSC",
    color = NULL
  ) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) +
  facet_wrap(name_of_water_body ~ season,
             scales = "free") +
  guides(color = guide_legend(nrow = 1, byrow = TRUE)) +
  theme(
    text = element_text(family = "serif", size = 20),
    strip.text.y = element_blank(),
    legend.position = "bottom"
  ) 

plt_ppc_psc

save_gg(
  plt            = plt_ppc_psc,
  save_location  = dir_plt_save,
  save_name      = "psc_vs_ppc",
  verbose        = TRUE,
  overwrite      = FALSE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "jpeg",
  height         = 20,
  width          = 30,
  dpi            = 600,
  units          = "in",
  bg             = "white"
 )

```


Autumn has the most scatter, but looks to have more PSC relating to more
production
Summer has areas with more PSC and PPC, but cannot distinquish as to where
Winter is also mostly PSC heavy, with a couple of points with high PPC
Spring is the season with the most PPC promience


