---
title: "Merge Metadata and CHEMTAX"
author: "Sebastian Di Geronimo"
date: "2024-01-30"
format: html
editor: source
---

# 1.0 ---- Summary of Document ----

+ setup 
  - load libraries and data

+ load-data
  - load data to be used in this script
  
: file-paths-chemtax-data
  - Search for CHEMTAX data, metadata, and create variables for phytoplankton 
    names and seasons
  
: load-chemtax-data
  - load CHEMTAX and meta data

: merge-meta-data
  - merge both with seasons

: save-merged-data
  - save CHEMTAX and metadata with/out duplicates
  - save phytoplankton names
  - save CHEMTAX PFT averages

# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
librarian::shelf(
  librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  forcats, lubridate, glue, fs, magrittr, here,
  
  # additional
  rlang, cli,
  rstatix, readxl
  )

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
    )

source(here("scripts", "misc_functions.R"))
```


# 3.0 Load Metadata and CHEMTAX Data and Setup Variables

Merged metadata is created in `01_load_data.Rmd`
- "combined_pig_dat.csv"

CHEMTAX data is created from CHEMTAX program
- uses: 
  - "chemtax_season_clust.csv" created in `01_load_data.Rmd` 
  - takes results from each cluster (1 - 4) 
  - creates one file per cluster as "1_mbonpig_cluster[1-4].xls"
  - manually merge into "1_mbonpig.xls"
- data includes:
  - Seasonal clusters for each run of CHEMTAX with tabs used in CHEMTAX program
    - cluster 1: Winter
    - cluster 2: Spring
    - cluster 3: Summer
    - cluster 4: Autumn
  - Manually merged data set (used in the rest of this project)

## 3.1 Search for File Paths and Setup Variables

```{r file-paths-chemtax-data, include=FALSE}
# ---- set save directory ---- #
dir_data_save <- here("data", "processed", "04_1_merge_meta_cmtx")

# ---- set season names and order ---- #
szn <- c("Winter", "Spring", "Summer", "Autumn")

# ---- phytoplankton full names ---- #
# NOTE: this is the first instance of phytoplankton names and will be saved for 
# preceding files
phyto_name <-
  tibble(
    names = c(
      "Chlorophyte", "Cryptophyte", "Cyanobacteria 2",
      "Cyanobacteria 4", "Diatom 1", "Diatom 2",
      "Dinoflagellate", "Haptophyte 6", "Haptophyte 8",
      "Prasinophyte", "Cyanobacteria", "Diatom", "Haptophyte"
    ),
    shrt = c(
      "chloro", "crypto", "cyano2", "cyano4", "diat1", "diat2", "dino", "hapt6",
      "hapt8", "pras", "cyano", "diat", "hapt"
    )
  ) %T>%
  print()

# ---- CHEMTAX file paths ---- #
file_chem <-
  here("data", "raw") %>%
  dir_ls(
    recurse = TRUE,
    regexp  = "mbonpig.*\\.xls$"
    ) %>% 
  str_sort() %>%
  tibble(files = .) %T>% print()
```


## 3.2 Load Metadata and CHEMTAX Data

```{r load-chemtax-data, include=FALSE}
# ---- load metadata ---- #
meta_data <-
  here("data", "processed", "load_data", "combined_pig_dat.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  select(hplc_gsfc_id:lat) %>%
  mutate(season = fct(season, szn)) %>%
  filter(
    str_detect(name_of_water_body, "Florida Keys")
    
    # sample was missed in original analysis
    & !str_detect(sample, "WS18008_43")
  ) %T>% 
  print()

# ---- load average seasonal CHEMTAX data ---- #
# *not used beyond first plot*
# average each season
cmtx_all <-
  file_chem %>%
  filter(str_detect(files, "cluster")) %>%
  mutate(
    filename = basename(files),
    filename = tools::file_path_sans_ext(filename),
    ranges   = c("D177:N178", "D273:N274", "D241:N242", "D305:N306"),
    cluster  = str_extract(filename, "\\d$"),
    season   = fct_inorder(szn),
    data     = map2(
      files, ranges,
      ~ read_xls(
        path  = .x,
        sheet = "Notes",
        range = .y,
        .name_repair = janitor::make_clean_names
      )
    )
  ) %>%
  select(-c(1:ranges)) %>%
  unnest(data) %T>% 
  print()


# ---- load merged cluster CHEMTAX data ---- #
cmtx_raw <-
  pull(file_chem, files) %>%
  # remove files with cluster in the name
  str_subset("cluster", negate = TRUE) %>%
  readxl::read_xls(
    .,
    sheet = "Avgs",
    .name_repair = janitor::make_clean_names
  ) %>%
  filter(!str_detect(sample_id, "WS19210-59")) %T>% 
  print()
```

## 3.3 Merge CHEMTAX Results with Metadata

```{r merge-meta-data}
# ---- join metadata and chemtax data
cmtx_all_w <-
  full_join(
    meta_data,
    cmtx_raw,
    by = join_by(
        "sample" == "sample_id",
        "hplc_gsfc_id" == "sample"
        ),
  ) %T>% print()

# ---- view duplicates  ---- #
# 2 reason:
# - 1. duplicate sample taken in field
#   - this is signified with the same location and is usually +1 of sample ID
# - 2. internal consistency at NASA where they use subset of sample 
#   - this is signified with a ###.5
janitor::get_dupes(cmtx_all_w, cruise_id, date_time_utc, station, depth) 

# removing duplicates by averaging
cmtx_w <-
  cmtx_all_w %>%
  reframe(
    .by = c(cruise_id, date_time_utc, station, depth),
    pick(1:cluster_code),
    across(
      chloro:pras,
      \(x) mean(x, na.rm = TRUE)
    )
  ) %>%
  distinct(cruise_id, date_time_utc, station, depth, .keep_all = TRUE)  %>%
  mutate(station = str_replace(station, "9B", "9.5")) %T>%
  # filter(!str_detect(station, "60|21.5|24|^3$|9B|9.5")) %T>% 
  print()
```

## 3.4 Save: CHEMTAX Data

```{r save-merged-data}
# ---- save chemtax data with metadata ---- #
if (
  !file_exists(here(dir_data_save, "chemtax_w_meta_all_wide.csv")) |
    !file_exists(here(dir_data_save, "chemtax_w_meta_wide.csv"))
) {
  # wide format of CHEMTAX and metadata with duplicates
  save_csv(
    .data          = cmtx_all_w,
    save_location  = dir_data_save,
    save_name      = "chemtax_w_meta_all_wide",
    overwrite      = FALSE,
    verbose        = TRUE,
    time_stamp_fmt = NULL
  )

  # wide format of CHEMTAX and metadata without duplicates
  save_csv(
    .data          = cmtx_w,
    save_location  = dir_data_save,
    save_name      = "chemtax_w_meta_wide",
    overwrite      = FALSE,
    verbose        = TRUE,
    time_stamp_fmt = NULL
  )
} else {
  cli::cli_alert_info(
    c(
      "Skipping saving of {.file chemtax_w_meta} ",
      "because exists ",
      "and/or overwrite is set {.var FALSE}"
    )
  )
}

# ---- save average seasonal CHEMTAX results ---- #
if (!file_exists(here(dir_data_save, "chemtax_avg_seasonal.csv"))) {
  save_csv(
    .data          = cmtx_all,
    save_location  = dir_data_save,
    save_name      = "chemtax_avg_seasonal",
    overwrite      = FALSE,
    verbose        = TRUE,
    time_stamp_fmt = NULL
  )
} else {
  cli::cli_alert_info(
    c(
      "Skipping saving of {.file chemtax_avg_seasonal.csv} ",
      "because exists ",
      "and/or overwrite is set {.var FALSE}"
    )
  )
}

# ---- save phytplankton names ---- #
if (!file_exists(here(dir_data_save, "phyto_names.csv"))) {
  save_csv(
    .data          = phyto_name,
    save_location  = dir_data_save,
    save_name      = "phyto_names",
    overwrite      = FALSE,
    verbose        = TRUE,
    time_stamp_fmt = NULL
  )
} else {
  cli::cli_alert_info(
    c(
      "Skipping saving of {.file phyto_names.csv} ",
      "because exists ",
      "and/or overwrite is set {.var FALSE}"
    )
  )
}
```



