---
title: "Plot for Seasonal CHEMTAX Averages"
author: "Sebastian Di Geronimo"
date: "2024-01-30"
format: html
editor: source
---

# 1.0 ---- Summary of Document ----

NOTE: not very useful for analysis, but a good exploratory start

+ setup 
  - load libraries and data

+ functions-line-box
  - create function to plot line graphs

+ load-data
  - load CHEMTAX data 
  
: line-plt
  - create line plot for average concentration per PFT
  
  
# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
  # additional
  rlang, cli,
  rstatix, readxl, ggtext, cowplot
  )

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
    )

source(here("scripts", "misc_functions.R"))

# set colors
# alternate color palette
# tol_muted <- 
# c('#88CCEE', '#44AA99', '#117733', '#332288', '#DDCC77', '#999933',
#   '#CC6677', '#882255', '#AA4499')

tol_muted <- khroma::color("muted", names = FALSE)(9)

```

## 2.2 Line and Box Plot Plotting Functions 

Default Line Plot and Box Plots

```{r functions-line-box}
# ---- line plot ---- #
#' Line Plot
#'
#' FUNCTION_DESCRIPTION
#'
#' @param .dat DESCRIPTION.
#' @param .var DESCRIPTION.
#' @param .title DESCRIPTION.
#'
#' @return RETURN_DESCRIPTION
#' @examples
#' # ADD_EXAMPLES_HERE
line_plt <- function(.dat, .var, .title) {
    
    # set limits
    ylims <- c(0, max(.dat[[.var]]) * 1.05)
    brake <- pretty(ylims, 5)
    
    # plot
    .dat %>%
    ggplot(aes(x = season, y = .data[[.var]])) +
    geom_point(size = 2) +
    geom_line(group = 1, 
              linewidth  = 1) +
    
    # labels
    labs(
     title = glue("Average Concentration per Season of {.title}"),
     x     = "Season",
     y     = "Concentration of Chlorophyll-a (mg m<sup>-3</sup>)"
     ) +
     
    # scale setup   
    scale_y_continuous(
      breaks = brake,
      limits = ylims,
      expand = expansion(mult = 0)
      ) +
    
    # theme setup   
    theme_bw(base_family = "serif") +
    theme(
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      plot.title         = element_text(size = 15),
      axis.ticks.length  = unit(-0.05, "in"),
      axis.text.y        = element_text(margin = unit(c(0.3,0.3,0.3,0.3), "cm")),
      axis.title.y       = element_markdown(size = 10),
      axis.text.x        = element_text(margin = unit(c(0.3,0.3,0.3,0.3), "cm")),
      axis.ticks.x       = element_blank(),
      legend.background  = element_rect(color = "black", fill = "white")
    )
}
```


# 3.0 ---- Load Dataset ----

```{r load-data}
# ---- set location to save data and plots ---- #
dir_data_save <- here("data", "processed", "04_2_cmtx_seasonal_avg")
dir_plt_save  <- here("data", "plots", "04_2_cmtx_seasonal_avg")

dir_create(here(dir_plt_save))

# ---- set season names and order ---- #
szn <- c("Winter", "Spring", "Summer", "Autumn")

# ---- phytoplankton full names ---- #
phyto_name <-
  here("data", "processed", "04_1_merge_meta_cmtx", "phyto_names.csv") %>%
  read_csv(show_col_types = FALSE)  %T>% 
  print()

# ---- load chemtax data in wide format ---- #
cmtx_all <-
  here("data", "processed", "04_1_merge_meta_cmtx", "chemtax_avg_seasonal.csv") %>%
  read_csv(show_col_types = FALSE) %T>% 
  print()
```




# 4.0 ---- Line Plot: Average Seasonal PFT vs Season ----

Not super useful

Average PFT in concentration of total chlorophyll-a (mg m^-3)

```{r line-plt}
cmtx_all %>%
  select(-rms) %>%
  pivot_longer(
    cols      = -c(1:2),
    names_to  = "phyto",
    values_to = "avg_conc"
  ) %>%
  group_by(phyto) %>%
  nest() %>%
  # merge PFT shortname with longname
  left_join(phyto_name, by = c("phyto" = "shrt")) %$%
    
  # plot average concentration per season
  walk2(
    data, 
    names, 
    ~ line_plt(
      .dat = .x,
      .var = "avg_conc",
      .title = .y
    ) %>%
      print()
  )

```