---
title: "Seasonal CHEMTAX Box Plots"
author: "Sebastian Di Geronimo"
date: "2024-01-30"
format: html
editor: source
---

# 1.0 ---- Summary of Document ----



# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
  # additional
  rlang, cli,
  rstatix, readxl, ggtext, cowplot
  )

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
    )

source(here("scripts", "misc_functions.R"))
```

## 2.2 Load Dataset

```{r load-data}
# ---- set location to save data and plots ---- #
dir_data_save <- here("data", "processed", "04_3_cmtx_box_plots")
dir_plt_save  <- here("data", "plots", "04_3_cmtx_box_plots")

dir_create(here(dir_plt_save))

# ---- set season names and order ---- #
szn <- c("Winter", "Spring", "Summer", "Autumn")

# ---- phytoplankton full names ---- #
phyto_name <-
  here("data", "processed", "04_1_merge_meta_cmtx", "phyto_names.csv") %>%
  read_csv(show_col_types = FALSE) %T>% 
  print()

# ---- load chemtax data in wide format ---- #
cmtx_w <-
  here("data", "processed", "04_1_merge_meta_cmtx", "chemtax_w_meta_wide.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  # filter(!str_detect(station, "60|21.5|24|^3$|9B|9.5")) %>%
  filter(!str_detect(station, "60|21.5|24|^3$|9.5")) %>%
  mutate(
    season = fct(season, levels = szn),
    # station = str_replace(station, "9B", "9.5"),
    station = fct_reorder(station, lon)
  ) %>% 
  select(-c(hplc_gsfc_id, cruise:bottle, cluster_code, water_type)) %>%
    
  # add PFT concentrations to get total chlorophyll-a
  rowwise() %>%
  mutate(
    total_chl_a = sum(c_across(chloro:pras)),
    .before = chloro
  ) %>%
  ungroup() %>%
  filter(
    .by = c(station, cruise_id),
    depth == min(depth)
    & depth < 20
  ) %T>% 
  print()

```


## 2.3 Convert to Long Format with Percent


```{r convert-long-format}
# ---- convert to long format ---- #
# i.e. one column for all PFT and one for concentrations
cmtx_l <-
  cmtx_w %>%
  pivot_longer(
    cols      = chloro:pras,
    names_to  = "phyto",
    values_to = "conc"
  ) %>%
  left_join(phyto_name, by = c("phyto" = "shrt")) %>%
  
  # add percent
  mutate(percent = conc/total_chl_a) %T>% 
  print()

# ---- convert to percent ---- #
cmtx_l_pct <-
  cmtx_l %>%
  select(-conc) %T>% 
  print()
```


## 2.4 Line and Box Plot Plotting Functions 

Default Line Plot and Box Plots

```{r functions-line-box}
#' Box Plot
#'
#' FUNCTION_DESCRIPTION
#'
#' @param .dat DESCRIPTION.
#' @param .var DESCRIPTION.
#' @param .title DESCRIPTION.
#'
#' @return RETURN_DESCRIPTION
#' @examples
#' # ADD_EXAMPLES_HERE
# box_plt <- function(.dat, .var, .color, .shape, .title) {
box_plt <- function(.dat, .var, .title, ...) {
    
    # limits for plot
    ylims <-  c(0, quantile(.dat[[.var]], 0.99))
    brake <- pretty(ylims, 6)

    # plotting
    .dat %>%
    ggplot() +
    geom_boxplot(
      aes(
        x = season, 
        y = .data[[.var]]
        ),
      outlier.shape = NA) +
    geom_jitter(
        aes(
          x = season, 
          y = .data[[.var]], 
          ...
          # color = .data[[.color]],
          # shape = .data[[.shape]]
          ),
        width = 0.4,
        size = 0.5
        ) +
    
    # labels
    labs(
      title = glue("Concentration of {.title} per Season "),
      x     = "Season",
      y     = "Concentration (mg m<sup>-3</sup>)"
      # color = "Year"
      ) +

    # scale settings
    scale_color_discrete(type = "viridis") +
    
    scale_shape_discrete() +
        
    scale_y_continuous(
      breaks = brake,
      expand = expansion(mult = 0),
      limits = ylims,
      oob    = scales::squish
      ) +
    
    # theme settings
    theme_bw(base_family = "serif", base_size = 12) +
    theme(
      panel.grid        = element_blank(),
      plot.title        = element_text(size = 15),
      axis.ticks.length = unit(-0.05, "in"),
      axis.text.y       = element_text(margin = unit(c(0.3,0.3,0.3,0.3), "cm")),
      axis.title.y      = element_markdown(size = 10),
      axis.text.x       = element_text(margin = unit(c(0.3,0.3,0.3,0.3), "cm")),
      axis.ticks.x      = element_blank(),
      legend.background = element_rect(color = "black", fill = "white")
    )
}
```


# 4.0 ---- Box Plots ----

## 4.2 Box Plot: Individuals for Each PFT per Season

```{r boxplots, fig.width=10, fig.asp=0.8}
# ---- avg season line plots
cmtx_avgs_plt <-
  cmtx_l %>%
  nest(.by = c(phyto, names)) %>%
  mutate(
    plt = map2(
      data,
      names,
      ~ .x %>%
        mutate(Year = factor(year)) %>%
        box_plt(
        .dat   = .,
        .var   = "conc",
        color = Year,
        shape = Year,
        .title = .y
      )
    )
  )

# cmtx_avgs_plt$plt[[1]]
cmtx_avgs_plt$plt
```

## 4.3 Box Plot: Base for Larger/Small Dynamic Range PFTs

```{r box-base, fig.height=15, fig.width=20}
box_filter <- 
  . %>%
  filter(.by = c(station, names), n() > 3) %>%
   mutate(
    .by       = c(station, names),
    .after    = station,
    stn_label = glue("{station} (N = {n()})")
   ) %>%
  mutate(
    stn_label = fct_reorder(stn_label, lon)
  )

box_plt_base <-
  list(
    aes(
      x    = season,
      y    = conc,
      fill = season
    ),
    geom_boxplot(
        linewidth = 0.15,
        outlier.shape = 20, 
        outlier.size = 0.15
        ),
    scale_y_continuous(
      expand = c(0, 0),
      breaks = scales::breaks_width(0.1),
      minor_breaks = NULL,
      oob    = scales::squish,
      # labels = scales::label_percent(accuracy = 1)
      labels = scales::label_number(accuracy = 0.01)
    ),
    labs(
      # x = "Season",
      x = NULL,
      y = "Chlorophyll-a Concentration (mg m<sup>-3</sup>)"
    ),
    # theme_classic(),
    theme_bw(),
    theme(
      text               = element_text(family = "serif", size = 4, color = "black"),
      line               = element_line(color = "black", linewidth = 0.2),
      rect               = element_rect(color = "black", linewidth = 0.2),
      axis.text          = element_markdown(size = 4),
      axis.title.y       = element_markdown(size = 4),
      legend.position    = "none",
      strip.text.x       = element_text(size = 5, margin = margin(t = 2, b = 2)),
      strip.background   = element_rect(fill = "white", colour = "grey20"),
      panel.grid.major.x = element_blank(),
      panel.grid.major.y = element_line(color = "grey20", linewidth = 0.25),
    )
  )
```

## 4.4 Box Plot: Larger Dynamic Range PFTs

```{r box-lg-dynm, fig.height=15, fig.width=20}
# box plot by season (1 x 4)
box_lg <-
  cmtx_l %>%
  filter(phyto %in% c("cyano2", "cyano4", "hapt6", "hapt8")) %>%
  ggplot() +
  box_plt_base +
  facet_wrap(facet = ~names, ncol = length(unique(cmtx_l$phyto)) - 6) +
  coord_cartesian(ylim = c(0, 0.45), clip = "off")

# box plot by season and station
box_lg_stn <-
  cmtx_l %>%
  filter(phyto %in% c("cyano2", "cyano4", "hapt6", "hapt8")) %>%
  box_filter %>%
  ggplot() +
  box_plt_base +
  # facet_wrap(facet = ~ names + stn_label, ncol = 8) +
  facet_grid(rows = stn_label ~ names) +
  coord_cartesian(ylim = c(0, 0.4))
  # coord_cartesian(ylim = c(0, 1))

# stack box plots  
box_lg_stcked <- 
  patchwork::wrap_plots(
    box_lg, 
    box_lg_stn, 
    ncol = 1, 
    heights = c(0.2, 0.8)
  )

# display plots
box_lg
box_lg_stn
box_lg_stcked

# box plot by season (2 x 2)
box_lg <-
  cmtx_l %>%
  filter(phyto %in% c("cyano2", "cyano4", "hapt6", "hapt8")) %>%
  ggplot() +
  box_plt_base +
  facet_wrap(facet = ~names) +
  coord_cartesian(ylim = c(0, 0.45), clip = "off") +
  scale_y_continuous(
      limits = c(0, 0.45), 
      expand = c(0, 0),
      breaks = scales::breaks_width(0.1),
      minor_breaks = NULL,
      oob    = scales::squish,
      # labels = scales::label_percent(accuracy = 1)
      labels = scales::label_number(accuracy = 0.01))

box_lg
```


## 4.5 Box Plot: Smaller Dynamic Ranged PFTs


```{r box-sm-dynm, fig.height=15, fig.width=20}
# box plot by season (1 x 6)
box_sm <-
  cmtx_l %>%
  filter(phyto %in% c("chloro", "crypto", "diat1", "diat2", "dino", "pras")) %>%
  ggplot() + 
  box_plt_base +
  facet_wrap(facet = ~names, ncol = length(unique(cmtx_l$phyto)) - 4) +
  coord_cartesian(ylim = c(0, 0.2))
 
  
# box plot by season and station
box_sm_stn <-
  cmtx_l %>%
  filter(phyto %in% c("chloro", "crypto", "diat1", "diat2", "dino", "pras")) %>%
  box_filter %>%
  ggplot() +
  box_plt_base +
  # facet_wrap(facet = ~ names + stn_label, ncol = 8) +
  facet_grid(rows = stn_label ~ names) +
  coord_cartesian(ylim = c(0, 0.2))

# stack box plots
box_sm_stcked <-
  patchwork::wrap_plots(
    box_sm,
    box_sm_stn,
    ncol = 1,
    heights = c(0.25, 0.75)
  )

# display plots
box_sm
box_sm_stn
box_sm_stcked

# box plot by season (2 x 3)
box_sm <-
  cmtx_l %>%
  filter(phyto %in% c("chloro", "crypto", "diat1", "diat2", "dino", "pras")) %>%
  ggplot() + 
  box_plt_base +
  facet_wrap(facet = ~names) +
  coord_cartesian(ylim = c(0, 0.2), clip = "off") +
  scale_y_continuous(
    limits = c(0, 0.2),
    expand = c(0, 0),
    breaks = scales::breaks_width(0.05),
    minor_breaks = NULL,
    oob    = scales::squish,
    # labels = scales::label_percent(accuracy = 1)
    labels = scales::label_number(accuracy = 0.01)
  )

box_sm
```



```{r box-lg-dynm-percent, fig.height=15, fig.width=20}
# box plot by station
box_lg_pct <-
  cmtx_l_pct %>%
  filter(phyto %in% c("cyano2", "cyano4", "hapt6", "hapt8")) %>%
  ggplot() +
  box_plt_base +
  facet_wrap(facet = ~names, ncol = length(unique(cmtx_l$phyto)) - 6) +
  labs(y = "Relative Abundance (%)") +
  scale_y_continuous(labels = scales::label_percent(), expand = c(0,0)) +
  coord_cartesian(ylim = c(0, 1))

# box plot by season and station
box_lg_stn_pct <-
  cmtx_l_pct %>%
  filter(phyto %in% c("cyano2", "cyano4", "hapt6", "hapt8")) %>%
  box_filter %>%
  ggplot() +
  box_plt_base +
  # facet_wrap(facet = ~ names + stn_label, ncol = 8) +
  facet_grid(rows = stn_label ~ names) +
  scale_y_continuous(labels = scales::label_percent(), expand = c(0,0)) +
  coord_cartesian(ylim = c(0, 1))

# stack box plots  
box_lg_stcked_pct <- 
  patchwork::wrap_plots(
    box_lg_pct, 
    box_lg_stn_pct, 
    ncol = 1, 
    heights = c(0.2, 0.8)
  )

# display plots
box_lg_pct
box_lg_stn_pct
box_lg_stcked_pct


# box plot by season (2 x 2)
box_lg_pct <-
  cmtx_l_pct %>%
  filter(phyto %in% c("cyano2", "cyano4", "hapt6", "hapt8")) %>%
  ggplot() +
  box_plt_base +
  # facet_wrap(facet = ~names, ncol = length(unique(cmtx_l$phyto)) - 6) +
  facet_wrap(facet = ~names) +
  labs(y = "Relative Abundance (%)") +
  scale_y_continuous(labels = scales::label_percent(), expand = c(0,0)) +
  coord_cartesian(ylim = c(0, 1), clip = "off")

box_lg_pct
```


```{r box-sm-dynm-pct, fig.height=15, fig.width=20}
# box plot by season (1 x 6)
box_sm_pct <-
  cmtx_l_pct %>%
  filter(phyto %in% c("chloro", "crypto", "diat1", "diat2", "dino", "pras")) %>%
  ggplot() + 
  box_plt_base +
  facet_wrap(facet = ~names, ncol = length(unique(cmtx_l$phyto)) - 4) +
  labs(y = "Relative Abundance (%)") +
  scale_y_continuous(labels = scales::label_percent(), expand = c(0,0)) +
  coord_cartesian(ylim = c(0, 1))
 
  
# box plot by season and station
box_sm_stn_pct <-
  cmtx_l_pct %>%
  filter(phyto %in% c("chloro", "crypto", "diat1", "diat2", "dino", "pras")) %>%
  box_filter %>%
  ggplot() +
  box_plt_base +
  # facet_wrap(facet = ~ names + stn_label, ncol = 8) +
  facet_grid(rows = stn_label ~ names) +
  labs(y = "Relative Abundance (%)") +
  scale_y_continuous(labels = scales::label_percent(), expand = c(0,0)) +
  coord_cartesian(ylim = c(0, 1))

# stack box plots
box_sm_stcked_pct <-
  patchwork::wrap_plots(
    box_sm_pct,
    box_sm_stn_pct,
    ncol = 1,
    heights = c(0.25, 0.75)
  )

# display plots
box_sm_pct
box_sm_stn_pct
box_sm_stcked_pct

# box plot by season (2 x 3)
box_sm_pct <-
  cmtx_l_pct %>%
  filter(phyto %in% c("chloro", "crypto", "diat1", "diat2", "dino", "pras")) %>%
  ggplot() + 
  box_plt_base +
  # facet_wrap(facet = ~names, ncol = length(unique(cmtx_l$phyto)) - 4) +
  facet_wrap(facet = ~names) +
  labs(y = "Relative Abundance (%)") +
  scale_y_continuous(labels = scales::label_percent(), expand = c(0,0)) +
  coord_cartesian(ylim = c(0, 1), clip = "off")

box_sm_pct
```

## 4.6 Save: Large and Small Range Boxplots

```{r save-plt}
# save box_plots
walk2(
  verbose   = TRUE,  # set to FALSE to suppress output
  # overwrite = FALSE, # set to TRUE to overwrite existing files
  overwrite = TRUE, # set to TRUE to overwrite existing files
  .x = list2(
    "cmtx_boxplot_lg",
    "cmtx_boxplot_sm",
    "cmtx_boxplot_lg_stn",
    "cmtx_boxplot_sm_stn",
    "cmtx_boxplot_lg_stcked",
    "cmtx_boxplot_sm_stcked",
    "cmtx_boxplot_lg_pct",
    "cmtx_boxplot_sm_pct",
    "cmtx_boxplot_lg_stn_pct",
    "cmtx_boxplot_sm_stn_pct",
    "cmtx_boxplot_lg_stcked_pct",
    "cmtx_boxplot_sm_stcked_pct"
  ),
  .y = list2(
      box_lg, box_sm,
      box_lg_stn, box_sm_stn,
      box_lg_stcked, box_sm_stcked,
      box_lg_pct, box_sm_pct,
      box_lg_stn_pct, box_sm_stn_pct,
      box_lg_stcked_pct, box_sm_stcked_pct
      ),
  \(.x, .y, ...) {
    .x %>%
      str_to_lower() %>%
      str_replace_all("\\s", "_") %>%
      # glue(., "common suffix", .sep = "_") %>% # remove if none needed
      save_gg(
        .y,
        save_location  = here(dir_plt_save, Sys.Date()),
        save_name      = .,
        # time_stamp_fmt = "%Y%m%d_%H%M%S",
        time_stamp_fmt = NULL,
        device         = "jpeg",
        width          = 90,
        height         = 90 * 0.5625,
        dpi            = 600,
        units          = "mm",
        bg             = "white",
        ...
      )
  }
)

```
# ------------------------------ DELETE BELOW ---------------------------------


# 6.0 ---- Combine PFT Sub Groups ----

## 6.1 Add: PFT Groups Together

Groups to be combined have multiple types but are merged for ease of use
- diatom:
  - diatom 1 
  - diatom 2
- cyanobacteria:
  - cyano 2
  - cyano 4
- haptophyte:
  - hapt 6
  - hapt 8

```{r station-seasonally-yearly-avgs}
# # combine same base PFT groups into main group
# cmtx_w_comb <-
#   cmtx_w %>%
#   rowwise() %>%
#   mutate(
#     # total = sum(c_across(chloro:pras), na.rm = TRUE),
#     diat  = diat1 + diat2,
#     cyano = cyano2 + cyano4,
#     hapt  = hapt6 + hapt8
#   ) %>%
#   ungroup() %>%
#   # relocate(total, .before = chloro) %T>% 
#   print()
# 
# # calculate PFT fractions (0 - 100%)
# cmtx_fract_w <-
#   # cmtx_w_comb %>%
#   cmtx_w %>%
#   mutate(
#     .keep = "unused",
#     # across(c(1:depth, water_depth:lat)),
#     across(
#       # c(pras:hapt, chloro, crypto, dino),
#       chloro:last_col(),
#       (\(x) x / total_chl_a),
#       .names = "frac_{.col}"
#     )
#   ) %T>% print()
# 
# # convert to long format
# cmtx_fract_l <- 
#   cmtx_fract_w %>%
#   pivot_longer(
#     contains("frac"),
#     names_to     = "phyto",
#     values_to    = "frac",
#     names_prefix = "frac_"
#   ) %>%
#     
#   # add long names
#   left_join(phyto_name, by = c("phyto" = "shrt")) %T>% print()
```


## 6.2 View Averages for Both Seasonal and per Station

```{r view-avgs}
# format_pct <- 
#   . %>%
#   mutate(
#     percent_avg = scales::label_percent(accuracy  = 0.01)(frac_avg)
#   )
# 
# # ---- average by season ---- #
# # only one used further
# cmtx_avg_season_fract_l <-
#   cmtx_fract_l %>%
#   summarise(
#     .by = c(names, phyto, season),
#     across(
#       .cols  = frac,
#       .fns   = list(
#           avg = (\(x) mean(x, na.rm = TRUE)),
#           sd  = (\(x) sd(x, na.rm = TRUE))
#           ),
#       .names = "{.col}_{.fn}"
#     )
#   ) %>%
#   mutate(
#     percent_avg = scales::label_percent(accuracy  = 0.01)(frac_avg),
#     percent_sd  = scales::label_percent(accuracy  = 0.01)(frac_sd)
#   ) %>%
#  arrange(season) %T>% print()
# 
# 
# # ---- average by station ---- #
# cmtx_fract_l %>%
#   mutate(station = fct_reorder(station, lon)) %>%
#   arrange(station) %>%
#   summarise(
#     .by = c(names, phyto, station),
#     across(
#       .cols  = frac,
#       .fns   = (\(x) mean(x, na.rm = TRUE)),
#       .names = "{.col}_avg"
#     )
#   ) %>%
#   mutate(
#     percent_avg = scales::label_percent(accuracy  = 0.01)(frac_avg)
#   ) %>%
#  arrange(station) %T>% print()
# 
# # ---- average by station and season ---- #
# cmtx_fract_l %>%
#   mutate(station = fct_reorder(station, lon)) %>%
#   arrange(station) %>%
#   summarise(
#     # .by = c(cruise_id, station, season, names, phyto), # i think this is wrong, remove cruise ID
#     .by = c(station, season, names, phyto),
#     across(
#       .cols  = frac,
#       .fns   = (\(x) mean(x, na.rm = TRUE)),
#       .names = "{.col}_avg"
#     )
#   )  %>%
#   mutate(
#     percent_avg = scales::label_percent(accuracy  = 0.01)(frac_avg)
#   ) %>%
#  arrange(season) %T>% print()
```

## 6.4  Plot: Seasonal Fraction of each PFT (Bar)

Stacked Bar Chart
```{r plt-fracs, fig.width=20, fig.height=10}
# tol_muted <- c(khroma::color("muted", names = FALSE)(9), "#BBBBBB")
# 
# new_dat <- 
#   summarise(
#     cmtx_w, 
#     .by = season, 
#     chl_avg = mean(total_chl_a), 
#     chl_sd  = sd(total_chl_a),
#     chl_se  = chl_sd / sqrt(n())
#     ) %T>% 
#   print() 
# 
# sec_axe <- 
#   new_dat %>%
#   ggh4x::help_secondary(
#     secondary = c(0, 0.75), 
#     name = bquote("Concentration"~(mg~L^-1))
#     )
# 
# 
# 
# cmtx_avg_season_fract_l %>%
#   mutate(names = fct_reorder(names, frac_avg)) %>%
#   ggplot(
#     aes(
#       fill  = names,
#       # color = names,
#       y     = frac_avg,
#       x     = season
#     )
#   ) +
#   geom_bar(
#     color = "black",
#     position = "fill",
#     stat     = "identity"
#   )  +    
#   geom_line(
#     data = new_dat,
#     aes(x = season, y = sec_axe$proj(chl_avg), group = 1),
#     inherit.aes = FALSE,
#     linewidth = 2
#     ) +
#   geom_point(
#     data = new_dat,
#     aes(x = season, y = sec_axe$proj(chl_avg)),
#     inherit.aes = FALSE,
#     size = 5
#     ) +
#   geom_linerange(
#     data = new_dat,
#     aes(
#       x = season,
#       ymin = sec_axe$proj(chl_avg) - sec_axe$proj(chl_se),
#       ymax = sec_axe$proj(chl_avg) + sec_axe$proj(chl_se)),
#     inherit.aes = FALSE,
#   ) +
#   labs(
#     x     = "Season",
#     # y     = "Proportion to Total Chl-a",
#     y     = "PFT:Total Chlorophyll-a (%)",
#     fill  = "Phytoplankton Functional Type",
#     color = "Phytoplankton Functional Type"
#   ) +
#   scale_fill_manual(values  = tol_muted) +
#   scale_color_manual(values = tol_muted) +
#   scale_y_continuous(
#     labels = scales::percent_format(accuracy = 1),
#     expand = c(0,0),
#     limits = c(0,1),
#     oob    = scales::squish,
#     sec.axis = sec_axe
#     ) +
#   # theme
#   theme_bw() +
#   theme(
#     text             = element_text(family = "serif", size = 20, color = "black"),
#     axis.title       = element_text(size = 20),
#     legend.position  = "bottom",
#     legend.direction = "horizontal",
#     plot.margin = margin(15,10,10,10, "pt")
#   )
# 
# # ---- save plot
# save_gg(
#   plt            = last_plot(), # edit if want specific plot
#   save_location  = dir_plt_save,
#   save_name      = "pft_season_avg_prcnt_stack",
#   verbose        = TRUE,
#   # overwrite      = FALSE,
#   overwrite      = TRUE,
#   time_stamp_fmt = "%Y%m%d_%H%M%S",
#   device         = "jpeg",
#   height         = 8,
#   width          = 16,
#   dpi            = 600,
#   units          = "in",
#   bg             = "white"
#   )
```

Separated Bar Chart

```{r plt-fracs, fig.width=20, fig.height=10}
# cmtx_avg_season_fract_l %>%
#   mutate(names = fct_reorder(names, frac_avg)) %>%
#   ggplot(
#     aes(
#       fill  = season,
#       # color = season,
#       y     = frac_avg,
#       x     = fct_reorder(names, frac_avg, .desc = TRUE)
#     )
#   ) +
#   geom_bar(
#     color = "black",
#     # position = "fill",
#     position = position_dodge(),
#     stat     = "identity"
#   ) +
#   labs(
#     x     = "Season",
#     # y     = "Proportion to Total Chl-a",
#     y     = "PFT:Total Chlorophyll-a (%)",
#     fill  = "Phytoplankton Functional Type",
#     color = "Phytoplankton Functional Type"
#   ) +
#   scale_fill_manual(values  = tol_muted) +
#   scale_color_manual(values = tol_muted) +
#   scale_y_continuous(
#     labels = scales::percent_format(accuracy = 1),
#     expand = c(0,0),
#     limits = c(0,1)
#     ) +
#   # theme
#   theme_classic() +
#   theme(
#     text             = element_text(family = "serif", size = 20),
#     axis.title       = element_text(size = 20),
#     legend.position  = "bottom",
#     legend.direction = "horizontal",
#     plot.margin = margin(15,10,10,10, "pt")
#   )
# 
# # ---- save plot ---- #
# save_gg(
#   plt            = last_plot(), # edit if want specific plot
#   save_location  = dir_plt_save,
#   save_name      = "pft_season_avg_prcnt_sep",
#   verbose        = TRUE,
#   overwrite      = FALSE,
#   # overwrite      = TRUE,
#   time_stamp_fmt = "%Y%m%d_%H%M%S",
#   device         = "jpeg",
#   height         = 8,
#   width          = 16,
#   dpi            = 600,
#   units          = "in",
#   bg             = "white"
#   )
 
```

## 3.2 Bar Plot: PFT per **Season** and **Station** as Stacked Percent

```{r chemtax-stacked-bar}
# # CHEMTAX stacked bar graph
# cmtx_l %>%
#   filter(!is.na(conc)) %>%
#   summarise(
#   # .by = c(station, season, func_type, names),
#   .by = c(station, season, phyto, names),
#   conc = mean(conc)
#   ) %>%
#   mutate(
#     names = fct_reorder(names, conc, .desc = TRUE)
#   ) %T>% 
#   print() %>%
#   ggplot(
#     aes(
#       x     = station,
#       y     = conc,
#       fill  = names,
#       # color = names
#     )
#   ) +
#   geom_bar(
#     color = "black",
#     position = "fill",
#     stat = "identity",
#     na.rm = TRUE,
#     linewidth = 0.5
#   ) +
#   labs(
#     x     = "Station",
#     y     = "Percent (%)",
#     fill  = "Phytoplankton Functional Type",
#     color = "Phytoplankton Functional Type"
#   ) +
#   scale_y_continuous(
#     labels = scales::label_percent(),
#     expand = c(0, 0)
#   ) +
#   scale_fill_manual(values = tol_muted) +
#   facet_wrap(~season)
# 
# # ---- save plot
# save_gg(
#   plt            = last_plot(), # edit if want specific plot
#   save_location  = dir_plt_save,
#   save_name      = "pft_bar_percent_season",
#   verbose        = TRUE,
#   overwrite      = FALSE,
#   # overwrite      = TRUE,
#   time_stamp_fmt = "%Y%m%d_%H%M%S",
#   device         = "jpeg",
#   height         = 9.7,
#   width          = 19,
#   dpi            = 600,
#   units          = "in",
#   bg             = "white"
#   )
```

## 2.3 Add: Labels for Season and Year

Format:
- Spring - Autumun is "{season} {year}" 
- Winter: 
  - Dec = "{season} {year}-{year+1}"
  - Jan and Feb = "{season} {year-1}-{year}"
  
When a season and year factor was not sampled at all, added a no data
- Summer 2016
- Winter 2016 - 2017
- Spring 2020 (COVID-19 delays)


```{r chemtax-add_labels}
# extra_fct <- 
#   c(
#     "Summer 2016 (No Data)",
#     "Winter 2016-2017 (No Data)", 
#     "Spring 2020 (No Data)"
#   )
# # cmtx_l <-
#   cmtx_l %>%
#   arrange(date_time_utc) %>%
#   mutate(
#     .after = season,
#     season_year = case_when(
#       str_detect(season, "Aut|Spr|Sum") ~ glue("{season} {year}"),
#       str_detect(month, "Jan|Feb")      ~ glue("{season} {year-1}-{year}"),
#       str_detect(month, "Dec")          ~ glue("{season} {year}-{year+1}")),
#     
#     season_year = fct_inorder(season_year),
#     season_year = fct_expand(season_year, extra_fct),
#     season_year = fct_relevel(season_year, extra_fct[1], after = 1),
#     season_year = fct_relevel(season_year, extra_fct[2], after = 3),
#     season_year = fct_relevel(season_year, extra_fct[3], after = 16)
#   ) %T>% print()
```

## 3.3 Plot: Bar Plot: PFT per **Season** and **Year** as Stacked Percent


```{r chemtax-stacked-bar-year-season}
# extra_fct <- 
#   c(
#     "Summer 2016 (No Data)",
#     "Winter 2016-2017 (No Data)", 
#     "Spring 2020 (No Data)"
#   )
# 
# add_season_year <-
#   . %>%
#   arrange(date_time_utc) %>%
#   mutate(
#     .after = season,
#     season_year = case_when(
#       str_detect(season, "Aut|Spr|Sum") ~ glue("{season} {year}"),
#       str_detect(month, "Jan|Feb")      ~ glue("{season} {year-1}-{year}"),
#       str_detect(month, "Dec")          ~ glue("{season} {year}-{year+1}")),
#     
#     season_year = fct_inorder(season_year),
#     season_year = fct_expand(season_year, extra_fct),
#     season_year = fct_relevel(season_year, extra_fct[1], after = 1),
#     season_year = fct_relevel(season_year, extra_fct[2], after = 3),
#     season_year = fct_relevel(season_year, extra_fct[3], after = 16)
#   ) 
# 
# 
# # CHEMTAX stacked bar graph with ratio to chlor-a, organized by station
# cmtx_l %>%
#   add_season_year %T>% print() %>%
#   # mutate(station = fct_reorder(station, lon, .na_rm = FALSE)) %>%
#   # filter(!is.na(station))
#   ggplot(
#     aes(
#       y     = conc,
#       x     = station,
#       fill  = names,
#       color = names
#     )
#   ) +
#   geom_bar(
#     position = "fill",
#     stat     = "identity",
#     na.rm    = TRUE
#   ) +
#   labs(
#     x     = "Station",
#     y     = "Relative Ratio",
#     fill  = "Phytoplankton Functional Type",
#     color = "Phytoplankton Functional Type"
#   ) +
#   scale_fill_manual(values = tol_muted) +
#   scale_color_manual(values = tol_muted) +
#   scale_y_continuous(
#     labels = scales::label_percent(),
#     expand = c(0, 0)
#   ) +
#   facet_wrap(~season_year, ncol = 4, drop = FALSE) +
# 
#   # theme
#   theme(
#     axis.title       = element_text(size = 18),
#     legend.position  = c(0.6, 0.05),
#     legend.direction = "horizontal"
#   )
# 
# # ---- save plot
# save_gg(
#   plt            = last_plot(), # edit if want specific plot
#   save_location  = dir_plt_save,
#   save_name      = "pft_bar_percent_season_year",
#   verbose        = TRUE,
#   overwrite      = FALSE,
#   time_stamp_fmt = "%Y%m%d_%H%M%S",
#   device         = "jpeg",
#   height         = 9.7,
#   width          = 19,
#   dpi            = 600,
#   units          = "in",
#   bg             = "white"
#   )
# 

```


## 3.4 Plot: Bar Plot: PFT per **Season** and **Year** as Absolute VAr

```{r chemtax-stacked-bar-conc-year-season}
# # chemtax stacked bar graph (not ratio), organized by station
# cmtx_l %>%
#   add_season_year %>%
#   summarize(
#     conc = mean(conc, na.rm = TRUE),
#     # .by  = c(season_year, station, func_type)
#     .by  = c(season_year, station, phyto)
#   ) %>%
#   ggplot(aes(
#     fill  = phyto,
#     color = phyto,
#     # fill  = func_type,
#     # color = func_type,
#     y     = conc,
#     x     = station
#   )) +
#   geom_bar(stat = "identity") +
#   labs(
#     x     = "Station",
#     y     = "Concentration of Chlorophyll-a (mg m<sup>-3</sup>)",
#     fill  = "Phytoplankton Functional Type",
#     color = "Phytoplankton Functional Type"
#   ) +
#   scale_fill_manual(values  = tol_muted) +
#   scale_color_manual(values = tol_muted) +
#   scale_y_continuous(expand = c(0, 0)) +
#   facet_wrap(~season_year, ncol = 4, drop = FALSE) +
# 
#   # theme
#   theme(
#     axis.title.y     = element_markdown(),
#     legend.position  = c(0.6, 0.05),
#     legend.direction = "horizontal"
#   )
# 
# 
# # ---- save plot
# save_gg(
#   plt            = last_plot(), # edit if want specific plot
#   save_location  = dir_plt_save,
#   save_name      = "pft_bar_absolute_season_year",
#   verbose        = TRUE,
#   overwrite      = FALSE,
#   time_stamp_fmt = "%Y%m%d_%H%M%S",
#   device         = "jpeg",
#   height         = 9.7,
#   width          = 19,
#   dpi            = 600,
#   units          = "in",
#   bg             = "white"
#   )
# 

```


## 6.5 Table: PFT Percentages per Season


```{r table-pft-proportions}
# shelf(gt)
# gt_pft <-
# # cmtx_l_pct
#  cmtx_avg_season_fract_l %>%
#   mutate(
#     names = fct_reorder(names, frac_avg, .fun = mean, .desc = TRUE),
#     percent_avg = str_remove(percent_avg, "%"),
#     percent_lbls = glue("{percent_avg} \u00B1 {percent_sd}")
#   ) %>%
#   arrange(names) %>%
#   pivot_wider(
#     data         = .,
#     id_cols      = c(names, phyto), 
#     names_from   = c(season), 
#     values_from  = c(percent_lbls)
#     )  %>%
#           
#   # ---- gt table for PFT average +/- std dev per season
#   gt(
#     data          = .,
#     groupname_col = "names"
#   ) %>%
#   cols_hide(phyto) %>%
#   tab_options(
#     table_body.hlines.width      = 0,
#     table_body.vlines.width      = 0,
#     stub.border.width            = 0,
#     row_group.as_column          = TRUE,
#     row_group.padding.horizontal = 20,
#     row_group.padding            = 200,
#     stub_row_group.font.weight   = "bold",
#     stub_row_group.border.width  = 0,
#     row_group.border.top.width   = px(5)
#   ) %>%
#   cols_align("center") %>%
#   tab_style(
#     cell_text(
#       v_align = "middle",
#       align   = "center"
#     ),
#     locations = cells_row_groups()
#   ) %>%
#   tab_spanner(
#     label   = "Average % \u00B1 S.D.", 
#     columns = everything()) %>%
#   opt_table_font("serif")
# 
# gt_pft
# 
# # ---- save tables as html and docx
# if (FALSE) {
#   file_expr(
#     loc            = dir_plt_save,
#     file_base      = "pft_fraction_seasons",
#     exts           = c("html", "docx"),
#     time_stamp_fmt = "%Y%m%d_%H%M%S"
#   ) %$%
#   eval(file_expr) %>%
#   walk(
#     .,
#     \(file_n) {
#       gtsave(
#         gt_pft,
#         filename = file_n
#       )
#     }
#   )
# 
#  
# }
#  

```

## 6.3 Plot: Average PFT per Season with Standard Deviation Error Bars

Calculate per PFT:
- average 
- standard deviation

```{r cmtx-summary}
# cmtx_w %>%
#   summarise(
#     .by = season,
#     across(
#       # chloro:hapt,
#       chloro:pras,
#       list(
#         avg = (\(x) mean(x, na.rm = TRUE)),
#         sd  = (\(x) sd(x, na.rm = TRUE)),
#         var = (\(x) var(x, na.rm = TRUE))
#       )
#     )
#   ) %>%
#   select(-contains(c("var"))) %>%
#   pivot_longer(
#     data      = .,
#     cols      = -season,
#     names_to  = "phyto",
#     values_to = "conc", 
#     ) %>%
#   separate(phyto, into = c("phyto", "type"), sep = "_") %>%
#   pivot_wider(
#     data         = .,
#     names_from   = type, 
#     values_from  = c(conc), 
#     names_repair = janitor::make_clean_names
#     ) %>%
#   left_join(phyto_name, by = c("phyto" = "shrt")) %T>% 
#   print() %>%
#   ggplot(aes(x = season, y = avg, color = season)) +
#   geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
#   geom_point() +
#   geom_linerange(aes(ymin = avg - sd, ymax = avg + sd )) +
#   scale_y_continuous(breaks = scales::breaks_pretty()) +
#   labs(
#     x = "Season",
#     y = bquote("Concentration"~(mg~L^-1)),
#     color = NULL
#   ) +
#   facet_wrap(~names, ncol = 3, scales = "free_y") +
#   theme_bw(base_family = "serif") +
#   theme(legend.position = "top")
# 
# # ---- save plot
# save_gg(
#   plt            = last_plot(), # edit if want specific plot
#   save_location  = dir_plt_save,
#   save_name      = "pft_avg_per_season",
#   verbose        = TRUE,
#   overwrite      = FALSE,
#   time_stamp_fmt = "%Y%m%d_%H%M%S",
#   device         = "jpeg",
#   height         = 6,
#   width          = 10,
#   dpi            = 600,
#   units          = "in",
#   bg             = "white"
#   )

```