---
title: "Seasonal CHEMTAX Box Plots"
author: "Sebastian Di Geronimo"
date: "2024-01-30"
format: html
editor: source
---

# 1.0 ---- Summary of Document ----



# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
  # additional
  rlang, cli,
  rstatix, readxl, ggtext, cowplot
  )

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
    )

source(here("scripts", "misc_functions.R"))
```

## 2.2 Load Dataset

```{r load-data}
# ---- set location to save data and plots ---- #
dir_data_save <- here("data", "processed", "04_3_cmtx_box_plots")
dir_plt_save  <- here("data", "plots", "04_3_cmtx_box_plots")

dir_create(here(dir_plt_save))

# ---- set season names and order ---- #
szn <- c("Winter", "Spring", "Summer", "Autumn")

# ---- phytoplankton full names ---- #
phyto_name <-
  here("data", "processed", "04_1_merge_meta_cmtx", "phyto_names.csv") %>%
  read_csv(show_col_types = FALSE) %T>% 
  print()

# ---- load chemtax data in wide format ---- #
cmtx_w <-
  here("data", "processed", "04_1_merge_meta_cmtx", "chemtax_w_meta_wide.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  # filter(!str_detect(station, "60|21.5|24|^3$|9B|9.5")) %>%
  filter(!str_detect(station, "60|21.5|24|^3$|9.5")) %>%
  mutate(
    season = fct(season, levels = szn),
    # station = str_replace(station, "9B", "9.5"),
    station = fct_reorder(station, lon)
  ) %>% 
  select(-c(hplc_gsfc_id, cruise:bottle, cluster_code, water_type)) %>%
    
  # add PFT concentrations to get total chlorophyll-a
  rowwise() %>%
  mutate(
    total_chl_a = sum(c_across(chloro:pras)),
    .before = chloro
  ) %>%
  ungroup() %>%
  filter(
    .by = c(station, cruise_id),
    depth == min(depth)
    & depth < 20
  ) %T>% 
  print()

```
## 2.2.1 Remove Outliers?

```{r remove-outliers}
remove_out <- menu(c("yes", "no"), title = "Remove Outliers?")

if (remove_out == 1) {
  cmtx_w <-
    cmtx_w %>%
    filter(
      total_chl_a <= 1
    )
  
  file_suffix <- "_no_out"
} else {
  file_suffix <- ""
}
```

## 2.3 Convert to Long Format with Percent


```{r convert-long-format}
# ---- convert to long format ---- #
# i.e. one column for all PFT and one for concentrations
cmtx_l <-
  cmtx_w %>%
  pivot_longer(
    cols      = chloro:pras,
    names_to  = "phyto",
    values_to = "conc"
  ) %>%
  left_join(phyto_name, by = c("phyto" = "shrt")) %>%
  
  # add percent
  mutate(percent = conc/total_chl_a) %T>% 
  print()

# ---- convert to percent ---- #
cmtx_l_pct <-
  cmtx_l %>%
  select(-conc) %T>% 
  print()
```


## 2.4 Line and Box Plot Plotting Functions 

Default Line Plot and Box Plots

```{r functions-line-box}
#' Box Plot
#'
#' FUNCTION_DESCRIPTION
#'
#' @param .dat DESCRIPTION.
#' @param .var DESCRIPTION.
#' @param .title DESCRIPTION.
#'
#' @return RETURN_DESCRIPTION
#' @examples
#' # ADD_EXAMPLES_HERE
# box_plt <- function(.dat, .var, .color, .shape, .title) {
box_plt <- function(.dat, .var, .title, ...) {
    
    # limits for plot
    ylims <-  c(0, quantile(.dat[[.var]], 0.99))
    brake <- pretty(ylims, 6)

    # plotting
    .dat %>%
    ggplot() +
    geom_boxplot(
      aes(
        x = season, 
        y = .data[[.var]]
        ),
      outlier.shape = NA) +
    geom_jitter(
        aes(
          x = season, 
          y = .data[[.var]], 
          ...
          # color = .data[[.color]],
          # shape = .data[[.shape]]
          ),
        width = 0.4,
        size = 0.5
        ) +
    
    # labels
    labs(
      title = glue("Concentration of {.title} per Season "),
      x     = "Season",
      y     = "Concentration (mg m<sup>-3</sup>)"
      # color = "Year"
      ) +

    # scale settings
    scale_color_discrete(type = "viridis") +
    
    scale_shape_discrete() +
        
    scale_y_continuous(
      breaks = brake,
      expand = expansion(mult = 0),
      limits = ylims,
      oob    = scales::squish
      ) +
    
    # theme settings
    theme_bw(base_family = "serif", base_size = 12) +
    theme(
      panel.grid        = element_blank(),
      plot.title        = element_text(size = 15),
      axis.ticks.length = unit(-0.05, "in"),
      axis.text.y       = element_text(margin = unit(c(0.3,0.3,0.3,0.3), "cm")),
      axis.title.y      = element_markdown(size = 10),
      axis.text.x       = element_text(margin = unit(c(0.3,0.3,0.3,0.3), "cm")),
      axis.ticks.x      = element_blank(),
      legend.background = element_rect(color = "black", fill = "white")
    )
}
```

# 3.0 Summarize Stations

May move to somewhere else

```{r}
cmtx_w %>%
  count(station, season, lon, lat) %>%
  pivot_wider(
    data         = .,
    id_cols      = c(station, lon, lat), # *optional* vector of unaffected columns,
    names_from   = c(season), # category column(s) to pivot wide
    values_from  = c(n), # value column(s) that hold data for each category column
    names_sep    = "_",
    names_repair = janitor::make_clean_names
    )

cmtx_w
```



# 4.0 ---- Box Plots ----

## 4.2 Box Plot: Individuals for Each PFT per Season

```{r boxplots, fig.width=10, fig.asp=0.8}
# ---- avg season line plots
cmtx_avgs_plt <-
  cmtx_l %>%
  nest(.by = c(phyto, names)) %>%
  mutate(
    plt = map2(
      data,
      names,
      ~ .x %>%
        mutate(Year = factor(year)) %>%
        box_plt(
        .dat   = .,
        .var   = "conc",
        color = Year,
        shape = Year,
        .title = .y
      )
    )
  )

# cmtx_avgs_plt$plt[[1]]
cmtx_avgs_plt$plt
```

## 4.3 Box Plot: Base for Larger/Small Dynamic Range PFTs

```{r box-base, fig.height=15, fig.width=20}
box_filter <- 
  . %>%
  filter(.by = c(station, names), n() > 3) %>%
   mutate(
    .by       = c(station, names),
    .after    = station,
    stn_label = glue("{station} (N = {n()})")
   ) %>%
  mutate(
    stn_label = fct_reorder(stn_label, lon)
  )

box_plt_base <-
  list(
    aes(
      x    = season,
      y    = conc,
      fill = season
    ),
    geom_boxplot(
        linewidth = 0.15,
        outlier.shape = 20, 
        outlier.size = 0.15
        ),
    scale_y_continuous(
      expand = c(0, 0),
      breaks = scales::breaks_width(0.1),
      minor_breaks = NULL,
      oob    = scales::squish,
      # labels = scales::label_percent(accuracy = 1)
      labels = scales::label_number(accuracy = 0.01)
    ),
    labs(
      # x = "Season",
      x = NULL,
      y = "Chlorophyll-a Concentration (mg m<sup>-3</sup>)"
    ),
    # theme_classic(),
    theme_bw(),
    theme(
      text               = element_text(family = "serif", size = 4, color = "black"),
      line               = element_line(color = "black", linewidth = 0.2),
      rect               = element_rect(color = "black", linewidth = 0.2),
      axis.text          = element_markdown(size = 4),
      axis.title.y       = element_markdown(size = 4),
      legend.position    = "none",
      strip.text.x       = element_text(size = 5, margin = margin(t = 2, b = 2)),
      strip.background   = element_rect(fill = "white", colour = "grey20"),
      panel.grid.major.x = element_blank(),
      panel.grid.major.y = element_line(color = "grey20", linewidth = 0.25),
    )
  )
```

## 4.4 Box Plot: Larger Dynamic Range PFTs

```{r box-lg-dynm, fig.height=15, fig.width=20}
# box plot by season (1 x 4)
box_lg <-
  cmtx_l %>%
  filter(phyto %in% c("cyano2", "cyano4", "hapt6", "hapt8")) %>%
  ggplot() +
  box_plt_base +
  facet_wrap(facet = ~names, ncol = length(unique(cmtx_l$phyto)) - 6) +
  coord_cartesian(ylim = c(0, 0.45), clip = "off")

# box plot by season and station
box_lg_stn <-
  cmtx_l %>%
  filter(phyto %in% c("cyano2", "cyano4", "hapt6", "hapt8")) %>%
  box_filter %>%
  ggplot() +
  box_plt_base +
  # facet_wrap(facet = ~ names + stn_label, ncol = 8) +
  facet_grid(rows = stn_label ~ names) +
  coord_cartesian(ylim = c(0, 0.4))
  # coord_cartesian(ylim = c(0, 1))

# stack box plots  
box_lg_stcked <- 
  patchwork::wrap_plots(
    box_lg, 
    box_lg_stn, 
    ncol = 1, 
    heights = c(0.2, 0.8)
  )

# display plots
box_lg
box_lg_stn
box_lg_stcked

# box plot by season (2 x 2)
box_lg <-
  cmtx_l %>%
  filter(phyto %in% c("cyano2", "cyano4", "hapt6", "hapt8")) %>%
  ggplot() +
  box_plt_base +
  facet_wrap(facet = ~names) +
  coord_cartesian(ylim = c(0, 0.45), clip = "off") +
  scale_y_continuous(
      limits = c(0, 0.45), 
      expand = c(0, 0),
      breaks = scales::breaks_width(0.1),
      minor_breaks = NULL,
      oob    = scales::squish,
      # labels = scales::label_percent(accuracy = 1)
      labels = scales::label_number(accuracy = 0.01))

box_lg
```


## 4.5 Box Plot: Smaller Dynamic Ranged PFTs


```{r box-sm-dynm, fig.height=15, fig.width=20}
# box plot by season (1 x 6)
box_sm <-
  cmtx_l %>%
  filter(phyto %in% c("chloro", "crypto", "diat1", "diat2", "dino", "pras")) %>%
  ggplot() + 
  box_plt_base +
  facet_wrap(facet = ~names, ncol = length(unique(cmtx_l$phyto)) - 4) +
  coord_cartesian(ylim = c(0, 0.2))
 
  
# box plot by season and station
box_sm_stn <-
  cmtx_l %>%
  filter(phyto %in% c("chloro", "crypto", "diat1", "diat2", "dino", "pras")) %>%
  box_filter %>%
  ggplot() +
  box_plt_base +
  # facet_wrap(facet = ~ names + stn_label, ncol = 8) +
  facet_grid(rows = stn_label ~ names) +
  coord_cartesian(ylim = c(0, 0.2))

# stack box plots
box_sm_stcked <-
  patchwork::wrap_plots(
    box_sm,
    box_sm_stn,
    ncol = 1,
    heights = c(0.25, 0.75)
  )

# display plots
box_sm
box_sm_stn
box_sm_stcked

# box plot by season (2 x 3)
box_sm <-
  cmtx_l %>%
  filter(phyto %in% c("chloro", "crypto", "diat1", "diat2", "dino", "pras")) %>%
  ggplot() + 
  box_plt_base +
  facet_wrap(facet = ~names) +
  coord_cartesian(ylim = c(0, 0.2), clip = "off") +
  scale_y_continuous(
    limits = c(0, 0.2),
    expand = c(0, 0),
    breaks = scales::breaks_width(0.05),
    minor_breaks = NULL,
    oob    = scales::squish,
    # labels = scales::label_percent(accuracy = 1)
    labels = scales::label_number(accuracy = 0.01)
  )

box_sm
```



```{r box-lg-dynm-percent, fig.height=15, fig.width=20}
# box plot by station
box_lg_pct <-
  cmtx_l_pct %>%
  filter(phyto %in% c("cyano2", "cyano4", "hapt6", "hapt8")) %>%
  ggplot() +
  box_plt_base +
  facet_wrap(facet = ~names, ncol = length(unique(cmtx_l$phyto)) - 6) +
  labs(y = "Relative Abundance (%)") +
  scale_y_continuous(labels = scales::label_percent(), expand = c(0,0)) +
  coord_cartesian(ylim = c(0, 1))

# box plot by season and station
box_lg_stn_pct <-
  cmtx_l_pct %>%
  filter(phyto %in% c("cyano2", "cyano4", "hapt6", "hapt8")) %>%
  box_filter %>%
  ggplot() +
  box_plt_base +
  # facet_wrap(facet = ~ names + stn_label, ncol = 8) +
  facet_grid(rows = stn_label ~ names) +
  scale_y_continuous(labels = scales::label_percent(), expand = c(0,0)) +
  coord_cartesian(ylim = c(0, 1))

# stack box plots  
box_lg_stcked_pct <- 
  patchwork::wrap_plots(
    box_lg_pct, 
    box_lg_stn_pct, 
    ncol = 1, 
    heights = c(0.2, 0.8)
  )

# display plots
box_lg_pct
box_lg_stn_pct
box_lg_stcked_pct


# box plot by season (2 x 2)
box_lg_pct <-
  cmtx_l_pct %>%
  filter(phyto %in% c("cyano2", "cyano4", "hapt6", "hapt8")) %>%
  ggplot() +
  box_plt_base +
  # facet_wrap(facet = ~names, ncol = length(unique(cmtx_l$phyto)) - 6) +
  facet_wrap(facet = ~names) +
  labs(y = "Relative Abundance (%)") +
  scale_y_continuous(labels = scales::label_percent(), expand = c(0,0)) +
  coord_cartesian(ylim = c(0, 1), clip = "off")

box_lg_pct
```


```{r box-sm-dynm-pct, fig.height=15, fig.width=20}
# box plot by season (1 x 6)
box_sm_pct <-
  cmtx_l_pct %>%
  filter(phyto %in% c("chloro", "crypto", "diat1", "diat2", "dino", "pras")) %>%
  ggplot() + 
  box_plt_base +
  facet_wrap(facet = ~names, ncol = length(unique(cmtx_l$phyto)) - 4) +
  labs(y = "Relative Abundance (%)") +
  scale_y_continuous(labels = scales::label_percent(), expand = c(0,0)) +
  coord_cartesian(ylim = c(0, 1))
 
  
# box plot by season and station
box_sm_stn_pct <-
  cmtx_l_pct %>%
  filter(phyto %in% c("chloro", "crypto", "diat1", "diat2", "dino", "pras")) %>%
  box_filter %>%
  ggplot() +
  box_plt_base +
  # facet_wrap(facet = ~ names + stn_label, ncol = 8) +
  facet_grid(rows = stn_label ~ names) +
  labs(y = "Relative Abundance (%)") +
  scale_y_continuous(labels = scales::label_percent(), expand = c(0,0)) +
  coord_cartesian(ylim = c(0, 1))

# stack box plots
box_sm_stcked_pct <-
  patchwork::wrap_plots(
    box_sm_pct,
    box_sm_stn_pct,
    ncol = 1,
    heights = c(0.25, 0.75)
  )

# display plots
box_sm_pct
box_sm_stn_pct
box_sm_stcked_pct

# box plot by season (2 x 3)
box_sm_pct <-
  cmtx_l_pct %>%
  filter(phyto %in% c("chloro", "crypto", "diat1", "diat2", "dino", "pras")) %>%
  ggplot() + 
  box_plt_base +
  # facet_wrap(facet = ~names, ncol = length(unique(cmtx_l$phyto)) - 4) +
  facet_wrap(facet = ~names) +
  labs(y = "Relative Abundance (%)") +
  scale_y_continuous(labels = scales::label_percent(), expand = c(0,0)) +
  coord_cartesian(ylim = c(0, 1), clip = "off")

box_sm_pct
```

## 4.6 Save: Large and Small Range Boxplots

```{r save-plt}
# save box_plots
walk2(
  verbose   = TRUE,  # set to FALSE to suppress output
  # overwrite = FALSE, # set to TRUE to overwrite existing files
  overwrite = TRUE, # set to TRUE to overwrite existing files
  .x = list2(
    "cmtx_boxplot_lg",
    "cmtx_boxplot_sm",
    "cmtx_boxplot_lg_stn",
    "cmtx_boxplot_sm_stn",
    "cmtx_boxplot_lg_stcked",
    "cmtx_boxplot_sm_stcked",
    "cmtx_boxplot_lg_pct",
    "cmtx_boxplot_sm_pct",
    "cmtx_boxplot_lg_stn_pct",
    "cmtx_boxplot_sm_stn_pct",
    "cmtx_boxplot_lg_stcked_pct",
    "cmtx_boxplot_sm_stcked_pct"
  ),
  .y = list2(
      box_lg, box_sm,
      box_lg_stn, box_sm_stn,
      box_lg_stcked, box_sm_stcked,
      box_lg_pct, box_sm_pct,
      box_lg_stn_pct, box_sm_stn_pct,
      box_lg_stcked_pct, box_sm_stcked_pct
      ),
  \(.x, .y, ...) {
    .x %>%
      str_to_lower() %>%
      str_replace_all("\\s", "_") %>%
      glue(., .env$file_suffix)  %>% #, .sep = "_") %>% # remove if none needed
      save_gg(
        .y,
        save_location  = here(dir_plt_save, Sys.Date(), str_remove(file_suffix, "^_")),
        save_name      = .,
        # time_stamp_fmt = "%Y%m%d_%H%M%S",
        time_stamp_fmt = NULL,
        device         = "jpeg",
        width          = 90,
        height         = 90 * 0.5625,
        dpi            = 600,
        units          = "mm",
        bg             = "white",
        ...
      )
  }
)

```

