---
title: "CHEMTAX Bar Plots"
author: "Sebastian Di Geronimo"
date: "2024-01-30"
format: html
editor: source
---

# 1.0 ---- Summary of Document ----



# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
    # additional
    rlang
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )

source(here("scripts", "misc_functions.R"))
```

## 2.2 Load Dataset

```{r load-data}
# run snippet after saving file
# set location to save data and plots
dir_data_save <- here("data", "processed", "04_4_bar_plots")
dir_plt_save  <- here("data", "plots", "04_4_bar_plots")

# ---- set season names and order ---- #
szn <- c("Winter", "Spring", "Summer", "Autumn")

# ---- phytoplankton full names ---- #
phyto_name <-
  here("data", "processed", "04_1_merge_meta_cmtx", "phyto_names.csv") %>%
  read_csv(show_col_types = FALSE) %T>% 
  print()

# ---- load chemtax data in wide format ---- #
cmtx_w <-
  here("data", "processed", "04_1_merge_meta_cmtx", "chemtax_w_meta_wide.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  # filter(!str_detect(station, "60|21.5|24|^3$|9B|9.5")) %>%
  filter(!str_detect(station, "60|21.5|24|^3$|9.5")) %>%
  mutate(
    season = fct(season, levels = szn),
    # station = str_replace(station, "9B", "9.5"),
    station = fct_reorder(station, lon)
  ) %>% 
  select(-c(hplc_gsfc_id, cruise:bottle, cluster_code, water_type)) %>%
    
  # add PFT concentrations to get total chlorophyll-a
  rowwise() %>%
  mutate(
    total_chl_a = sum(c_across(chloro:pras)),
    .before = chloro
  ) %>%
  ungroup() %>%
  filter(
    .by = c(station, cruise_id),
    depth == min(depth)
    & depth < 20
  ) %T>% 
  print()

```


## 2.3 Convert to Long Format


Pivot longer by taking PFT columns and pivot to two columns of `phyto` for name
and `conc` for concentration.

Next, convert to percentage by `conc` / `total_chl_a`


```{r convert-long-format}
# one column for all PFT and one for concentrations
cmtx_l <-
  cmtx_w %>%
  pivot_longer(
    cols      = chloro:pras,
    names_to  = "phyto",
    values_to = "conc"
  ) %>%
  left_join(phyto_name, by = c("phyto" = "shrt")) %>%
  
  # calc percent
  mutate(percent = conc/total_chl_a) %T>% 
  print()

# extract percent column
cmtx_l_pct <-
  cmtx_l %>%
  select(-conc) %T>% 
  print()
```


# 3.0 ---- Combine PFT Sub Groups ----

## 3.1 Add: PFT Groups Together

Groups to be combined have multiple types but are merged for ease of use
- diatom:
  - diatom 1 
  - diatom 2
- cyanobacteria:
  - cyano 2
  - cyano 4
- haptophyte:
  - hapt 6
  - hapt 8

```{r station-seasonally-yearly-avgs}
# combine same base PFT groups into main group
cmtx_w_comb <-
  cmtx_w %>%
  rowwise() %>%
  mutate(
    .keep = "unused",       # remove used columns
    diat  = diat1  + diat2,
    cyano = cyano2 + cyano4,
    hapt  = hapt6  + hapt8
  ) %>%
  ungroup() %T>%
  print()

# calculate PFT fractions (0 - 100%)
cmtx_fract_w <-
  cmtx_w %>%
  mutate(
    .keep = "unused",
    across(
      chloro:last_col(),
      (\(x) x / total_chl_a),
      .names = "frac_{.col}"
    )
  ) %T>% print()

# calculate PFT fractions (0 - 100%)
cmtx_fract_w_comb <-
  cmtx_w_comb %>%
  mutate(
    .keep = "unused",
    across(
      c(pras:hapt, chloro, crypto, dino),
      (\(x) x / total_chl_a),
      .names = "frac_{.col}"
    )
  ) %T>% print()

# convert to long format function
cmtx_fract_l_base <- 
  . %>%
  pivot_longer(
    contains("frac"),
    names_to     = "phyto",
    values_to    = "frac",
    names_prefix = "frac_"
  ) %>%
    
  # add long names
  left_join(phyto_name, by = c("phyto" = "shrt")) %T>% print()

# ---- without combining diat, cyano and hapt ---- #
cmtx_fract_l <- 
  cmtx_fract_w %>%
  cmtx_fract_l_base

# ---- without combining diat, cyano and hapt ---- #
cmtx_fract_l_comb <- 
  cmtx_fract_w_comb %>%
  cmtx_fract_l_base

```

## 3.2 View Averages for Both Seasonal and per Station

```{r view-avgs}
# ---- average by season ---- #
# only one used further
cmtx_avg_season_fract_l <-
  cmtx_fract_l %>%
  summarise(
    .by = c(names, phyto, season),
    across(
      .cols  = frac,
      .fns   = list(
          avg = \(x) { mean(x, na.rm = TRUE) },
          sd  = \(x) { sd(x, na.rm = TRUE) },
          se  = \(x) { sd(x, na.rm = TRUE) / sqrt(n()) },
          num = ~ n()
          ),
      .names = "{.col}_{.fn}"
    )
  ) %>%
  mutate(
    percent_avg = scales::label_percent(accuracy  = 0.01)(frac_avg),
    percent_sd  = scales::label_percent(accuracy  = 0.01)(frac_sd),
    percent_se  = scales::label_percent(accuracy  = 0.01)(frac_se),
    season1 = glue("{season} (N = {frac_num})"),
    season1 = fct_reorder(season1, as.numeric(season))
  ) %>%
 arrange(season) %T>% print()


# ---- average by station ---- #
cmtx_fract_l %>%
  mutate(station = fct_reorder(station, lon)) %>%
  arrange(station) %>%
  summarise(
    .by = c(names, phyto, station),
    across(
      .cols  = frac,
      .fns   = (\(x) mean(x, na.rm = TRUE)),
      .names = "{.col}_avg"
    )
  ) %>%
  mutate(
    percent_avg = scales::label_percent(accuracy  = 0.01)(frac_avg)
  ) %>%
 arrange(station) %T>% print()

# ---- average by station and season ---- #
cmtx_fract_l %>%
  mutate(station = fct_reorder(station, lon)) %>%
  arrange(station) %>%
  summarise(
    .by = c(station, season, names, phyto),
    across(
      .cols  = frac,
      .fns   = (\(x) mean(x, na.rm = TRUE)),
      .names = "{.col}_avg"
    )
  )  %>%
  mutate(
    percent_avg = scales::label_percent(accuracy  = 0.01)(frac_avg)
  ) %>%
 arrange(season) %T>% print()
```

## 3.3  Plot: Seasonal Fraction of each PFT (Bar)

### 3.3.1 Stacked Bar Chart
```{r plt-fracs, fig.width=20, fig.height=10}
tol_muted <- c(khroma::color("muted", names = FALSE)(9), "#BBBBBB")

new_dat <- 
  summarise(
    cmtx_w, 
    .by = season, 
    chl_avg = mean(total_chl_a), 
    chl_sd  = sd(total_chl_a),
    chl_se  = chl_sd / sqrt(n()),
    num = n()
    ) %>%
  mutate(season = glue("{season} (N = {num})")) %T>% 
  print() 

sec_axe <- 
  new_dat %>%
  ggh4x::help_secondary(
    # secondary = c(0, 1.5), 
    secondary = c(0, 1.0), 
    name = bquote("Chlorophyll-a Concentration"~(mg~m^-3))
    )



cmtx_avg_season_fract_l %>%
  mutate(names = fct_reorder(names, frac_avg)) %>%
  ggplot(
    aes(
      fill  = names,
      y     = frac_avg,
      x     = season1
    )
  ) +
  geom_bar(
    color     = "black",
    position  = "fill",
    stat      = "identity",
    linewidth = 0.2
  )  +    
  geom_line(
    data = new_dat,
    aes(x = season, y = sec_axe$proj(chl_avg), group = 1),
    inherit.aes = FALSE,
    linewidth = 0.2
    ) +
  geom_linerange(
    data = new_dat,
    aes(
      x = season,
      ymin = sec_axe$proj(chl_avg) - sec_axe$proj(chl_sd), # standard deviation
      ymax = sec_axe$proj(chl_avg) + sec_axe$proj(chl_sd)  # standard deviation
      
      # ymin = sec_axe$proj(chl_avg) - sec_axe$proj(chl_se), # standard error
      # ymax = sec_axe$proj(chl_avg) + sec_axe$proj(chl_se)),  # standard error
  ),
    inherit.aes = FALSE,
  linewidth = 0.3
  ) +
  geom_point(
    data = new_dat,
    aes(x = season, y = sec_axe$proj(chl_avg)),
    inherit.aes = FALSE,
    size  = 1,
    shape = 20
    ) +
  labs(
    x = NULL,
    # y     = "Proportion to Total Chl-a",
    # y     = "PFT:Total Chlorophyll-a (%)",
    y     = "Relative Abundance (%)",
    fill  = "Phytoplankton\nFunctional Type",
    color = "Phytoplankton\nFunctional Type"
  ) +
  scale_fill_manual(values  = tol_muted) +
  scale_color_manual(values = tol_muted) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = c(0,0),
    limits = c(0,1),
    oob    = scales::squish,
    sec.axis = sec_axe
    ) +
  scale_x_discrete(labels = distinct(cmtx_avg_season_fract_l, season, season1)) +
  coord_cartesian(clip = "off") +
    
  # theme
  theme_bw() +
  theme(
    text              = element_text(family = "serif", size = 6, color = "black"),
    line              = element_line(color = "black", linewidth = 0.2),
    rect              = element_rect(color = "black", linewidth = 0.2),
    axis.title        = element_text(size = 5),
    axis.text         = element_text(color = "black"),
    axis.ticks.x      = element_blank(),
    legend.position   = "bottom",
    legend.direction  = "horizontal",
    legend.text       = element_text(size = 4),
    legend.title      = element_text(size = 5),
    legend.key.size   = unit(2, "mm"),
    legend.box.margin = margin(t = -11, b = -5)
  )

# ---- save plot
save_gg(
  plt            = last_plot(), # edit if want specific plot
  save_location  = here(dir_plt_save, "season_stacked"),
  save_name      = "pft_season_avg_prcnt_stack",
  verbose        = TRUE,
  overwrite      = FALSE,
  # overwrite      = TRUE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "png",
  height         = 90 * 0.5625,
  width          = 90,
  dpi            = 600,
  units          = "mm",
  bg             = "white",
  )
```

### 3.3.1 Separated Bar Chart

```{r plt-fracs, fig.width=20, fig.height=10}
cmtx_avg_season_fract_l %>%
  mutate(names = fct_reorder(names, frac_avg)) %>%
  ggplot(
    aes(
      fill  = season,
      y     = frac_avg,
      x     = fct_reorder(names, frac_avg, .desc = TRUE)
    )
  ) +
  geom_bar(
    color = "black",
    # position = "fill",
    position = position_dodge(),
    stat     = "identity"
  ) +
    
  labs(
    x     = "Season",
    # y     = "Proportion to Total Chl-a",
    # y     = "PFT:Total Chlorophyll-a (%)",
    y     = "Relative Abundance (%)",
    fill  = "Phytoplankton Functional Type",
    color = "Phytoplankton Functional Type"
  ) +
  scale_fill_manual(values  = tol_muted) +
  scale_color_manual(values = tol_muted) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = c(0,0),
    limits = c(0,1),
    sec.axis = sec_axe
    ) +
  # theme
  theme_classic() +
  theme(
    text             = element_text(family = "serif", size = 20),
    axis.title       = element_text(size = 20),
    legend.position  = "bottom",
    legend.direction = "horizontal",
    plot.margin = margin(15,10,10,10, "pt")
  )

# ---- save plot ---- #
save_gg(
  plt            = last_plot(), # edit if want specific plot
  save_location  = here(dir_plt_save, "separated"),
  save_name      = "pft_season_avg_prcnt_sep",
  verbose        = TRUE,
  overwrite      = FALSE,
  # overwrite      = TRUE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "png",
  height         = 90 * 0.5625,
  width          = 90,
  dpi            = 600,
  units          = "mm",
  bg             = "white"
  )
 
```

## 3.4 Bar Chart: PFT per **Season** and **Station** as Stacked Percent


Figure 6

```{r chemtax-stacked-bar}
tol_muted <- c(khroma::color("muted", names = FALSE)(9), "#BBBBBB")

new_dat <- 
  cmtx_w  %>%
  mutate(
    .by = season,
    season1 = glue("{season} (N = {n()})")) %>%
  mutate(
    season = fct_reorder(season1, as.numeric(season))
  ) %>%
  summarise(
    .by = c(season, station), 
    chl_avg = mean(total_chl_a), 
    chl_sd  = sd(total_chl_a),
    chl_se  = chl_sd / sqrt(n())
    )  %T>% 
  print()

sec_axe <- 
  new_dat %>%
  ggh4x::help_secondary(
    secondary = c(0, 2), 
    name = bquote("Chlorophyll-a Concentration"~(mg~m^-3))
    )


# CHEMTAX stacked bar graph
cmtx_stck_bar_szn_stn <- 
  cmtx_l %>%
  filter(!is.na(conc)) %>%
  mutate(
    .by = season,
    season1 = glue("{season} (N = {n()/10})")
    ) %>%
  mutate(
    season = fct_reorder(season1, as.numeric(season))
  ) %>%
  summarise(
  .by = c(station, season, phyto, names),
  conc = mean(conc)
  ) %>%
  mutate(
    names = fct_reorder(names, conc, .desc = FALSE)
  ) %T>% 
  print() %>%
  ggplot(
    aes(
      x     = station,
      y     = conc,
      fill  = names
    )
  ) +
  geom_bar(
    color     = "black",
    position  = "fill",
    stat      = "identity",
    na.rm     = TRUE,
    linewidth = 0.1
  ) +
  geom_rect(
    data = tibble(
      station = c("16", "10", "7"), 
      x  = c(2.6, 4.6, 6.6), 
      x2 = c(3.4, 5.4, 7.4)), 
    aes(xmin = x, xmax = x2, ymin = 0.01, ymax = 0.99),
    fill        = NA,
    color       = "red",
    linewidth   = 0.4,
    inherit.aes = FALSE
    ) +
  geom_linerange(
    data = new_dat,
    aes(
      x = station,
      ymin = sec_axe$proj(chl_avg) - sec_axe$proj(chl_sd), # standard deviation
      ymax = sec_axe$proj(chl_avg) + sec_axe$proj(chl_sd)  # standard deviation
      
      # ymin = sec_axe$proj(chl_avg) - sec_axe$proj(chl_se), # standard error
      # ymax = sec_axe$proj(chl_avg) + sec_axe$proj(chl_se)),  # standard error
  ),
    inherit.aes = FALSE,
  linewidth = 0.2
  ) +
  geom_point(
    data = new_dat,
    aes(x = station, y = sec_axe$proj(chl_avg)),
    inherit.aes = FALSE,
    fill  = "white",
    size  = 0.5,
    shape = 20,
  ) +
  labs(
    # x     = "Station",
    x     = NULL,
    y     = "Relative Abundance (%)",
    fill  = "Phytoplankton\nFunctional Type",
    color = "Phytoplankton\nFunctional Type"
  ) +
  scale_y_continuous(
    labels = scales::label_percent(),
    expand = c(0, 0),
    limits = c(0, 1),
    oob    = scales::squish,
    sec.axis = sec_axe
  ) +
  scale_fill_manual(values = tol_muted) +
  coord_cartesian(clip = "off") +
  facet_wrap(~ season) +
    
  # theme
  theme_bw() +
  theme(
    # text               = element_text(family = "serif", size = 20, color = "black"),
    text               = element_text(family = "serif", size = 6, color = "black"),
    line               = element_line(color = "black", linewidth = 0.2),
    rect               = element_rect(color = "black", linewidth = 0.2),
    # axis.title       = element_text(size = 20),
    axis.title         = element_text(size = 5),
    axis.text          = element_text(color = "black"),
    axis.ticks.x       = element_blank(),
    legend.position    = "bottom",
    legend.direction   = "horizontal",
    legend.text        = element_text(size = 4),
    legend.title       = element_text(size = 5),
    legend.key.size    = unit(2, "mm"),
    # legend.box.spacing = unit(2, "mm"),
    legend.box.margin  = margin(t = -11, b = -5),
    strip.text.x       = element_text(size = 4, margin = margin(t = 2, b = 2)),
    strip.background   = element_rect(fill = "white", colour = "black"),
    
  )

cmtx_stck_bar_szn_stn %>%
  egg::tag_facet(
    family   = "serif", 
    tag_pool = LETTERS, 
    fontface = 1, 
    vjust    = -0.5,
    hjust    = 0, 
    size     = 2
    )

# ---- save plot
save_gg(
  plt            = last_plot(), # edit if want specific plot
  save_location  = here(dir_plt_save, "season_station"),
  save_name      = "pft_bar_percent_season",
  verbose        = TRUE,
  overwrite      = FALSE,
  # overwrite      = TRUE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "png",
  height         = 90 * 0.5625,
  width          = 90,
  dpi            = 600,
  units          = "mm",
  bg             = "white"
  )
```


## 3.5 Table: PFT Percentages per Season

Can be either standard deviation or standard error. 
Decide which to use.
2 Changes will need to be made by commenting and uncommenting below

```{r table-pft-proportions}
shelf(gt)
gt_pft <-
# cmtx_l_pct
 cmtx_avg_season_fract_l %>%
  mutate(
    names = fct_reorder(names, frac_avg, .fun = mean, .desc = TRUE),
    percent_avg = str_remove(percent_avg, "%"),
    # percent_lbls = glue("{percent_avg} \u00B1 {percent_se}")
    percent_lbls = glue("{percent_avg} \u00B1 {percent_sd}")
  ) %>%
  arrange(names) %>%
  pivot_wider(
    data         = .,
    id_cols      = c(names, phyto), 
    names_from   = c(season1), 
    values_from  = c(percent_lbls)
    )  %>%
          
  # ---- gt table for PFT average +/- std dev per season
  gt(
    data          = .,
    groupname_col = "names"
  ) %>%
  cols_hide(phyto) %>%
  tab_options(
    table_body.hlines.width      = 0,
    table_body.vlines.width      = 0,
    stub.border.width            = 0,
    row_group.as_column          = TRUE,
    row_group.padding.horizontal = 20,
    row_group.padding            = 200,
    stub_row_group.font.weight   = "bold",
    stub_row_group.border.width  = 0,
    row_group.border.top.width   = px(5)
  ) %>%
  cols_align("center") %>%
  tab_style(
    cell_text(
      v_align = "middle",
      align   = "center"
    ),
    locations = cells_row_groups()
  ) %>%
  tab_spanner(
    # label   = "Average % \u00B1 S.E.",
    label   = "Average % \u00B1 S.D.",
    columns = everything()) %>%
  opt_table_font("serif")

gt_pft

# ---- save tables as html and docx
if (FALSE) {
  file_expr(
    loc            = dir_plt_save,
    file_base      = "pft_fraction_seasons",
    exts           = c("html", "docx"),
    time_stamp_fmt = "%Y%m%d_%H%M%S"
  ) %$%
  eval(file_expr) %>%
  walk(
    .,
    \(file_n) {
      gtsave(
        gt_pft,
        filename = file_n
      )
    }
  )

 
}
```




# 4.0 ---- Additional Plots ----

## 4.1 Add: Labels for Season and Year

Format:
- Spring - Autumun is "{season} {year}" 
- Winter: 
  - Dec = "{season} {year}-{year+1}"
  - Jan and Feb = "{season} {year-1}-{year}"
  
When a season and year factor was not sampled at all, added a no data
- Summer 2016
- Winter 2016 - 2017
- Spring 2020 (COVID-19 delays)


```{r chemtax-add_labels}
# ---- CHEMTAX Season-Year Labels ---- #
extra_fct <- 
  c(
    "Summer 2016 (No Data)",
    "Winter 2016-2017 (No Data)", 
    "Spring 2020 (No Data)"
  )

add_season_year <-
  . %>%
  arrange(date_time_utc) %>%
  mutate(
    .after = season,
    season_year = case_when(
      str_detect(season, "Aut|Spr|Sum") ~ glue("{season} {year}"),
      str_detect(month, "Jan|Feb")      ~ glue("{season} {year-1}-{year}"),
      str_detect(month, "Dec")          ~ glue("{season} {year}-{year+1}")),
    
    season_year = fct_inorder(season_year),
    season_year = fct_expand(season_year, extra_fct),
    season_year = fct_relevel(season_year, extra_fct[1], after = 1),
    season_year = fct_relevel(season_year, extra_fct[2], after = 3),
    season_year = fct_relevel(season_year, extra_fct[3], after = 16)
  ) %T>% print()
```

## 4.2 Plot: Bar Plot: PFT per **Season** and **Year** as Stacked Percent


```{r chemtax-stacked-bar-year-season}
# CHEMTAX stacked bar graph with ratio to chlor-a, organized by station
cmtx_l %>%
  add_season_year %T>%
  print() %>%
  # mutate(station = fct_reorder(station, lon, .na_rm = FALSE)) %>%
  # filter(!is.na(station))
  ggplot(
    aes(
      y     = conc,
      x     = station,
      fill  = names,
      color = names
    )
  ) +
  geom_bar(
    position = "fill",
    stat     = "identity",
    na.rm    = TRUE
  ) +
  labs(
    x     = "Station",
    y     = "Relative Abundance (%)",
    fill  = "Phytoplankton Functional Type",
    color = "Phytoplankton Functional Type"
  ) +
  scale_fill_manual(values = tol_muted) +
  scale_color_manual(values = tol_muted) +
  scale_y_continuous(
    labels = scales::label_percent(),
    expand = c(0, 0)
  ) +
  facet_wrap(~season_year, ncol = 4, drop = FALSE) +

  # theme
  theme(
    axis.title       = element_text(size = 18),
    legend.position  = c(0.6, 0.05),
    legend.direction = "horizontal"
  )

# ---- save plot
save_gg(
  plt            = last_plot(), # edit if want specific plot
  save_location  = here(dir_plt_save, "percent_season_year"),
  save_name      = "pft_bar_percent_season_year",
  verbose        = TRUE,
  overwrite      = FALSE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "jpeg",
  height         = 9.7,
  width          = 19,
  dpi            = 600,
  units          = "in",
  bg             = "white"
  )


```


## 4.3 Plot: Bar Plot: PFT per **Season** and **Year** as Absolute VAr

```{r chemtax-stacked-bar-conc-year-season}
# chemtax stacked bar graph (not ratio), organized by station
cmtx_l %>%
  add_season_year  %T>% 
  print() %>%
  summarize(
    conc = mean(conc, na.rm = TRUE),
    .by  = c(season_year, station, phyto)
  ) %>%
  ggplot(aes(
    fill  = phyto,
    color = phyto,
    # fill  = func_type,
    # color = func_type,
    y     = conc,
    x     = station
  )) +
  geom_bar(stat = "identity") +
  labs(
    x     = "Station",
    y     = "Concentration of Chlorophyll-a (mg m<sup>-3</sup>)",
    fill  = "Phytoplankton Functional Type",
    color = "Phytoplankton Functional Type"
  ) +
  scale_fill_manual(values  = tol_muted) +
  scale_color_manual(values = tol_muted) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(~season_year, ncol = 4, drop = FALSE) +

  # theme
  theme(
    axis.title.y     = ggtext::element_markdown(),
    legend.position  = c(0.6, 0.05),
    legend.direction = "horizontal"
  )


# ---- save plot
save_gg(
  plt            = last_plot(), # edit if want specific plot
  save_location  = here(dir_plt_save, "absolute_season_year"),
  save_name      = "pft_bar_absolute_season_year",
  verbose        = TRUE,
  overwrite      = FALSE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "jpeg",
  height         = 9.7,
  width          = 19,
  dpi            = 600,
  units          = "in",
  bg             = "white"
  )


```




## 4.4 Plot: Average PFT per Season with Standard Deviation Error Bars

Calculate per PFT:
- average 
- standard deviation

```{r cmtx-summary}
cmtx_w %>%
  summarise(
    .by = season,
    across(
      # chloro:hapt,
      chloro:pras,
      list(
        avg = \(x) { mean(x, na.rm = TRUE)},
        sd  = \(x) { sd(x, na.rm = TRUE)},
        var = \(x) { var(x, na.rm = TRUE)}
      )
    )
  ) %>%
  select(-contains(c("var"))) %>%
  pivot_longer(
    data      = .,
    cols      = -season,
    names_to  = "phyto",
    values_to = "conc", 
    ) %>%
  separate(phyto, into = c("phyto", "type"), sep = "_") %>%
  pivot_wider(
    data         = .,
    names_from   = type, 
    values_from  = c(conc), 
    names_repair = janitor::make_clean_names
    ) %>%
  left_join(phyto_name, by = c("phyto" = "shrt")) %T>% 
  print() %>%
  ggplot(aes(x = season, y = avg, color = season)) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_point() +
  geom_linerange(aes(ymin = avg - sd, ymax = avg + sd ), key_glyph = draw_key_rect) +
  scale_y_continuous(breaks = scales::breaks_pretty()) +
  labs(
    x = "Season",
    y = bquote("Concentration"~(mg~m^-3)),
    color = NULL
  ) +
  facet_wrap(~names, ncol = 3, scales = "free_y") +
  theme_bw(base_family = "serif") +
  theme(legend.position = "top")

# ---- save plot
save_gg(
  plt            = last_plot(), # edit if want specific plot
  save_location  = here(dir_plt_save, "pft_average_season"),
  save_name      = "pft_avg_per_season",
  verbose        = TRUE,
  # overwrite      = FALSE,
  overwrite      = TRUE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "jpeg",
  height         = 6,
  width          = 10,
  dpi            = 600,
  units          = "in",
  bg             = "white"
  )

```