---
title: "Seasonal CHEMTAX Anova Tests"
author: "Sebastian Di Geronimo"
date: "2024-01-30"
format: html
editor: source
---

# 1.0 ---- Summary of Document ----



# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
  # additional
  rlang, cli,
  rstatix, readxl, ggtext, cowplot,
  
  ggstatsplot, vegan
  )

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
    )

source(here("scripts", "misc_functions.R"))
```

## 2.2 Load Dataset

```{r load-data}
# run snippet after saving file
# set location to save data and plots
dir_data_save <- here("data", "processed", "04_6_anova")
dir_plt_save  <- here("data", "plots", "04_6_anova")

dir_create(here(dir_plt_save))

# ---- set season names and order ---- #
szn <- c("Winter", "Spring", "Summer", "Autumn")

# ---- phytoplankton full names ---- #
phyto_name <-
  here("data", "processed", "04_1_merge_meta_cmtx", "phyto_names.csv") %>%
  read_csv(show_col_types = FALSE) %T>% 
  print()

# ---- load chemtax data in wide format ---- #
cmtx_w <-
  here("data", "processed", "04_1_merge_meta_cmtx", "chemtax_w_meta_wide.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  filter(!str_detect(station, "60|21.5|24|^3$|9.5")) %>%
  mutate(
    season = fct(season, levels = szn),
    # station = str_replace(station, "9B", "9.5"),
    station = fct_reorder(station, lon),
    labels = 
      if_else(
        str_detect(station, "10|16|7"), 
        "Nearshore", 
        "Offshore"
        )
  ) %>% 
  select(-c(hplc_gsfc_id, cruise:bottle, cluster_code)) %>%
    
  # add PFT concentrations to get total chlorophyll-a
  rowwise() %>%
  mutate(
    total_chl_a = sum(c_across(chloro:pras)),
    .before = chloro
  ) %>%
  ungroup() %>%
  filter(
    .by = c(station, cruise_id),
    depth == min(depth)
    & depth < 20
  ) %T>% 
  print()

```

```{r convert-long-format}
# ---- convert to long format ---- #
# i.e. one column for all PFT and one for concentrations
cmtx_l <-
  cmtx_w %>%
  pivot_longer(
    cols      = chloro:pras,
    names_to  = "phyto",
    values_to = "conc"
  ) %>%
  left_join(phyto_name, by = c("phyto" = "shrt")) %>%
  
  # add percent
  mutate(percent = conc/total_chl_a) %T>% 
  print()

# ---- convert to percent ---- #
cmtx_l_pct <-
  cmtx_l %>%
  select(-conc) %T>% 
  print()

```

# 3.0 ---- Perform Kruskal-Wallis and Dunn Test ----

Here Kruskal-Wallis was performed to identify if there was any significant 
difference between each season and a PFT.

If p-value <= 0.05, next would be pairwise Dunn's test to test season-phyto 
combination for significance.


## 3.1 Abs. ---- K-W PFT on season

Absolute total chlorophyll-a


```{r absolute-kruskal-and-dunn}
# concentration on season grouped by PFT
cmtx_kruskal_season <-
  cmtx_l %>% 
  group_by(phyto) %>% 
  kruskal_test(conc ~ season) %>%
  mutate(p2_adj = scales::label_pvalue(0.05)(p), .before = 3) %>%
  arrange(p)

# concentration on stn grouped by PFT
cmtx_kruskal_stn <-
  cmtx_l %>% 
  group_by(phyto) %>% 
  kruskal_test(conc ~ station) %>%
  mutate(p2_adj = scales::label_pvalue(0.05)(p), .before = 3) %>%
  arrange(p)

# extract PFT where p value > 0.05
p_gt_05_season <- filter(cmtx_kruskal_season, p > 0.05)
p_gt_05_stn    <- filter(cmtx_kruskal_stn, p > 0.05)

# Seasonal Dunn's test 
cmtx_dunn_season <-
  cmtx_l %>%
  # remove PFT with no signficant difference
  filter(!(phyto %in% p_gt_05_season$phyto)) %>%
  group_by(phyto) %>%
  dunn_test(
    conc ~ season, 
    p.adjust.method = "bonferroni",
    detailed = TRUE
    ) %>%
  mutate(p2_adj = scales::label_pvalue(0.05)(p.adj), .before = 3) %>%
  arrange(phyto, p.adj)

# Station Dunn's test
cmtx_dunn_stn <-
  cmtx_l %>%
  # remove PFT with no signficant difference
  filter(!(phyto %in% p_gt_05_stn$phyto)) %>%
  group_by(phyto) %>%
  dunn_test(
    conc ~ station, 
    p.adjust.method = "bonferroni",
    detailed = TRUE
    ) %>%
  mutate(p2_adj = scales::label_pvalue(0.05)(p.adj), .before = 3) %>%
  arrange(phyto, p.adj)
```


### 3.1.1 Abs. Summary: Season

Overall not sig. diff.:
- diat1
- dino

Based on the results, Diatom1 and Dinoflagellates did not have any significant 
difference between each season. 

From the pairwise Dunn's test, only 25/48 pairs were significantly different.

```{r season-absolute-kruskal-and-dunn}
cmtx_kruskal_season

# remove diat1 and dino
p_gt_05_season

cmtx_dunn_season

# sig Dunn
cmtx_dunn_season %>%
  filter(p.adj <= 0.05) %>% # remove PFTs with p>0.05
  unite("groups", group1, group2, sep = "-") 

# no sig Dunn
cmtx_dunn_season %>%
  filter(p.adj > 0.05) %>% 
  unite("groups", group1, group2, sep = "-")
```

### 3.1.2 Abs. Summary: Station

Overall not sig. diff.:
- chloro
- cyano2 
- hapt6 
- pras

Based on the results, Chlorophytes, Cyanobacteria2, Haptophyte6 and Prasinophyte
did not have any significant difference between each season. 

From the pairwise Dunn's test, only 79/168 pairs were significantly different.

```{r station-absolute-kruskal-and-dunn}
cmtx_kruskal_stn

# remove chloro, cyano2, hapt6 and pras
p_gt_05_stn

cmtx_dunn_stn

# sig Dunn
cmtx_dunn_stn %>%
  filter(p.adj <= 0.05) %>% 
  unite("groups", group1, group2, sep = "-") %T>% 
  print() %>%
  count(phyto)

# no sig Dunn
cmtx_dunn_stn %>%
  filter(p.adj > 0.05) %>% 
  unite("groups", group1, group2, sep = "-")  %T>% 
  print() %>%
  count(phyto)

```


## 3.2 Abs. Statsplot: Kruskal-Wallis on PFT Concentrations

```{r k-w-statplot, fig.width=25, fig.asp=0.5}
cmtx_stat_plt <- 
  cmtx_l %>%
  nest(.by = names) %>%
    
  mutate(
    stat_plt_year =
      map2(
        .x = data,
        .y = names,
        ~ ggbetweenstats(
          data = .x,
          x = year,
          y = conc,
          xlab = "Season",
          ylab = "Concentration",
          title = glue("Year: {.y}"),
          type = "nonparametric"
        )
      ),
    stat_plt_season =
      map2(
        .x = data,
        .y = names,
        ~ ggbetweenstats(
          data = .x,
          x = season,
          y = conc,
          xlab = "Season",
          ylab = "Concentration",
          title = glue("Season: {.y}"),
          type = "nonparametric"
        )
      )
  ) 

pull(cmtx_stat_plt, stat_plt_season)
pull(cmtx_stat_plt, stat_plt_year)

```

## 3.3 Rel. ---- K-W Abundance of PFTs

```{r relative-kruskal-and-dunn}
# concentration on season grouped by PFT
cmtx_kruskal_season <-
  cmtx_l %>% 
  group_by(phyto) %>% 
  kruskal_test(percent ~ season) %>%
  mutate(p2_adj = scales::label_pvalue(0.05)(p), .before = 3) %>%
  arrange(p)

# concentration on stn grouped by PFT
cmtx_kruskal_stn <-
  cmtx_l %>% 
  group_by(phyto) %>% 
  kruskal_test(percent ~ station) %>%
  mutate(p2_adj = scales::label_pvalue(0.05)(p), .before = 3) %>%
  arrange(p)

# concentration on stn grouped by PFT
cmtx_kruskal_dist <-
  cmtx_l %>% 
  group_by(phyto) %>% 
  kruskal_test(percent ~ labels) %>%
  mutate(p2_adj = scales::label_pvalue(0.05)(p), .before = 3) %>%
  arrange(p)

# extract PFT where p value > 0.05
p_gt_05_season <- filter(cmtx_kruskal_season, p > 0.05)
p_gt_05_stn    <- filter(cmtx_kruskal_stn, p > 0.05)
p_gt_05_dist   <- filter(cmtx_kruskal_dist, p > 0.05)

# Seasonal Dunn's test 
cmtx_dunn_season <-
  cmtx_l %>%
  # remove PFT with no signficant difference
  filter(!(phyto %in% p_gt_05_season$phyto)) %>%
  group_by(phyto) %>%
  dunn_test(
    percent ~ season, 
    p.adjust.method = "bonferroni",
    detailed = TRUE
    ) %>%
  mutate(p2_adj = scales::label_pvalue(0.05)(p.adj), .before = 3) %>%
  arrange(phyto, p.adj)

# Station Dunn's test
cmtx_dunn_stn <-
  cmtx_l %>%
  # remove PFT with no signficant difference
  filter(!(phyto %in% p_gt_05_stn$phyto)) %>%
  group_by(phyto) %>%
  dunn_test(
    percent ~ station, 
    p.adjust.method = "bonferroni",
    detailed = TRUE
    ) %>%
  mutate(p2_adj = scales::label_pvalue(0.05)(p.adj), .before = 3) %>%
  arrange(phyto, p.adj)

# Distance Dunn's test
cmtx_dunn_dist <-
  cmtx_l %>%
  # remove PFT with no signficant difference
  filter(!(phyto %in% p_gt_05_dist$phyto)) %>%
  group_by(phyto) %>%
  dunn_test(
    percent ~ labels, 
    p.adjust.method = "bonferroni",
    detailed = TRUE
    ) %>%
  mutate(p2_adj = scales::label_pvalue(0.05)(p.adj), .before = 3) %>%
  arrange(phyto, p.adj)
```

### 3.3.1 Rel. Summary: Season

Overall not sig. diff.:
- Diatom1

Based on the results, Diatom1 did not have any significant difference between 
each season. 

From the pairwise Dunn's test, only 29/54 pairs were significantly different.

```{r season-relative-kruskal-and-dunn}
cmtx_kruskal_season

# remove diat1
p_gt_05_season 

cmtx_dunn_season

# sig Dunn
cmtx_dunn_season %>% 
  filter(p.adj <= 0.05) %>% 
  unite("groups", group1, group2, sep = "-") %T>%
  print() %>%
  count(phyto)

# no sig Dunn
cmtx_dunn_season %>% 
  filter(p.adj > 0.05) %>% 
  unite("groups", group1, group2, sep = "-") %T>%
  print() %>%
  count(phyto)
```

### 3.3.2 Rel. Summary: Station

Overall not sig. diff.:
- chloro
- pras

Based on the results, Chlorophytes and Prasinophytes did not have any 
significant difference between each season. 

From the pairwise Dunn's test, only 89/224 pairs were significantly different.

```{r station-relative-kruskal-and-dunn}
cmtx_kruskal_stn

# remove chloro and pras
p_gt_05_stn 

cmtx_dunn_stn

# sig Dunn
cmtx_dunn_stn %>%
  filter(p.adj <= 0.05) %>%
  unite("groups", group1, group2, sep = "-") %T>%
  print() %>%
  count(phyto)

# no sig Dunn
cmtx_dunn_stn %>%
  filter(p.adj > 0.05) %>%
  unite("groups", group1, group2, sep = "-")  %T>%
  print() %>%
  count(phyto)
```


### 3.3.3 Rel. Summary: Distance from Shore

Overall not sig. diff.:
- chloro
- pras

Based on the results, Chlorophytes and Prasinophytes did not have any significant 
difference between each season. 

From the pairwise Dunn's test, all 8 pairs were significantly different.

```{r distance-relative-kruskal-and-dunn}
cmtx_kruskal_dist

# remove chloro and pras
p_gt_05_dist 

cmtx_dunn_dist

# sig Dunn
cmtx_dunn_dist %>%
  filter(p.adj <= 0.05) %>%
  unite("groups", group1, group2, sep = "-") %T>%
  print() %>%
  count(phyto)

# no sig Dunn
cmtx_dunn_dist %>%
  filter(p.adj > 0.05) %>%
  unite("groups", group1, group2, sep = "-")  %T>%
  print() %>%
  count(phyto)

```

Based on the results, Diatom 1 did not have any significant difference between 
each season. 

From the pairwise Dunn's test, only 30 pairs were significantly different.


## 3.4 Rel. Statsplot: Kruskal-Wallis on PFT Concentrations

```{r relative-kruskal-wallis, fig.width=25, fig.asp=0.5}
cmtx_stat_plt <- 
  cmtx_l %>%
  nest(.by = names) %>%
  # slice(1) %>%
    
  mutate(
    stat_plt_year =
      map2(
        .x = data,
        .y = names,
        ~ ggbetweenstats(
          data = .x,
          x = year,
          y = percent,
          xlab = "Season",
          ylab = "Concentration",
          title = glue("Year: {.y}"),
          type = "nonparametric"
        )
      ),
    stat_plt_season =
      map2(
        .x = data,
        .y = names,
        ~ ggbetweenstats(
          data = .x,
          x = season,
          y = percent,
          xlab = "Season",
          ylab = "Concentration",
          title = glue("Season: {.y}"),
          type = "nonparametric"
        )
      )
  ) 

pull(cmtx_stat_plt, stat_plt_season)
pull(cmtx_stat_plt, stat_plt_year)

```


# 4.0 ---- Pairwise Permanova Test ----

- CTD vars (temp, par and sal) vs season
- PFTs vs season
  - pairwise seasons

Notes about PERMANOVA (use `betadisper`):
- homogeniety of dispersion:
  - needs to be check under the assumption of PERMANOVA
  - dispersion per grouping factor (i.e. variance within groups) is not 
    significantly different
  - however, this may be able to be ignored if when plotted, there is clear 
    group separation
  - if there is a significant difference, then you can use the "strata" argument
    in the `adonis2` function to account for this
  - this test is to check if the differences we see between groups is due to 
    differences in dispersion and not due to differences in the mean 
    (i.e. centroid, or "location")     
    - we want it to be non-significant so we can check differences between 
      groups
    
- R-squared:
  - This is the proportion of sum of squares from the total
  - don't over interpret this!
  - wrong: this shows the influence of the factor on explained variation, this 
    will tell you the proportion of total variance explained by the grouping 
    factor
    
- the p-value
  - the probability of observing the effect size by chance
  
- pairwise comparison:
  - performed if the p-value is significant to determine which groups are 
    significantly different from each other




## 4.1 Check Homogeneity of Dispersion


```{r dispersion}
set.seed(1082002)
# PFT conc vs season
cat("\n-----------------------\n\n---- PFT conc vs season ----\n\n")
manova_pft_season <-
  vegdist(
    select(cmtx_w, chloro:pras),
    method = "bray"
  ) %>%
  betadisper(
    group = cmtx_w$season
  ) %T>% print()

TukeyHSD(manova_pft_season)
anova(manova_pft_season)
plot(manova_pft_season)
boxplot(manova_pft_season)


cat("\n-----------------------\n\n---- PFT conc vs distance from shore ----\n\n")
manova_pft_season_dist <-
  vegdist(
    select(cmtx_w, chloro:pras),
    method = "bray"
  ) %>%
  betadisper(
    group = cmtx_w$labels
  ) %T>% print()

TukeyHSD(manova_pft_season_dist)
anova(manova_pft_season_dist)
plot(manova_pft_season_dist)
boxplot(manova_pft_season_dist)

cat("\n-----------------------\n\n---- PFT perc vs season ----\n\n")

manova_pft_season_pct <-
  cmtx_w %>%
  mutate(across(chloro:pras, ~ .x/total_chl_a)) %>%
  select(chloro:pras) %>%
  vegdist(
    .,
    method = "bray"
  ) %>%
  betadisper(
    group = cmtx_w$season
  ) %T>% print()    

TukeyHSD(manova_pft_season_pct)
anova(manova_pft_season_pct)
plot(manova_pft_season_pct)
boxplot(manova_pft_season_pct)

```

### 4.2 Function: Permutational MANOVA

```{r season-permanova}
#' Pairwise Permutation MANOVA
#'
#' FUNCTION_DESCRIPTION
#'
#' @param .data DESCRIPTION.
#' @param pair_col DESCRIPTION.
#' @param pairs DESCRIPTION.
#' @param formula DESCRIPTION.
#' @param dist_method DESCRIPTION.
#' @param strata DESCRIPTION.
#'
#' @return RETURN_DESCRIPTION
#' @examples
#' # ADD_EXAMPLES_HERE
szn_adonis2 <- function(
        .data, 
        pair_col, 
        pairs, 
        formula = NULL, # something like "season * station" 
        dist_method = "bray",
        strata = NULL) {

    dat_pair  <- filter(.data, str_detect( {{ pair_col }}, pairs))
    value     <- select(dat_pair, chloro:pras) 
    
    if (!is.null(strata)) strata <- dat_pair[[strata]]
    
    permanova_results <- 
      adonis2(
        formula  = as.formula(paste("value ~", formula)), 
        data     = dat_pair, 
        method   = dist_method,
        strata   =  strata 
        )   
    
    
    return(
      list(
        "r_squared"  = permanova_results[["R2"]][1],
        "pairwise_p" = permanova_results[["Pr(>F)"]][1],
        "results"    = permanova_results
        )
    )
    
  # ---- end of szn_adonis2 function    
}
```



### 4.3 Perform Permutational MANOVA

```{r season-permanova}
set.seed(1082002)
# PFT vs season
cat("\n-----------------------\n\n---- PFT conc vs season ----\n\n")
adonis2(
  select(cmtx_w, chloro:pras) ~ season,
  data   = select(cmtx_w, season),
  method = "bray"
)

# PFT vs distance
cat("\n-----------------------\n\n---- PFT conc vs distance ----\n\n")
adonis2(
  select(cmtx_w, chloro:pras) ~ labels,
  data   = select(cmtx_w, labels),
  method = "bray"
)

cat("\n-----------------------\n\n---- PFT conc vs distance + season ----\n\n")
adonis2(
  select(cmtx_w, chloro:pras) ~ labels + season,
  data   = select(cmtx_w, labels, season),
  method = "bray"
)

cat("\n-----------------------\n\n---- Pairwise Test ----\n\n")
# pairwise test
permanova_r <-
  combn(szn, 2) %>%
  t() %>%
  as_tibble() %>%
  set_colnames(c("szn1", "szn2")) %>%
  unite(col = "season_pairs", szn1, szn2, sep = "|") %>%
  mutate(
    pairwise_p =
      map(
        .x = season_pairs,
        .f =
          \(.x) {
             
           cat("------\n\n", .x, "\n\n") 
            
          szn_adonis2(
            .data    = select(.env$cmtx_w, season, chloro:pras),
            pairs    = .x,
            pair_col = season,
            formula  = "season",
            strata   = NULL
            # strata   = "station"
          ) %T>% 
              print()
          }
      ),
  ) %>%
  unnest_wider(pairwise_p) %>%
    
  # ---- correct p-values
  mutate(
    p_adjust = p.adjust(pairwise_p, method = "bonferroni"),
    .before = results
  ) %T>% print()


unnest(permanova_r, results)
permanova_r$results


```



# 5.0 *WIP*: Mann-Kendall Test

Probably won't use this 

Null hypothesis is that the data come from a population with independent 
realizations and are identically distributed

```{r}
trend::mk.test()


cmtx_l %>%
  filter(phyto == "chloro") %>%
  arrange(date_time_utc) %$%
  trend::mk.test(conc) 

```

# ---- 6.0 *WIP* MANOVA ----

Doesn't really work


```{r}
# # ---- CHEMTAX Season-Year Labels ---- #
# extra_fct <- 
#   c(
#     "Summer 2016 (No Data)",
#     "Winter 2016-2017 (No Data)", 
#     "Spring 2020 (No Data)"
#   )
# 
# add_season_year <-
#   . %>%
#   arrange(date_time_utc) %>%
#   mutate(
#     .after = season,
#     season_year = case_when(
#       str_detect(season, "Aut|Spr|Sum") ~ glue("{season} {year}"),
#       str_detect(month, "Jan|Feb")      ~ glue("{season} {year-1}-{year}"),
#       str_detect(month, "Dec")          ~ glue("{season} {year}-{year+1}")),
#     
#     season_year = fct_inorder(season_year),
#     season_year = fct_expand(season_year, extra_fct),
#     season_year = fct_relevel(season_year, extra_fct[1], after = 1),
#     season_year = fct_relevel(season_year, extra_fct[2], after = 3),
#     season_year = fct_relevel(season_year, extra_fct[3], after = 16)
#   ) 

```


```{r dunn-season-year}
# cmtx_dunn2 <-
#   cmtx_l %>%
#   add_season_year %>%
#   group_by(phyto) %>%
#   dunn_test(
#     conc ~ season_year,
#     p.adjust.method = "bonferroni",
#     detailed = TRUE
#   ) %>%
#   mutate(
#     p     = num(p, digits = 3),
#     p.adj = num(p.adj, digits = 3),
#   ) %>%
#   nest(.by = phyto) %$%
#   walk2(
#     .x = phyto,
#     .y = data,
#     (\(x, y) mutate(y,
#       phyto = x,
#       .before = 1
#     ) %>%
#       filter(p < 0.1) %>%
#       print())
#   )
```


## 5.2 MANOVA Test


```{r manova}
# cmtx_man <- 
#   cmtx_w %>%
#   select(chloro:last_col()) %>%
#   as.matrix() 
# 
# manova(cmtx_man ~ season, data = cmtx_w, season) %>%
#     summary()
# 
# cmtx_w %>%
#   mutate(season_year = str_c(season, year(date_time_utc), sep = " "),
#          .before = 2) %>%
#   manova(cmtx_man ~ season_year, data = .) %>%
#   summary()
```

