---
title: "Time Series for Stations CHEMTAX"
author: "Sebastian Di Geronimo"
date: "2024-01-30"
format: html
editor: source
---

# 1.0 ---- Summary of Document ----



# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
  # additional
  rlang, cli,
  rstatix, 
  # readxl, 
  ggtext, cowplot
  )

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
    )

source(here("scripts", "misc_functions.R"))

# set colors
# alternate color palette
# tol_muted <- 
# c('#88CCEE', '#44AA99', '#117733', '#332288', '#DDCC77', '#999933',
#   '#CC6677', '#882255', '#AA4499')

tol_muted <- khroma::color("muted", names = FALSE)(9)

```

## 2.2 Load Dataset

```{r load-data}
# set location to save data and plots
dir_data_save <- here("data", "processed", "04_7_time_series_stations")
dir_plt_save  <- here("data", "plots", "04_7_time_series_stations")

dir_create(here(dir_plt_save))

# ---- set season names and order ---- #
szn <- c("Winter", "Spring", "Summer", "Autumn")

# ---- phytoplankton full names ---- #
phyto_name <-
  here("data", "processed", "04_1_merge_meta_cmtx", "phyto_names.csv") %>%
  read_csv(show_col_types = FALSE) %T>% 
  print()

# ---- load chemtax data in wide format ---- #
cmtx_w <-
  here("data", "processed", "04_1_merge_meta_cmtx", "chemtax_w_meta_wide.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  # filter(!str_detect(station, "60|21.5|24|^3$|9B|9.5")) %>%
  filter(!str_detect(station, "60|21.5|24|^3$|9.5")) %>%
  mutate(
    season = fct(season, levels = szn),
    # station = str_replace(station, "9B", "9.5"),
    station = fct_reorder(station, lon)
  ) %>% 
  select(-c(hplc_gsfc_id, cruise:bottle, cluster_code, water_type)) %>%
    
  # add PFT concentrations to get total chlorophyll-a
  rowwise() %>%
  mutate(
    total_chl_a = sum(c_across(chloro:pras)),
    .before = chloro
  ) %>%
  ungroup() %>%
  filter(
    .by = c(station, cruise_id),
    depth == min(depth)
    & depth < 20
  ) %T>% 
  print()

```

```{r convert-long-format}
# ---- convert to long format ---- #
# i.e. one column for all PFT and one for concentrations
cmtx_l <-
  cmtx_w %>%
  pivot_longer(
    cols      = chloro:pras,
    names_to  = "phyto",
    values_to = "conc"
  ) %>%
  left_join(phyto_name, by = c("phyto" = "shrt")) %>%
  
  # add percent
  mutate(percent = conc/total_chl_a) %T>% 
  print()

# ---- convert to percent ---- #
cmtx_l_pct <-
  cmtx_l %>%
  select(-conc) %T>% 
  print()

```

## 2.3 Combine PFT Sub Groups

Groups to be combined have multiple types but are merged for ease of use
- diatom:
  - diatom 1 
  - diatom 2
- cyanobacteria:
  - cyano 2
  - cyano 4
- haptophyte:
  - hapt 6
  - hapt 8

```{r combine-pft}
# combine same base PFT groups into main group
cmtx_w_comb <-
  cmtx_w %>%
  rowwise() %>%
  mutate(
    .keep = "unused",       # remove used columns
    diat  = diat1  + diat2,
    cyano = cyano2 + cyano4,
    hapt  = hapt6  + hapt8
  ) %>%
  ungroup() %T>%
  print()

# calculate PFT fractions (0 - 100%)
cmtx_fract_w_comb <-
  cmtx_w_comb %>%
  mutate(
    .keep = "unused",
    across(
      c(pras:hapt, chloro, crypto, dino),
      (\(x) x / total_chl_a),
      .names = "percent_{.col}"
    )
  ) %T>% print()

# ---- with combining diat, cyano and hapt ---- #
cmtx_fract_l_comb <- 
  cmtx_fract_w_comb %>%
  pivot_longer(
    contains("percent"),
    names_to     = "phyto",
    values_to    = "percent",
    names_prefix = "percent_"
  ) %>%
    
  # add long names
  left_join(phyto_name, by = c("phyto" = "shrt")) %T>% print()
```


# 3.0 ---- Time Series of 3 Stations with 2 PFTs -----

## 3.1 Plot: Calculate Fraction and Plot 3 Stations with 2 PFTs (Diatom and Cyano)

Stations 10, 16, WS: Plot Cyano and Diat vs time and 

Stations:
- 10
- 16
- WS

PFT:
- cyanobacteria (cyano2 and cyano4)
- diatom (diat1 and diat2)

Shows opposing signals where lower diatoms shows increase in cyano relatively

Issue is this doesn't show abosolute, where cyano could have incrased thus the
ratio is farther off.

```{r plt-cyano-diat, fig.height=8, fig.width=16}
time_series_base <- 
  . %>%
  # filter for 2 PFTs and 3 stations
  filter(
    str_detect(phyto, "diat|cyano")
    & station %in% c("WS", "10", "16")
   ) %T>% print()
  
cmtx_l_pct %>%
  time_series_base %>%
  # plot
  ggplot(aes(
    x     = lubridate::ymd_hms(date_time_utc),
    # y     = frac,
    y     = percent,
    color = names,
    fill  = names
  )) +
  geom_point() +
  geom_line() +
  
  labs(
    x     = NULL,
    y     = "Relative Abundance (%)",
    color = "Functional Type",
    fill  = "Functional Type"
  ) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     expand = c(0,0),
                     limits = c(0,1)) +
  facet_wrap(facets = "station") +
  theme_classic() +
  theme(
    text                = element_text(family = "serif", size = 40),
    axis.text           = element_text(size = 30),
    axis.text.x         = element_text(angle = 45, hjust = 1), 
    axis.ticks.length.x = unit(10, "pt"),
    legend.position     = c(0.5, 0.8),
    panel.spacing       = unit(20, "pt") 
  )

# ---- save plot
save_gg(
  plt            = last_plot(), # edit if want specific plot
  save_location  = dir_plt_save,
  save_name      = "stn_subset_frac_sep_diat_cyano",
  verbose        = TRUE,
  overwrite      = FALSE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "jpeg",
  height         = 8,
  width          = 16,
  dpi            = 600,
  units          = "in",
  bg             = "white"
  )
```


## 3.2 Plot: Commbined Fraction and Plot 3 Stations with 2 PFTs (Diatom and Cyano)

```{r plt-comb-cyano-diat, fig.height=8, fig.width=16}
cmtx_fract_l_comb %>%
  time_series_base %>%
  # plot
  ggplot(aes(
    x     = lubridate::ymd_hms(date_time_utc),
    y     = percent,
    color = names,
    fill  = names
  )) +
  geom_point() +
  geom_line() +
  
  labs(
    x     = NULL,
    y     = "Relative Abundance (%)",
    color = "Functional Type",
    fill  = "Functional Type"
  ) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     expand = c(0,0),
                     limits = c(0,1)) +
  facet_wrap(facets = "station") +
  theme_classic() +
  theme(
    text                = element_text(family = "serif", size = 40),
    axis.text           = element_text(size = 30),
    axis.text.x         = element_text(angle = 45, hjust = 1), 
    axis.ticks.length.x = unit(10, "pt"),
    legend.position     = c(0.5, 0.8),
    panel.spacing       = unit(20, "pt") 
  )

# ---- save plot
save_gg(
  plt            = last_plot(), # edit if want specific plot
  save_location  = dir_plt_save,
  save_name      = "stn_subset_frac_comb_diat_cyano",
  verbose        = TRUE,
  overwrite      = FALSE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "jpeg",
  height         = 8,
  width          = 16,
  dpi            = 600,
  units          = "in",
  bg             = "white"
  )


```

## 3.3 Timeseries per Station


```{r diat-ts}
# time series for diatom at specific stations
cmtx_w_comb %>%
  pivot_longer(
    cols      = chloro:hapt,
    names_to  = "func_type",
    values_to = "conc"
  ) %>%
  print() %>%
  ggplot(
    aes(
      x     = date_time_utc, 
      y     = conc, 
      color = func_type, 
      fill  = func_type
      )
    ) +
  geom_point() +
  geom_line() +
  facet_wrap(facets = ~ station) +
  ggthemes::theme_clean()
```
