---
title: "Time Series Plots"
author: "Sebastian Di Geronimo"
date: "2024-01-30"
format: html
editor: source
---

# 1.0 ---- Summary of Document ----



# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
    # additional
    ggthemes
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )
  
theme_set(theme_calc(base_family  = "serif"))
```

## 2.2 Load Dataset

```{r load-data}
# set location to save data and plots
dir_data_save <- here("data", "processed", "05_1_time_series_plots")
dir_plt_save  <- here("data", "plots", "05_1_time_series_plots")

dir_create(here(dir_plt_save))

# ---- set season names and order ---- #
szn <- c("Winter", "Spring", "Summer", "Autumn")

# ---- phytoplankton full names ---- #
phyto_name <-
  here("data", "processed", "04_1_merge_meta_cmtx", "phyto_names.csv") %>%
  read_csv(show_col_types = FALSE) %T>% 
  print()

# ---- load chemtax data in wide format ---- #
cmtx_w_og <-
  here("data", "processed", "04_1_merge_meta_cmtx", "chemtax_w_meta_wide.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(
    .after = cluster_code,
    across(
      chloro:pras,
      ~ if_else(
        str_detect(sample, "WS20278-32"),
        .x / 10,
        .x
      )
    )
  ) %>%
  # filter(!str_detect(station, "60|21.5|24|^3$|9B|9.5")) %>%
  filter(!str_detect(station, "60|21.5|24|^3$|9.5")) %>%
  mutate(
    season = fct(season, levels = szn),
    # station = str_replace(station, "9B", "9.5"),
    station = fct_reorder(station, lon)
  ) %>%
  select(-c(hplc_gsfc_id, cruise:bottle, cluster_code, water_type)) %>%
  # add PFT concentrations to get total chlorophyll-a
  rowwise() %>%
  mutate(
    total_chl_a = sum(c_across(chloro:pras)),
    .before = chloro
  ) %>%
  ungroup() %>%
  filter(
    .by = c(station, cruise_id),
    depth == min(depth) &
      depth < 20
  ) %T>%
  print()

```


## 2.2.1 Remove Outliers?

```{r remove-outliers}
remove_out <- menu(c("yes", "no"), title = "Remove Outliers?")

if (remove_out == 1) {
  message("Filtering out outliers!")
  cmtx_w <-
    cmtx_w_og %>%
    filter(
      total_chl_a <= 1
    ) %T>% 
    print()
  
  file_suffix <- "_no_out"
} else {
  message("Keeping outliers!")
  cmtx_w <-
    cmtx_w_og 
  file_suffix <- ""
}
```


## 2.3 Convert to Long Format


Pivot longer by taking PFT columns and pivot to two columns of `phyto` for name
and `conc` for concentration.

Next, convert to percentage by `conc` / `total_chl_a`


```{r convert-long-format}
# ---- convert to long format ---- #
# i.e. one column for all PFT and one for concentrations
cmtx_l <-
  cmtx_w %>%
  pivot_longer(
    cols      = chloro:pras,
    names_to  = "phyto",
    values_to = "conc"
  ) %>%
  left_join(phyto_name, by = c("phyto" = "shrt")) %>%
  
  # add percent
  mutate(percent = conc/total_chl_a) %T>% 
  print()

# ---- convert to percent ---- #
cmtx_l_pct <-
  cmtx_l %>%
  select(-conc) %T>% 
  print()

cmtx_w_pct <- 
  cmtx_l_pct %>%
  select(-names) %>%
  pivot_wider(
    data         = .,
    names_from   = c(phyto),
    values_from  = c(percent),
    names_sep    = "_",
    names_repair = janitor::make_clean_names
    )

# maybe delete
# # ---- convert to long format ---- #
# # i.e. one column for all PFT and one for concentrations
# cmtx_l <-
#   cmtx_w %>%
#   pivot_longer(
#     cols      = chloro:pras,
#     names_to  = "phyto",
#     values_to = "conc"
#   ) %>%
#   left_join(phyto_name, by = c("phyto" = "shrt"))
# 
# # ---- convert to percent ---- #
# cmtx_l_pct <-
#   cmtx_w %>%
#   mutate(across(chloro:pras, ~ .x/total_chl_a)) %>%
#   
#   pivot_longer(
#     cols      = chloro:pras,
#     names_to  = "phyto",
#     values_to = "conc"
#   ) %>%
#   left_join(phyto_name, by = c("phyto" = "shrt"))
# 
# # rm(file_chem, meta_data, cmtx_raw, cmtx_all_w)
```



# 4.0 Timeseries Plots

## 4.1 *WIP* Line Plot: Timesseries PFT with LOESS Smoother

This was an attempt to see if a smoother could show seasonal trends for each
PFT.

The 3 plots:
- for all y = concentration, facet = PFT
  1. entire series, x = datetime
  2. day of year, x = doy
  3. month, x = month

```{r pft-ts-loess, fig.width=30, fig.height=15}
# base plot list
plt_fun <-
  list(
    aes(
      y     = conc,
      color = names
    ),
    geom_point(
      show.legend = FALSE, ,
      na.rm       = TRUE
    ),
    # geom_smooth(
    #   formula     = y ~ s(x),
    #   method      = "gam",
    #   se          = FALSE,
    #   show.legend = FALSE
    # ),
    geom_smooth(
      formula     = y ~ x,
      method      = "loess",
      span        = 0.3,
      se          = FALSE,
      show.legend = FALSE,
      na.rm       = TRUE
    ),
    labs(
      x = NULL,
      y = "Pigment:Chlorophyll-a",
      color = NULL
    ),
    scale_y_continuous(expand = c(0, 0)),
    scale_color_brewer(palette = "Paired"),
    facet_grid(names ~ .,
      # scales = "free_y"
      scales = "fixed"
    )
  )


# point plot with smoothing line for 1 functional group, colored by season
plt_smooth <-
  cmtx_l_pct %>%
  rename("conc" = percent) %>%
  filter(!is.na(conc)) %>%
  ggplot() +
  plt_fun


# combine the three versions
shelf(patchwork)
(plt_smooth +
  aes(x = date_time_utc) +
        labs(x = "Date")
    ) +
  (plt_smooth +
    aes(x = yday(date_time_utc)) +
    labs(y = NULL,
         x = "Day of Year")
   ) +
  (plt_smooth +
    aes(x = month(date_time_utc)) +
    labs(y = NULL,
         x = "Month")
   ) 
```




```{r, fig.width=40, fig.asp=0.5}
cmtx_l_pct %>%
  rename("conc" = percent) %>%
  mutate(
    names2 = names
  ) %>%
  filter(
    .by = c(station, names),
    n() > 3,
    !is.na(conc)
  ) %>%
  nest(.by = names2) %>%
  # slice(1) %>%

  mutate(
    plt =
      map2(
        data,
        names2,
        ~ ggplot(
          .x,
          aes(
            x     = date_time_utc,
            y     = conc,
          )
          ) +
          geom_hline(yintercept = 0, color = "red", alpha = 0.2) +
          geom_smooth(
            color       = "grey70",
            alpha       = 0.5,
            formula     = y ~ s(x, bs = "tp", k = 10),
            method      = "gam",
            se          = FALSE,
            show.legend = FALSE,
            na.rm       = TRUE
          ) +
          geom_smooth(
            color       = "grey70",
            alpha       = 0.5,
            linetype    = "dashed",
            formula     = y ~ x,
            method      = "loess",
            span        = 0.3,
            se          = FALSE,
            show.legend = FALSE,
            na.rm       = TRUE
          ) +
          geom_smooth(
           aes(y = total_chl_a),
            color       = "red",
            alpha       = 0.8,
            linetype    = "dashed",
            formula     = y ~ x,
            method      = "loess",
            span        = 0.3,
            se          = FALSE,
            show.legend = FALSE,
            na.rm       = TRUE
          ) +
          geom_point(
            aes(y = total_chl_a),
            fill = "red",
            size = 3,
            shape = 24,
            na.rm = TRUE
          ) +
          geom_point(
            aes(fill = season),
            shape = 23,
            size  = 4,
            na.rm = TRUE
          ) +
          labs(title = .y) +
          scale_x_datetime(
            date_labels = "%Y",
            breaks = "1 year",
            minor_breaks = "3 months"
          ) +
          scale_y_continuous(
            expand = c(0, 0),
            limits = c(0, NA),
            breaks = seq(0.0, 1, 0.1),
            minor_breaks = seq(0.05, 1, 0.15)
          ) +
          facet_wrap(
            station ~ .,
            ncol   = 4,
            scales = "fixed"
          ) +
          coord_cartesian(ylim = c(0, 1)) +
          theme(
            panel.grid.major.x = element_line(color = "grey", linewidth = 0.5),
            panel.grid.minor = element_line(color = "grey", linewidth = 0.2,
                                              linetype = "dashed")
            )
      )
  ) %>%
  pull(plt)
```



What trends are visible?

Choloro
- range 0 - 15% of total chl-a, average ~5 - 7%
- cyclic pattern
- station WS and LK have peaks in Spring and Winter
  - lowest in Autumn
  
Crypto
- most stations 0 - 10% of total chl-a
- station 16 has largest range from 0 - 45% of total chl-a
  - nearer shore, near channel to West Florida Shelf
- no clear seasonal trend  


Cyano2
- range 0 - 50% of total chl-a
- always above 10% of total chl-a
- no clear trend
- potential bloom in station 16 in winter 2019 with chloro and diat2

Cyano 4 (prochlorococcus)
- station 16, 7, and 10 were lowest, with 0 - 5% of total chl-a
  - station 16 especially low for most of it with one spike in winter 2019
  - station 7 had more variabiliy with peaks ~8% in autumn and spring
- all other stations had 5 - 30% of total chl-a
- WS and LK seem to share similar trend
- MR

Diat 1 
- potential bloom in station 16 in winter 2018 autumn

```{r, fig.width=40, fig.asp=0.5}
cmtx_l_pct %>%
  rename("conc" = percent) %>%
  mutate(
    # conc = conc / total_chl_a,
    names2 = names
  ) %>%
  filter(
    .by = c(station, names),
    n() > 3,
    !is.na(conc)
  ) %>%
  nest(.by = station) %>%
  # slice(1) %>%

  mutate(
    plt =
      map2(
        data,
        station,
        \(.x, .y) {    
            
        sec <- ggh4x::help_secondary(
            .x,
            primary   = c(0, 1), 
            secondary = total_chl_a,
            method = "range",
            name = "Total Chlor a"
            )
            
        plts <-     
        ggplot(
          .x,
          aes(
            x = date_time_utc,
            y = conc
          )
          ) +
          geom_hline(yintercept = c(0, 0.5), color = "red", alpha = 0.2) +
          # geom_hline(
          #     aes(yintercept = ave(conc, names, FUN = range),
          #                        
          #         color = names), 
          #     alpha = 0.8,
          #     linetype = "dashed",
          #     show.legend = FALSE
          #     ) +
          geom_point(
            # aes(y = total_chl_a),
            aes(y = sec$proj(total_chl_a)),
            fill = "grey50",
            alpha = 0.5,
            size = 3,
            shape = 24,
            na.rm = TRUE
          ) +
          geom_line(
            # aes(y = total_chl_a),
            aes(y = sec$proj(total_chl_a)),
            # fill = "grey50",
            color = "grey40",
            alpha = 0.5,
            linetype = "twodash",
            # size = 3,
            # shape = 24,
            na.rm = TRUE
          ) +
          # geom_smooth(
          #  # aes(y = total_chl_a),
          #   aes(y = sec$proj(total_chl_a)),
          #   color       = "grey50",
          #   alpha       = 0.5,
          #   linetype    = "dashed",
          #   formula     = y ~ x,
          #   method      = "loess",
          #   span        = 0.3,
          #   se          = FALSE,
          #   show.legend = FALSE,
          #   na.rm       = TRUE
          # ) +
          geom_smooth(
            color       = "grey70",
            alpha       = 0.5,
            formula     = y ~ s(x, bs = "tp", k = 10),
            method      = "gam",
            se          = FALSE,
            show.legend = FALSE,
            na.rm       = TRUE
          ) +
          # geom_smooth(
          #   color       = "grey70",
          #   alpha       = 0.5,
          #   # linetype    = "dashed",
          #   formula     = y ~ x,
          #   method      = "loess",
          #   span        = 0.3,
          #   se          = FALSE,
          #   show.legend = FALSE,
          #   na.rm       = TRUE
          # ) +
          geom_point(
            aes(fill = season),
            shape = 23,
            size  = 4,
            na.rm = TRUE
          ) +
          geom_line(
            na.rm = TRUE
          ) +
          labs(title = .y) +
          scale_y_continuous(
            expand = c(0, 0),
            limits = c(0, NA),
            breaks = seq(0.0, 1, 0.1),
            minor_breaks = seq(0.05, 1, 0.1),
            sec.axis = sec
          ) +  
          scale_x_datetime(
            date_labels = "%Y",
            breaks = "1 year",
            # minor_breaks = "6 months"
          ) +
          facet_wrap(
            names ~ .,
            ncol   = 4,
            scales = "fixed"
          ) +
          # coord_cartesian(ylim = c(0, 1)) +
          theme(
            panel.grid.major.x = element_line(color = "grey80", linewidth = 0.5),
            panel.grid.minor = element_line(color = "grey80", linewidth = 0.4,
                                              # linetype = ""
                                            )
            )
        
        # plotly::ggplotly(plts)
        }
      )
  ) %>%
    # slice(1:3) %>%
  pull(plt)


```
What trends are visible?

MR
- cyano2 has largest variability from 10 - 60% of total chl-a
  - early autumn tends to be higher then later autumn (Sep - Nov)
  - maximal usually in Summer (Jun - Aug)
  - depths are pretty close except spring and summer 2019 
- cyano4 is close second with 10 - 30% of total chl-a
  - tends to peak in Autumn 
- Hapto6 consistently 5 - 20% of total chl-a with peaks in Autumn
  - mostly around 10% of total chl-a for most seasons
- hapt8 is usually 7 - 15% of total chl-a
  - highest in winter 2018 - 2019 and 2020 - 2021 >17% of total chl-a
- chloro  tend to between 0 - 10% of total chl-a
  - peaks in winter then decreases towards 0% for the rest of the year
  - seems like boom and bust cycles
- crypto never exceed 10% of total chl-a with similar peaks in winter
  - spring and summer tend to be below 5% of total chl-a
- pras tends to be between 0 - 15%
  - some peaks in autumn, but seems to be fairly similar each season around 5%
- diat1, diat2 and dino all tend to have <5% of total chl-a
  - diat1 only peak during bloom in winter 2019 at ~7%
  - diat2 has 3 peaks >=10%  during summer 2018, winter and summer 2019  
    - seems to be during overall chl-a peaks

7 
- cyano2 
  - similar to MR, largest contribution of 20 - 45% of total chl-a
  - peak in autumn 2017 with high chl-a (0.85) 
- cyano4 
  - much lower than MR, with 0 - 10% of total chl-a
  - typically lower than 5% of total chl-a
- chloro and crypto
  - have similar pattern and percents of 5 - 15%
- hapt8 and pras 
  - lower than 10% of total chl-a
- diat1 and diat2
  - usually max ~20% but not at the same time
  - diat1 lowest summer 1019 to summer 2020 while diat2 is consistent 5 - 20%
- dino 
  - tend to be low 5 - 10% of total chl-a
  - summer has small peaks
- hapt6
  - similar to dino with 5 - 20% of total chl-a
  - peaks tend to be in spring with lowest in summer

12
- offshore station
- cyano2 
  - typically highest overall with 20 - 55% of total chl-a
  - lowest in winters with peaks in spring - summer
  - high chlo-a looks to be inverted
- cyano4
  - ranges 5 - 40% of total chl-a
  - has peaks in autumn with lows during summer
- diat1, diat2 and dino
  - practically non existent with <5% of total chl-a
- hapt6 and hapt8
  - have similar contributions 0f 5 - 20% of total chl-a
  - follow similar patterns but hapt6 lags behind hapt8 in terms of peaks
- pras
  - mostly 0 - 10% of total chl-a
  - no consistent pattern
- chloro 
  - seems to have lowest in autumn and highest in winter - spring

10
- nearshore station
- cyano2 
  - still highest contributor with 15 - 45% of total chl-a
  - a bloom seems possible in winter 2019 with ~80% of total chl-a
  - seems like early season is lower with later season is higher?
- cyano2
  - very low compared to the other stations <5% for the most part
- chloro
  - between 5 - 10% of total chl-a
  - two low points between summer and autumn of 2019
- 

16
- nearshore station

18
- offshore station

LK



WS

### 4.1. Each Station Facet by PFT


```{r pft-by-station, fig.width=40, fig.asp=0.5}
cmtx_l_pct %>%
  rename("conc" = percent) %>%
  mutate(
    names2 = names,
    names = fct_reorder(
      names,
      conc,
      .fun  = median,
      .desc = TRUE
    )
  ) %>%
  filter(
    .by = c(station, names),
    n() > 3,
    !is.na(conc)
  ) %>%
  nest(.by = station) %>%
  # slice(1) %>%

  mutate(
    plt =
      map2(
        data,
        station,
        \(.x, .y) {    
            
        sec <- ggh4x::help_secondary(
            .x,
            primary   = c(0, 1), 
            secondary = total_chl_a,
            method = "range",
            name = "Total Chlor a"
            )
        
        .x %>%
        # mutate(
        #   names = fct_reorder(
        #       names, 
        #       conc, 
        #       .fun = median, 
        #       .desc = TRUE)
        #   ) %>%
        
        ggplot(
         .,
         aes(
           x = date_time_utc,
           y = conc, fill = names
         )
       ) +
        geom_area(alpha = 0.6, linewidth = 1, colour = "black")   +
        ggnewscale::new_scale_fill() +
        ggnewscale::new_scale_color() +
          # geom_vline(aes(xintercept = date_time_utc)) +
          geom_line(
            aes(y = sec$proj(total_chl_a)),
            color = "firebrick1",
            alpha = 0.5,
            # linetype = "twodash",
            na.rm = TRUE
          ) +
          geom_point(
            aes(
              y = sec$proj(total_chl_a), 
              color = season, 
              fill  = season, 
              shape = season
              ),
            # fill = "grey50",
            alpha = 0.5,
            size = 5,
            # shape = 24,
            na.rm = TRUE
          ) +
          # geom_smooth(
          #   color       = "grey70",
          #   alpha       = 0.5,
          #   formula     = y ~ s(x, bs = "tp", k = 10),
          #   method      = "gam",
          #   se          = FALSE,
          #   show.legend = FALSE,
          #   na.rm       = TRUE
          # ) +
          # geom_point(
          #   aes(fill = names),
          #   shape = 23,
          #   size  = 4,
          #   na.rm = TRUE
          # ) +
          # geom_line(
          #     aes(color = names),
          #   na.rm = TRUE
          # ) +
          # geom_area(
          #   aes(fill = names),
          #   alpha = 0.6, linetype = 1, colour = "black")
          
          labs(title = .y) +
          scale_y_continuous(
            expand = c(0, 0),
            limits = c(0, NA),
            breaks = seq(0.0, 1, 0.1),
            minor_breaks = seq(0.05, 1, 0.1),
            sec.axis = sec
          ) +
          scale_x_datetime(
            date_labels = "%Y",
            breaks = "1 year",
          ) +
             facet_wrap(
            names ~ .,
            ncol   = 4,
            scales = "fixed"
          ) +
          theme(
            panel.grid.major.x = element_line(color = "grey80", linewidth = 0.5),
            panel.grid.minor = element_line(color = "grey80", linewidth = 0.4,
                                              # linetype = ""
                                            )
            )
        }
      )
  ) %>%
    # slice(1) %>%
  pull(plt) #-> test
    # save_gg(
    #     plt = last_plot(), 
    #     save_location  = here(),
    #     save_name      = "test",
    #     overwrite      = T,
    #     verbose        = TRUE,
    #     time_stamp_fmt = NULL,
    #     device         = "jpeg",
    #     width = 18.27, height = 11.51, dpi = 300
    # )
```
- cyano4 and hapt8 are low nearshore
- crypto only explode at station 16
- diat2 lowest off shore
  - highest near shore (16, 10, 7)
- dino and diat1
  - sames as diat2
 
### 4.1. Each PFT Facet by Station
  
```{r station-by-pft, fig.width=40, fig.asp=0.5}
cmtx_l_pct %>%
  rename("conc" = percent) %>%
  mutate(
    names2 = names,
    names = fct_reorder(
      names,
      conc,
      .fun  = median,
      .desc = TRUE
    )
  ) %>%
  filter(
    .by = c(station, names),
    n() > 3,
    !is.na(conc)
  ) %>%
  nest(.by = names) %>%
  mutate(
    plt =
      map2(
        data,
        names,
        \(.x, .y) {    
            
        sec <- ggh4x::help_secondary(
            .x,
            primary   = c(0, 1), 
            secondary = total_chl_a,
            method = "range",
            name = "Total Chlor a"
            )
        
        .x %>%
        # mutate(
        #   names = fct_reorder(
        #       names, 
        #       conc, 
        #       .fun = median, 
        #       .desc = TRUE)
        #   ) %>%
        
        ggplot(
         .,
         aes(
           x = date_time_utc,
           y = conc, fill = station
         )
       ) +
        geom_area(alpha = 0.6, linewidth = 1, colour = "black")   +
        ggnewscale::new_scale_fill() +
        ggnewscale::new_scale_color() +
          # geom_vline(aes(xintercept = date_time_utc)) +
          geom_line(
            aes(y = sec$proj(total_chl_a)),
            color = "firebrick1",
            alpha = 0.5,
            # linetype = "twodash",
            na.rm = TRUE
          ) +
          geom_point(
            aes(
              y = sec$proj(total_chl_a), 
              color = season, 
              fill  = season, 
              shape = season
              ),
            # fill = "grey50",
            alpha = 0.5,
            size = 5,
            # shape = 24,
            na.rm = TRUE
          ) +
          geom_smooth(
            color       = "black",
            # alpha       = 0.5,
            formula     = y ~ x,
            method      = "loess",
            span = 0.3,
            se          = FALSE,
            show.legend = FALSE,
            na.rm       = TRUE
          ) +
          # geom_smooth(
          #   color       = "grey70",
          #   alpha       = 0.5,
          #   formula     = y ~ s(x, bs = "tp", k = 10),
          #   method      = "gam",
          #   se          = FALSE,
          #   show.legend = FALSE,
          #   na.rm       = TRUE
          # ) +
          # geom_point(
          #   aes(fill = names),
          #   shape = 23,
          #   size  = 4,
          #   na.rm = TRUE
          # ) +
          # geom_line(
          #     aes(color = names),
          #   na.rm = TRUE
          # ) +
          # geom_area(
          #   aes(fill = names),
          #   alpha = 0.6, linetype = 1, colour = "black")
          
          labs(title = .y) +
          scale_y_continuous(
            expand = c(0, 0),
            limits = c(0, NA),
            breaks = seq(0.0, 1, 0.1),
            minor_breaks = seq(0.05, 1, 0.1),
            sec.axis = sec
          ) +
          scale_x_datetime(
            date_labels = "%Y",
            breaks = "1 year",
          ) +
             facet_wrap(
            station ~ .,
            ncol   = 4,
            scales = "fixed"
          ) +
          theme(
            # panel.grid.major.x = element_line(color = "grey80", linewidth = 0.5),
            panel.grid.major.x = element_line(color = "black", linewidth = 0.5),
            panel.grid.minor = element_line(color = "grey80", linewidth = 0.4,
                                              # linetype = ""
                                            )
            )
        }
      )
  ) %>%
  pull(plt)

```

```{r tchla-by-station, fig.width=40, fig.asp=0.5}
cmtx_w %>%
  filter(
    .by = c(station),
    n() > 3,
    !is.na(total_chl_a)
  ) %>%
   ggplot(
         .,
         aes(
           x = date_time_utc,
           y = total_chl_a, 
           fill = station
         )
       ) +
        geom_area(alpha = 0.6, linewidth = 1, colour = "black")   +
        ggnewscale::new_scale_fill() +
        ggnewscale::new_scale_color() +
          geom_hline(
            yintercept = 1,
            color = "firebrick1",
            alpha = 0.5,
            linetype = "dashed",
            linewidth = 1.5,
            na.rm = TRUE
          ) +
          geom_point(
            aes(
              # y = sec$proj(total_chl_a), 
              color = season, 
              fill  = season, 
              shape = season
              ),
            # fill = "grey50",
            alpha = 0.75,
            size = 5,
            # shape = 24,
            na.rm = TRUE
          ) +
          geom_smooth(
            color       = "blue",
            # alpha       = 0.5,
            formula     = y ~ x,
            method      = "loess",
            span = 0.3,
            se          = FALSE,
            show.legend = FALSE,
            na.rm       = TRUE,
            linetype = "dashed"
          ) +

          labs(title = NULL) +
          scale_y_continuous(
            expand = c(0, 0),
            limits = c(0, 2.0),
            breaks = seq(0.0, 2, 0.25),
            # minor_breaks = seq(0.05, 1, 0.1),
            oob = scales::oob_squish
            # sec.axis = sec
          ) +
          scale_x_datetime(
            date_labels = "%Y",
            breaks = "1 year",
          ) +
             facet_wrap(
            station ~ .,
            ncol   = 4,
            scales = "fixed"
          ) +
          theme(
            # panel.grid.major.x = element_line(color = "grey80", linewidth = 0.5),
            panel.grid.major.x = element_line(color = "black", linewidth = 0.5),
            panel.grid.minor = element_line(color = "grey80", linewidth = 0.4,
                                              # linetype = ""
                                            )
            )

```


  
```{r, fig.height=15}
cmtx_l %>%
  mutate(pft_pct = conc/total_chl_a) %>%
  summarise(
    .by = c(cruise_id, phyto, names),
    avg = mean(pft_pct),
    sd  = sd(pft_pct),
    dates = mean(date_time_utc)
  ) %>%
  ggplot(aes(x = dates, y = avg)) +
    geom_vline(
      data = 
        tibble(
          x_yr = as_datetime(glue("{c(2016, 2017, 2018, 2019, 2020, 2021)}-01-01")) 
          ),
      aes(xintercept = x_yr)
      ) +
    geom_line(aes(color = names)) +
    geom_point(aes(color = names)) +
    geom_errorbar(aes(ymax = avg + sd, ymin = avg - sd, color = names)) +
    geom_point(
      data = mutate(cmtx_l, pft_pct = conc/total_chl_a), 
      aes(x = date_time_utc, y = pft_pct), 
      color = "grey70"
      ) +
    facet_grid(names ~ .) +
    scale_x_datetime(date_breaks = "1 year") +
    ylim(0, 1)
    


```


# Plot cyano2 with chla over time

```{r}
cyano_time <- 
  cmtx_w_pct %>%
  mutate(
    labels = case_when(
  str_detect(station, "10|16") ~ "Nearshore",
  .default = "Offshore"
    )
  ) %>%
  ggplot() +
  geom_point(
      data = ~ filter(.x, str_detect(station, "10|16|12|18")),
    aes(x = date_time_utc, 
        y = cyano2,
        color = labels
        )
  )  +
  geom_smooth(
    aes(
      x = date_time_utc, 
      y = cyano2),
    method = "gam",
    # span = 0.15,
    se = FALSE
  ) +
  geom_smooth(
    data = ~ filter(.x, str_detect(station, "10|16|12|18")),
    aes(
      x = date_time_utc, 
      y = cyano2,
      color = labels),
    method = "gam",
    # span = 0.2,
    se = FALSE
  ) +
  scale_x_datetime(breaks = "1 year",
                   ) +
  theme(
    panel.grid.major.x = element_line(color = "black")    
    )


tchl_time <- 
  cmtx_w %>%
  ggplot() + 
  geom_point(
    aes(x = date_time_utc, 
        y = total_chl_a,
        color = season
        )
  ) +
  geom_smooth(
    aes(
      x = date_time_utc, 
      y = total_chl_a),
    method = "loess",
    span = 0.15,
    se = FALSE
  ) +
  scale_x_datetime(breaks = "1 year")  +
  coord_cartesian(
    # ylim = c(NA, 2)
  ) +
  theme(
    panel.grid.major.x = element_line(color = "black")    
    )

cyano_time /
tchl_time
```


```{r}
cmtx_w_pct %>%
  ggplot() + 
  geom_point(
    aes(
      y = cyano2, 
      x = total_chl_a,
      color = season
        ),
    # color = "red"
  ) +
  coord_cartesian(
    ylim = c(NA, 1)
  ) +
  theme(
    panel.grid.major.x = element_line(color = "black")    
    ) +
  facet_wrap(
    ~ season
  )
  cmtx_w %>%
  ggplot() + 
  geom_point(
    aes(
      y = cyano2, 
      x = total_chl_a,
      color = season
        ),
    # color = "red"
  ) +
  coord_cartesian(
    ylim = c(NA, 1)
  ) +
  theme(
    panel.grid.major.x = element_line(color = "black")    
    ) +
  facet_wrap(
    ~ year
  )
```

```{r}
lm(total_chl_a ~ season, data = cmtx_w) %>%
    summary()
```

```{r station-by-pft, fig.width=40, fig.asp=0.5}
test1 <-   
  cmtx_l_pct %>%
  rename("conc" = percent) %>%
  mutate(
    names2 = names,
    names = fct_reorder(
      names,
      conc,
      .fun  = median,
      .desc = TRUE
    )
  ) %>%
  filter(
    .by = c(station, names),
    n() > 3,
    !is.na(conc)
  )
sec <- ggh4x::help_secondary(
  test1,
  primary   = c(0, 1),
  secondary = total_chl_a,
  method    = "range",
  name      = "Total Chlor a"
)
```
### Hurricane Dates
from: <https://coast.noaa.gov/hurricanes>
```{r hurricane-dates}
hurricane_dates <-
  tribble(
    ~date, ~name,
    # "2015-08-30", "Erika", ## 
    
    # "2016-10-12", "Nicole", ##
    
    # "2017-07-31", "Emily", ##
    "2017-09-10", "Irma",
    # "2017-10-29", "Philippe", ##
    
    "2018-09-03", "Gordon", 
    
    # "2019-09-01", "Dorian", ##
    
    # "2020-05-25", "Bertha", ##
    "2020-08-02", "Isaias",
    # "2020-08-26", "Laura", ##
    "2020-09-11", "Sally",
    "2020-11-09", "Eta", 
    
  ) %>%
  mutate(date = as_datetime(date)) %T>% 
  print()


```

```{r stack-station-by-pft, fig.width=40, fig.asp=0.5}
test1 %>%
  mutate(
    # date_time_utc = as_date(date_time_utc)
    names = fct_reorder(
      names,
      conc,
      .fun = median,
      .desc = FALSE
    )
  ) %>%
  ggplot(
    .,
    aes(
      x = date_time_utc,
      y = conc,
      fill = names
    )
  ) +
  geom_area(alpha = 0.6, linewidth = 1, colour = "black") +
  facet_wrap(~station, ncol = 4) +
  ggnewscale::new_scale_fill() +
  ggnewscale::new_scale_color() +
  geom_line(
    aes(y = sec$proj(total_chl_a)),
    color = "firebrick1",
    alpha = 0.5,
    na.rm = TRUE
  ) +
  geom_point(
    aes(y = sec$proj(total_chl_a)),
    color = "firebrick1",
    na.rm = TRUE
  ) +
  geom_vline(xintercept = hurricane_dates$date, color = "red", linetype = "dashed") +
  scale_y_continuous(
    expand       = c(0, 0),
    limits       = c(0, NA),
    breaks       = seq(0.0, 1, 0.1),
    minor_breaks = seq(0.05, 1, 0.1),
    sec.axis     = sec
  ) +
  scale_x_datetime(
    date_labels = "%Y",
    breaks = "1 year",
  ) +
  theme(
    panel.grid.major.x = element_line(color = "black", linewidth = 0.5),
    panel.grid.minor = element_line(color = "grey80", linewidth = 0.4)
  )
```

```{r station-by-pft, fig.width=40, fig.asp=0.5}
test2 <-
  cmtx_l %>%
  mutate(
    names2 = names,
    names = fct_reorder(
      names,
      conc,
      .fun  = median,
      .desc = TRUE
    )
  ) %>%
  filter(
    .by = c(station, names),
    n() > 3,
    !is.na(conc)
  )

sec2 <- ggh4x::help_secondary(
  test2,
  primary = c(0, 1),
  secondary = total_chl_a,
  method = "range",
  name = "Total Chlor a"
)
```

```{r}
test2 %>%
  mutate(
    # date_time_utc = as_date(date_time_utc)
    names = fct_reorder(
      names,
      conc,
      .fun = median,
      .desc = FALSE
    )
  ) %>%
  ggplot(
    .,
    aes(
      x = date_time_utc,
      y = conc,
      fill = names
    )
  ) +
  geom_area(alpha = 0.6, linewidth = 1, colour = "black") +
  facet_wrap(~station, ncol = 4) +
  ggnewscale::new_scale_fill() +
  ggnewscale::new_scale_color() +
  # geom_line(
  #   aes(y = total_chl_a),
  #   # aes(y = sec2$proj(total_chl_a)),
  #   color = "firebrick1",
  #   alpha = 0.5,
  #   na.rm = TRUE
  # ) +
  geom_point(
    aes(y = total_chl_a, color = season),
    # aes(y = sec$proj(total_chl_a)),
    # color = "firebrick1",
    na.rm = TRUE
  ) +
  geom_vline(xintercept = hurricane_dates$date, color = "red", linetype = "dashed") +
  scale_y_continuous(
    expand       = c(0, 0),
    limits       = c(0, NA),
    breaks       = seq(0.0, 1, 0.1),
    minor_breaks = seq(0.05, 1, 0.1),
    # sec.axis     = sec2,
    oob = scales::oob_keep
  ) +
  scale_x_datetime(
    date_labels = "%Y",
    breaks = "1 year",
  ) +
  coord_cartesian(ylim = c(NA, 1.5), clip = "on") +
  theme(
    panel.grid.major.x = element_line(color = "black", linewidth = 0.5),
    panel.grid.minor = element_line(color = "grey80", linewidth = 0.4)
  )
```

