---
title: "Merge CHEMTAX and Nutrient Data"
author: "Sebastian Di Geronimo"
date: "2024-01-30"
format: html
editor: source
---

# 1.0 ---- Summary of Document ----



# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
    # additional
    
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )

source(here("scripts", "misc_functions.R"))
```

## 2.2 Load Dataset

### 2.2.1 CHEMTAX Data

Load CHEMTAX data and metadata. Filter data to only include samples collected at
the surface.

```{r load-data}
# set location to save data and plots
dir_data_save <- here("data", "processed", "06_1_merge_chemtax_nutrient")
dir_plt_save  <- here("data", "plots", "06_1_merge_chemtax_nutrient")

# ---- set season names and order ---- #
szn <- c("Winter", "Spring", "Summer", "Autumn")

# ---- phytoplankton full names ---- #
phyto_name <-
  here("data", "processed", "04_1_merge_meta_cmtx", "phyto_names.csv") %>%
  read_csv(show_col_types = FALSE) %T>% 
  print()

# ---- load chemtax data in wide format ---- #
cmtx_w <-
  here("data", "processed", "04_1_merge_meta_cmtx", "chemtax_w_meta_wide.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  # filter(!str_detect(station, "60|21.5|24|^3$|9B|9.5")) %>%
  filter(!str_detect(station, "60|21.5|24|^3$|9.5")) %>%
  mutate(
    season = fct(season, levels = szn),
    # station = str_replace(station, "9B", "9.5"),
    station = fct_reorder(station, lon)
  ) %>% 
  select(-c(hplc_gsfc_id, cruise:bottle, cluster_code, water_type)) %>%
    
  # add PFT concentrations to get total chlorophyll-a
  rowwise() %>%
  mutate(
    total_chl_a = sum(c_across(chloro:pras)),
    .before = chloro
  ) %>%
  ungroup() %>%
  filter(
    .by = c(station, cruise_id),
    depth == min(depth)
    & depth < 20
  ) %T>% 
  print()

```
### 2.2.2 NOAA Data

#### 2.2.2.1 NOAA `NA` Strings

These are all the variations of text contained in columns that should be 
numeric. This text will be converted to `NA` when read in using 
`openxlsx::read.xlsx`.


```{r noaa-na-strings}
noaa_bad <- c(
  "NA", "Skip too shallow",
  "97/98", "109/110",
  "113/114", "115/116",
  "117/118", "-",
  " Aboard R/V Savannah", "Lost ",
  "Lost", "Do Not Use (Discrete Chl samples not usable)",
  "Broken Vial", "Missing Vial ",
  "Missed", "Double labeled",
  "Missing", "RE-CAST NOT SAMPLED",
  "Missing ", "NOT COLLECTED",
  "Broken", "Do Not Use (Discrete Chl samples not usable",
  "Sensor bad", "Missing Tube ",
  "Bottles never tripped - recast", "Sensor Drift ",
  "Gain far too high unusable", "Sensor Cut out",
  "Excluded", "Data lost from Acetone Spill",
  "77,78", "79,80",
  "81,82", "Bad analysis",
  "Acidified prior to Fo", "173/174",
  "175/176", "177/178","179/180", "182/183","184/185", "186/187","188/189", 
  "190/191","192/193", "107/108","112/113", "152/153","155/156", "162/163",
  "165/166", "193/194","128, 129", "131, 132","163, 164", "165, 166","167, 168",
  "169, 170","171, 172", "175, 176","158/159", "160/161","164/165", "172/173",
  "174/175", "178/179","180/181", "189/190","7/8", "43/44","51/52", "85/86",
  "89/90", "135/136","138/139", "153/154","170/171", "69,70","71,72", 
  "Unusable", "Spilled"
)
```

#### 2.2.2.2 Load NOAA 

```{r load-noaa-data}
noaa_data <-
  here("data", "raw") %>%
  dir_ls(regexp = "WSMasterSampleLog") %>%
  str_subset("~", negate = TRUE)

noaa_dat <-
  openxlsx::getSheetNames(noaa_data) %>% 
  str_subset("(?i)all") %>% 
  openxlsx::read.xlsx(
    noaa_data, 
    sheet       = .,
    detectDates = TRUE,
    na.strings  = noaa_bad
  ) %>% 
  janitor::clean_names(replace = c(janitor:::mu_to_u, "#" = "number")) %>% 
  rename(
    "lon" = longitude_decimal,
    "lat" = latitude_decimal) %>%
  mutate(
    .after    = date_gmt,
    time_gmt  = as.numeric(time_gmt),
    time_gmt  = time_gmt + 1,
    time_gmt  = janitor::excel_numeric_to_date(time_gmt, include_time = TRUE),
    time_gmt  = hms::as_hms(time_gmt),
    date_time = ymd_hms(paste(date_gmt, time_gmt))
  ) %>% 
  mutate(
    .by = cruise,
    .after  = time_gmt,
    lon_lag = dplyr::lead(lon),
    lat_lag = dplyr::lead(lat),
    cruise = case_when(
      and(
        str_detect(cruise, "WS18120"),
        between(
          date_gmt, 
          as_date("2018-06-01"), 
          as_date("2018-09-01"))
        ) ~ "WS18218",
      str_detect(cruise, "SAV1803") ~ "SV18067",
      str_detect(cruise, "SAV18173") ~ "SV18173",
      str_detect(cruise, "WS20231") ~ "WS20230",
      .default = cruise),
       # SV18067	= SAV1803			
       # SV18173   = SAV18173				
       # WS20230   = WS20231 
     station2 = station, 
     station = case_match(
        station,
        c("21/LK", "21.5/LK", "21") ~ "LK",
        .default = station
      )
    ) %>% 
    mutate(
    .by = c(cruise, station),
    stn_samp = n(),
    depth = as.numeric(depth),
    .before = station) %>%
    filter(
      .by = c(cruise, station),
      min(depth) == depth 
      & depth < 20
    ) %>%
      arrange(-depth) %T>% print()

```

### 2.2.3 Load CTD Data 

```{r ctd-data}
ctd_dat_summary <- 
  here("data", "processed", "ctd", "ctd_all.csv") %>%
  read_csv(show_col_types = FALSE)  %>%
  mutate(time = ymd_hms(time)) %T>% print()
```



## 2.3 Merge Data

```{r merge-chemtax-nutrient}
cmtx_nutr <- 
  left_join(
    cmtx_w,
    noaa_dat,
    by = c("cruise_id" = "cruise", "station")
  ) %T>% print()

env_cmtx <- 
  cmtx_nutr %>%
  left_join(
    ctd_dat_summary,
    by = join_by(
      "cruise_id" == "cruise_id",
      "station" == "station"
    )
  ) %>%
  distinct(sample, .keep_all = TRUE) %T>% print()

# ---- filter PFT-CTD data for joint pairs without missing data
env_cmtx_filt <-
  env_cmtx  %>%
  drop_na(chloro:pras, pressure, temp, par, sal) %>%
  mutate(group = season) %T>% print()



# ---- show unmatched PFT-CTD rows
naniar::vis_miss(cmtx_nutr)
naniar::vis_miss(env_cmtx)
naniar::vis_miss(ctd_dat_summary)

filter(env_cmtx, is.na(file)) %>%
  select(cruise_id, station, date_time_utc) %>%
  arrange(date_time_utc) %T>% print() %>% 
  distinct(., cruise_id) %T>% print()

```

```{r save-merged-data}
# ---- save clean NOAA Nutrient Data ---- #
save_csv(
  .data          = noaa_dat,
  save_location  = dir_data_save,
  save_name      = "noaa_nutrient_tidy",
  overwrite      = FALSE,
  verbose        = TRUE,
  time_stamp_fmt = NULL,
  utf_8          = FALSE
)
  
# ---- save merged CHEMTAX-Nutrient Data ---- #
save_csv(
  .data          = cmtx_nutr,
  save_location  = dir_data_save,
  save_name      = "chemtax_nutrient",
  overwrite      = FALSE,
  verbose        = TRUE,
  time_stamp_fmt = NULL,
  utf_8          = FALSE
)

# ---- save merged CHEMTAX-Nutrient-CTD Data ---- #
save_csv(
  .data          = env_cmtx,
  save_location  = dir_data_save,
  save_name      = "chemtax_nutrient_ctd",
  overwrite      = FALSE,
  verbose        = TRUE,
  time_stamp_fmt = NULL,
  utf_8          = FALSE
)


```



## 2.4 Cruise Table

```{r cruise-table} 
# using cmtx_nutr, create a summary table of the cruises 
cruise_table <-
  cmtx_nutr %>% 
  summarise(
    .by = cruise_id,
    sample_count = n(),
    start_date = min(date_gmt, na.rm = TRUE),
    end_date = max(date_gmt, na.rm = TRUE),
  ) %>%
  arrange(start_date) %>%
  mutate(
    .keep = "unused",
    .before = sample_count,
    cruise_dates = (
      paste(
        format(start_date, "%B %d"), 
        format(end_date, "%B %d %Y"), 
        sep = " - "
      )
    )
  ) %T>% 
  print()

shelf(gt)
test_table <-
  gt::gt(cruise_table) %>%
  
  # rename columns using gt
  gt::cols_label(
    cruise_id = "Cruise ID",
    sample_count = "Sample Count",
    cruise_dates = "Cruise Dates"
  ) %>%
    
  # center justify columns
  gt::tab_style(
    style = gt::cell_text(
      font = "serif",
      align = "center"
    ),
    locations = gt::cells_body(
      columns = c(cruise_id, sample_count, cruise_dates)
    )
  ) %T>%
  print()

  # save gt table 
  # gt::gtsave(
  #   test_table,
  #   "cruise_table.docx",
  #   here("data", "plots"),
  #   # width = 6,
  #   # height = 4,
  #   # units = "in",
  #   # dpi = 300
  # )
    
cmtx_nutr %>%
  count(station, name = "station_count") %>%
  # removed stations with only one sample
  filter(station_count > 1) %>% 
  gt::gt() %>%
  gt::cols_label(
    station = "Station",
    station_count = "Station Count"
  )
``` 