---
title: "Plots for CHEMTAX and Nutrients"
author: "Sebastian Di Geronimo"
date: "2024-01-30"
format: html
editor: source
---

# 1.0 ---- Summary of Document ----



# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
    # additional
    patchwork
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )
  

```

## 2.2 Load Dataset

```{r load-data}
# set location to save data and plots
dir_data_save <- here("data", "processed", "06_2_cmtx_nutrient_plots")
dir_plt_save  <- here("data", "plots", "06_2_cmtx_nutrient_plots")

dir_create(dir_data_save)

# ---- set season names and order ---- #
szn <- c("Winter", "Spring", "Summer", "Autumn")

# ---- seasonal stripes ---- #
stripe <-
  expand.grid(2015:2021, paste0("-", c(12, 3, 6, 9))) %>%
  mutate(
    date = paste0(Var1, Var2),
    date = ym(date),
    date_max = date + months(3)
  ) %>%
  mutate(season = rep(c("Winter", "Spring", "Summer", "Autumn"), each = 7))

# ---- phytoplankton full names ---- #
phyto_name <-
  here("data", "processed", "04_1_merge_meta_cmtx", "phyto_names.csv") %>%
  read_csv(show_col_types = FALSE) %T>% 
  print()

# ---- load chemtax data in wide format ---- #
cmtx_nutr <-
  here("data", "processed", "06_1_merge_chemtax_nutrient", "chemtax_nutrient.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(
    season  = fct(season, levels = szn),
    station = fct_reorder(station, lon.x),
    dist    = case_when(
      str_detect(station, "10|16|7") ~ "Nearshore",
      .default = "Offshore"
    )
  ) %>%
  relocate(dist, .after = station) %>%
  mutate(
    across(
      phyto_name$shrt[1:10], 
      list(pct = \(.x) .x / total_chl_a)
      )
  ) %T>% 
  print()

```
### Min Time for Each Cruise

```{r cruise-min-time}
cruise_min <-
  cmtx_nutr %>%
  select(cruise_id, date_time_utc) %>%
  filter(
    .by = cruise_id,
    date_time_utc == min(date_time_utc)
  ) %>%
  arrange(date_time_utc) %T>% 
  print()
```



### Hurricane Dates
from: <https://coast.noaa.gov/hurricanes>
```{r hurricane-dates}
hurricane_dates <-
  tribble(
    ~date, ~name,
    # "2015-08-30", "Erika", ## 
    
    # "2016-10-12", "Nicole", ##
    
    # "2017-07-31", "Emily", ##
    "2017-09-10", "Irma",
    # "2017-10-29", "Philippe", ##
    
    "2018-09-03", "Gordon", 
    
    # "2019-09-01", "Dorian", ##
    
    # "2020-05-25", "Bertha", ##
    "2020-08-02", "Isaias",
    # "2020-08-26", "Laura", ##
    "2020-09-11", "Sally",
    "2020-11-09", "Eta", 
    
  ) %>%
  mutate(date = ymd(date)) %T>% 
  print()


```



### Average Water Depth

```{r average-water-depth, fig.height=20}
cmtx_nutr %>%
  select(
    cruise_id, date_time_utc, season, station, dist, water_depth
  ) %>% 
    filter(
      !(str_detect(station, "18")
      & water_depth < 5)
    ) %>%
  summarise(
    .by = dist,
    mean_water = mean(water_depth),
    min_water = min(water_depth),
    max_water = max(water_depth)
  )


cmtx_nutr %>%
  select(
    cruise_id, date_time_utc, season, station, dist, water_depth
  ) %>%
    
  summarise(
    .by = c(station, dist),
    mean_water = mean(water_depth),
    min_water = min(water_depth),
    max_water = max(water_depth)
  )
```

# Summarize Data

```{r select-vars, fig.height=20}
# select subset of variables
cmtx_sel_var <- 
  cmtx_nutr %>%
  select(
    c(cruise_id, date_time_utc, season, station, dist, chloro:pras, chloro_pct:pras_pct, 
      temperature, salinity, 
      total_chl_a, avg_chl_a_ug_l:si_u_m),
      -nutrients_tube_number
  ) %T>% 
  print()

cmtx_filt_stn <- 
  cmtx_sel_var %>%
  filter(str_detect(station, "10|12|16|18"))
```




```{r stat-summ}
stat_summ <-
  list(
    n   = \(.x) n(),
    avg = \(.x) mean(.x, na.rm = TRUE),
    sd  = \(.x) sd(.x, na.rm = TRUE),
    med = \(.x) median(.x, na.rm = TRUE),
    min = \(.x) min(.x, na.rm = TRUE),
    max = \(.x) max(.x, na.rm = TRUE)
  )


chmx_nut_long <- 
  cmtx_sel_var %>%
  mutate(
      nit_pho_ratio = no3_no2_u_m / (po4_u_m + 0.001),
      nit_pho_ratio = if_else(nit_pho_ratio == Inf, NA, nit_pho_ratio)
      ) %>%
  select(-c(chloro:pras_pct)) %>%
  pivot_longer(
    data      = .,
    # cols      = c(chloro:pras_pct, temperature:nit_pho_ratio), 
    cols      = c(temperature:nit_pho_ratio), 
    names_to  = "name", 
    values_to = "value", 
  ) %T>% 
  print()

dat_summary <-
  list(
    # by name of parameter
    avg_param = 
      chmx_nut_long %>%
      summarise(
        .by = c(name),
        across(
          value,
          stat_summ
        )
      ),

    # by cruise
    avg_cruise =
      chmx_nut_long %>%
      summarise(
        .by = c(name, cruise_id),
        across(
          value,
          stat_summ
        )
      ) %>%
      arrange(name, cruise_id) %>%
      left_join(cruise_min),

    # by distance from shore
    avg_dist = 
      chmx_nut_long %>%
      summarise(
        .by = c(dist, name),
        across(
          value,
          stat_summ
        )
      ) %>%
      arrange(name),

    # by season
    avg_season =
      chmx_nut_long %>%
      summarise(
        .by = c(season, name),
        across(
          value,
          stat_summ
        )
      ) %>%
      arrange(name, season),

    # by cruise and dist
    avg_cruise_dist =
      chmx_nut_long %>%
      summarise(
        .by = c(cruise_id, dist, name),
        across(
          value,
          stat_summ
        )
      ) %>%
      arrange(name, cruise_id) %>%
      left_join(cruise_min),

    # by season and dist
    avg_season_dist =
      chmx_nut_long %>%
      summarise(
        .by = c(season, dist, name),
        across(
          value,
          stat_summ
        )
      ) %>%
      arrange(name, season)
  )

dat_summary



# save file
if (!file_exists(here(dir_data_save, "summarized_parameter.xlsx"))) {
  writexl::write_xlsx(
    x = dat_summary,
    path = here(dir_data_save, "summarized_parameter.xlsx")
  )

  shell.exec(dir_data_save)
}
```

```{r}
cmtx_sel_var %>%
  select(station, season, dist, c(chloro:pras_pct)) %>%
  pivot_longer(
    data      = .,
    # cols      = c(chloro:pras_pct, temperature:nit_pho_ratio), 
    cols      = c(chloro:pras_pct), 
    names_to  = "name", 
    values_to = "value", 
  ) %T>% 
  print() %>%
  summarise(
        .by = c(station, name, dist),
        across(
          value,
          stat_summ
        )
      ) %>%
      arrange(name, dist) %T>% 
  print() %>%
  filter(str_detect(station, "10|12|16|18|7"))

cmtx_sel_var %>%
  select(station, season, dist, c(chloro:pras_pct)) %>%
  pivot_longer(
    data      = .,
    # cols      = c(chloro:pras_pct, temperature:nit_pho_ratio), 
    cols      = c(chloro:pras_pct), 
    names_to  = "name", 
    values_to = "value", 
  ) %>% 
  # print() %>%
  summarise(
        .by = c(station, name, season, dist),
        across(
          value,
          stat_summ
        )
      ) %>%
      arrange(name, dist, station, season) %T>% 
  print() %>%
  filter(str_detect(station, "10|12|16|18|7"))

cmtx_sel_var %>%
  select(station, season, dist, date_time_utc, c(chloro:pras_pct)) %>%
  mutate(year = year(date_time_utc %m+% months(1)),
         .after = date_time_utc) %>%
  pivot_longer(
    data      = .,
    # cols      = c(chloro:pras_pct, temperature:nit_pho_ratio), 
    cols      = c(chloro:pras_pct), 
    names_to  = "name", 
    values_to = "value", 
  ) %>% 
  # print() %>%
  summarise(
        .by = c(name, year, season),
        across(
          value,
          stat_summ
        )
      ) %>%
      arrange(name, season, year) 

cmtx_sel_var %>%
  select(station, season, dist, date_time_utc, c(chloro:pras_pct)) %>%
  mutate(year = year(date_time_utc %m+% months(1)),
         .after = date_time_utc) %>%
  pivot_longer(
    data      = .,
    # cols      = c(chloro:pras_pct, temperature:nit_pho_ratio), 
    cols      = c(chloro:pras_pct), 
    names_to  = "name", 
    values_to = "value", 
  ) %>% 
  # print() %>%
  summarise(
        .by = c(name, year, season, dist),
        across(
          value,
          stat_summ
        )
      ) %>%
      arrange(name, season, dist, year)
```


# Plot Monthly Data

```{r plot-monthly, fig.height=10}
cmtx_nutr_plt_l <-
  cmtx_sel_var %>%
  pivot_longer(
    data      = .,
    cols      = c(chloro:last_col()), # columns to pivot long,
    names_to  = "name", # desired name for category column
    values_to = "value", # desired name for value column
  ) %>%
  mutate(
    .by = c(cruise_id, dist),
    .after = date_time_utc,
    # month_year = format(date_time_utc, format = "%Y-%m"),
    # month_year = ym(month_year),
    month_year = min(date_time_utc),
    month_year = as_date(month_year)
    # date_time_utc2 = as_date(date_time_utc),
    # .keep = "used"
  ) %T>% 
  print()
# nest(cmtx_nutr_plt_l, .by = name, .key = "data_raw")
```


```{r plot-monthly, fig.height=10}
cmtx_nutr_plt <-
  cmtx_nutr_plt_l %>% 
  summarise(
    .by = c(name, dist, month_year),
    across(
      value,
      stat_summ
    )
  )  %T>% 
  print() %>%
  nest(.by = name) %T>%
  print() %>%
  left_join(
    nest(cmtx_nutr_plt_l, .by = name, .key =  "data_raw"),
    by = join_by(name)
    ) %>%
  # slice(1:2) %>%
  # filter(
  #   str_detect(name, "temperature")
  # ) %>%
  mutate(
  plt = pmap(
    list(
      .x = data,
      .y = name,
      .z = data_raw
      ),
    .hurr = hurricane_dates,
    stripe = stripe,
    color_val = c(
            "Offshore"  = "blue",
            "Nearshore" = "red",
            "Winter"    = "#88CCEE",
            "Spring"    = "#44AA99",
            "Summer"    = "#DDCC77",
            "Autumn"    = "#CC6677"
            ),
    \(.x, .y, .hurr, .z, stripe, color_val) {

      plt <-
        .x %>%
        ggplot(aes(x = month_year, y = value_avg, color = dist, fill = dist)) +
        geom_rect(
          data = stripe,
          aes(
            xmin = date, xmax = date_max,
            ymin = -Inf, ymax = Inf,
            fill = season
          ),
          alpha = 0.5,
          inherit.aes = FALSE,
          show.legend = FALSE
        ) +
        # scale_fill_manual(
        #   values = c(
        #     "Offshore"  = "blue",
        #     "Nearshore" = "red",
        #     "Winter"    = "grey",
        #     "Spring"    = "white",
        #     "Summer"    = "grey",
        #     "Autumn"    = "white"
        #     ),
        #   guide = "none") +
        scale_fill_manual(
          values = color_val,
          guide = "none") +
        geom_line() +
        geom_point(shape = 22, color = "black") +
        
        # ---- add raw data ---- #
        geom_point(
          data = .z,
          aes(
            x     = as_date(date_time_utc),
            y     = value,
            color = dist,
            shape = dist),
          position = position_dodge(0.9)
          ) +
        geom_errorbar(
          aes(ymin = value_avg - value_sd, ymax = value_avg + value_sd)
        ) +
        geom_vline(xintercept = .hurr$date, color = "red", linetype = "dashed") +
        geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
        geom_text(
          data = .hurr,
          aes(
            x     = date, 
            y     = max(.x$value_avg, na.rm = TRUE) * 1.25, 
            label = name),
          angle = 90,
          vjust = -0.5,
          hjust = 1.5,
          inherit.aes = FALSE,
        ) +
        labs(
          title = .y,
          x = NULL,
          color = NULL,
          shape = NULL
        ) +
        scale_x_date(breaks = "1 year", expand = c(0, 0)) +
        scale_y_continuous(expand = expansion(add = c(0.005, 0))) +
        coord_cartesian(
          # ylim = c(0, NA),
          xlim = c(as_date("2016-01-01"), as_date("2021-05-01"))
        ) +
        theme_bw() +
        theme(
          legend.position = "top"
        )
    }
  )
) 
```


# ----- Start HERE if NEEDED -----

## Monthly Average PFT

```{r pft-plot, fig.height=10}
cmtx_nutr_plt %>%
  slice(1:20) %>%
  # slice(1) %>%
  arrange(name) %>%
  pull(plt) %>%
  map(
    \(.x) {
      .x +
      scale_y_continuous(expand = expansion(add = c(0.005, 0))) +
          coord_cartesian(
            ylim = c(0, NA), 
            xlim = c(as_date("2016-01-01"), as_date("2021-05-01"))
            ) 
    }
  )
```

```{r pft-interactive, fig.height=10}
cmtx_nutr_plt %>%
  # slice(1) %>%
  slice(1:20) %>%
  arrange(name) %>%
  pull(plt) %>%
  map(
    \(.x) {
      plotly::ggplotly(.x)
    }
  )
```



## Average Monthly Environment

```{r env-plot, fig.height=10}
cmtx_nutr_plt %>%
  slice(21:31) %>%
  # slice(1) %>%
  pull(plt) 
```


```{r env-interactive, fig.height=10}
cmtx_nutr_plt %>%
  slice(21:31) %>%
  # slice(1) %>%
  pull(plt) %>%
  map(
    \(.x) {
      plotly::ggplotly(.x)
    }
  )
```

```{r temp-pft, fig.height=15}
# shelf(patchwork)

cmtx_nutr_plt %>%
  slice(1:20) %$%
  imap(
    plt,
    .env = pull(slice(cmtx_nutr_plt, 21), plt)[[1]],
    \(.x, .env, idx) {
      print(idx)    
      (
        (.x / .env) +
          patchwork::plot_layout(guides = "collect") &
          theme(legend.position = "top")
          )
    }
  )
```


```{r temp-chla-pft, fig.height=15}
# shelf(patchwork)
plot_stck <- \(
  .x,
  .top,
  .bot,
  limits = list(
    mid = c(0, 1),
    top = c(NA, NA),
    bot = c(0, 1.2)
  ),
  idx) {
  print(idx)

  (
    (
      (.top +
        scale_y_continuous(
         limits = limits$top,
         oob    = scales::oob_keep,
         breaks = seq(0, 1, 0.1),
         expand = expansion(add = c(0.02, 0))
         )
      ) /
        (.x +
          scale_y_continuous(
            # limits = c(0, 1),
            limits = limits$mid,
            oob    = scales::oob_keep,
            breaks = seq(0, 1, 0.1),
            expand = expansion(add = c(0.02, 0))
          )
        ) /
        (.bot +
          scale_y_continuous(
            limits = limits$bot,
            oob    = scales::oob_keep,
            expand = expansion(add = c(0.02, 0))
          )
        )
    ) +
      patchwork::plot_layout(guides = "collect") &
      theme(legend.position = "top")
  )
}

imap(
  .x = cmtx_nutr_plt$plt[1:20],
  .top = cmtx_nutr_plt$plt[[21]], # temp
  .bot = cmtx_nutr_plt$plt[[23]], # total_chl_A
  limits = list(
    mid = c(0, 1),
    top = c(NA, NA),
    bot = c(0, 1.2)
  ),
  plot_stck
)

```
```{r, fig.height=15}
mutate(cmtx_nutr_plt, row = row_number(), .before = 1)
imap(
  .x  = cmtx_nutr_plt$plt[11],
  .top = cmtx_nutr_plt$plt[[18]],
  .bot = cmtx_nutr_plt$plt[[13]],
  limits = list(
    mid = c(0, 0.5),
    top = c(0, 0.5),
    bot = c(0, 0.7)
  ),
   plot_stck
)

cmtx_nutr_plt$plt[[15]]

imap(
  .x  = cmtx_nutr_plt$plt[12],
  .top = cmtx_nutr_plt$plt[[19]],
  .bot = cmtx_nutr_plt$plt[[13]],
  limits = list(
    mid = c(0, 0.5),
    top = c(0, 0.5),
    bot = c(0, 0.7)
  ),
   plot_stck
)
```


```{r}
cmtx_nutr_plt$plt[[23]] +
    scale_y_continuous(limits = c(0, 1.2), oob = scales::oob_keep,
                       expand = expansion(add = c(0.02, 0)))
```

# Stack Across Months
```{r, fig.height=15}
# function used to set all date to same year
  same_year <- function(x, .year = 2000) {
    year(x) <- .year
    x
  }

dat_month_stk <- 
  cmtx_nutr_plt_l %>%
  summarise(
    .by = c(name, dist, month_year),
    across(
      value,
      stat_summ
    )
  )   %>%
  mutate(
    .after = 1,
    date = as_date(month_year),
    year = year(date),
    year = as.character(year),
    year_cat = paste(year, dist),
    date2 = map(
        .x = date,
        .f = same_year
      )
    ) %>%
  unnest(date2)  %>%
  arrange(date2) %T>%
  print() %>%
  nest(.by = name) %T>%
  # slice(1:20) %T>%
  print() %>%
  mutate(
    plt_stk = 
      map2(
        .x = data,
        .y = name,
        \(.x, .y) {
          
          .x %>%
          ggplot(aes(x = date2, y = value_avg, color = year)
                 ) +
          geom_errorbar(
            aes(ymin = value_avg - value_sd,
                ymax = value_avg + value_sd)
          ) +
          geom_point() +
          geom_path() +
          scale_x_date(date_labels = "%b", date_breaks = "1 month") +
          scale_y_continuous(
            limits = c(-0.005, NA),
            expand = expansion()
            ) +
          labs(
            title = .y,
            color = NULL
          ) +
          facet_wrap(
            ~dist,
            ncol = 1) +
          theme_bw() +
          theme(
            legend.position = "top"
          )
        }
      )
  ) %T>% 
  print()
```


```{r, fig.height=15}
dat_month_stk  %>%
  slice(c(11:20)) %>%
  pull(plt_stk)
```
```{r, fig.height=25}
imap(
  .x  = dat_month_stk$plt_stk[12],
  .top = dat_month_stk$plt_stk[[19]],
  .bot = dat_month_stk$plt_stk[[13]],
  limits = list(
    mid = c(0, 0.3),
    top = c(0, 0.3),
    bot = c(0, 0.6)
  ),
   plot_stck
)
```

```{r, fig.height=15}
dat_month_stk  %>%
  slice(c(1:10)) %>%
  pull(plt_stk)

```

```{r, fig.height=15}

dat_month_stk  %>%
  slice(-c(1:20)) %>%
  slice(1:2) %>%
  pull(plt_stk) %>%
  .[[1]] +
  scale_y_continuous(limits = c(NA, NA))

dat_month_stk  %>%
  slice(-c(1:20)) %>%
  slice(1:2) %>%
  pull(plt_stk) %>%
  .[[2]] +
  scale_y_continuous(limits = c(NA, NA))

dat_month_stk  %>%
  slice(-c(1:20)) %>%
  slice(3) %>%
  pull(plt_stk) %>%
  .[[1]] +
  scale_y_continuous(limits = c(0, 0.8),
                     oob = scales::oob_keep)# +
  facet_wrap(~dist,
            ncol = 1, scales = "free_y")


dat_month_stk  %>%
  slice(-c(1:20)) %>%
  slice(-c(1:2)) %>%
  pull(plt_stk)
```


# PCA Nuritents

```{r pca-nutr}
cmtx_pca <- 
  cmtx_sel_var %>%
  drop_na() %T>% 
  print()

nutr_pca <- 
  cmtx_pca %>%
  select(-c(chloro:pras_pct)) %>%
  mutate(
    # .by = dist,
    across(
      c(temperature:last_col()), 
      \(x) scale(x)
      )
  ) %>%
  select(-c(1:5, avg_chl_a_ug_l)) %>%
  
  prcomp(~ ., data = .)


summary(nutr_pca)

# plot pca
nutr_pca_plt_bi <-
  nutr_pca %>%
  factoextra::fviz_pca_biplot(
    geom = "point",
    # col.ind  = month(cmtx_pca$date_time_utc, label = TRUE),
    # fill.ind = month(cmtx_pca$date_time_utc, label = TRUE),
    col.ind  = cmtx_pca$season,
    fill.ind = cmtx_pca$dist,
    fill.var = "red",
    col.var  = "red",
    addEllipses = FALSE,
    legend.title = list(color = "Month", fill = "Dist", shape = "Month"),
    ggtheme = ggthemes::theme_calc()
  )

nutr_pca_plt_ind <-
  nutr_pca %>%
  factoextra::fviz_pca_ind(
    geom = "point",
    # col.ind  = month(cmtx_pca$date_time_utc, label = TRUE),
    # fill.ind = month(cmtx_pca$date_time_utc, label = TRUE),
    col.ind  = cmtx_pca$season,
    fill.ind = cmtx_pca$dist,
    fill.var = "red",
    col.var  = "red",
    addEllipses = FALSE,
    legend.title = list(color = "Month", fill = "Dist", shape = "Month"),
    ggtheme = ggthemes::theme_calc()
  )

nutr_pca_plt_bi  
nutr_pca_plt_ind
plotly::ggplotly(nutr_pca_plt_ind)
```

# Plot PFT vs Envrionmental Parameters

## NOT USED:


Warning: Using an external vector in selections was deprecated in tidyselect 1.1.0.
ℹ Please use `all_of()` or `any_of()` instead.
  # Was:
  data %>% select(.x)

  # Now:
  data %>% select(all_of(.x))


```{r stacked-plot-env-pft, fig.height=20}
# nut_plt <- 
# cmtx_filt_stn %>%
#   mutate(nit_pho_ratio = no3_no2_u_m / po4_u_m,
#          diat1_pct = diat1 / total_chl_a) %>%
#   pivot_longer(
#     data      = .,
#     cols      = c(diat1, diat1_pct, temperature, salinity, nh4_u_m:si_u_m, nit_pho_ratio), # columns to pivot long,
#     names_to  = "name", # desired name for category column
#     values_to = "value", # desired name for value column
#   ) %>%
#   ggplot(aes(y = value, x = date_time_utc, color = station, station = station)) +
#     geom_line() + 
#     geom_point() +
#     # geom_line(aes(y = diat1)) +
#     facet_wrap(~ name, ncol = 1, scale = "free_y") +
#     theme_minimal()
# 
# plotly::ggplotly(nut_plt)


# plt_phyto_fun <- function(data, .phyto) {
#     
#   data %>%
#   rename(phyto = {{ .phyto }}) %>%
#   mutate(nit_pho_ratio = no3_no2_u_m / po4_u_m,
#          phyto_pct = phyto / total_chl_a) %>%
#   pivot_longer(
#     data      = .,
#     cols      = c(phyto, phyto_pct, temperature, salinity, nh4_u_m:si_u_m, nit_pho_ratio), # columns to pivot long,
#     names_to  = "name", # desired name for category column
#     values_to = "value", # desired name for value column
#   ) %>%
#   ggplot(aes(y = value, x = date_time_utc, color = station, station = station)) +
#     geom_line() + 
#     geom_point() +
#     labs(title = .phyto) +
#     facet_wrap(~ name, ncol = 1, scale = "free_y") +
#     theme_minimal()
# }
# 
# 
# # plt_phyto_fun(cmtx_filt_stn, "diat1")
# 
# map(
#   .x = phyto_name$shrt[1:10],
#   \(.x, .data) plt_phyto_fun(.data, .x)  %>%
#     plotly::ggplotly() ,
#   .data = cmtx_filt_stn
# )
```

## Functions

```{r pft-env-func, fig.width=15}
plt_param <- function(.data, .x, .y, ...,  .y_unit = .y, .x_unit = .x, 
                      .smooth =  geom_smooth(method = "lm", se = FALSE),
                      total_smooth = FALSE) {
   plt <-  
  .data %>%
  rename(.x_var = {{ .x }}, .y_var = {{ .y }}) %>%
  ggplot(aes(y = .y_var, x = .x_var,  ...)) +
    geom_point(aes(group = date_time_utc, shape = season)) +
    labs(title = glue("{.x} vs {.y}"), 
         x = .x_unit,
         y = .y_unit
         ) +
    theme_bw() +
    .smooth
   
   if (total_smooth) {
       plt <- plt +
    geom_smooth(aes(y = .y_var, x = .x_var), method = "lm", se = FALSE,
                inherit.aes = FALSE,
                color = "black") #+
   }
   
    # facet_wrap(~ season, nrow = 2)
   return(plt)
}

plot_func <- function(.x, .y, .data, .color, .fill, .interact = FALSE,
                      .smooth =  geom_smooth(method = "lm", se = FALSE),
                      ...) {
  plt_param(.data, .x, .y, color = {{ .color }}, .smooth = .smooth, ...) %>%
    {
      if (.interact) {
        plotly::ggplotly()
      } else {
        .
      }
    }
}

plot_func_szn <- function(.x, .y, .data, .color, .fill) {
    (plt_param(.data, .x, .y, color = {{ .color }})  + 
    facet_wrap(~ season, nrow = 2)) %>%
      plotly::ggplotly()
  }

# select phyto names
phyto_sel <- 
    c(phyto_name$shrt[1:10], paste0(phyto_name$shrt[1:10], "_pct")) %>%
    sort()

names(cmtx_filt_stn)
plt_param(
  .data   = cmtx_filt_stn, 
  .x      = "po4_u_m", 
  .x_unit = "umol",
  .y      = "cyano4", 
  .y_unit = "mg / m^3",
  # fill   = station,
  color = station
  )
plt_param(
  .data   = cmtx_filt_stn, 
  .x      = "po4_u_m", 
  .x_unit = "umol",
  .y      = "cyano4", 
  .y_unit = "mg / m^3",
  # fill   = station,
  color = station
  )  + 
    facet_wrap(~ season, nrow = 2)
```

## Phosphate

```{r phos-plot, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "po4_u_m",
  .data  = cmtx_filt_stn,
  .color = dist,
  .interact = FALSE,
  plot_func
)

```

```{r phos-plot-filt-out, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "po4_u_m",
  .data  = filter(cmtx_filt_stn, po4_u_m < 0.4),
  .color = dist,
  plot_func
)
```

Season

```{r phos-plot-season, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "po4_u_m",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func_szn
)
```


## Temperature

```{r temp-plot, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "temperature",
  .data  = cmtx_filt_stn,
  .color = dist,
  total_smooth = TRUE,
  plot_func
)
```


```{r temp-plot-filt-out, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "temperature",
  .data  = filter(cmtx_filt_stn, total_chl_a < 1),
  .color = dist,
  # .interact = FALSE,
  # .smooth =  geom_smooth(method = "loess", se = FALSE),
  # total_smooth = FALSE
  plot_func
)
```

Season

```{r temp-plot-season, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "temperature",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func_szn
)
```


Temp vs Total Chla
```{r hplc-chl-plot, fig.width=15}
map2(
  .y     = "total_chl_a",
  .x     = "temperature",
  .data  = filter(cmtx_filt_stn, total_chl_a < 1),
  .color = dist,
  plot_func
)
map2(
  .y     = "total_chl_a",
  .x     = "temperature",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func
)

```

## Salinity

```{r salinity-plot, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "salinity",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func
)

```
```{r salinity-plot-filt-out, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "salinity",
  .data  = filter(cmtx_filt_stn, between(salinity, 33.15, 38.24)),
  .color = dist,
  plot_func
)

```

Season

```{r salinity-plot-season, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "salinity",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func_szn
)

```

## Ammonmium

```{r nh4-plot, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "nh4_u_m",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func
)

```
```{r nh4-plot-filt-out, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "nh4_u_m",
  .data  = filter(cmtx_filt_stn, nh4_u_m < 2),
  .color = dist,
  plot_func
)

```

Season

```{r nh4-plot-season, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "nh4_u_m",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func_szn
)

```

## Silicate

```{r silicate-plot, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "si_u_m",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func
)

```

```{r silicate-plot-filt-out, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "si_u_m",
  .data  = filter(cmtx_filt_stn, si_u_m < 10),
  .color = dist,
  plot_func
)

```

Season

```{r silicate-plot-season, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "si_u_m",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func_szn
)

```

## N+N

```{r n-n-plot, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "no3_no2_u_m",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func
)
```
```{r n-n-plot-filt-out, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "no3_no2_u_m",
  .data  = filter(cmtx_filt_stn, no3_no2_u_m < 1),
  .color = dist,
  plot_func
)
```

Season

```{r n-n-plot-season, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "no3_no2_u_m",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func_szn
)

```

## NO3

```{r no3-plot, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "no3_u_m",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func
)

```
```{r no3-plot-filt-out, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "no3_u_m",
  .data  = filter(cmtx_filt_stn, no3_u_m < 1),
  .color = dist,
  plot_func
)

```

Season

```{r no3-plot-season, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "no3_u_m",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func_szn
)

```

## NO2

```{r no2-plot, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "no2_u_m",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func
)

```

Season

```{r no2-plot-season, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "no2_u_m",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func_szn
)

```

```{r}
cmtx_filt_stn

```

## HPLC Total Chlorophyll-a
```{r hplc-chl-plot, fig.width=15}
# map2(
#   .y     = phyto_sel[6],
#   .x     = phyto_sel[8],
#   .data  = cmtx_filt_stn,
#   .color = dist,
#   total_smooth = TRUE,
#   plot_func
# )

```


```{r hplc-chl-plot, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "total_chl_a",
  .data  = cmtx_filt_stn,
  .color = dist,
  total_smooth = TRUE,
  plot_func
)

```
```{r hplc-chl-plot, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "total_chl_a",
  .data  = filter(cmtx_filt_stn, total_chl_a < 1),
  .color = dist,
  total_smooth = TRUE,
  .interact = TRUE,
  plot_func
)

```

Season

```{r hplc-chl-plot-season, fig.width=15}
map2(
  .y     = phyto_sel,
  .x     = "total_chl_a",
  .data  = cmtx_filt_stn,
  .color = dist,
  plot_func_szn
)

```



## DIN:DIP

```{r n-p-ratio-plot, fig.width=15}
cmtx_filt_stn %>%
  mutate(
    n_p_ratio = no3_no2_u_m / po4_u_m,
    n_p_ratio2 = (no3_no2_u_m + nh4_u_m) / po4_u_m
  ) %>%
  map2(
    .y     = phyto_sel,
    .x     = "n_p_ratio2",
    .data  = .,
    .color = dist,
    plot_func
  )

```

Season

```{r n-p-ratioplot-season, fig.width=15}
cmtx_filt_stn %>%
  mutate(
    n_p_ratio = no3_no2_u_m / po4_u_m,
    n_p_ratio2 = (no3_no2_u_m + nh4_u_m) / po4_u_m
  ) %>%
  map2(
    .y     = phyto_sel,
    .x     = "n_p_ratio2",
    .data  = .,
    .color = dist,
    plot_func_szn
  )

```



- event
"date_time_utc", "season", "station", "dist"                 

- PFT
"chloro", "crypto", "cyano2", "cyano4", "diat1", "diat2", "dino", "hapt6", 
"hapt8", "pras"                 

- environment
"temperature", "salinity", 
"total_chl_a", "avg_chl_a_ug_l", "avg_phaeophytin_ug_l", 
"nh4_u_m", "po4_u_m", "no3_no2_u_m", "no3_u_m", "no2_u_m", "si_u_m"



# 3.0 ---- Data Analysis ----


## 3.1 Plot: PFT vs Environmental Variables

Per station

```{r plt-pft-env-station, fig.width=25, fig.asp=0.5}
cmtx_nutr %>%
  select(
    season, station, lon = lon.x, chloro:pras, temperature, salinity, avg_chl_a_ug_l:si_u_m,
    -nutrients_tube_number
  ) %>%
  filter(
    not(
        cyano2 > 2 
      | chloro > 0.4
      | crypto > 0.2
      | diat1 > 0.6
      | diat2 > 0.3
      )
  ) %>%
  pivot_longer(
    data      = .,
    cols      = c(chloro:pras),
    names_to  = "pft",
    values_to = "conc",
  ) %>%
  nest(.by = c(pft)) %>%
  # slice(1) %>%
  # nest(.by = c(season, pft)) %>%
  mutate(
    plt = map2(
      pft,
      data,
      (\(.x, .y) {
        .y %>%
          pivot_longer(
            data      = .,
            cols      = c(temperature:si_u_m),  
            names_to  = "env_param",
            values_to = "value", 
            ) %>%
          mutate(station = fct_reorder(station, lon)) %>%
         # ggplot(aes(x = conc, y = value, color = season)) +
         ggplot(aes(x = conc, y = value, color = season)) +
         # geom_hline(yintercept = 0, color = "grey50", linetype = "dashed") +
         # geom_vline(xintercept = 0, color = "grey50", linetype = "dashed") +
         geom_point(na.rm = TRUE) +
              
              
         # linear model
         # geom_smooth(
         #     method = "lm", se = FALSE, na.rm = TRUE, formula = "y ~ x",
         #     color = "grey50", linetype = "dashed"
         #     ) +
         # geom_smooth(method = "lm", se = FALSE, na.rm = TRUE, formula = "y ~ x") +
         
         scale_x_continuous(breaks = scales::pretty_breaks()) +
         labs(title = .x) +
         theme_bw() +
         # facet_wrap(~env_param, ncol = 5, scales = "free") +
         facet_grid(env_param ~ station, scales = "free") +
         theme(panel.grid.major.x = element_line(color = "grey50", linetype = "dashed"))

      })
    )
  ) %>%
  pull(plt)
```


```{r, fig.width=25, fig.asp=0.5}
cmtx_nutr %>%
  select(
    station, season, chloro:pras, temperature, salinity, avg_chl_a_ug_l:si_u_m,
    -nutrients_tube_number
  ) %>%
  rowwise() %>%
  mutate(
    .before = chloro,
    sums    = sum(c_across(chloro:pras), na.rm = TRUE)
  ) %>%
 ungroup() %>%
 mutate(
   .after = sums,
   across(chloro:pras, \(x) (x / sums))
   ) %>%
  filter(
    not(
        cyano2 > 2 
      | chloro > 0.4
      | crypto > 0.2
      | diat1 > 0.6
      | diat2 > 0.3)
    # & !str_detect(station, "24|21.5|3|9.5|9B")
  ) %>%
  pivot_longer(
    data      = .,
    cols      = c(chloro:pras),
    names_to  = "pft",
    values_to = "conc",
  )  %T>% print() %>%
  nest(.by = c(pft, station)) %>%
    # slice(1) %>%
  # nest(.by = c(season, pft)) %>%
  mutate(
    pft = glue("{pft}, station {station}"),
    plt = map2(
      pft,
      data,
      (\(.x, .y) {
        .y %>%
          pivot_longer(
            data      = .,
            cols      = c(temperature:si_u_m),
            names_to  = "env_param", 
            values_to = "value",
            ) %>%
         ggplot(aes(y = conc, x = value, color = season
                    )) +
         # geom_hline(yintercept = 0, color = "grey50", linetype = "dashed") +
         # geom_vline(xintercept = 0, color = "grey50", linetype = "dashed") +
         geom_point(na.rm = TRUE) +
         scale_y_continuous(breaks = scales::pretty_breaks(),
                            labels = scales::label_percent(0.01)) +
         labs(title = .x) +
         facet_wrap(~env_param, ncol = 4, scales = "free") +
         # facet_wrap(env_param ~ station, ncol = 8, scales = "free") +
         theme_bw() +
         theme(panel.grid.major.x = element_line(color = "grey50", linetype = "dashed"))

      })
    )
  ) %>%
  pull(plt)

```


## 3. Plot: Environmental Variables vs Time

Add LOESS smoother

```{r plt-env-time, fig.width=20, fig.asp=0.5}
plt_nut_time <- 
  cmtx_nutr %>%
  select(
    date_time_utc, season, station, dist, chloro:pras, temperature, salinity, avg_chl_a_ug_l:si_u_m,
    -nutrients_tube_number
  ) %>%
  pivot_longer(
    data      = .,
    cols      = temperature:last_col(),     # columns to pivot long,
    names_to  = "env",  # desired name for category column
    values_to = "value", # desired name for value column
    ) %>%
    
  # filter(
  #     .by = "env",
  #     value < quantile(value, 0.95, na.rm = TRUE)
  #     ) %>%
    
  ggplot(aes(x = date_time_utc, y = value, color = dist)) +
  geom_hline(yintercept = 0, color = "grey50", linetype = "dashed") +
  geom_point(na.rm = TRUE) +
  # geom_smooth(color = "gray50", formula = "y ~ x", method = "loess", span = 0.15,
  #             na.rm = TRUE, se = FALSE) +
  theme_bw() +
  theme(panel.grid.major.x = element_line(color = "grey50", linetype = "dashed"))


plt_nut_time +
  facet_grid(env ~ station, scales = "free") 
plt_nut_time +
  facet_wrap(~ env, ncol = 4, scales = "free") +
  geom_smooth(color = "gray50", formula = "y ~ x", method = "loess", span = 0.15,
              na.rm = TRUE, se = FALSE)
plt_nut_time +
  facet_wrap( ~ env + dist, ncol = 4, scales = "free")# +
  geom_smooth(color = "gray50", formula = "y ~ x", method = "loess", span = 0.15,
              na.rm = TRUE, se = FALSE)
```


```{r plt-env-time, fig.width=20, fig.asp=0.5}
plt_nut_time <- 
  cmtx_nutr %>%
  select(
    date_time_utc, season, station, chloro:pras, temperature, salinity, avg_chl_a_ug_l:si_u_m,
    -nutrients_tube_number
  ) %>%
  pivot_longer(
    data      = .,
    cols      = temperature:last_col(),     # columns to pivot long,
    names_to  = "env",  # desired name for category column
    values_to = "value", # desired name for value column
    ) %>%
    
  filter(
      .by = "env",
      value < quantile(value, 0.95, na.rm = TRUE)
      ) %>%
    
  ggplot(aes(x = date_time_utc, y = value, color = season)) +
  # geom_hline(yintercept = 0, color = "grey50", linetype = "dashed") +
  geom_point(na.rm = TRUE) +
  # geom_smooth(color = "gray50", formula = "y ~ x", method = "loess", span = 0.15,
  #             na.rm = TRUE, se = FALSE) +
  theme_bw() +
  theme(panel.grid.major.x = element_line(color = "grey50", linetype = "dashed"))


plt_nut_time +
  facet_grid(env ~ station, scales = "free") 
plt_nut_time +
  facet_wrap(~ env, ncol = 4, scales = "free") +
  geom_smooth(color = "gray50", formula = "y ~ x", method = "loess", span = 0.15,
              na.rm = TRUE, se = FALSE) 
```



