---
title: "Ordination Plots"
author: "Sebastian Di Geronimo"
date: "2024-02-20"
format: html
editor: source
---

# 1.0 ---- Summary of Document ----

This document create nMDS plots for the Florida Keys stations.

Create 4 plots:
- *seasonal* and *nearshore* vs *offshore* for:
  - all stations
  - filtered stations 10, 12, 16, 18


# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
    # additional
    vegan, 
    gavinsimpson/ggvegan, 
    patchwork,
    prismatic
    # fawda123/ggord, 
    # jfq3/ggordiplots,
    # GGally, 
    # cowplot, 
    # ggfortify, 
    # ape,
    # ecodist # not sure if needed
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )
  
source(here("scripts", "misc_functions.R"))

# color palette
tol_muted <-
  c(
    "#88CCEE",
    "#44AA99",
    "#DDCC77",
    "#882255",
    "#AA4499",
    "#BBBBBB",
    "#FFAABB",
    "#117733",
    "#332288",
    "#999933",
    "#CC6677"
  )
```

## 2.2 Load Dataset

```{r load-data}
# set location to save data and plots
dir_data_save <- here("data", "processed", "06_3_ordination")
dir_plt_save  <- here("data", "plots", "06_3_ordination")

# ---- set season names and order ---- #
szn <- c("Winter", "Spring", "Summer", "Autumn")

# ---- phytoplankton full names ---- #
phyto_name <-
  here("data", "processed", "04_1_merge_meta_cmtx", "phyto_names.csv") %>%
  read_csv(show_col_types = FALSE) %T>% 
  print()

# ---- load chemtax data in wide format ---- #
cmtx_w <-
  here("data", "processed", "04_1_merge_meta_cmtx", "chemtax_w_meta_wide.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  filter(!str_detect(station, "60|21.5|24|^3$|9.5")) %>%
  mutate(
    season = fct(season, levels = szn),
    station = fct_reorder(station, lon)
  ) %>% 
  select(-c(hplc_gsfc_id, cruise:bottle, cluster_code, water_type)) %>%
    
  # add PFT concentrations to get total chlorophyll-a
  rowwise() %>%
  mutate(
    total_chl_a = sum(c_across(chloro:pras)),
    .before = chloro
  ) %>%
  ungroup() %>%
  filter(
    .by = c(station, cruise_id),
    depth == min(depth)
    & depth < 20
  ) %T>% 
  print()

```

## 2.2.1 Remove Outliers?

```{r remove-outliers}
remove_out <- menu(c("yes", "no"), title = "Remove Outliers?")

if (remove_out == 1) {
  cmtx_w <-
    cmtx_w %>%
    filter(
      total_chl_a <= 1
    ) %T>% 
    print()
  
  file_suffix <- "_no_out"
} else {
  file_suffix <- ""
}
```

## 2.3 Convert to Long Format

```{r convert-long-format}
# ---- convert to long format ---- #
# i.e. one column for all PFT and one for concentrations
cmtx_l <-
  cmtx_w %>%
  pivot_longer(
    cols      = chloro:pras,
    names_to  = "phyto",
    values_to = "conc"
  ) %>%
  left_join(phyto_name, by = c("phyto" = "shrt")) %>%
  
  # add percent
  mutate(percent = conc/total_chl_a) %T>% 
  print()

# ---- convert to percent ---- #
cmtx_l_pct <-
  cmtx_l %>%
  select(-conc) %T>% 
  print()

cmtx_w_pct <- 
  cmtx_l_pct %>%
  select(-names) %>%
  pivot_wider(
    data         = .,
    # id_cols      = c(), # *optional* vector of unaffected columns,
    names_from   = c(phyto), # category column(s) to pivot wide
    values_from  = c(percent), # value column(s) that hold data for each category column
    names_sep    = "_",
    names_repair = janitor::make_clean_names
    )
```

# ---- 3.0 nMDS ----


## 3.1 nMDS Calculation

Run nMDS:
1. All stations

2. 2x transects 
  - show near/offshore differences
  - stations 10, 12, 16, 18 


```{r nmds-type}
set.seed(1355453)

nmds_type <- c("abs", "rel")

dat_nmds <- 
  switch(
    nmds_type[2],
    abs = cmtx_w,
    rel = cmtx_w_pct
  ) %T>% print()

dist_nmds <- 
  switch(
    nmds_type[2],
    abs = "euclidean",
    rel = "bray"
  ) %T>% print()
```


```{r nmds-calc}
nmds <-
  dat_nmds %>%
  # cmtx_w %>%
  # cmtx_w_pct %>%
  select(chloro:last_col()) %>%
  # add(., 0.0001) %>%
  # log(.) %>%
  metaMDS(., distance = dist_nmds, try = 100, trymax = 100) %T>%
  # metaMDS(., distance = "bray", try = 100, trymax = 100) %T>%
  print()

# Call:
# metaMDS(comm = ., distance = "bray") 
# 
# global Multidimensional Scaling using monoMDS
# 
# Data:     . 
# Distance: bray 
# 
# Dimensions: 2 
# Stress:     0.130862 
# Stress type 1, weak ties
# Best solution was repeated 17 times in 20 tries
# The best solution was from try 4 (random start)
# Scaling: centring, PC rotation, halfchange scaling 
# Species: expanded scores based on ‘.’ 

# extract NMDS scores (x and y coordinates) for sites
data_scores <-
  dat_nmds %>%
  # cmtx_w %>%
  # cmtx_w_pct %>%
  bind_cols(scores(nmds)$sites) %>%
  mutate(
    labels = 
      if_else(
        str_detect(station, "10|16|7"), 
        "Nearshore", 
        "Offshore"
        )
    ) %T>%
  print()


# near/offshore pairs
set.seed(1355453)
nmds_filt <- 
  dat_nmds %>%
  # cmtx_w %>%
  # cmtx_w_pct %>%
  filter(str_detect(station, "10|12|16|18")) %>%
  select(chloro:last_col()) %>%
  metaMDS(., distance = dist_nmds, try = 100, trymax = 100) %T>%
  print()

# Call:
# metaMDS(comm = ., distance = "jaccard") 
# 
# global Multidimensional Scaling using monoMDS
# 
# Data:     . 
# Distance: jaccard 
# 
# Dimensions: 2 
# Stress:     0.1248301 
# Stress type 1, weak ties
# Best solution was repeated 15 times in 20 tries
# The best solution was from try 6 (random start)
# Scaling: centring, PC rotation, halfchange scaling 
# Species: expanded scores based on ‘.’

# extract NMDS scores (x and y coordinates) for sites
data_scores_filt <-
  dat_nmds %>%
  # cmtx_w %>%
  # cmtx_w_pct %>%
  filter(str_detect(station, "10|12|16|18")) %>%
  bind_cols(scores(nmds_filt)$sites)  %>%
  mutate(
    labels = 
      if_else(
        str_detect(station, "10|16|7"), 
        "Nearshore", 
        "Offshore"
        )
    ) %T>%
  print()

```


## 3.2 nMDS Plot

```{r base-plt-function}
nmds_plt_func <- function(.x) {
  
  ordiggplot(.x) +

  geom_ordi_axis(
    lty       = "dashed",
    color     = "black",
    linewidth = 0.2
  ) +
  
  # scale options
  scale_shape_manual(values = c(15:18)) +
  scale_color_manual(values = tol_muted) +
  scale_x_continuous(limits = c(-1.25, NA)) +
  # scale_x_continuous(limits = c(-2.5, NA)) +
  # labels
  labs(
    shape = "Season",
    color = "Season"
  ) +  
    
  # theme
  theme_bw(base_family = "serif") +
  theme(
    # overall
    text = element_text(family = "serif", size = 6, color = "black"),
    line = element_line(color = "black", linewidth = 0.2),
    rect = element_rect(color = "black", linewidth = 0.2),
    
    # axis
    axis.text.y  = element_text(
      colour = "black",
      size   = 5,
      face   = "bold"
    ),
    axis.text.x  = element_text(
      colour = "black",
      face   = "bold",
      size   = 5
    ),
    axis.title.y = element_text(
      face   = "bold",
      size   = 5,
      colour = "black"
    ),
    axis.title.x = element_text(
      face   = "bold",
      size   = 5,
      colour = "black"
    ),
    
    # legend
    legend.text  = element_text(
      size   = 3,
      face   = "bold",
      colour = "black"
    ),
    legend.byrow = TRUE,
    legend.position      = "top",
    legend.justification = "left",
    legend.title         = element_text(
      size   = 5,
      colour = "black",
      face   = "bold"
    ),
    legend.margin    = margin(b = -10),
    legend.key.size  = unit(2, "mm"),
    
    # panel
    panel.background = element_rect(color = "white"),
    panel.border     = element_rect(
      colour    = "black",
      fill      = NA,
      linewidth = 0.2
    )
  )
}
```

### 3.2.1 All Station Plots

```{r extract-nmds-info}
# data_centroid_season <- 
#   data_scores %>%
#   select(NMDS1, NMDS2, season) %>%
#   summarise(
#     .by = season,
#     NMDS1_mean = mean(NMDS1),
#     NMDS2_mean = mean(NMDS2)
#   ) %T>% 
#   print()
# 
# data_centroid_dist <- 
#   data_scores %>%
#   select(NMDS1, NMDS2, labels) %>%
#   summarise(
#     .by = labels,
#     NMDS1_mean = mean(NMDS1),
#     NMDS2_mean = mean(NMDS2)
#   ) %T>% 
#   print()
```


```{r nmds}
# shelf(camcorder)
# camcorder::gg_record(
#     dir    = here::here("data", "cam_test", "nmds_season_station"),
#     device = "png",
#     width  = 90,
#     height = 90 * 0.5625,
#     dpi    = 600,
#     units  = "mm"
# )

# shell.exec(here::here("data", "cam_test", "nmds_filt_season_station"))
# gg_stop_recording()

# gg_resize_film(width = 80, height = 90)

# season only
nmds_season <-
  nmds_plt_func(nmds) +
  stat_ellipse(
    data = data_scores,
    aes(x = NMDS1, y = NMDS2, color = season, shape = season),
    level     = 0.95,
    linetype  = 2,
    geom      = "polygon",
    linewidth = 0.25,
    alpha     = 0.01,
    show.legend = FALSE,
    inherit.aes = FALSE
    ) +
  geom_point(
    data = data_scores,
    aes(x = NMDS1, y = NMDS2, color = season, shape = season),
    size  = 0.5,
    alpha = 0.5,
    inherit.aes = FALSE
  ) +
  # geom_point(
  #   data = data_centroid_season,
  #   aes(
  #     x     = NMDS1_mean, 
  #     y     = NMDS2_mean, 
  #     fill  = season, 
  #     color = after_scale(clr_darken(fill))
  #     ),
  #   shape = 8,
  #   size  = 2,
  #   alpha = 1,
  #   inherit.aes = FALSE,
  #   show.legend = FALSE
  # ) +
  labs(color = "Season", fill = "Season") +
  scale_fill_manual(values = tol_muted[1:4]) +
  geom_ordi_arrow(
    "species", 
    linewidth = 0.2,
    text.params = list(
        size = 1, color = "black",
        fontface = "bold")
    ) +
  guides(
    fill  = guide_legend(nrow = 2, byrow = TRUE),
    shape = guide_legend(nrow = 2, byrow = TRUE)
  ) +
  theme(
    legend.key.spacing.y = unit(0.1, "mm"),
    legend.key.size      = unit(1, "mm"),
    plot.margin          = margin(b = 10)
    )

nmds_season

# station only
nmds_station <- 
  nmds_plt_func(nmds) +
  stat_ellipse(
    data = data_scores,
    aes(x = NMDS1, y = NMDS2, color = labels, shape = labels),
    level     = 0.95,
    linetype  = 2,
    geom      = "polygon",
    linewidth = 0.25,
    alpha     = 0.01,
    show.legend = FALSE,
    inherit.aes = FALSE
  ) +
  geom_point(
    data = data_scores,
    aes(x = NMDS1, y = NMDS2, color = labels, shape = labels),
    size  = 0.5,
    alpha = 0.75,
    inherit.aes = FALSE
  ) + 
  # geom_point(
  #   data = data_centroid_dist,
  #   aes(
  #     x     = NMDS1_mean, 
  #     y     = NMDS2_mean, 
  #     fill  = labels, 
  #     color = after_scale(clr_darken(fill))
  #     ),
  #   shape = 8,
  #   size  = 2,
  #   alpha = 1,
  #   inherit.aes = FALSE,
  #   show.legend = FALSE
  # ) +
  labs(
    shape = "Station",
    color = "Station",
    fill  = "Station"
   ) +
  scale_fill_manual(values = tol_muted[1:2]) +
  geom_ordi_arrow(
    "species", 
    linewidth   = 0.2,
    text.params = list(
      size     = 1, 
      color    = "black",
      fontface = "bold")
      )

nmds_station

# both season and station
nmds_season_station <- 
  (nmds_season / nmds_station) + 
  plot_annotation(
    tag_levels = "A",
    tag_prefix = "(",
    tag_suffix = ")") &
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    plot.tag.position = c(0.1, 0.9),
    plot.margin = margin(b = 1)
    )

nmds_season_station


```


Save plots


```{r save-all-plot}
ovrwrt <- FALSE
# ovrwrt <- TRUE

nmds_loc <- here(dir_plt_save, "nmds_all") 

# ---- save season and station ---- #
save_gg(
  plt            = nmds_season_station,
  save_location  = here(nmds_loc, str_remove(file_suffix, "^_")),
  save_name      = glue("nmds_all_both_station_season", file_suffix),
  verbose        = TRUE,
  overwrite      = ovrwrt,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "png",
  width          = 90 * 0.5,
  height         = 70,
  dpi            = 600,
  units          = "mm",
  bg             = "white"
)


# ---- save season and station separately ---- #
save_gg(
  plt            = nmds_station,
  save_location  = here(nmds_loc, str_remove(file_suffix, "^_")),
  save_name      = glue("nmds_all_station", file_suffix),
  verbose        = TRUE,
  overwrite      = ovrwrt,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "png",
  width          = 80,
  height         = 90 * 0.5,
  dpi            = 600,
  units          = "mm",
  bg             = "white"
)

save_gg(
  plt            = nmds_season,
  save_location  = here(nmds_loc, str_remove(file_suffix, "^_")),
  save_name      = glue("nmds_all_season", file_suffix),
  verbose        = TRUE,
  overwrite      = ovrwrt,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "png",
  width          = 80,
  height         = 90 * 0.5,
  dpi            = 600,
  units          = "mm",
  bg             = "white"
)

rm(nmds_loc)
```

### 3.2.2 Near/Offshore Station Plots

```{r extract-nmds-info}
# data_centroid_season <- 
#   data_scores_filt %>%
#   select(NMDS1, NMDS2, season) %>%
#   summarise(
#     .by = season,
#     NMDS1_mean = mean(NMDS1),
#     NMDS2_mean = mean(NMDS2)
#   ) %T>% 
#   print()
# 
# data_centroid_dist <- 
#   data_scores_filt %>%
#   select(NMDS1, NMDS2, labels) %>%
#   summarise(
#     .by = labels,
#     NMDS1_mean = mean(NMDS1),
#     NMDS2_mean = mean(NMDS2)
#   ) %T>% 
#   print()
```

```{r filt-nmds}
# season only
nmds_season <- 
  nmds_plt_func(nmds_filt) +
  stat_ellipse(
    data = data_scores_filt,
    aes(x = NMDS1, y = NMDS2, color = season, shape = season),
    level     = 0.95,
    linetype  = 2,
    geom      = "polygon",
    linewidth = 0.25,
    alpha     = 0.01,
    show.legend = FALSE,
    inherit.aes = FALSE
  ) +
  geom_point(
    data = data_scores_filt,
    aes(x = NMDS1, y = NMDS2, color = season, shape = season),
    size  = 0.5,
    alpha = 0.75,
    inherit.aes = FALSE
  ) +
  # geom_point(
  #   data = data_centroid_season,
  #   aes(
  #     x     = NMDS1_mean, 
  #     y     = NMDS2_mean, 
  #     fill  = season, 
  #     color = after_scale(clr_darken(fill))
  #     ),
  #   shape = 8,
  #   size  = 2,
  #   alpha = 1,
  #   inherit.aes = FALSE,
  #   show.legend = FALSE
  # ) +
  labs(color = "Season", fill = "Season") +
  scale_fill_manual(values = tol_muted[1:4]) +
  geom_ordi_arrow(
    "species", 
    linewidth   = 0.2,
    text.params = list(
      size     = 1, 
      color    = "black",
      fontface = "bold")
     ) +
  guides(
     fill  = guide_legend(nrow = 2, byrow = TRUE),
     shape = guide_legend(nrow = 2, byrow = TRUE)
  ) +
  theme(
    legend.key.spacing.y = unit(0.1, "mm"),
    legend.key.size      = unit(1, "mm"),
    plot.margin          = margin(b = 10)
    )


nmds_season

# station only
nmds_station <- 
  nmds_plt_func(nmds_filt) +
  stat_ellipse(
    data = data_scores_filt,
    aes(x = NMDS1, y = NMDS2, color = labels, shape = labels),
    level     = 0.95,
    linetype  = 2,
    geom      = "polygon",
    linewidth = 0.25,
    alpha     = 0.01,
    show.legend = FALSE,
    inherit.aes = FALSE
  ) +
  geom_point(
    data = data_scores_filt, 
    aes(x = NMDS1, y = NMDS2, color = labels, shape = labels),
    size  = 0.5,
    alpha = 0.75,
    inherit.aes = FALSE
  ) + 
  # geom_point(
  #   data = data_centroid_dist,
  #   aes(
  #     x     = NMDS1_mean, 
  #     y     = NMDS2_mean, 
  #     fill  = labels, 
  #     color = after_scale(clr_darken(fill))
  #     ),
  #   shape = 8,
  #   size  = 2,
  #   alpha = 1,
  #   inherit.aes = FALSE,
  #   show.legend = FALSE
  # ) +
  labs(
    shape = "Station",
    color = "Station"
   ) +
  scale_fill_manual(values = tol_muted[1:2]) +
  geom_ordi_arrow(
    "species", 
    linewidth   = 0.2,
    text.params = list(
      size     = 1, 
      color    = "black",
      fontface = "bold")
    )

nmds_station

# shelf(camcorder)
# camcorder::gg_record(
#     dir    = here::here("data", "cam_test", "nmds_filt_season_station"),
#     device = "png",
#     width  = 90 * 0.54,
# #     height = 90 * 0.5,
#     height = 70,
#     dpi    = 600,
#     units  = "mm"
# )

# shell.exec(here::here("data", "cam_test", "nmds_filt_season_station"))
# gg_stop_recording()

# gg_resize_film(width = 90 * 0.5, height = 70)
# gg_resize_film(width = 90 * 0.5, height = 70, dpi = 600, units = "mm")

# both 
nmds_season_station <-
  (nmds_season / nmds_station) +
  plot_annotation(
    tag_levels = "A",
    tag_prefix = "(",
    tag_suffix = ")") &
    theme(
      plot.background = element_rect(fill = "white", color = NA),
      plot.tag.position = c(0.1, 0.9),
      plot.margin = margin(b = 1)
      )

nmds_season_station
```


Save plots


```{r save-filt-nmds}
ovrwrt <- FALSE
# ovrwrt <- TRUE

nmds_loc_filt <- here(dir_plt_save, "filt") 

# ---- save season and station ---- #
save_gg(
  plt            = nmds_season_station,
  save_location  = here(nmds_loc_filt, str_remove(file_suffix, "^_")),
  save_name      = glue("nmds_filt_both_station_season", file_suffix),
  verbose        = TRUE,
  overwrite      = ovrwrt,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "png",
  width          = 90 * 0.5,
  height         = 70,
  dpi            = 600,
  units          = "mm",
  bg             = "white"
)

# ---- save season and station separately ---- #
save_gg(
  plt            = nmds_station,
  save_location  = here(nmds_loc_filt, str_remove(file_suffix, "^_")),
  save_name      = glue("nmds_filt_station", file_suffix),
  verbose        = TRUE,
  overwrite      = ovrwrt,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "png",
  width          = 80,
  height         = 90 * 0.5,
  dpi            = 600,
  units          = "mm",
  bg             = "white"
)

save_gg(
  plt            = nmds_season,
  save_location  = here(nmds_loc_filt, str_remove(file_suffix, "^_")),
  save_name      = glue("nmds_filt_season", file_suffix),
  verbose        = TRUE,
  overwrite      = ovrwrt,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "png",
  width          = 80,
  height         = 90 * 0.5,
  dpi            = 600,
  units          = "mm",
  bg             = "white"
)

rm(nmds_loc_filt)
```


# ---- 4.0 ANOSIM ----

```{r lib}
shelf(indicspecies)
```
## 4.1 ANOSIM Calc

```{r anosim}
temp <- 
cmtx_w_pct %>%
  mutate(
    .before = total_chl_a,
    labels  = 
      if_else(
        str_detect(station, "10|16|7"), 
        "Nearshore", 
        "Offshore"
        )
    ) 

# seasons
## absolute
anosim(
  select(cmtx_w, chloro:last_col()),
  cmtx_w$season,
  distance     = "bray",
  permutations = 999
)

## relative
anosim(
  select(cmtx_w_pct, chloro:last_col()),
  cmtx_w$season,
  distance     = "bray",
  permutations = 999
)

# near vs off
## absolute
anosim(
  select(cmtx_w, chloro:last_col()),
  temp$labels,
  distance     = "bray",
  permutations = 999
)

## relative
anosim(
  select(cmtx_w_pct, chloro:last_col()),
  temp$labels,
  distance     = "bray",
  permutations = 999
)


```

## 4.3 Similarity Percent

```{r simper}
vegan::simper(
  select(cmtx_w_pct, chloro:last_col()),
  temp$labels
) %>%
  summary(digits = 3)

vegan::simper(
  select(cmtx_w_pct, chloro:last_col()),
  cmtx_w_pct$season
) %>%
  summary(digits = 3)

```


## 4.2 Indicator Species Analysis

<https://emf-creaf.github.io/indicspecies/articles/IndicatorSpeciesAnalysis.html>

NOTES on using `multipatt`:
- the `.g` suffix is a correction for unequal grouped samples
- Indicator values calculated using `func = "IndVal"` or `func = "IndVal.g"`
  - used for predictive values of species as indicators of the conditions 
  - can display in summary Components 'A' and 'B' using 
    `summary(indval, indvalcomp = TRUE)`
  - 'A' = probability of site/sample belonging to target site group
    - aka specificity/positive predictive value
    - I think it means if 'A' is high for a specific species in group x, there 
      is a higher probability that if you see that specie within a sample, it is 
      more likely to be within in group 'x'
      ex: if 'A' is 1.0 for species 'x' in 'Group 1', then if species 'x' is 
          observed, it is a 100% probability to be in 'Group 1' based on the 
          input data. Maybe this extends to if include a new site, if this 
          specie is observed, you can say this site is probably in 'Group 1'
          
    
  - 'B' = probability of finding species in sites belonging to the site group
    - aka fidelity/sensitivity
    - I think it means that if 'B' is high for a specific species in a group, 
      then if you picked a specific group, high 'B' values would be the species 
      more likely to be observed
      ex: if 'B' is 1.0, then if you wanted to target group 2, the species with
          1.0 would be 100% to be seen in that group. This might be useful when
          trying to target specific species, you can go to the groups with high
          'B' values

  - Also, if 'B' is low and 'A' is high, then it means species x is a good 
    indicator for group 1, but you may not always see that species in group 1 
    but it was only seen in group 1. So if you observe it, it definitely in 
    group 1, but if you sampled all group 1, you may not see it, and you won't 
    see it in any other group
  - if high 'B' and low 'A', species x will indicate group 2 because it 
    definitely appears in all group 2, but it may also appear in other groups

- correlation indice using `func = "r"` or `func = "r.g"` 
  - represents Pearson's *phi coefficient of association*, to determine 
    ecological preferences of species among a set of alternative site groups
  - measures the correlation between two binary vectors
  - uses presence-absence
  - can use negative values and represents an avoidance of environmental 
    condition

### 4.2.1 IndVal: Absolute Season

```{r abs-season}
# Indicator Value
# multipatt(
#   select(cmtx_w, chloro:last_col()),
#   cmtx_w$season,
#   func = "IndVal",
#   control = how(nperm = 999)
# ) %>%
#   summary()

multipatt(
  select(cmtx_w, chloro:last_col()),
  cmtx_w$season,
  func    = "IndVal.g",
  control = how(nperm = 999)
) %>%
  summary()

# Pearson Coefficient Association
# multipatt(
#   select(cmtx_w, chloro:last_col()),
#   cmtx_w$season,
#   func = "r",
#   control = how(nperm = 999)
# ) %>%
#   summary()

multipatt(
  select(cmtx_w, chloro:last_col()),
  cmtx_w$season,
  func    = "r.g",
  control = how(nperm = 999)
) %>%
  summary()
```

### 4.2.2 IndVal: Absolute Nearshore/Offshore

```{r abs-nearoff}
# Indicator Value
# multipatt(
#   select(cmtx_w, chloro:last_col()),
#   temp$labels,
#   func = "IndVal",
#   control = how(nperm = 999),
# ) %>%
#   summary()

multipatt(
  select(cmtx_w, chloro:last_col()),
  temp$labels,
  func    = "IndVal.g",
  control = how(nperm = 999)
) %>%
  summary()

# Pearson Coefficient Association
# multipatt(
#   select(cmtx_w, chloro:last_col()),
#   temp$labels,
#   func = "r",
#   control = how(nperm = 999)
# ) %>%
#   summary()

multipatt(
  select(cmtx_w, chloro:last_col()),
  temp$labels,
  func    = "r.g",
  control = how(nperm = 999)
) %>%
  summary()
```


### 4.2.1 IndVal: Relative Season

```{r rel-season}
# Indicator Value
# multipatt(
#   select(cmtx_w_pct, chloro:last_col()),
#   cmtx_w$season,
#   func = "IndVal",
#   control = how(nperm = 999)
# ) %>%
#   summary()

multipatt(
  select(cmtx_w_pct, chloro:last_col()),
  cmtx_w$season,
  func    = "IndVal.g",
  control = how(nperm = 999)
) %>%
  summary()

# Pearson Coefficient Association
# multipatt(
#   select(cmtx_w_pct, chloro:last_col()),
#   cmtx_w$season,
#   func = "r",
#   control = how(nperm = 999)
# ) %>%
#   summary()

multipatt(
  select(cmtx_w_pct, chloro:last_col()),
  cmtx_w$season,
  func    = "r.g",
  control = how(nperm = 999)
) %>%
  summary(alpha = 1)
```


### 4.2.4 IndVal: Relative Nearshore/Offshore 

```{r rel-nearoff}
# Indicator Value
# multipatt(
#   select(cmtx_w_pct, chloro:last_col()),
#   temp$labels,
#   func = "IndVal",
#   control = how(nperm = 999)
# ) %>%
#   summary()

multipatt(
  select(cmtx_w_pct, chloro:last_col()),
  temp$labels,
  func    = "IndVal.g",
  control = how(nperm = 999)
) %>%
  summary()

# Pearson Coefficient Association
# multipatt(
#   select(cmtx_w_pct, chloro:last_col()),
#   temp$labels,
#   func = "r",
#   control = how(nperm = 999)
# ) %>%
#   summary()

multipatt(
  select(cmtx_w_pct, chloro:last_col()),
  temp$labels,
  func    = "r.g",
  control = how(nperm = 999)
) %>%
  summary(alpha = 1)

```





