---
title: "Template for Mapping"
author: "Sebastian DiGeronimo"
date: '2022-06-14'
output: html_document
---

# 1.0 ---- Summary of Document ----

- Map
  - NOA topography and coastline data
  - Site locations for IMaRS MBON
  - Inset map of Florida


- Kriging Maps
  - Pigment data 
  - kriged maps using the `fields` package 
    - modified code from Enrique Montes

# 2.0 ---- Setup ----

## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
    # additional 
    cowplot, metR, sf, shadowtext,
    # TODO: need to move away from raster and rgdal for terra
    fields, rgdal, ncdf4, raster 
)

conflicts_prefer(
    dplyr::filter(), 
    dplyr::select()
    )
knitr::opts_chunk$set(echo = TRUE)


# source functions
source(here("scripts", "map_files_dwnlod.R")) # map download function
source(here("scripts", "krig_fun.R"))         # Kriging
source(here("scripts", "misc_functions.R")) 

# fixes issue with reading file
sf::sf_use_s2(FALSE)
```


AI attempt at plotting general currents
```{r}
shelf(maps)
## 1. Simple basemap for W Atlantic / Gulf (replace with your own map)
world <- ggplot2::map_data("usa")

p_base <- ggplot() +
  geom_polygon(
    data = world,
    aes(long, lat, group = group),
    fill  = "grey85",
    color = "grey40",
    linewidth = 0.2
  ) +
  coord_quickmap(
    xlim = c(-90, -70),
    ylim = c(20, 30)
  ) +
  theme_minimal()

# Loop Current (Caribbean -> Yucatan Channel -> Gulf interior -> FL Straits)
loop_current <- tibble::tribble(
  ~lon,  ~lat,
  -84,   18,   # Caribbean north of Honduras
  -86.5, 20,   # into Yucatan Channel
  -87.5, 23,   # northward into central Gulf
  -87.5, 26,   # extended Loop
  -85,   24,   # turning east
  -82,   24    # exiting towards Florida Straits
)

# Florida Current (Straits of Florida)
florida_current <- tibble::tribble(
  ~lon,  ~lat,
  -82, 24,    # entry from Loop Current
  -80.5, 25,  # along Keys
  -80, 26.5   # off Miami
)

# Gulf Stream (up the U.S. East Coast)
gulf_stream <- tibble::tribble(
  ~lon,  ~lat,
  -80, 26.5,  # from Florida Current
  -79, 30,    # off NE Florida / Georgia
  -77, 33,    # off Carolinas
  -74, 36     # off mid‑Atlantic
)

p_base +
  # Loop Current
  geom_path(
    data = loop_current,
    aes(lon, lat),
    colour    = "dodgerblue3",
    linewidth = 1,
    lineend   = "round"
  ) +
  geom_segment(
    data = loop_current[nrow(loop_current), , drop = FALSE],
    aes(
      x = lon - 1, y = lat - 0.5,
      xend = lon,  yend = lat
    ),
    arrow = arrow(length = unit(0.3, "cm"), type = "closed"),
    colour = "dodgerblue3",
    linewidth = 1
  ) +
  annotate(
    "text",
    x = -87, y = 24.5,
    label = "Loop Current",
    colour = "dodgerblue3",
    fontface = "italic",
    size = 3.2
  ) +

  # Florida Current
  geom_path(
    data = florida_current,
    aes(lon, lat),
    colour    = "dodgerblue3",
    linewidth = 1,
    lineend   = "round"
  ) +
  annotate(
    "text",
    x = -81, y = 26.5,
    label = "Florida Current",
    colour = "dodgerblue3",
    fontface = "italic",
    size = 3.2
  ) +

  # Gulf Stream
  geom_path(
    data = gulf_stream,
    aes(lon, lat),
    colour    = "dodgerblue3",
    linewidth = 1,
    lineend   = "round"
  ) +
  geom_segment(
    data = gulf_stream[nrow(gulf_stream), , drop = FALSE],
    aes(
      x = lon - 1, y = lat - 0.5,
      xend = lon,  yend = lat
    ),
    arrow = arrow(length = unit(0.3, "cm"), type = "closed"),
    colour = "dodgerblue3",
    linewidth = 1
  ) +
  annotate(
    "text",
    x = -77.5, y = 34.5,
    label = "Gulf Stream",
    colour = "dodgerblue3",
    fontface = "italic",
    size = 3.2
  )

currents <- data.frame(
  lon = c(-88,-86,-84,-82,-80),
  lat = c(24,25,26,27,28),
  u = c(0.3,0.4,0.6,0.8,1.0),
  v = c(0.1,0.2,0.3,0.5,0.7)
)

scale_factor <- 1.5   # controls arrow length

ggplot() +

  geom_polygon(data = world,
               aes(long, lat, group = group),
               fill = "grey85",
               color = "black",
               linewidth = 0.2) +

  geom_segment(data = currents,
               aes(x = lon,
                   y = lat,
                   xend = lon + u * scale_factor,
                   yend = lat + v * scale_factor),
               arrow = arrow(length = unit(0.15,"cm")),
               linewidth = 0.5,
               color = "black") +

  coord_quickmap(xlim = c(-95,-70), ylim = c(20,40)) +

  theme_classic() +
  labs(x = "Longitude", y = "Latitude")
```
```{r WIP2}
# shelf(grid, smoothr, maps)
# ## 1. Simple basemap for W Atlantic / Gulf (replace with your own map)
# world <- ggplot2::map_data("usa")
# 
# p_base <- ggplot() +
#   geom_polygon(
#     data = world,
#     aes(long, lat, group = group),
#     fill  = "grey85",
#     color = "grey40",
#     linewidth = 0.2
#   ) +
#   coord_quickmap(
#     xlim = c(-90, -70),
#     ylim = c(20, 30)
#   ) +
#   theme_minimal()
# 
# 
# 
# 
# 
# p_base +
#   geom_sf(
#     data = loop_smooth,
#     color = "dodgerblue3",
#     linewidth = 1
#   ) +
#   # add an arrow at the downstream end
#   geom_segment(
#     data = loop_pts[nrow(loop_pts), , drop = FALSE],
#     aes(
#       x = lon - 0.7, y = lat - 0.3,
#       xend = lon,   yend = lat
#     ),
#     arrow = arrow(length = unit(0.3, "cm"), type = "closed"),
#     color = "dodgerblue3",
#     linewidth = 1,
#     inherit.aes = FALSE
#   ) +
#   annotate(
#     "text",
#     x = -87, y = 24.5,
#     label = "Loop Current",
#     colour = "dodgerblue3",
#     fontface = "italic",
#     size = 3.2
#   ) +
#   geom_point(
#     data = loop_pts,
#     aes(x = lon, y = lat)
#   ) +
#   coord_sf(xlim = c(-90,-80), ylim = c(20,30))
```


## 2.2 Set Map Information

Set paths for download and search:
  - land 
  - topographic 
    - Note: they don't have to be the same location
    
Set spatial extend as a bounding box
- Florida Keys
  - zoomed in for stations specific
  - Min lon/lat: 
    * -82.00 deg W
    * 24.25 deg N
  - Max lon/lat: 
    * -80.00 deg W
    * 24.25 deg N

- Florida
  - zoomed out for inset map
  - Min lon/lat: 
    * -90.00 deg W
    * 20.00 deg N   
  - Max lon/lat: 
    * -75.00 deg W
    * 35.00 deg N

```{r map-info, include=FALSE}
# path for map to find/download to, they can be different depending on
# your computer
path_land <- here("data", "map")
path_topo <- here("data", "map")

# Region of Interest Bounding Box                   ======== edit this if needed
# Change the spatial extent
exnt <- c(xmin = -82,   # West
          xmax = -80,   # East
          ymin = 24.25, # South
          ymax = 25.75  # North 
        )

# inset extent (i.e. smaller map on top)            ======== edit this if needed
exnt2 <-  c(-90, # West
            -75, # East
            20,  # South
            35   # North
            ) 

names(exnt2) <- names(exnt)
```


## 2.3 Load Dataset

Load data sets:
- Combined Pigment Data created in `01_load_data.Rmd`
 
TODO: *load additonal data*
- CHEMTAX Seasonal Averages created in `04_chemtax_analysis.Rmd`
- Surface Averaged CTD Data created in `02_ctd_download.Rmd`
- Merge CHETMAX-CTD

```{r load-pigment-data}
dir_plt_save <- here("data", "plots", "map")

pig_dat  <- 
  here("data", "processed", "load_data", "combined_pig_dat.csv") %>%
  read_csv(show_col_types = FALSE) %T>% print()
```


## 2.4 Download and Load Bathymetry and Coastline Files

**Note:** this will be done automatically if files **do not exists**

Topography, ETOPO1, 0.0166667 degrees, Global (longitude -180 to 180), (Ice Sheet Surface) from https://coastwatch.pfeg.noaa.gov/erddap/griddap/

Global Self-consistent, Hierarchical, High-resolution Geography Database (GSHHG) from https://www.ngdc.noaa.gov/mgg/shorelines/
```{r maps-download-load}
world_download(
    path_land = path_land,
    path_topo = path_topo,
    extent    = exnt,
    .timeout = 1000
)

# Load coastline and topography data
map_obj <- 
    load_map_obj( 
    .map_coast = path_land, 
    .map_state = path_land,
    .map_bath  = path_topo, 
    .map_file  = "etopo1.nc",
    .extent    = exnt2
    )

# create base map
(base_map <-
  with(
    map_obj,
    base_map_plot(
      coast_topo,
      bathy,
      exnt,
      .set_breaks = c(10, 50, 100, 200),
      familyface = NULL,
      # contour_labels = TRUE
    )
  )
)
```
```{r}
shelf(seascapeR)

ply_fknms <-
  get_url_ply(
    sanctuary = "fknms",
    dir_ply = here("data", "map", "sanctuary")
  ) %>%
  nngeo::st_remove_holes()

```



# 3.0 ---- Plot Mapped Station ----

## 3.1 Plot FL Keys stations on map

Use station locations from pigment lon/lat and add to base map created in 
section 2.4

```{r plt-fl-keys, fig.height=6, fig.width=10}
lat_lon_dat <- 
  pig_dat %>%
  filter(
      name_of_water_body == "Florida Keys"
      & !str_detect(station, "21.5|9B|24|^3$|9.5")
      # & str_detect(station, "MR|LK|WS")
          ) %>%
  select(station, lat, lon) %>%
  distinct(station, lat, lon) %>% 
  mutate(
    colors = 
      if_else(
        str_detect(station, "10|16|7|^3"),
        "Nearshore",
        "Offshore"
        )
    ) 


stn_map <-
  base_map +
  geom_sf(data = ply_fknms, fill = NA, color = "red", linewidth = rel(1.1)) +
  # geom_sf(data = nngeo::st_remove_holes(ply_fknms), fill = NA, color = "red", alpha = 0.05) +
  
  annotate(
    geom        = "shadowtext",
    # x           = c(-80.52, -81.1, -81.85, -81.95),
    # y           = c(24.84, 24.61, 24.455, 24.49),
    # label       = c(100, 50, 25, 10),
    # angle       = c(35, 24, 15, -3),
    x           = c(-80.48, -80.52, -81.1, -81.95),
    y           = c(24.7, 24.84, 24.61, 24.49),
    angle       = c(35, 35, 24, -3),
    label       = c(200, 100, 50, 10),
    size        = 7,
    # fontface    = "bold",
    # label.color = NA,
    # family      = "serif",
    # family      = "TT Arial",
    color = "grey",
    bg.color = "white"
  ) +
  geom_point(
    data = lat_lon_dat,
    aes(
      x = lon,
      y = lat,
      fill = colors,
      shape = colors,
    ),
    stroke      = 0.5,
    size        = 6,
    # alpha     = 0.5,
    # show.legend = FALSE
  ) +

  ggrepel::geom_label_repel(
    aes(x = lon, 
        y = lat, 
        label = station),
    data   = lat_lon_dat,
    size   = 6,
    hjust  = 0,
    # family = "serif",
    # family = "TT Arial",
    point.padding = 10,
    min.segment.length = 0.01
    # direction = "y"
  ) +    

  annotate(
    geom        = "text",
    x           = c(-80.73, -80.7, -80.75, -80.25),
    y           = c(25.06, 24.5, 25.55, 25.55),
    label       = c("Florida Bay", "Atlantic Ocean", "Everglades", "Biscayne Bay"),
    size        = 7,
    fontface    = "bold",
    angle = c(30, 30, 0, 72)
  ) +
    
  labs(
    shape = "Station",
    fill  = "Station"
  ) +
  coord_sf(
    xlim = exnt[1:2], 
    ylim = exnt[3:4]) +
  scale_shape_manual(values = c(25, 21)) +
  scale_fill_manual(values = c("#88CCEE", "#882255")) +
  theme(
    legend.position        = "inside",
    legend.position.inside = c(0.85, 0.1), 
    legend.background      = element_rect(color = "black"),
    plot.background        = element_rect(fill = "white"),
    legend.text            = element_text(size = 15),
    legend.title           = element_text(size = 17),
    legend.key.height = unit(10, "mm")
  )

stn_map



```
```{r}
(stn_map2 <- 
  stn_map +
  geom_segment(
    # data = ,
    aes(x = c(-81.2, -81.3, -80.95),    y = c(24.99, 25.3, 25.1),
        xend = c(-81.17, -81.25, -81.1), yend = c(24.8, 25.11, 25)
        ),
    arrow = arrow(length = unit(0.3, "cm"), type = "closed"),
    color = "black",
    linewidth = 1.2,
    inherit.aes = FALSE
  ))
```



## 3.2 Create Inset Florida Map

Zoomed out map of Florida to be added on top of the station map

```{r plt-large, fig.width=8, fig.height=10}
inset_map <-
  with(
    map_obj,
    map_inset(coast_topo, state,
      c(
        xmin = min(lat_lon_dat$lon), # West
        xmax = max(lat_lon_dat$lon), # East
        ymin = min(lat_lon_dat$lat), # South
        ymax = max(lat_lon_dat$lat) # North
      ),
      add_tol = 0.25
    )
  )

inset_map
```
```{r WIP2}
# Example: schematic Loop Current waypoints (lon/lat)
loop_pts <- tibble::tribble(
  ~lon,  ~lat,
  -80.7,   18, # 1
  -85., 20, # 2
  -87.2, 23, # 3
  -87.5, 26, # 4
  -86, 26.5, # 5
  -84.5, 24.2, # 6
  -82,   23.9 # 7
) %>%
  mutate(type = "loop")

# 1) Make an sf LINESTRING
loop_line <- st_as_sf(loop_pts, coords = c("lon", "lat"), crs = 4326) %>%
  summarise(do_union = FALSE) %>%
  st_cast("LINESTRING")

# 2) Smooth it (Chaikin or spline)
loop_smooth <- smoothr::smooth(
  loop_line,
  method = "spline",   # or "chaikin"
  vertex_factor = 5       # increase for more curvature
)  %>%
  mutate(type = "loop")

# Example: schematic Florida Current waypoints (lon/lat)
fl_pts <- tibble::tribble(
  ~lon,  ~lat,
  -82,   23.9, # 1
  -80,   24.5, # 2
  -79.65,   25.5, # 3 
) %>%
  mutate(type = "fl")

# 1) Make an sf LINESTRING
fl_line <- st_as_sf(fl_pts, coords = c("lon", "lat"), crs = 4326) %>%
  summarise(do_union = FALSE) %>%
  st_cast("LINESTRING")

# 2) Smooth it (Chaikin or spline)
fl_smooth <- smoothr::smooth(
  fl_line,
  method = "spline",   # or "chaikin"
  vertex_factor = 5       # increase for more curvature
)  %>%
  mutate(type = "fl")

# Example: schematic Gulf Stream waypoints (lon/lat)
gs_pts <- tibble::tribble(
  ~lon,  ~lat,
  -79.65,   25.5, # 1
  -79.65,   28, # 2
  -80,   31, # 3
  -75,   34.5, # 4
) %>%
  mutate(type = "gs")

# 1) Make an sf LINESTRING
gs_line <- st_as_sf(gs_pts, coords = c("lon", "lat"), crs = 4326) %>%
  summarise(do_union = FALSE) %>%
  st_cast("LINESTRING")

# 2) Smooth it (Chaikin or spline)
gs_smooth <- smoothr::smooth(
  gs_line,
  method = "spline",   # or "chaikin"
  vertex_factor = 5       # increase for more curvature
)  %>%
  mutate(type = "gs")

inset_map2 <- 
  inset_map +
   geom_sf(
    data      = bind_rows(gs_smooth, fl_smooth, loop_smooth),
    color     = c("green", "red", "dodgerblue3"),
    fill      = c("green", "red", "dodgerblue3"),
    linewidth = 1,
     arrow    = arrow(
       ends = "last", 
       type = "closed", 
       length = unit(0.3, "cm"))
  ) +

  annotate(
    "text",
    x        = c(-86.1, -79.5, -78.5), 
    y        = c(24.5, 24.25, 30),
    angle    = c(0, 32, 85),
    label    = c("Loop\nCurrent", "Florida Current", "Gulf Stream"),
    colour   = c("dodgerblue3", "red", "green"),
    fontface = "italic",
    size     = 3.2
  ) +
  # geom_point(
  #   data = loop_pts,
  #   aes(x = lon, y = lat)
  # ) +
  labs(x = NULL, y = NULL) +
  coord_sf(ylim = c(min(lat_lon_dat$lat) - 3.5, NA))
```

## 3.3 Merge Station Map and Florida Map

Combine the two maps created in sectin 3.2 and 3.3

May need to play with draw_plot(east.map, x, y, width and height)


```{r plt-lg-sm, fig.height=6, fig.width=10}
fullmap <- 
  ggdraw() +
  draw_plot(stn_map2) +
  draw_plot(inset_map2, x = 0.06, y = 0.52, width = 0.3, height = 0.5)
fullmap

# ---- save plot
walk(
  # c("png", "svg"),
  c("jpeg"),
  ~ save_gg(
    plt            = fullmap,
    save_location  = dir_plt_save,
    save_name      = "map_station_locations",
    verbose        = TRUE,
    # overwrite      = TRUE,
    # time_stamp_fmt = NULL,
    time_stamp_fmt = "%Y%m%d",
    device         = .x,
    height         = 8.5,
    width          = 10,
    dpi            = 600,
    units          = "in",
    bg             = "white"
  )
)

# shelf(camcorder)
# camcorder::gg_record(
#   dir    = here::here("data", "cam_test", "map_of_station"),
#   device         = "png",
#   height         = 8.5,
#   width          = 10,
#   dpi            = 600,
#   units          = "in",
#   bg             = "white"
# )

# shell.exec(here::here("data", "cam_test", "map"))
# camcorder::gg_stop_recording()

# camcorder::gg_resize_film(width = 90, height = 90 * 0.5625,  dpi = 600, units = "mm")

```


# 4.0 ---- Kriging for All Pigments ----

```{r clean workspace}
# clean up workspace
suppressWarnings(rm(stn_map, inset_map, lat_lon_dat, exnt, exnt2))
```

## 4.1 Prepare Data for Spatial Krig

Steps:
1. Select necesary columns
2. Filter data to the `loc` variable
  - Florida Keys
  - Southwest Florida Shelf
  - West Florid Shelf
3. Pivot long pigment concentrations 
4. Calculate the pigment concentartion:chlor-a ratio
5. Calculate an average per season and pigment
6. Add columns for map titles and file names

```{r prep-data}
# select subregion
loc <- "Florida Keys"

# filter for pigment and season
pigm_dat <-
  pig_dat %>%
  select(
    hplc_gsfc_id, sample,                                     # <-- id
    station, lat, lon, name_of_water_body,                    # <-- spatial
    year, date_time_utc, season,                              # <-- temporal
    tot_chl_a, tot_chl_b:dp                                   # <-- measurements
  ) %>% 
    
  # select region
  filter(str_detect(name_of_water_body, loc)) %>%
  mutate(
    .before    = tot_chl_a,
    tot_chl_a2 = tot_chl_a) %>%
  pivot_longer(
    tot_chl_a:gyro,
    names_to  = "conc_name",
    values_to = "conc"
  ) %>%
  
  # calc ratio of each pigment to total chlor-a
  mutate(ratio_pig_chlora = conc / tot_chl_a2) %>%
      
  # calc averages per season and station
  summarise(
    .by       = c(season, station, conc_name),
    conc_avg  = mean(conc, na.rm = TRUE),
    ratio_avg = mean(ratio_pig_chlora, na.rm = TRUE),
    lat       = median(lat, na.rm = TRUE),
    lon       = median(lon, na.rm = TRUE),
  ) %>%
      
  # nest by season and pigment
  nest(.by = c(season, conc_name)) %>%
  
  # add title and file names     
  mutate(
    title_conc    = glue("{season}: Average {conc_name}"),
    title_ratio   = glue("{season}: {conc_name}:TChla"),
    sv_name_conc  = glue("mbon_pigm_{conc_name}_conc_{str_to_lower(season)}"),
    sv_name_ratio = glue("mbon_pigm_{conc_name}_ratio_{str_to_lower(season)}")
  )
  
```

## 4.2 Calculate: Spatial Krig for Each Pigment

Uses pigment concentrations or pigment ratios to calculate an estimated spatial kirged map using either pigment concentrations  or pigment:chlor-a ratio


Some samples may not produce maps due to how the math works and will be 
ignored. This usually happens when there's no variation in data or something 
else. These samples will be shown if it needs to be looked into.


```{r cal-krig, fig.width=30, fig.height=15}
# select ratio or concentrations to krig
pick_rat_con <-
  tribble(
      ~col_name_dat, ~save_name,        ~folder,
      "conc_avg",  "sv_name_conc", "concentration",
      "ratio_avg", "sv_name_ratio", "ratio"
  ) %>%
  filter(str_detect(col_name_dat, "(?i)conc")) %T>% print()
  # filter(str_detect(col_name_dat, "(?i)ratio")) %T>% print()

cli_alert_info(
  "{.var {col_red(pick_rat_con$folder)}} was selected to perform Kriging.")

# perfrom Krig
krig_pigments <-
  pigm_dat %>%
  nest(.by = conc_name) %>%
      
  # select proportion of pigment cocentrations as decimal 0.0 - 1.0
  slice_sample(prop = 1.0) %>%
  unnest(data) %>%
  
  mutate(
    krig = pmap(
      list(data, conc_name, season, title_ratio, row_number()),
      .progress = TRUE,
      possibly(
        \(x, y, z, title, row) {
          cli::cli_alert_info("Map {row} of {nrow(pigm_dat)}")
          map_data(
            .data     = x,
            conc_name = y,
            col_name  = pick_rat_con$col_name_dat,
            grp_name  = z,
            .title    = title,
            base_map  = base_map,
          )
        },
        otherwise = NULL
      )
    )
  ) 

# show season pigment pair with errors
filter(krig_pigments, map_lgl(krig, is.null))

# remove rows with errors to plot and save
krig_pigments <- 
  krig_pigments %>%
  # filter out `NULL`
  filter(!map_lgl(krig, is.null)) %>%
  hoist(krig, "dat", "map")

# sample 5 pigments season combos to view data and map 
krig_pigments %>%
  slice_sample(n = 5) %$%
  walk2(data, map, (\(x, y) { print(x); print(y)} ))

# save krig_pigments
save_map(
  n_files         = 0,    # edit to include more files 0 - nrow(krig_pigments)
  prop_files      = NULL, # edit to include more files 0 - 1%
  .data           = krig_pigments,
  folder_location = here(dir_plt_save, pick_rat_con$folder),
  save_column     = pick_rat_con$save_name,
  time_stamp_fmt  = "%Y%m%d_%H%M%S",
  verbose         = TRUE,
  overwrite       = FALSE,
  device          = "jpeg",
  height          = 9.7,
  width           = 19,
  dpi             = 600,
  units           = "in",
  bg              = "white"
)

```



# Don't Use: Old plotting method
```{r base-plot-fuco}
# if (FALSE) {
# png(file, height = 11, width = 4, units = "in", res = 300)
# par(
#     mfrow = c(3, 1),
#     mar = c(4.5, 4, 2, 1),
#     oma = c(4, 1, 4, 1)
# )
# 
# imagePlot(
#     ratio_kriged_fuco$x,
#     ratio_kriged_fuco$y,
#     ratio_kriged_fuco$z,
#     col = cols_fuco,
#     breaks = breaks_fuco,
#     asp = 1,
#     xlab = '',
#     ylab = '',
#     las = 1,
#     xlim = xlims_fuco,
#     ylim = ylims_fuco,
#     nlevel = length(cols_fuco),
#     legend.width = 1,
#     legend.mar = 3
# )
# 
# contour(
#     ratio_kriged_fuco$x,
#     ratio_kriged_fuco$y,
#     ratio_kriged_fuco$z,
#     levels = breaks_fuco,
#     add = TRUE
# )
# 
# image(
#     ratio_SE_fuco,
#     add = TRUE,
#     breaks = quantile(ratio_SE_fuco$z, c(1, 1), na.rm = TRUE),
#     col = 'white'
# )
# 
# image(
#     topo_lon,
#     topo_lat,
#     topo,
#     breaks = c(-1, 100),
#     col = 'white',
#     add = TRUE
# )
# 
# # plot(FL,col='gray70',add=TRUE)
# plot(world2, col = 'gray70', add = TRUE)
# contour(
#     topo_lon,
#     topo_lat,
#     topo,
#     add = TRUE,
#     levels = c(-100, -50, -25, -10),
#     col = 'gray40'
# )
# 
# points(df_fuco$lon,
#        df_fuco$lat,
#        pch = 20,
#        col = 'yellow')
# 
# mtext(expression(paste('Longitude (\u00B0W)')), 1, line = 3, cex = 0.75)
# 
# mtext(expression(paste('Latitude (\u00B0N)')), 2, line = 3, cex = 0.75)
# mtext('Fucoxanthin:TChla: Winter', adj = 1, cex = 0.75)
# }
```