---
title: "Load Data"
author: "Sebastian DiGeronimo"
date: '2022-06-13'
output: html_document
---

# Small lists of things to do
* rename headers to be `tidy`
    can use: .name_repair = janitor::make_clean_names inside read_csv
    
* combine date and time columns to a standard date_time
    - Something like this
     mutate(
    date_time_utc = ymd_hms(paste(`year col`, `month col`, `day col`, `time col`), tz = "utc"),
    .before = lat
  )
  
* set -8888 to 0
* address notes from file to potentially ignore certain rows 
* start plotting
    * histogram, per stations, per pigments, etc
    * simple plots over time
    * maybe some heatmap
    * maybe map

```{r setup, include=FALSE}
root <- rprojroot::find_rstudio_root_file()
knitr::opts_chunk$set(echo = TRUE)
library("ggplot2")
library("tibble")
library("tidyr")
library("readr")
library("purrr")
library("dplyr")
library("stringr")
library("forcats")
library("lubridate")
library("glue")
library("fs")
library("magrittr")
```

```{r create-dir}
# will create a set directory if does not exists
# useful for new projects
mainDir <- rprojroot::find_rstudio_root_file()
subDir <-
    c("data/raw",
      "data/processed",
      "data/plots",
      "data/metadata",
      "Rmd",
      "scripts")

fs::dir_create(path = paste0(mainDir,"/",subDir))
rm(mainDir, subDir)
```


```{r file-paths}
# find carbonate data
dir <- "//data//raw//"

# pigment files
file.pig <-
  fs::dir_ls(path = paste0(root,dir),
             recurse = TRUE,
             regexp = "\\.xlsx$")

```

```{r}
df <- readxl::read_xlsx(file.pig[1], sheet = 2, skip = 8, na = "-9999", .name_repair = janitor::make_clean_names)
df[df == -8888] <- 0
df$gmt_time <- gsub("1899-12-31 ", "", as.character(df$gmt_time))
df <- mutate(df, date_time_utc = ymd_hms(paste(year_of_sample, gregorian_month, day_of_gregorian_month, gmt_time), tz = "utc"), .before = longitude)

# janitor functions
# str(df)

# df %>%
#     filter(`[Allo]` >= 0) %>%
#     ggplot(aes(`Sequential Day of Year`,`[Allo]`)) +
#     geom_point()
```

```{r}
df2 <- 
    readxl::read_xlsx(file.pig[2], sheet = 2, skip = 8, na = "-9999", 
                      .name_repair = janitor::make_clean_names)
df2[df2 == -8888] <- 0
df2$time <- gsub("1899-12-31 ", "", as.character(df2$time))
df2 <-
    mutate(df2,
           date_time_utc = ymd_hms(paste(year, month, day, time), tz = "utc"),
           .before = lon)
df2$indicate_if_filters_are_replicates <- as.character(df2$indicate_if_filters_are_replicates)


df3 <-readxl::read_xlsx(file.pig[3], sheet = "Report", skip = 8, na = "-9999", 
                      .name_repair = janitor::make_clean_names) 
df3[df3 == -8888] <- 0
df3$time <- gsub("1899-12-31 ", "", as.character(df3$time))
df3 <- mutate(df3, date_time_utc = ymd_hms(paste(year, month, day, time), tz = "utc"), 
              water_depth = case_when(water_depth == "flow trough" ~ 0,
                                      TRUE ~ as.numeric(water_depth)),
              .before = lon) 

df4 <- 
    readxl::read_xlsx(file.pig[4], sheet = "Report", skip = 8, na = "-9999", 
                      .name_repair = janitor::make_clean_names)
df4[df4 == -8888] <- 0
df4$time <- gsub("1899-12-31 ", "", as.character(df4$time))
df4 <- mutate(df4, date_time_utc = ymd_hms(paste(year, month, day, time), tz = "utc"), .before = lon)

df234 <- full_join(df3, df2)
df234 <- mutate(df234, hplc_gsfc_id = case_when(is.na(hplc_gsfc_id) ~ gsfc_sample_code,
                                                TRUE ~ hplc_gsfc_id))
```
```{r}
df5 <- 
    readxl::read_xlsx(file.pig[5], sheet = "Report", skip = 8, na = "-9999", 
                      .name_repair =janitor::make_clean_names)
df5[df5 == -8888] <- 0
df5$gmt_time <- gsub("1899-12-31 ", "", as.character(df5$gmt_time))
df <- mutate(df, date_time_utc = ymd_hms(paste(year_of_sample, gregorian_month, day_of_gregorian_month, gmt_time), tz = "utc"), .before = longitude)




#for looking at similarities in the column names between different files
# x <- cbind(names(df),names(df2),names(df3),names(df4),names(df5))
```

