---
title: "Template for Mapping"
author: "Sebastian DiGeronimo"
date: '2022-06-14'
output: html_document
---

# USage

## Creates map of site locations for IMaRS MBON
Uses topography and coastline data from NOAA.
Creates a map of station locations with an inset map of Florida


## Kriging Maps
Uses pigment data and creates kriged maps using the `fields` package and some 
modified code from Enrique Montes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
librarian::shelf(
    librarian, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here, conflicted,
    # broom # optional
     
    # additional 
    fields, cowplot, metR, rgdal, ncdf4, sf, 
    raster
)

# prefer select and filter 
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

# source map download function
source(here("scripts", "map_files_dwnlod.R"))

# fixes issue with reading file
sf::sf_use_s2(FALSE)
```


```{r map-info, include=FALSE}
# path for map to find/download to, they can be different depending on
# your computer
path_land <- here("data", "map")
path_topo <- here("data", "map")

# Region of Interest Bounding Box                   ======== edit this if needed
# Change the spatial extent
exnt <- c(xmin = -82,   # West
          xmax = -80,   # East
          ymin = 24.25, # South
          ymax = 25.75  # North 
        )

# inset extent (i.e. smaller map on top)            ======== edit this if needed
exnt2 <-  c(-90, # West
            -75, # East
            20,  # South
            35   # North
            ) 

names(exnt2) <- names(exnt)
```

# Load Combined Pigment Data
```{r load-pigment-data}
pig_dat  <- 
  here("data", "processed", "combined_pig_dat.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  select(
    -c(pi, 
       bottle, indicate_if_filters_are_replicates, volfilt,
       filter_type:filter_storage_before_shipment_to_gfsc,
       comments:other, collected_with_positive_pressure_or_vaccuum,
       date_extracted_month_day_year,
       cruise, vx_ml)
    )
```


# Download and Load Bathymetry and Coastline Files
**Note:** this will be done automatically if files **do not exists**

Topography, ETOPO1, 0.0166667 degrees, Global (longitude -180 to 180), (Ice Sheet Surface) from https://coastwatch.pfeg.noaa.gov/erddap/griddap/

Global Self-consistent, Hierarchical, High-resolution Geography Database (GSHHG) from https://www.ngdc.noaa.gov/mgg/shorelines/
```{r maps-download-load}
world_download(
    path_land = path_land,
    path_topo = path_topo,
    extent    = exnt
)

# Load coastline and topography data
map_obj <- 
    load_map_obj( 
    .map_coast = path_land, 
    .map_state = path_land,
    .map_bath  = path_topo, 
    .map_file  = "etopo1.nc",
    .extent    = exnt2
    )
```

# Plot FL Keys stations on map
```{r plt-fl-keys, fig.height=6, fig.width=10}
lat_lon_dat <- 
  pig_dat %>%
  filter(
      name_of_water_body == "Florida Keys"  
      & !str_detect(station, "60")) %>%
  select(station, lat, lon) %>% 
  distinct(station, lat, lon)

# ---- create base map
base_map <- with(map_obj, base_map_plot(coast_topo, bathy, exnt))

stn_map <-
  base_map +
  geom_point(
    data = lat_lon_dat,
    aes(
      x = lon,
      y = lat
      # shape = station,
    ),
    size        = 2,
    # alpha     = 0.5,
    stroke      = 1.5,
    show.legend = FALSE
  ) +

  labs(
    shape = "Station",
    fill  = "Station"
  ) +
  coord_sf(xlim = exnt[1:2], 
           ylim = exnt[3:4]) +
  scale_shape_manual(values = c(24, 25, 21))

stn_map
```

# Create Inset Map (i.e. Zoomed Out Map) 
```{r plt-large, fig.width=8, fig.height=10}
inset_map <-
  with(
    map_obj,
    map_inset(coast_topo, state,
      c(
        xmin = min(lat_lon_dat$lon), # West
        xmax = max(lat_lon_dat$lon), # East
        ymin = min(lat_lon_dat$lat), # South
        ymax = max(lat_lon_dat$lat) # North
      ),
      add_tol = 0.25
    )
  )

inset_map
```

# putting both of the maps together
May need to play with draw_plot(east.map, x, y, width and height)
```{r plt-lg-sm, fig.height=6, fig.width=10}
fullmap <- 
  ggdraw() +
  draw_plot(stn_map) +
  draw_plot(inset_map, x = 0.04, y = 0.57, width = 0.19, height = 0.4)
fullmap
```


```{r save-plt-map}
if (FALSE) {
  {    
  map_save_name <-
    here(
      "data", "plots",
      glue::glue(
        "map",
        format(Sys.time(), "_%Y%m%d_%H%M%S"),
        ".{c('jpeg', 'svg')}"
      )
    )


  save_plot(
    filename    = map_save_name[1],
    plot        = fullmap,
    base_height = 6,
    base_width  = 10,
    dpi         = 600,
    units       = "in",
    device      = "jpeg"
  )

  save_plot(
    filename    = map_save_name[2],
    plot        = fullmap,
    base_height = 6,
    base_width  = 10,
    dpi         = 600,
    units       = "in",
    device      = "svg"
  )
  
  rm(map_save_name)
  }
}

# clean up workspace
suppressWarnings(rm(stn_map, inset_map, lat_lon_dat, exnt, exnt2))
```

# ---- Kriging for All Pigments ----
Test to see if it works
```{r map-pigments}
source(here("scripts", "krig_fun.R"))

# filter for pigment and season
loc <- "Florida Keys"

pigm_dat <-
  pig_dat %>%
  pivot_longer(
    tot_chl_b:dp,
    names_to  = "conc_name",
    values_to = "conc"
  ) %>%
  filter(str_detect(name_of_water_body, loc)
         & !str_detect(station, "60")
         ) %>%
  select(
    hplc_gsfc_id, sample,                                     # <-- id
    station, lat, lon, name_of_water_body,                    # <-- spatial
    year, date_time_utc, season,                              # <-- temporal
    tot_chl_a, conc_name, conc                                # <-- measurements
  ) %>% 
  nest(.by = c(season, conc_name)) %>%
  
  # calc ratio of each pigment to total chlor-a
  mutate(
    data = pmap(
      list(season, conc_name, data),
      function(x, y, z) {
        z %>%
          mutate(ratio = conc / tot_chl_a) %>%
          # group_by(station) %>%
          summarise(
              .by = station,
            across(
              c(ratio, conc, tot_chl_a),
              # .fns = list(avg = (\(x) mean(x, na.rm = TRUE))),
              .fns   = (\(x) mean(x, na.rm = TRUE)),
              .names = "{.col}_avg"
            ),
            across(c(lat, lon), .fn = mean)
          ) %>%
          # ungroup() %>%
          mutate(
            season    = x,
            conc_name = y,
            .before   = 1
          )
       # ---- end of function
       }
    ),
    szn_name = case_when(
      str_detect(season, "(?i)sp") ~ "Spring",
      str_detect(season, "(?i)su") ~ "Summer",
      str_detect(season, "^(?i)f") ~ "Autumn",
      str_detect(season, "^(?i)w") ~ "Winter",
      .default = season
    ),
    title   = glue("{szn_name}: {conc_name}:TChla"),
    title2  = glue("{szn_name}: Average {conc_name}"),
    sv_name = glue("MBON_pigm_{conc_name}_{str_to_lower(szn_name)}")
  )
```

2 Samples will cause and error:
W	tot_chl_b
F	dv_chl_b

Not sure why these don't work, but is ignored when running. 

Here are offending samples: 
```{r}
pigm_dat %>%
    slice(c(67, 115)) %>%
    unnest(
        data, 
        names_sep = "_") 
```

```{r map-pigments, fig.width=30, fig.height=15}
pigm_dat1 <-
  pigm_dat %>%
    
  # remove two samples that cause issues, don't need anymore
  # slice(-c(67, 115)) %>%
  slice_sample(n = 10) %>%
 mutate(
    krig = pmap(
      list(data, conc_name, szn_name, title),
      .progress = TRUE,
      (\(x, y, z, title) map_data(
        .data     = x,
        conc_name = y,
        col_name  = "ratio_avg",
        # col_name  = "conc_avg",
        loc_name  = z,
        .title    = title,
        base_map  = base_map,
      ))
    )
  ) %>%
  hoist(krig, "dat", "map")

pigm_dat1 %>%
    slice_sample(n = 5) %$% map
pigm_dat1 %>%
    slice_sample(n = 5) %$% dat

if (FALSE) {
   # save krig maps
    pigm_dat1 %$%
    pwalk(list(map, sv_name),
          save_maps,
          sv = TRUE) 
}

```



# ---- ERROR: tot_chl_a doesn't exist in chemtax_with_metadata ----
```{r}
#Longer chemtax dataframe with combined functional types 
cmtx_longer3 <-
    fs::dir_ls(here(), 
               regexp  = "chemtax_with_metadata", 
               recurse = TRUE) %>% 
    read.csv() %>%
    
    mutate(diat  = diat1 + diat2,
           cyano = cyano2 + cyano4,
           hapt  = hapt6 + hapt8,
           .keep = "unused"
              ) #%>% 
    cmtx_longer3 
    pivot_longer(cols = chloro:hapt, 
                 names_to = "conc_name", 
                 values_to = "conc") %>%
    
    filter(station != 60) #%>%
  
  
      select(
       hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  conc_name, conc, tot_chl_a 
    )
```

# Don't Use: Old plotting method
```{r base-plot-fuco}
# if (FALSE) {
# png(file, height = 11, width = 4, units = "in", res = 300)
# par(
#     mfrow = c(3, 1),
#     mar = c(4.5, 4, 2, 1),
#     oma = c(4, 1, 4, 1)
# )
# 
# imagePlot(
#     ratio_kriged_fuco$x,
#     ratio_kriged_fuco$y,
#     ratio_kriged_fuco$z,
#     col = cols_fuco,
#     breaks = breaks_fuco,
#     asp = 1,
#     xlab = '',
#     ylab = '',
#     las = 1,
#     xlim = xlims_fuco,
#     ylim = ylims_fuco,
#     nlevel = length(cols_fuco),
#     legend.width = 1,
#     legend.mar = 3
# )
# 
# contour(
#     ratio_kriged_fuco$x,
#     ratio_kriged_fuco$y,
#     ratio_kriged_fuco$z,
#     levels = breaks_fuco,
#     add = TRUE
# )
# 
# image(
#     ratio_SE_fuco,
#     add = TRUE,
#     breaks = quantile(ratio_SE_fuco$z, c(1, 1), na.rm = TRUE),
#     col = 'white'
# )
# 
# image(
#     topo_lon,
#     topo_lat,
#     topo,
#     breaks = c(-1, 100),
#     col = 'white',
#     add = TRUE
# )
# 
# # plot(FL,col='gray70',add=TRUE)
# plot(world2, col = 'gray70', add = TRUE)
# contour(
#     topo_lon,
#     topo_lat,
#     topo,
#     add = TRUE,
#     levels = c(-100, -50, -25, -10),
#     col = 'gray40'
# )
# 
# points(df_fuco$lon,
#        df_fuco$lat,
#        pch = 20,
#        col = 'yellow')
# 
# mtext(expression(paste('Longitude (\u00B0W)')), 1, line = 3, cex = 0.75)
# 
# mtext(expression(paste('Latitude (\u00B0N)')), 2, line = 3, cex = 0.75)
# mtext('Fucoxanthin:TChla: Winter', adj = 1, cex = 0.75)
# }
```