---
title: "Template for Mapping"
author: "Sebastian DiGeronimo"
date: '2022-06-14'
output: html_document
---

## Creates map of site locations for IMaRS MBON
Uses topography and coastline data from NOAA.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("fields")
library("raster")
library("rerddap")
library("rgdal")
library("ncdf4")
library("magrittr")
library("ggplot2")
library("sf")
library("metR")
library("tidyr")
library("dplyr")
library("cowplot")
library("conflicted")
library("here")
library("stringr")
library("glue")

# specify select and filter 
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

# fixes issue with reading file
sf::sf_use_s2(FALSE)

# path for map to find/download to, they can be different depending on
# your computer
path_land <- here("data", "map")
path_topo <- here("data", "map")

# Region of Interest Bounding Box    ------------- edit this
# Change the spatial extent
exnt <- c(xmin = -82, # West
          xmax = -80, # East
          ymin = 24.25,# South
          ymax = 25.75   # North 
        )

# inset extent (i.e. smaller map on top of zoomed in map)
exnt2 <-  c(-90, # West
            -75, # East
            20,  # South
            35   # North
            ) 

names(exnt2) <- names(exnt)
```

# Load Combined Pigment Data
```{r load-pigment-data}
path_out <- here("data", "processed")
pig_dat  <- read.csv(here(path_out,"combined_pig_dat.csv")) %>%
    select(-c(pi, 
              bottle, indicate_if_filters_are_replicates, volfilt,
              filter_type:filter_storage_before_shipment_to_gfsc,
              comments:other, collected_with_positive_pressure_or_vaccuum,
              date_extracted_month_day_year,
              cruise, vx_ml))
```


# Download bathymetry and coastline files
**Note:** this will be done automatically if files **do not exists**

Topography, ETOPO1, 0.0166667 degrees, Global (longitude -180 to 180), (Ice Sheet Surface) from https://coastwatch.pfeg.noaa.gov/erddap/griddap/

Global Self-consistent, Hierarchical, High-resolution Geography Database (GSHHG) from https://www.ngdc.noaa.gov/mgg/shorelines/
```{r download-maps}
# source map download function
source(here("scripts", "map_files_dwnlod.R"))

world_download(
    # path_land = path_land,
    path_land = path_land,
    path_topo = path_topo,
    extent = exnt
)
```

# Load coastline and topography data
```{r load-shapefiles}
# select and read coastline file from GSHHS then crop
world2 <- fs::dir_ls(path = path_land,
               recurse = TRUE ,
               regexp = "GSHHS_h_L1.shp") %>%
               st_read(.) %>%
               st_crop(., st_bbox(exnt2))

# read state lines and crop
state <- fs::dir_ls(path = path_land,
               recurse = TRUE ,
               regexp = "WDBII_border_h_L2.shp") %>%
               st_read(.) %>%
               st_crop(., st_bbox(exnt2))


bathy <- nc_open(here(path_topo, "etopo1.nc"))
topo <- ncvar_get(bathy, 'altitude')
topo_lat <- ncvar_get(bathy, 'latitude')
topo_lon <- ncvar_get(bathy, 'longitude')
nc_close(bathy)

# read topography data
bathy <- raster::raster(here(path_topo,"etopo1.nc")) %>%
  as.data.frame(xy = T)

rm(path_topo, path_land)
```

# Plot FL Keys stations on map
```{r plt-fl-keys, fig.height=6, fig.width=10}
lat_lon_dat <- pig_dat %>%
    filter(name_of_water_body == "Florida Keys" & 
           !str_detect(station, "60")) %>%
    select(station, lat, lon) %>% 
    distinct(station, lat, lon)

stn.map <- 
    ggplot() +
    geom_contour2(data = bathy, 
                  aes(x = x, 
                      y = y, 
                      z = -Altitude), 
                  col = "grey40",
                  breaks = c(100, 50, 25, 10, 0)
                  ) +
    geom_sf(data = world2) +
    geom_point(data = lat_lon_dat,
               aes(
                   x = lon,
                   y = lat
                   # shape = station,
                   ),
              size = 2,
              # alpha = 0.5,
              stroke = 1.5,
              show.legend = FALSE) +
      
    # geom_text(data = lat_lon_dat, 
    #           aes(x = lon, y = lat, label = station),
    #           vjust = -0.5,
    #           hjust = -0.0001,
    #           size = 5) +
    
    scale_shape_manual(values = c(24,25,21)) +
    scale_x_continuous(expand = c(0,0)) + 
    scale_y_continuous(expand = c(0,0), 
                       breaks = seq(floor(exnt[3]), ceiling(exnt[4]), 0.5)
                       ) +
    coord_sf(xlim = exnt[1:2], ylim = exnt[3:4]) +
    labs(
      x = "",
      y = "",
      shape = "Station",
      fill = "Station") +
    theme_bw() +
    theme(text = element_text(family = "serif", size = 20),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()
          # legend.position = c(0.85, 0.25),
          # legend.title = element_text(size = 10), 
          # legend.text = element_text(size = 15),
          # legend.key.size = unit(0.6, "cm"),
          # legend.margin=margin(c(0.1,0.1,0.1,0.1), unit='cm')
          )

stn.map
# plotly::ggplotly(
#     ggplot() +
#   geom_point(data = pig_dat,
#              aes(
#                x = lon,
#                y = lat,
#                fill = station,
#                ##shape = station,
#              ),
#              size = 2,
#              alpha = 0.5,
#              stroke = 1.5,
#              show.legend = FALSE)
# )
```
# Zoomed out map 
```{r plt-large, fig.width=8, fig.height=10}
east.map <- 
      
    ggplot() +
    geom_sf(data = world2) +  
    geom_sf(data = state, color = "grey40", linewidth = 0.2) +
  
    # create red box on map to show sampling locations
    geom_rect(                  
        aes(
          xmin = min(lat_lon_dat$lon) - 0.25, # West
          xmax = max(lat_lon_dat$lon) + 0.25, # East
          ymin = min(lat_lon_dat$lat) - 0.25, # South
          ymax = max(lat_lon_dat$lat) + 0.25  # North
        ),
        color = "red",
        fill = NA
    ) +
    
    scale_x_continuous(expand = c(0,0)) + 
    scale_y_continuous(expand = c(0,0)) +
    coord_sf(xlim = exnt2[1:2], ylim = exnt2[3:4], 
            ) +
    
    theme_bw() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          plot.margin = grid::unit(c(0,0,0,0), "in"),
          panel.border = element_rect(linetype = "solid", 
                                      color = "black", 
                                      fill = NA)
        )

east.map
```

# putting both of the maps together
May need to play with draw_plot(east.map, x, y, width and height)
```{r plt-lg-sm, fig.height=6, fig.width=10}
ggdraw() +
  draw_plot(stn.map) +
  draw_plot(east.map, x = 0.04, y = 0.57, width = 0.1908108, height = 0.4)
```
```{r save-plt-map}
if (FALSE) {
    save_plot(
        filename = here("data", "plots",
                        glue::glue("map",
                                   format(Sys.time(), '_%Y%m%d_%H%M%S'),
                                   ".jpeg")), 
        plot = last_plot(),
        base_height = 6,
        base_width = 10,
        dpi = 600, 
        units = "in", 
        device = 'jpeg')
    
    save_plot(
        filename = here("data", "plots",
                        glue::glue("map",
                                   format(Sys.time(), '_%Y%m%d_%H%M%S'),
                                   ".svg")), 
        plot = last_plot(),
        base_height = 6,
        base_width = 10,
        dpi = 600, 
        units = "in", 
        device = 'svg')
}
```

# Kriging for Fucoxanthin
```{r filter-fuco-winter}
## testing out mapping one pigment during one season
pig <- "fuco"
fuco_data <-
    pig_dat %>%
    filter(
        # season == "Winter" 
        season == "W" 
        & name_of_water_body == "Florida Keys"
        ) %>%
    dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  tot_chl_a, fuco
        ) %>%
    mutate(fuco_ratio = fuco/tot_chl_a, 
           .after = fuco)

df_fuco <-
    fuco_data %>%
    group_by(station) %>%
    summarise(
        across(c(fuco_ratio, fuco, tot_chl_a), 
               .fn = list(avg = mean), na.rm = TRUE), 
        across(c(lat, lon), .fn = mean)) %>%
    ungroup()
```
```{r commented-out}
# sta_list_fuco <- unique(fuco_data$station)
# df_fuco <- matrix(ncol=6, nrow=length(sta_list_fuco))
# for (i in 1:length(sta_list_fuco)) {
#     sta_val <- fuco_data[fuco_data$lat_lon == sta_list_fuco[i], ]
#     chla_avg <- mean(sta_val$tot_chl_a)
#     fuco_avg <- mean(sta_val$fuco)
#     fuco_rat_avg <- mean(sta_val$fuco_ratio)
#     
#     df_fuco[i,] <- cbind(sta_list_fuco[i], sta_val$lon[1], sta_val$lat[1], chla_avg,
#                          fuco_avg, fuco_rat_avg)
# }
#     
# df_fuco <- as.data.frame(df_fuco)
# colnames(df_fuco) <- c("station", "longitude", "latitude", "chla", "fuco", "fuco_ratio") # change column names
# cols.num_fuco <- c("longitude", "latitude", "chla", "fuco", "fuco_ratio") # make these columns numeric vectors (double-check with "sapply(df, class)")
# df_fuco[cols.num_fuco] <- sapply(df_fuco[cols.num_fuco],as.numeric)

```
```{r kriging}
# -------------- kriging FUCO --------------
# kriging resolution
resolution <- 0.01
adj        <- 0.1

# krigging locations
df_fuco.loc   <- select(df_fuco, lat, lon)

loc.grid_fuco <- 
    list(lon = seq(min(df_fuco$lon, na.rm = T) - adj,
                   max(df_fuco$lon, na.rm = T) + adj, 
                   resolution),
         
         lat = seq(min(df_fuco$lat, na.rm = T) - adj, 
                   max(df_fuco$lat, na.rm = T) + adj, 
                   resolution))

# limits for plotting
xlims_fuco    <- range(loc.grid_fuco$lon)
ylims_fuco    <- range(loc.grid_fuco$lat)

# spatialProcess()
# ----------------- Pigment FUCO ratio krig -----------------
my.krig_fuco      <- 
    spatialProcess(as.matrix(df_fuco.loc), 
                                    df_fuco$fuco_ratio_avg)
my.krig_fuco      <- 
    spatialProcess((df_fuco.loc), 
                                    df_fuco$fuco_ratio_avg)

```
```{r label}

ratio_kriged_fuco <- predictSurface(my.krig_fuco, 
                                    loc.grid_fuco, 
                                    extrap = FALSE)

ratio_SE_fuco     <- predictSurfaceSE(my.krig_fuco, 
                                      loc.grid_fuco, 
                                      extrap = FALSE)

# corrects the kriged min and max to observed min and max
if (max(ratio_kriged_fuco$z, na.rm = TRUE) > max(df_fuco$fuco_rat_avg, na.rm = TRUE)) {
  ratio_kriged_fuco$z[which(ratio_kriged_fuco$z > max(df_fuco$fuco_rat_avg, na.rm = TRUE))] <- max(df_fuco$fuco_rat_avg, na.rm = TRUE)
}

if (min(ratio_kriged_fuco$z, na.rm = T) < min(df_fuco$fuco_rat_avg, na.rm = TRUE)) {
  ratio_kriged_fuco$z[which(ratio_kriged_fuco$z < min(df_fuco$fuco_rat_avg, na.rm = TRUE))] <- min(df_fuco$fuco_rat_avg, na.rm = TRUE)
}

### color and contour breaks FUCO
breaks_fuco <- pretty(df_fuco_$fuco_rat_avg, n = 10)
breaks_fuco <- seq(0, 0.6, by = 0.05) ### use for fixed breaks.

chl_col <-
    colorRampPalette(c("honeydew2", 
                       "darkseagreen3", 
                       "forestgreen", 
                       "darkslategrey"))

cols_fuco <- chl_col(length(breaks_fuco) - 1)

# krig file out
file <- here("data", "map", 
             glue("MBON_pigm_", pig, "_winter_",
                  format(Sys.time(), "%y%m%d_%H%M%S") ,
                  ".png"))

if (FALSE) png(file, height = 11, width = 4, units = "in", res = 300)

```
```{r}

### FUCO ###

par(mfrow=c(3,1),mar=c(4.5,4,2,1),oma=c(4,1,4,1))

imagePlot(ratio_kriged_fuco$x,
          ratio_kriged_fuco$y,
          ratio_kriged_fuco$z,
          col=cols_fuco,breaks=breaks_fuco,asp=1,
          xlab='',ylab='',las=1,
          xlim=xlims_fuco,ylim=ylims_fuco,
          nlevel=length(cols_fuco),legend.width=1,legend.mar=3)
contour(ratio_kriged_fuco$x,
        ratio_kriged_fuco$y,
        ratio_kriged_fuco$z,
        levels=breaks_fuco,add=T)
image(ratio_SE_fuco,add=T,breaks=quantile(ratio_SE_fuco$z,c(1,1),na.rm = TRUE),col='white')
image(topo_lon,topo_lat,topo,breaks=c(-1,100),col='white',add=T)
# plot(FL,col='gray70',add=T)
plot(world2,col='gray70',add=T)
contour(topo_lon,topo_lat,topo,add=T,levels=c(-100,-50,-25,-10),col='gray40')
points(df_fuco$lon,df_fuco$lat,pch=20,col='yellow')
mtext(expression(paste('Longitude (',degree,'W)')),1,line=3,cex=.75)
mtext(expression(paste('Latitude (',degree,'N)')),2,line=3,cex=.75)
mtext('Fucoxanthin:TChla: Winter',adj=1,cex=.75)
```

```{r function-for-mapping-pigments}
pig_dat_longer <- pig_dat %>% 
    pivot_longer(tot_chl_b:dp, names_to = "pigment", values_to = "conc")

map_pigment <- function(.pig, .long_file, .season) {
    pig_data <- .long_file %>%
        filter(season == .season, name_of_water_body == "Florida Keys",
               pigment == .pig
        ) %>%
        dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  pigment, conc, tot_chl_a
        ) %>%
        group_by(hplc_gsfc_id) %>% 
    mutate(ratio = conc/tot_chl_a, .after=conc)
    
    df_pigm <- pig_data %>% 
        group_by(station) %>% 
        summarize(ratio_avg = mean(ratio), conc_avg = mean(conc), chla_avg = mean(tot_chl_a), across()) %>% 
        distinct(station, .keep_all = T) %>% 
        ungroup()
    #krigging
    resolution <- .01
    adj <- .1
    ### krigging locations
    df_pigm.loc <- cbind(lon=df_pigm$lon,lat=df_pigm$lat)
    loc.grid_pigm <- list(lon=seq(min(df_pigm$lon, na.rm = TRUE) - adj, max(df_pigm$lon, na.rm = TRUE) + adj, resolution),
                 lat=seq(min(df_pigm$lat, na.rm = TRUE) - adj, max(df_pigm$lat, na.rm = TRUE) + adj, resolution))
    ### limits for plotting
    xlims_pigm <- range(loc.grid_pigm$lon)
    ylims_pigm <- range(loc.grid_pigm$lat)
    
    #pigment krig
    my.krig_pigm <- spatialProcess(df_pigm.loc, df_pigm$ratio_avg)
    ratio_kriged_pigm <- predictSurface(my.krig_pigm, loc.grid_pigm, extrap=F)
    ratio_SE_pigm <- predictSurfaceSE(my.krig_pigm, loc.grid_pigm, extrap=F)
    
    ### ----------------- Pigment pigm ratio krig -----------------

    ### corrects the krigged min and max to observed min and max
    if(max(ratio_kriged_pigm$z, na.rm = TRUE) > max(df_pigm$ratio_avg, na.rm = TRUE)){
        ratio_kriged_pigm$z[which(ratio_kriged_pigm$z > max(df_pigm$ratio_avg, na.rm = TRUE))] <- max(df_pigm$ratio_avg, na.rm = TRUE)
    }
    if(min(ratio_kriged_pigm$z, na.rm = TRUE) < min(df_pigm$ratio_avg, na.rm = TRUE)){
        ratio_kriged_pigm$z[which(ratio_kriged_pigm$z < min(df_pigm$ratio_avg, na.rm = TRUE))] <- min(df_pigm$ratio_avg, na.rm = TRUE)
    }
    ### color and contour breaks pigm
    breaks_pigm <- pretty(df_pigm$ratio_avg, n=10)
    breaks_pigm <- seq(0, 0.6, by=.05) ### use for fixed breaks.
    chl_col <- colorRampPalette(c('honeydew2','darkseagreen3','forestgreen','darkslategrey'))
    cols_pigm <- chl_col(length(breaks_pigm)-1)
    root <- rprojroot::find_rstudio_root_file()
    dir5 <- "//data//plots//"
    path_map = paste0(root,dir5)
    file <-
    paste0(pathmap,'MBON_pigm_',
           .pig,
           .season,
           format(Sys.time(), "%y%m%d_%H%M%S") ,
           '.png')
    png(file, height = 11, width = 4, units = 'in', res=300)
    
    par(mfrow=c(3,1),mar=c(4.5,4,2,1),oma=c(4,1,4,1))

    imagePlot(ratio_kriged_pigm$x,
              ratio_kriged_pigm$y,
              ratio_kriged_pigm$z,
              col=cols_pigm,breaks=breaks_pigm,asp=1,
              xlab='',ylab='',las=1,
              xlim=xlims_pigm,ylim=ylims_pigm,
              nlevel=length(cols_pigm),legend.width=1,legend.mar=3)
contour(ratio_kriged_pigm$x,
        ratio_kriged_pigm$y,
        ratio_kriged_pigm$z,
        levels=breaks_pigm,add=T)
image(ratio_SE_pigm,add=T,breaks=quantile(ratio_SE_pigm$z,c(1,1),na.rm = TRUE),col='white')
image(topo_lon,topo_lat,topo,breaks=c(-1,100),col='white',add=T)
plot(world2,col='gray70',add=T)
contour(topo_lon,topo_lat,topo,add=T,levels=c(-100,-50,-25,-10),col='gray40')
points(df_pigm$lon,df_pigm$lat,pch=20,col='yellow')
mtext(expression(paste('Longitude (',degree,'W)')),1,line=3,cex=.75)
mtext(expression(paste('Latitude (',degree,'N)')),2,line=3,cex=.75)
mtext(paste(.pig,':TChla:',.season),adj=1,cex=.75)

}


map_pigment("fuco", pig_dat_longer, "Winter")
map_pigment("fuco", pig_dat_longer, "Summer")
map_pigment("fuco", pig_dat_longer, "Spring")
map_pigment("fuco", pig_dat_longer, "Autumn")



```

```{r function-for-mapping}

map_phyto <- function(.phyto, .long_file, .season) {
    phyto_data <- .long_file %>%
        filter(season == .season, name_of_water_body == "Florida Keys",
               func_type == .phyto
        ) %>%
        dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  func_type, conc, tot_chl_a
        ) %>%
        group_by(hplc_gsfc_id) %>% 
    mutate(ratio = conc/tot_chl_a, .after=conc)
    
    df_phyto <- phyto_data %>% 
        group_by(station) %>% 
        summarize(ratio_avg = mean(ratio), conc_avg = mean(conc), chla_avg = mean(tot_chl_a), across()) %>% 
        distinct(station, .keep_all = T) %>% 
        ungroup()
    #krigging
    resolution <- .01
    adj <- .1
    ### krigging locations
    df_phyto.loc <- cbind(lon=df_phyto$lon,lat=df_phyto$lat)
    loc.grid_phyto <- list(lon=seq(min(df_phyto$lon, na.rm = TRUE) - adj, max(df_phyto$lon, na.rm = TRUE) + adj, resolution),
                 lat=seq(min(df_phyto$lat, na.rm = TRUE) - adj, max(df_phyto$lat, na.rm = TRUE) + adj, resolution))
    ### limits for plotting
    xlims_phyto <- range(loc.grid_phyto$lon)
    ylims_phyto <- range(loc.grid_phyto$lat)
    
    #phyto krig
    my.krig_phyto <- spatialProcess(df_phyto.loc, df_phyto$ratio_avg)
    ratio_kriged_phyto <- predictSurface(my.krig_phyto, loc.grid_phyto, extrap=F)
    ratio_SE_phyto <- predictSurfaceSE(my.krig_phyto, loc.grid_phyto, extrap=F)
    
    ### ----------------- phyto ratio krig -----------------

    ### corrects the krigged min and max to observed min and max
    if(max(ratio_kriged_phyto$z, na.rm = TRUE) > max(df_phyto$ratio_avg, na.rm = TRUE)){
        ratio_kriged_phyto$z[which(ratio_kriged_phyto$z > max(df_phyto$ratio_avg, na.rm = TRUE))] <- max(df_phyto$ratio_avg, na.rm = TRUE)
    }
    if(min(ratio_kriged_phyto$z, na.rm = TRUE) < min(df_phyto$ratio_avg, na.rm = TRUE)){
        ratio_kriged_phyto$z[which(ratio_kriged_phyto$z < min(df_phyto$ratio_avg, na.rm = TRUE))] <- min(df_phyto$ratio_avg, na.rm = TRUE)
    }
    ### color and contour breaks phyto
    breaks_phyto <- pretty(df_phyto$ratio_avg, n=10)
    breaks_phyto <- seq(0, 0.6, by=.05) ### use for fixed breaks.
    chl_col <- colorRampPalette(c('honeydew2','darkseagreen3','forestgreen','darkslategrey'))
    cols_phyto <- chl_col(length(breaks_phyto)-1)
    root <- rprojroot::find_rstudio_root_file()
    dir5 <- "//data//plots//"
    path_map <- paste0(root,dir5)
    file <-
    paste0(path_map,'MBON_pigments_',
           .phyto,
           .season,
           format(Sys.time(), "%y%m%d_%H%M%S") ,
           '.png')
    png(file, height = 11, width = 4, units = 'in', res=300)
    
    par(mfrow=c(3,1),mar=c(4.5,4,2,1),oma=c(4,1,4,1))

    imagePlot(ratio_kriged_phyto$x,
              ratio_kriged_phyto$y,
              ratio_kriged_phyto$z,
              col=cols_phyto,breaks=breaks_phyto,asp=1,
              xlab='',ylab='',las=1,
              xlim=xlims_phyto,ylim=ylims_phyto,
              nlevel=length(cols_phyto),legend.width=1,legend.mar=3)
    contour(ratio_kriged_phyto$x,
            ratio_kriged_phyto$y,
            ratio_kriged_phyto$z,
            levels=breaks_phyto,add=T)
    image(ratio_SE_phyto,add=T,breaks=quantile(ratio_SE_phyto$z,c(1,1),na.rm = TRUE),col='white')
    image(topo_lon,topo_lat,topo,breaks=c(-1,100),col='white',add=T)
    plot(world2,col='gray70',add=T)
    contour(topo_lon,topo_lat,topo,add=T,levels=c(-100,-50,-25,-10),col='gray40')
    points(df_phyto$lon,df_phyto$lat,pch=20,col='yellow')
    mtext(expression(paste('Longitude (',degree,'W)')),1,line=3,cex=.75)
    mtext(expression(paste('Latitude (',degree,'N)')),2,line=3,cex=.75)
    mtext(paste(.phyto,':TChla:',.season),adj=1,cex=.75)

}


map_phyto("cyano4", cmtx_longer2, "Winter")
map_phyto("fuco", cmtx_longer2, "Summer")
map_phyto("fuco", cmtx_longer2, "Spring")
map_phyto("fuco", cmtx_longer2, "Autumn")

```
```{r}
#Longer chemtax dataframe with combined functional types 
cmtx_longer3 <- cmtx_avgs %>% 
    mutate(diat = diat1 + diat2,
           cyano = cyano2 + cyano4,
           hapt = hapt6 + hapt8,
           .keep = "unused"
              ) %>% 
    pivot_longer(cols = chloro:hapt, names_to = "func_type", values_to = "conc") %>%
    filter(station != 60)

#function for mapping multiple seasons in one file
map_phyto <- function(.phyto, .long_file) {
    ### ----------------- SPRING -----------------
     phyto_spring_data <- .long_file %>%
        filter(season == "Spring", name_of_water_body == "Florida Keys",
               func_type == .phyto
        ) %>%
        dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  func_type, conc, tot_chl_a
        ) %>%
        group_by(hplc_gsfc_id) %>% 
    mutate(ratio = conc/tot_chl_a, .after=conc)
    
    df_phyto_spring <- phyto_spring_data %>% 
        group_by(station) %>% 
        summarize(ratio_avg = mean(ratio), conc_avg = mean(conc), chla_avg = mean(tot_chl_a), across()) %>% 
        distinct(station, .keep_all = T) %>% 
        ungroup()
    #krigging
    resolution <- .01
    adj <- .1
    ### krigging locations
    df_phyto_spring.loc <- cbind(lon=df_phyto_spring$lon,lat=df_phyto_spring$lat)
    loc.grid_phyto_spring <- list(lon=seq(min(df_phyto_spring$lon, na.rm = TRUE) - adj, max(df_phyto_spring$lon, na.rm = TRUE) + adj, resolution),
                 lat=seq(min(df_phyto_spring$lat, na.rm = TRUE) - adj, max(df_phyto_spring$lat, na.rm = TRUE) + adj, resolution))
    ### limits for plotting
    xlims_phyto_spring <- range(loc.grid_phyto_spring$lon)
    ylims_phyto_spring <- range(loc.grid_phyto_spring$lat)
    
    #phyto krig
    my.krig_phyto_spring <- spatialProcess(df_phyto_spring.loc, df_phyto_spring$ratio_avg)
    ratio_kriged_phyto_spring <- predictSurface(my.krig_phyto_spring, loc.grid_phyto_spring, extrap=F)
    ratio_SE_phyto_spring <- predictSurfaceSE(my.krig_phyto_spring, loc.grid_phyto_spring, extrap=F)
    
    ### ----------------- Spring Phyto ratio krig -----------------

    ### corrects the krigged min and max to observed min and max
    if(max(ratio_kriged_phyto_spring$z, na.rm = TRUE) > max(df_phyto_spring$ratio_avg, na.rm = TRUE)){
        ratio_kriged_phyto_spring$z[which(ratio_kriged_phyto_spring$z > max(df_phyto_spring$ratio_avg, na.rm = TRUE))] <- max(df_phyto_spring$ratio_avg, na.rm = TRUE)
    }
    if(min(ratio_kriged_phyto_spring$z, na.rm = TRUE) < min(df_phyto_spring$ratio_avg, na.rm = TRUE)){
        ratio_kriged_phyto_spring$z[which(ratio_kriged_phyto_spring$z < min(df_phyto_spring$ratio_avg, na.rm = TRUE))] <- min(df_phyto_spring$ratio_avg, na.rm = TRUE)
    }
    
    ### color and contour breaks phyto spring
    breaks_phyto_spring <- pretty(df_phyto_spring$ratio_avg, n=10)
    breaks_phyto_spring <- seq(0, 0.6, by=.05) ### use for fixed breaks.
    chl_col <- colorRampPalette(c('honeydew2','darkseagreen3','forestgreen','darkslategrey'))
    cols_phyto_spring <- chl_col(length(breaks_phyto_spring)-1)
    
    
    
    ### ----------------- SUMMER -----------------
     phyto_summer_data <- .long_file %>%
        filter(season == "Summer", name_of_water_body == "Florida Keys",
               func_type == .phyto
        ) %>%
        dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  func_type, conc, tot_chl_a
        ) %>%
        group_by(hplc_gsfc_id) %>% 
    mutate(ratio = conc/tot_chl_a, .after=conc)
    
    df_phyto_summer <- phyto_summer_data %>% 
        group_by(station) %>% 
        summarize(ratio_avg = mean(ratio), conc_avg = mean(conc), chla_avg = mean(tot_chl_a), across()) %>% 
        distinct(station, .keep_all = T) %>% 
        ungroup()
    ### krigging locations
    df_phyto_summer.loc <- cbind(lon=df_phyto_summer$lon,lat=df_phyto_summer$lat)
    loc.grid_phyto_summer <- list(lon=seq(min(df_phyto_summer$lon, na.rm = TRUE) - adj, max(df_phyto_summer$lon, na.rm = TRUE) + adj, resolution),
                 lat=seq(min(df_phyto_summer$lat, na.rm = TRUE) - adj, max(df_phyto_summer$lat, na.rm = TRUE) + adj, resolution))
    ### limits for plotting
    xlims_phyto_summer <- range(loc.grid_phyto_summer$lon)
    ylims_phyto_summer <- range(loc.grid_phyto_summer$lat)
    
    #phyto krig
    my.krig_phyto_summer <- spatialProcess(df_phyto_summer.loc, df_phyto_summer$ratio_avg)
    ratio_kriged_phyto_summer <- predictSurface(my.krig_phyto_summer, loc.grid_phyto_summer, extrap=F)
    ratio_SE_phyto_summer <- predictSurfaceSE(my.krig_phyto_summer, loc.grid_phyto_summer, extrap=F)
    
    ### ----------------- summer Phyto ratio krig -----------------

    ### corrects the krigged min and max to observed min and max
    if(max(ratio_kriged_phyto_summer$z, na.rm = TRUE) > max(df_phyto_summer$ratio_avg, na.rm = TRUE)){
        ratio_kriged_phyto_summer$z[which(ratio_kriged_phyto_summer$z > max(df_phyto_summer$ratio_avg, na.rm = TRUE))] <- max(df_phyto_summer$ratio_avg, na.rm = TRUE)
    }
    if(min(ratio_kriged_phyto_summer$z, na.rm = TRUE) < min(df_phyto_summer$ratio_avg, na.rm = TRUE)){
        ratio_kriged_phyto_summer$z[which(ratio_kriged_phyto_summer$z < min(df_phyto_summer$ratio_avg, na.rm = TRUE))] <- min(df_phyto_summer$ratio_avg, na.rm = TRUE)
    }
    
    ### color and contour breaks phyto summer
    breaks_phyto_summer <- pretty(df_phyto_summer$ratio_avg, n=10)
    breaks_phyto_summer <- seq(0, 0.6, by=.05) ### use for fixed breaks.
    cols_phyto_summer <- chl_col(length(breaks_phyto_summer)-1)
    
    
    ###----------------- FALL -----------------------
     phyto_autumn_data <- .long_file %>%
        filter(season == "Autumn", name_of_water_body == "Florida Keys",
               func_type == .phyto
        ) %>%
        dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  func_type, conc, tot_chl_a
        ) %>%
        group_by(hplc_gsfc_id) %>% 
    mutate(ratio = conc/tot_chl_a, .after=conc)
    
    df_phyto_autumn <- phyto_autumn_data %>% 
        group_by(station) %>% 
        summarize(ratio_avg = mean(ratio), conc_avg = mean(conc), chla_avg = mean(tot_chl_a), across()) %>% 
        distinct(station, .keep_all = T) %>% 
        ungroup()
    ### krigging locations
    df_phyto_autumn.loc <- cbind(lon=df_phyto_autumn$lon,lat=df_phyto_autumn$lat)
    loc.grid_phyto_autumn <- list(lon=seq(min(df_phyto_autumn$lon, na.rm = TRUE) - adj, max(df_phyto_autumn$lon, na.rm = TRUE) + adj, resolution),
                 lat=seq(min(df_phyto_autumn$lat, na.rm = TRUE) - adj, max(df_phyto_autumn$lat, na.rm = TRUE) + adj, resolution))
    ### limits for plotting
    xlims_phyto_autumn <- range(loc.grid_phyto_autumn$lon)
    ylims_phyto_autumn <- range(loc.grid_phyto_autumn$lat)
    
    #phyto krig
    my.krig_phyto_autumn <- spatialProcess(df_phyto_autumn.loc, df_phyto_autumn$ratio_avg)
    ratio_kriged_phyto_autumn <- predictSurface(my.krig_phyto_autumn, loc.grid_phyto_autumn, extrap=F)
    ratio_SE_phyto_autumn <- predictSurfaceSE(my.krig_phyto_autumn, loc.grid_phyto_autumn, extrap=F)
    
    ### ----------------- autumn Phyto ratio krig -----------------

    ### corrects the krigged min and max to observed min and max
    if(max(ratio_kriged_phyto_autumn$z, na.rm = TRUE) > max(df_phyto_autumn$ratio_avg, na.rm = TRUE)){
        ratio_kriged_phyto_autumn$z[which(ratio_kriged_phyto_autumn$z > max(df_phyto_autumn$ratio_avg, na.rm = TRUE))] <- max(df_phyto_autumn$ratio_avg, na.rm = TRUE)
    }
    if(min(ratio_kriged_phyto_autumn$z, na.rm = TRUE) < min(df_phyto_autumn$ratio_avg, na.rm = TRUE)){
        ratio_kriged_phyto_autumn$z[which(ratio_kriged_phyto_autumn$z < min(df_phyto_autumn$ratio_avg, na.rm = TRUE))] <- min(df_phyto_autumn$ratio_avg, na.rm = TRUE)
    }
    
    ### color and contour breaks phyto autumn
    breaks_phyto_autumn <- pretty(df_phyto_autumn$ratio_avg, n=10)
    breaks_phyto_autumn <- seq(0, 0.6, by=.05) ### use for fixed breaks.
    cols_phyto_autumn <- chl_col(length(breaks_phyto_autumn)-1)
    
    
    ###----------------- WINTER -----------------------
    phyto_winter_data <- .long_file %>%
        filter(season == "Winter", name_of_water_body == "Florida Keys",
               func_type == .phyto
        ) %>%
        dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  func_type, conc, tot_chl_a
        ) %>%
        group_by(hplc_gsfc_id) %>% 
    mutate(ratio = conc/tot_chl_a, .after=conc)
    
    df_phyto_winter <- phyto_winter_data %>% 
        group_by(station) %>% 
        summarize(ratio_avg = mean(ratio), conc_avg = mean(conc), chla_avg = mean(tot_chl_a), across()) %>% 
        distinct(station, .keep_all = T) %>% 
        ungroup()
    ### krigging locations
    df_phyto_winter.loc <- cbind(lon=df_phyto_winter$lon,lat=df_phyto_winter$lat)
    loc.grid_phyto_winter <- list(lon=seq(min(df_phyto_winter$lon, na.rm = TRUE) - adj, max(df_phyto_winter$lon, na.rm = TRUE) + adj, resolution),
                 lat=seq(min(df_phyto_winter$lat, na.rm = TRUE) - adj, max(df_phyto_winter$lat, na.rm = TRUE) + adj, resolution))
    ### limits for plotting
    xlims_phyto_winter <- range(loc.grid_phyto_winter$lon)
    ylims_phyto_winter <- range(loc.grid_phyto_winter$lat)
    
    #phyto krig
    my.krig_phyto_winter <- spatialProcess(df_phyto_winter.loc, df_phyto_winter$ratio_avg)
    ratio_kriged_phyto_winter <- predictSurface(my.krig_phyto_winter, loc.grid_phyto_winter, extrap=F)
    ratio_SE_phyto_winter <- predictSurfaceSE(my.krig_phyto_winter, loc.grid_phyto_winter, extrap=F)
    
    ### ----------------- winter Phyto ratio krig -----------------

    ### corrects the krigged min and max to observed min and max
    if(max(ratio_kriged_phyto_winter$z, na.rm = TRUE) > max(df_phyto_winter$ratio_avg, na.rm = TRUE)){
        ratio_kriged_phyto_winter$z[which(ratio_kriged_phyto_winter$z > max(df_phyto_winter$ratio_avg, na.rm = TRUE))] <- max(df_phyto_winter$ratio_avg, na.rm = TRUE)
    }
    if(min(ratio_kriged_phyto_winter$z, na.rm = TRUE) < min(df_phyto_winter$ratio_avg, na.rm = TRUE)){
        ratio_kriged_phyto_winter$z[which(ratio_kriged_phyto_winter$z < min(df_phyto_winter$ratio_avg, na.rm = TRUE))] <- min(df_phyto_winter$ratio_avg, na.rm = TRUE)
    }
    
    ### color and contour breaks phyto winter
    breaks_phyto_winter <- pretty(df_phyto_winter$ratio_avg, n=10)
    breaks_phyto_winter <- seq(0, 0.6, by=.05) ### use for fixed breaks.
    cols_phyto_winter <- chl_col(length(breaks_phyto_winter)-1)
    
    
    root <- rprojroot::find_rstudio_root_file()
    dir5 <- "//data//plots//"
    path_map <- paste0(root,dir5)
    file <-
    paste0(path_map,'MBON_pigments_',
           .phyto,
           format(Sys.time(), "%y%m%d_%H%M%S") ,
           '.png')
    png(file, height = 11, width = 4, units = 'in', res=300)
    
    par(mfrow=c(4,1),mar=c(4.5,4,2,1),oma=c(4,1,4,1))

    ###SPRING
    imagePlot(ratio_kriged_phyto_spring$x,
              ratio_kriged_phyto_spring$y,
              ratio_kriged_phyto_spring$z,
              col=cols_phyto_spring,breaks=breaks_phyto_spring,asp=1,
              xlab='',ylab='',las=1,
              xlim=xlims_phyto_spring,ylim=ylims_phyto_spring,
              nlevel=length(cols_phyto_spring),legend.width=1,legend.mar=3)
    contour(ratio_kriged_phyto_spring$x,
            ratio_kriged_phyto_spring$y,
            ratio_kriged_phyto_spring$z,
            levels=breaks_phyto_spring,add=T)
    image(ratio_SE_phyto_spring,add=T,breaks=quantile(ratio_SE_phyto_spring$z,c(1,1),na.rm = TRUE),col='white')
    image(topo_lon,topo_lat,topo,breaks=c(-1,100),col='white',add=T)
    plot(world2,col='gray70',add=T)
    contour(topo_lon,topo_lat,topo,add=T,levels=c(-100,-50,-25,-10),col='gray40')
    points(df_phyto_spring$lon,df_phyto_spring$lat,pch=20,col='yellow')
    mtext(expression(paste('Longitude (',degree,'W)')),1,line=3,cex=.75)
    mtext(expression(paste('Latitude (',degree,'N)')),2,line=3,cex=.75)
    mtext(paste(.phyto,':TChla:Spring'),adj=1,cex=.75)
    
    ###SUMMER
    imagePlot(ratio_kriged_phyto_summer$x,
              ratio_kriged_phyto_summer$y,
              ratio_kriged_phyto_summer$z,
              col=cols_phyto_summer,breaks=breaks_phyto_summer,asp=1,
              xlab='',ylab='',las=1,
              xlim=xlims_phyto_summer,ylim=ylims_phyto_summer,
              nlevel=length(cols_phyto_summer),legend.width=1,legend.mar=3)
    contour(ratio_kriged_phyto_summer$x,
            ratio_kriged_phyto_summer$y,
            ratio_kriged_phyto_summer$z,
            levels=breaks_phyto_summer,add=T)
    image(ratio_SE_phyto_summer,add=T,breaks=quantile(ratio_SE_phyto_summer$z,c(1,1),na.rm = TRUE),col='white')
    image(topo_lon,topo_lat,topo,breaks=c(-1,100),col='white',add=T)
    plot(world2,col='gray70',add=T)
    contour(topo_lon,topo_lat,topo,add=T,levels=c(-100,-50,-25,-10),col='gray40')
    points(df_phyto_summer$lon,df_phyto_summer$lat,pch=20,col='yellow')
    mtext(expression(paste('Longitude (',degree,'W)')),1,line=3,cex=.75)
    mtext(expression(paste('Latitude (',degree,'N)')),2,line=3,cex=.75)
    mtext(paste(.phyto,':TChla:Summer'),adj=1,cex=.75)
    
    ###FALL
    imagePlot(ratio_kriged_phyto_autumn$x,
              ratio_kriged_phyto_autumn$y,
              ratio_kriged_phyto_autumn$z,
              col=cols_phyto_autumn,breaks=breaks_phyto_autumn,asp=1,
              xlab='',ylab='',las=1,
              xlim=xlims_phyto_autumn,ylim=ylims_phyto_autumn,
              nlevel=length(cols_phyto_autumn),legend.width=1,legend.mar=3)
    contour(ratio_kriged_phyto_autumn$x,
            ratio_kriged_phyto_autumn$y,
            ratio_kriged_phyto_autumn$z,
            levels=breaks_phyto_autumn,add=T)
    image(ratio_SE_phyto_autumn,add=T,breaks=quantile(ratio_SE_phyto_autumn$z,c(1,1),na.rm = TRUE),col='white')
    image(topo_lon,topo_lat,topo,breaks=c(-1,100),col='white',add=T)
    plot(world2,col='gray70',add=T)
    contour(topo_lon,topo_lat,topo,add=T,levels=c(-100,-50,-25,-10),col='gray40')
    points(df_phyto_autumn$lon,df_phyto_autumn$lat,pch=20,col='yellow')
    mtext(expression(paste('Longitude (',degree,'W)')),1,line=3,cex=.75)
    mtext(expression(paste('Latitude (',degree,'N)')),2,line=3,cex=.75)
    mtext(paste(.phyto,':TChla:Autumn'),adj=1,cex=.75)
    
    ###WINTER
    imagePlot(ratio_kriged_phyto_winter$x,
              ratio_kriged_phyto_winter$y,
              ratio_kriged_phyto_winter$z,
              col=cols_phyto_winter,breaks=breaks_phyto_winter,asp=1,
              xlab='',ylab='',las=1,
              xlim=xlims_phyto_winter,ylim=ylims_phyto_winter,
              nlevel=length(cols_phyto_winter),legend.width=1,legend.mar=3)
    contour(ratio_kriged_phyto_winter$x,
            ratio_kriged_phyto_winter$y,
            ratio_kriged_phyto_winter$z,
            levels=breaks_phyto_winter,add=T)
    image(ratio_SE_phyto_winter,add=T,breaks=quantile(ratio_SE_phyto_winter$z,c(1,1),na.rm = TRUE),col='white')
    image(topo_lon,topo_lat,topo,breaks=c(-1,100),col='white',add=T)
    plot(world2,col='gray70',add=T)
    contour(topo_lon,topo_lat,topo,add=T,levels=c(-100,-50,-25,-10),col='gray40')
    points(df_phyto_winter$lon,df_phyto_winter$lat,pch=20,col='yellow')
    mtext(expression(paste('Longitude (',degree,'W)')),1,line=3,cex=.75)
    mtext(expression(paste('Latitude (',degree,'N)')),2,line=3,cex=.75)
    mtext(paste(.phyto,':TChla:Winter'),adj=1,cex=.75)

    dev.off()

}

map_phyto("chloro", cmtx_longer3)

map_phyto("cyano", cmtx_longer3)

map_phyto("crypto", cmtx_longer3)

map_phyto("diat", cmtx_longer3)

map_phyto("dino", cmtx_longer3)

map_phyto("hapt", cmtx_longer3)

map_phyto("pras", cmtx_longer3)


```
