---
title: "Template for Mapping"
author: "Sebastian DiGeronimo"
date: '2022-06-14'
output: html_document
---

## Creates map of site locations for Natalia Lopez Figueroa Zooplankton Paper
Uses topography and coastline data from NOAA.
Dataframe of station location from Natalia, programmed here.

```{r setup, include=FALSE}
root <- rprojroot::find_rstudio_root_file()
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = root)
library("fields")
library("raster")
library("rerddap")
library("rgdal")
library("ncdf4")
library("magrittr")
library("ggplot2")
library("sf")
library("metR")
library("tidyr")
library("dplyr")
library("cowplot")

# fixes issue with reading file?
sf::sf_use_s2(FALSE)

root <- rprojroot::find_rstudio_root_file()

path <- paste0(root,"/data/map/")

# Region of Interest Bounding Box    ------------- edit this
# Change the spatial extent
exnt <- c(xmin = -82, # West
          xmax = -80, # East
          ymin = 24.25,# South
          ymax = 25.75   # North 
        )

# inset extent (i.e. smaller map on top of zoomed in map)
exnt2 <-  c(-90, # West
            -75, # East
            20,  # South
            35   # North
            ) 
names(exnt2) <- names(exnt)
```

Load some dataset here!!!
```{r}
dir2 <- "//data//processed//"
path_out = paste0(root,dir2)
pig_dat <- read.csv(paste0(path_out,"combined_pig_dat.csv"))

```


# download bathymetry and coastline files
**Note:** this will be done automatically if files **do not exists**
```{r download-maps}
fs::dir_create(path)

# Topography, ETOPO1, 0.0166667 degrees, Global (longitude -180 to 180), (Ice Sheet Surface)
# from https://coastwatch.pfeg.noaa.gov/erddap/griddap/
# download if not already 
if (!file.exists(paste0(path,"etopo1.nc"))){
  message("Getting Topography Data")
  # ERDDAP method
  etopo    <- info('etopo180')

  # ERDDAP extract and save
  griddap(etopo, latitude=exnt[3:4], longitude=exnt[1:2], stride=c(1,1), fields='altitude',
          store = disk(paste0(root,"/data/map/")))
  
  # Rename file 
  file.rename(fs::dir_ls(path, regexp = "\\.nc$"),
              paste0(root,"/data/map/etopo1.nc"))
  
  rm(etopo)
}

# Global Self-consistent, Hierarchical, High-resolution Geography Database (GSHHG)
# from https://www.ngdc.noaa.gov/mgg/shorelines/
# download GSHHS shapefile if not already downloaded
coast         <- fs::dir_ls(path = paste0(path),
               recurse = TRUE ,
               regexp = "GSHHS_h_L1.shp")

if (rlang::is_empty(coast)){
  message("Getting Coastline Data")
  temp        <- tempfile()
      download.file(
          "ftp://ftp.soest.hawaii.edu/gshhg/gshhg-shp-2.3.7.zip",
          temp,
          method = "libcurl",
          mode = "wb"
      )
      unzip(exdir = paste0(path), temp)
      unlink(temp)
      rm(temp)
}

rm(coast)
```

# ggplot way of mapping
```{r load-map}
# select and read coastline file from GSHHS then crop
world2 <- fs::dir_ls(path = paste0(path),
               recurse = TRUE ,
               regexp = "GSHHS_h_L1.shp") %>%
               st_read(.) %>%
               st_crop(., st_bbox(exnt2))

# read state lines and crop
state <- fs::dir_ls(path = paste0(path),
               recurse = TRUE ,
               regexp = "WDBII_border_h_L2.shp") %>%
               st_read(.) %>%
               st_crop(., st_bbox(exnt2))


bathy <- nc_open(paste0(root,"/data/map/etopo1.nc"))
topo <- ncvar_get(bathy, 'altitude')
topo_lat <- ncvar_get(bathy, 'latitude')
topo_lon <- ncvar_get(bathy, 'longitude')
nc_close(bathy)

# read bahtymetry data
bathy <- raster::raster(paste0(path,"etopo1.nc")) %>%
  as.data.frame(xy = T)


```

```{r plots, fig.height=6, fig.width=10}
lat_lon_dat <- pig_dat %>%
    filter(name_of_water_body == "Florida Keys", station != "60") %>%
    dplyr::select(station, lat, lon) %>% 
    distinct(station, lat, lon)

stn.map <- ggplot() +
  geom_contour2(data = bathy, 
                aes(x = x, y = y, z = -Altitude), 
                col = "grey40",
                breaks = c(100, 40, 25, 10, 0)
                ) +
  geom_sf(data = world2) +
  # geom_point(data = pig_dat,
    geom_point(data = lat_lon_dat,
             aes(
               x = lon,
               y = lat
               ##shape = station,
             ),
             size = 2,
             # alpha = 0.5,
             stroke = 1.5,
             show.legend = FALSE) +
    # geom_text(data = lat_lon_dat, 
    #           aes(x = lon, y = lat, label = station),
    #           vjust = -0.5,
    #           hjust = -0.0001,
    #           size = 5) +
  scale_shape_manual(values = c(24,25,21))+
  scale_x_continuous(expand = c(0,0)) + 
  scale_y_continuous(expand = c(0,0), breaks = seq(30.5,32,0.5)) +
  coord_sf(xlim = exnt[1:2], ylim = exnt[3:4]) +
  labs(
    x = "",
    y = "",
    shape = "Station",
    fill = "Station") +
  theme_bw() +
  theme(text=element_text(family="serif", size=20),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()
        # legend.position = c(0.85, 0.25),
        # legend.title = element_text(size = 10), 
        # legend.text = element_text(size = 15),
        # legend.key.size = unit(0.6, "cm"),
        # legend.margin=margin(c(0.1,0.1,0.1,0.1), unit='cm')
        )

stn.map
# plotly::ggplotly(
#     ggplot() +
#   geom_point(data = pig_dat,
#              aes(
#                x = lon,
#                y = lat,
#                fill = station,
#                ##shape = station,
#              ),
#              size = 2,
#              alpha = 0.5,
#              stroke = 1.5,
#              show.legend = FALSE)
# )
```
# Zoomed out map 
```{r us-map-plot, fig.width=2.862162, fig.height=6}
east.map <- ggplot() +
  geom_sf(data = world2) +  
  geom_sf(data = state) +
  geom_rect(                  # create red box on map to show sampling locations
    aes(
      xmin = min(lat_lon_dat$lon) - 0.25, # West
      # xmax = max(pig_dat$lon) + 0.25, # East
      xmax = -79.83275, # East
      ymin = min(lat_lon_dat$lat) - 0.25, # South
      # ymax = max(pig_dat$lat) + 0.25  # North
      ymax = 28.25 # North, incorrect coordinate was messing up box size
    ),
    color = "red",
    fill = NA
  ) +
  scale_x_continuous(expand = c(0,0)) + 
  scale_y_continuous(expand = c(0,0)) +
  coord_sf(xlim = exnt2[1:2], ylim = exnt2[3:4], 
           ) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        plot.margin=grid::unit(c(0,0,0,0), "in"),
        panel.border = element_rect(linetype = "solid", color = "black", fill = NA)
        )

east.map
```

# putting both of the maps together
May need to play with draw_plot(east.map, x, y, width and height)
```{r fig.height=6, fig.width=10}
library("cowplot")

ggdraw() +
  draw_plot(stn.map) +
  draw_plot(east.map, x = 0.04, y = 0.57, width = 0.1908108, height = 0.4)



save_plot(filename = paste0(path,"../plots/map", 
           format(Sys.time(), '_%Y%m%d_%H%M%S'),
           ".jpg"), 
          plot = last_plot(),
          base_height = 6,
          base_width = 10,
          dpi = 600, units = "in", device='png')

save_plot(filename = paste0(path,"../plots/map", 
           format(Sys.time(), '_%Y%m%d_%H%M%S'),
           ".svg"), 
          plot = last_plot(),
          base_height = 6,
          base_width = 10,
          dpi = 600, units = "in", device='svg')
```

```{r fuco-mapping-winter}
## testing out mapping one pigment during one season
pig <- "fuco"
fuco_data <- pig_dat %>%
    filter(
        season == "Winter", name_of_water_body == "Florida Keys"
        ) %>%
    dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  tot_chl_a, fuco
        ) %>%
    mutate(fuco_ratio = fuco/tot_chl_a, .after=fuco)

df_fuco <- fuco_data %>%
    group_by(station) %>%
    summarize(fuco_rat_avg = mean(fuco_ratio), fuco_avg = mean(fuco), chla_avg = mean(tot_chl_a), across()) %>%
    distinct(station, .keep_all = T) %>% 
    ungroup()
# ----    
# sta_list_fuco <- unique(fuco_data$station)
# df_fuco <- matrix(ncol=6, nrow=length(sta_list_fuco))
# for (i in 1:length(sta_list_fuco)) {
#     sta_val <- fuco_data[fuco_data$lat_lon == sta_list_fuco[i], ]
#     chla_avg <- mean(sta_val$tot_chl_a)
#     fuco_avg <- mean(sta_val$fuco)
#     fuco_rat_avg <- mean(sta_val$fuco_ratio)
#     
#     df_fuco[i,] <- cbind(sta_list_fuco[i], sta_val$lon[1], sta_val$lat[1], chla_avg,
#                          fuco_avg, fuco_rat_avg)
# }
#     
# df_fuco <- as.data.frame(df_fuco)
# colnames(df_fuco) <- c("station", "longitude", "latitude", "chla", "fuco", "fuco_ratio") # change column names
# cols.num_fuco <- c("longitude", "latitude", "chla", "fuco", "fuco_ratio") # make these columns numeric vectors (double-check with "sapply(df, class)")
# df_fuco[cols.num_fuco] <- sapply(df_fuco[cols.num_fuco],as.numeric)

#### -------------- kriging FUCO --------------
### krigging resolution
resolution <- .01
adj <- .1
### krigging locations
df_fuco.loc <- cbind(lon=df_fuco$lon,lat=df_fuco$lat)
loc.grid_fuco <- list(lon=seq(min(df_fuco$lon, na.rm=T) - adj, max(df_fuco$lon, na.rm=T) + adj, resolution),
                 lat=seq(min(df_fuco$lat, na.rm=T) - adj, max(df_fuco$lat, na.rm=T) + adj, resolution))
### limits for plotting
xlims_fuco <- range(loc.grid_fuco$lon)
ylims_fuco <- range(loc.grid_fuco$lat)

#spatialProcess()
### ----------------- Pigment FUCO ratio krig -----------------
my.krig_fuco <- spatialProcess(df_fuco.loc, df_fuco$fuco_rat_avg)
ratio_kriged_fuco <- predictSurface(my.krig_fuco, loc.grid_fuco, extrap=F)
ratio_SE_fuco <- predictSurfaceSE(my.krig_fuco, loc.grid_fuco, extrap=F)

### corrects the krigged min and max to observed min and max
if(max(ratio_kriged_fuco$z, na.rm=T) > max(df_fuco$fuco_rat_avg, na.rm=T)){
  ratio_kriged_fuco$z[which(ratio_kriged_fuco$z > max(df_fuco$fuco_rat_avg, na.rm=T))] <- max(df_fuco$fuco_rat_avg, na.rm=T)
}
if(min(ratio_kriged_fuco$z, na.rm=T) < min(df_fuco$fuco_rat_avg, na.rm=T)){
  ratio_kriged_fuco$z[which(ratio_kriged_fuco$z < min(df_fuco$fuco_rat_avg, na.rm=T))] <- min(df_fuco$fuco_rat_avg, na.rm=T)
}

### color and contour breaks FUCO
breaks_fuco <- pretty(df_fuco_$fuco_rat_avg, n=10)
breaks_fuco <- seq(0, 0.6, by=.05) ### use for fixed breaks.
chl_col <- colorRampPalette(c('honeydew2','darkseagreen3','forestgreen','darkslategrey'))
cols_fuco <- chl_col(length(breaks_fuco)-1)


dir4 <- "//data//map//"
path_map = paste0(root,dir4)
setwd(path_map)


file <-
    paste0('MBON_pigm_',
           pig,
           '_winter_',
           format(Sys.time(), "%y%m%d_%H%M%S") ,
           '.png')

png(file, height = 11, width = 4, units = 'in', res=300)

### FUCO ###

par(mfrow=c(3,1),mar=c(4.5,4,2,1),oma=c(4,1,4,1))

imagePlot(ratio_kriged_fuco$x,
          ratio_kriged_fuco$y,
          ratio_kriged_fuco$z,
          col=cols_fuco,breaks=breaks_fuco,asp=1,
          xlab='',ylab='',las=1,
          xlim=xlims_fuco,ylim=ylims_fuco,
          nlevel=length(cols_fuco),legend.width=1,legend.mar=3)
contour(ratio_kriged_fuco$x,
        ratio_kriged_fuco$y,
        ratio_kriged_fuco$z,
        levels=breaks_fuco,add=T)
image(ratio_SE_fuco,add=T,breaks=quantile(ratio_SE_fuco$z,c(1,1),na.rm=T),col='white')
image(topo_lon,topo_lat,topo,breaks=c(-1,100),col='white',add=T)
# plot(FL,col='gray70',add=T)
plot(world2,col='gray70',add=T)
contour(topo_lon,topo_lat,topo,add=T,levels=c(-100,-50,-25,-10),col='gray40')
points(df_fuco$lon,df_fuco$lat,pch=20,col='yellow')
mtext(expression(paste('Longitude (',degree,'W)')),1,line=3,cex=.75)
mtext(expression(paste('Latitude (',degree,'N)')),2,line=3,cex=.75)
mtext('Fucoxanthin:TChla: Winter',adj=1,cex=.75)
```

```{r function-for-mapping-pigments}
pig_dat_longer <- pig_dat %>% 
    pivot_longer(tot_chl_b:dp, names_to = "pigment", values_to = "conc")

map_pigment <- function(.pig, .long_file, .season) {
    pig_data <- .long_file %>%
        filter(season == .season, name_of_water_body == "Florida Keys",
               pigment == .pig
        ) %>%
        dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  pigment, conc, tot_chl_a
        ) %>%
        group_by(hplc_gsfc_id) %>% 
    mutate(ratio = conc/tot_chl_a, .after=conc)
    
    df_pigm <- pig_data %>% 
        group_by(station) %>% 
        summarize(ratio_avg = mean(ratio), conc_avg = mean(conc), chla_avg = mean(tot_chl_a), across()) %>% 
        distinct(station, .keep_all = T) %>% 
        ungroup()
    #krigging
    resolution <- .01
    adj <- .1
    ### krigging locations
    df_pigm.loc <- cbind(lon=df_pigm$lon,lat=df_pigm$lat)
    loc.grid_pigm <- list(lon=seq(min(df_pigm$lon, na.rm=T) - adj, max(df_pigm$lon, na.rm=T) + adj, resolution),
                 lat=seq(min(df_pigm$lat, na.rm=T) - adj, max(df_pigm$lat, na.rm=T) + adj, resolution))
    ### limits for plotting
    xlims_pigm <- range(loc.grid_pigm$lon)
    ylims_pigm <- range(loc.grid_pigm$lat)
    
    #pigment krig
    my.krig_pigm <- spatialProcess(df_pigm.loc, df_pigm$ratio_avg)
    ratio_kriged_pigm <- predictSurface(my.krig_pigm, loc.grid_pigm, extrap=F)
    ratio_SE_pigm <- predictSurfaceSE(my.krig_pigm, loc.grid_pigm, extrap=F)
    
    ### ----------------- Pigment pigm ratio krig -----------------

    ### corrects the krigged min and max to observed min and max
    if(max(ratio_kriged_pigm$z, na.rm=T) > max(df_pigm$ratio_avg, na.rm=T)){
        ratio_kriged_pigm$z[which(ratio_kriged_pigm$z > max(df_pigm$ratio_avg, na.rm=T))] <- max(df_pigm$ratio_avg, na.rm=T)
    }
    if(min(ratio_kriged_pigm$z, na.rm=T) < min(df_pigm$ratio_avg, na.rm=T)){
        ratio_kriged_pigm$z[which(ratio_kriged_pigm$z < min(df_pigm$ratio_avg, na.rm=T))] <- min(df_pigm$ratio_avg, na.rm=T)
    }
    ### color and contour breaks pigm
    breaks_pigm <- pretty(df_pigm$ratio_avg, n=10)
    breaks_pigm <- seq(0, 0.6, by=.05) ### use for fixed breaks.
    chl_col <- colorRampPalette(c('honeydew2','darkseagreen3','forestgreen','darkslategrey'))
    cols_pigm <- chl_col(length(breaks_pigm)-1)
    root <- rprojroot::find_rstudio_root_file()
    dir5 <- "//data//plots//"
    path_map = paste0(root,dir5)
    file <-
    paste0(pathmap,'MBON_pigm_',
           .pig,
           .season,
           format(Sys.time(), "%y%m%d_%H%M%S") ,
           '.png')
    png(file, height = 11, width = 4, units = 'in', res=300)
    
    par(mfrow=c(3,1),mar=c(4.5,4,2,1),oma=c(4,1,4,1))

    imagePlot(ratio_kriged_pigm$x,
              ratio_kriged_pigm$y,
              ratio_kriged_pigm$z,
              col=cols_pigm,breaks=breaks_pigm,asp=1,
              xlab='',ylab='',las=1,
              xlim=xlims_pigm,ylim=ylims_pigm,
              nlevel=length(cols_pigm),legend.width=1,legend.mar=3)
contour(ratio_kriged_pigm$x,
        ratio_kriged_pigm$y,
        ratio_kriged_pigm$z,
        levels=breaks_pigm,add=T)
image(ratio_SE_pigm,add=T,breaks=quantile(ratio_SE_pigm$z,c(1,1),na.rm=T),col='white')
image(topo_lon,topo_lat,topo,breaks=c(-1,100),col='white',add=T)
plot(world2,col='gray70',add=T)
contour(topo_lon,topo_lat,topo,add=T,levels=c(-100,-50,-25,-10),col='gray40')
points(df_pigm$lon,df_pigm$lat,pch=20,col='yellow')
mtext(expression(paste('Longitude (',degree,'W)')),1,line=3,cex=.75)
mtext(expression(paste('Latitude (',degree,'N)')),2,line=3,cex=.75)
mtext(paste(.pig,':TChla:',.season),adj=1,cex=.75)

}


map_pigment("fuco", pig_dat_longer, "Winter")
map_pigment("fuco", pig_dat_longer, "Summer")
map_pigment("fuco", pig_dat_longer, "Spring")
map_pigment("fuco", pig_dat_longer, "Autumn")



```

```{r function-for-mapping}

map_phyto <- function(.phyto, .long_file, .season) {
    phyto_data <- .long_file %>%
        filter(season == .season, name_of_water_body == "Florida Keys",
               func_type == .phyto
        ) %>%
        dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  func_type, conc, tot_chl_a
        ) %>%
        group_by(hplc_gsfc_id) %>% 
    mutate(ratio = conc/tot_chl_a, .after=conc)
    
    df_phyto <- phyto_data %>% 
        group_by(station) %>% 
        summarize(ratio_avg = mean(ratio), conc_avg = mean(conc), chla_avg = mean(tot_chl_a), across()) %>% 
        distinct(station, .keep_all = T) %>% 
        ungroup()
    #krigging
    resolution <- .01
    adj <- .1
    ### krigging locations
    df_phyto.loc <- cbind(lon=df_phyto$lon,lat=df_phyto$lat)
    loc.grid_phyto <- list(lon=seq(min(df_phyto$lon, na.rm=T) - adj, max(df_phyto$lon, na.rm=T) + adj, resolution),
                 lat=seq(min(df_phyto$lat, na.rm=T) - adj, max(df_phyto$lat, na.rm=T) + adj, resolution))
    ### limits for plotting
    xlims_phyto <- range(loc.grid_phyto$lon)
    ylims_phyto <- range(loc.grid_phyto$lat)
    
    #phyto krig
    my.krig_phyto <- spatialProcess(df_phyto.loc, df_phyto$ratio_avg)
    ratio_kriged_phyto <- predictSurface(my.krig_phyto, loc.grid_phyto, extrap=F)
    ratio_SE_phyto <- predictSurfaceSE(my.krig_phyto, loc.grid_phyto, extrap=F)
    
    ### ----------------- phyto ratio krig -----------------

    ### corrects the krigged min and max to observed min and max
    if(max(ratio_kriged_phyto$z, na.rm=T) > max(df_phyto$ratio_avg, na.rm=T)){
        ratio_kriged_phyto$z[which(ratio_kriged_phyto$z > max(df_phyto$ratio_avg, na.rm=T))] <- max(df_phyto$ratio_avg, na.rm=T)
    }
    if(min(ratio_kriged_phyto$z, na.rm=T) < min(df_phyto$ratio_avg, na.rm=T)){
        ratio_kriged_phyto$z[which(ratio_kriged_phyto$z < min(df_phyto$ratio_avg, na.rm=T))] <- min(df_phyto$ratio_avg, na.rm=T)
    }
    ### color and contour breaks phyto
    breaks_phyto <- pretty(df_phyto$ratio_avg, n=10)
    breaks_phyto <- seq(0, 0.6, by=.05) ### use for fixed breaks.
    chl_col <- colorRampPalette(c('honeydew2','darkseagreen3','forestgreen','darkslategrey'))
    cols_phyto <- chl_col(length(breaks_phyto)-1)
    root <- rprojroot::find_rstudio_root_file()
    dir5 <- "//data//plots//"
    path_map <- paste0(root,dir5)
    file <-
    paste0(path_map,'MBON_pigments_',
           .phyto,
           .season,
           format(Sys.time(), "%y%m%d_%H%M%S") ,
           '.png')
    png(file, height = 11, width = 4, units = 'in', res=300)
    
    par(mfrow=c(3,1),mar=c(4.5,4,2,1),oma=c(4,1,4,1))

    imagePlot(ratio_kriged_phyto$x,
              ratio_kriged_phyto$y,
              ratio_kriged_phyto$z,
              col=cols_phyto,breaks=breaks_phyto,asp=1,
              xlab='',ylab='',las=1,
              xlim=xlims_phyto,ylim=ylims_phyto,
              nlevel=length(cols_phyto),legend.width=1,legend.mar=3)
    contour(ratio_kriged_phyto$x,
            ratio_kriged_phyto$y,
            ratio_kriged_phyto$z,
            levels=breaks_phyto,add=T)
    image(ratio_SE_phyto,add=T,breaks=quantile(ratio_SE_phyto$z,c(1,1),na.rm=T),col='white')
    image(topo_lon,topo_lat,topo,breaks=c(-1,100),col='white',add=T)
    plot(world2,col='gray70',add=T)
    contour(topo_lon,topo_lat,topo,add=T,levels=c(-100,-50,-25,-10),col='gray40')
    points(df_phyto$lon,df_phyto$lat,pch=20,col='yellow')
    mtext(expression(paste('Longitude (',degree,'W)')),1,line=3,cex=.75)
    mtext(expression(paste('Latitude (',degree,'N)')),2,line=3,cex=.75)
    mtext(paste(.phyto,':TChla:',.season),adj=1,cex=.75)

}


map_phyto("cyano4", cmtx_longer2, "Winter")
map_phyto("fuco", cmtx_longer2, "Summer")
map_phyto("fuco", cmtx_longer2, "Spring")
map_phyto("fuco", cmtx_longer2, "Autumn")

```
```{r}
#Longer chemtax dataframe with combined functional types 
cmtx_longer3 <- cmtx_avgs %>% 
    mutate(diat = diat1 + diat2,
           cyano = cyano2 + cyano4,
           hapt = hapt6 + hapt8,
           .keep = "unused"
              ) %>% 
    pivot_longer(cols = chloro:hapt, names_to = "func_type", values_to = "conc") %>%
    filter(station != 60)

#function for mapping multiple seasons in one file
map_phyto <- function(.phyto, .long_file) {
    ### ----------------- SPRING -----------------
     phyto_spring_data <- .long_file %>%
        filter(season == "Spring", name_of_water_body == "Florida Keys",
               func_type == .phyto
        ) %>%
        dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  func_type, conc, tot_chl_a
        ) %>%
        group_by(hplc_gsfc_id) %>% 
    mutate(ratio = conc/tot_chl_a, .after=conc)
    
    df_phyto_spring <- phyto_spring_data %>% 
        group_by(station) %>% 
        summarize(ratio_avg = mean(ratio), conc_avg = mean(conc), chla_avg = mean(tot_chl_a), across()) %>% 
        distinct(station, .keep_all = T) %>% 
        ungroup()
    #krigging
    resolution <- .01
    adj <- .1
    ### krigging locations
    df_phyto_spring.loc <- cbind(lon=df_phyto_spring$lon,lat=df_phyto_spring$lat)
    loc.grid_phyto_spring <- list(lon=seq(min(df_phyto_spring$lon, na.rm=T) - adj, max(df_phyto_spring$lon, na.rm=T) + adj, resolution),
                 lat=seq(min(df_phyto_spring$lat, na.rm=T) - adj, max(df_phyto_spring$lat, na.rm=T) + adj, resolution))
    ### limits for plotting
    xlims_phyto_spring <- range(loc.grid_phyto_spring$lon)
    ylims_phyto_spring <- range(loc.grid_phyto_spring$lat)
    
    #phyto krig
    my.krig_phyto_spring <- spatialProcess(df_phyto_spring.loc, df_phyto_spring$ratio_avg)
    ratio_kriged_phyto_spring <- predictSurface(my.krig_phyto_spring, loc.grid_phyto_spring, extrap=F)
    ratio_SE_phyto_spring <- predictSurfaceSE(my.krig_phyto_spring, loc.grid_phyto_spring, extrap=F)
    
    ### ----------------- Spring Phyto ratio krig -----------------

    ### corrects the krigged min and max to observed min and max
    if(max(ratio_kriged_phyto_spring$z, na.rm=T) > max(df_phyto_spring$ratio_avg, na.rm=T)){
        ratio_kriged_phyto_spring$z[which(ratio_kriged_phyto_spring$z > max(df_phyto_spring$ratio_avg, na.rm=T))] <- max(df_phyto_spring$ratio_avg, na.rm=T)
    }
    if(min(ratio_kriged_phyto_spring$z, na.rm=T) < min(df_phyto_spring$ratio_avg, na.rm=T)){
        ratio_kriged_phyto_spring$z[which(ratio_kriged_phyto_spring$z < min(df_phyto_spring$ratio_avg, na.rm=T))] <- min(df_phyto_spring$ratio_avg, na.rm=T)
    }
    
    ### color and contour breaks phyto spring
    breaks_phyto_spring <- pretty(df_phyto_spring$ratio_avg, n=10)
    breaks_phyto_spring <- seq(0, 0.6, by=.05) ### use for fixed breaks.
    chl_col <- colorRampPalette(c('honeydew2','darkseagreen3','forestgreen','darkslategrey'))
    cols_phyto_spring <- chl_col(length(breaks_phyto_spring)-1)
    
    
    
    ### ----------------- SUMMER -----------------
     phyto_summer_data <- .long_file %>%
        filter(season == "Summer", name_of_water_body == "Florida Keys",
               func_type == .phyto
        ) %>%
        dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  func_type, conc, tot_chl_a
        ) %>%
        group_by(hplc_gsfc_id) %>% 
    mutate(ratio = conc/tot_chl_a, .after=conc)
    
    df_phyto_summer <- phyto_summer_data %>% 
        group_by(station) %>% 
        summarize(ratio_avg = mean(ratio), conc_avg = mean(conc), chla_avg = mean(tot_chl_a), across()) %>% 
        distinct(station, .keep_all = T) %>% 
        ungroup()
    ### krigging locations
    df_phyto_summer.loc <- cbind(lon=df_phyto_summer$lon,lat=df_phyto_summer$lat)
    loc.grid_phyto_summer <- list(lon=seq(min(df_phyto_summer$lon, na.rm=T) - adj, max(df_phyto_summer$lon, na.rm=T) + adj, resolution),
                 lat=seq(min(df_phyto_summer$lat, na.rm=T) - adj, max(df_phyto_summer$lat, na.rm=T) + adj, resolution))
    ### limits for plotting
    xlims_phyto_summer <- range(loc.grid_phyto_summer$lon)
    ylims_phyto_summer <- range(loc.grid_phyto_summer$lat)
    
    #phyto krig
    my.krig_phyto_summer <- spatialProcess(df_phyto_summer.loc, df_phyto_summer$ratio_avg)
    ratio_kriged_phyto_summer <- predictSurface(my.krig_phyto_summer, loc.grid_phyto_summer, extrap=F)
    ratio_SE_phyto_summer <- predictSurfaceSE(my.krig_phyto_summer, loc.grid_phyto_summer, extrap=F)
    
    ### ----------------- summer Phyto ratio krig -----------------

    ### corrects the krigged min and max to observed min and max
    if(max(ratio_kriged_phyto_summer$z, na.rm=T) > max(df_phyto_summer$ratio_avg, na.rm=T)){
        ratio_kriged_phyto_summer$z[which(ratio_kriged_phyto_summer$z > max(df_phyto_summer$ratio_avg, na.rm=T))] <- max(df_phyto_summer$ratio_avg, na.rm=T)
    }
    if(min(ratio_kriged_phyto_summer$z, na.rm=T) < min(df_phyto_summer$ratio_avg, na.rm=T)){
        ratio_kriged_phyto_summer$z[which(ratio_kriged_phyto_summer$z < min(df_phyto_summer$ratio_avg, na.rm=T))] <- min(df_phyto_summer$ratio_avg, na.rm=T)
    }
    
    ### color and contour breaks phyto summer
    breaks_phyto_summer <- pretty(df_phyto_summer$ratio_avg, n=10)
    breaks_phyto_summer <- seq(0, 0.6, by=.05) ### use for fixed breaks.
    cols_phyto_summer <- chl_col(length(breaks_phyto_summer)-1)
    
    
    ###----------------- FALL -----------------------
     phyto_autumn_data <- .long_file %>%
        filter(season == "Autumn", name_of_water_body == "Florida Keys",
               func_type == .phyto
        ) %>%
        dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  func_type, conc, tot_chl_a
        ) %>%
        group_by(hplc_gsfc_id) %>% 
    mutate(ratio = conc/tot_chl_a, .after=conc)
    
    df_phyto_autumn <- phyto_autumn_data %>% 
        group_by(station) %>% 
        summarize(ratio_avg = mean(ratio), conc_avg = mean(conc), chla_avg = mean(tot_chl_a), across()) %>% 
        distinct(station, .keep_all = T) %>% 
        ungroup()
    ### krigging locations
    df_phyto_autumn.loc <- cbind(lon=df_phyto_autumn$lon,lat=df_phyto_autumn$lat)
    loc.grid_phyto_autumn <- list(lon=seq(min(df_phyto_autumn$lon, na.rm=T) - adj, max(df_phyto_autumn$lon, na.rm=T) + adj, resolution),
                 lat=seq(min(df_phyto_autumn$lat, na.rm=T) - adj, max(df_phyto_autumn$lat, na.rm=T) + adj, resolution))
    ### limits for plotting
    xlims_phyto_autumn <- range(loc.grid_phyto_autumn$lon)
    ylims_phyto_autumn <- range(loc.grid_phyto_autumn$lat)
    
    #phyto krig
    my.krig_phyto_autumn <- spatialProcess(df_phyto_autumn.loc, df_phyto_autumn$ratio_avg)
    ratio_kriged_phyto_autumn <- predictSurface(my.krig_phyto_autumn, loc.grid_phyto_autumn, extrap=F)
    ratio_SE_phyto_autumn <- predictSurfaceSE(my.krig_phyto_autumn, loc.grid_phyto_autumn, extrap=F)
    
    ### ----------------- autumn Phyto ratio krig -----------------

    ### corrects the krigged min and max to observed min and max
    if(max(ratio_kriged_phyto_autumn$z, na.rm=T) > max(df_phyto_autumn$ratio_avg, na.rm=T)){
        ratio_kriged_phyto_autumn$z[which(ratio_kriged_phyto_autumn$z > max(df_phyto_autumn$ratio_avg, na.rm=T))] <- max(df_phyto_autumn$ratio_avg, na.rm=T)
    }
    if(min(ratio_kriged_phyto_autumn$z, na.rm=T) < min(df_phyto_autumn$ratio_avg, na.rm=T)){
        ratio_kriged_phyto_autumn$z[which(ratio_kriged_phyto_autumn$z < min(df_phyto_autumn$ratio_avg, na.rm=T))] <- min(df_phyto_autumn$ratio_avg, na.rm=T)
    }
    
    ### color and contour breaks phyto autumn
    breaks_phyto_autumn <- pretty(df_phyto_autumn$ratio_avg, n=10)
    breaks_phyto_autumn <- seq(0, 0.6, by=.05) ### use for fixed breaks.
    cols_phyto_autumn <- chl_col(length(breaks_phyto_autumn)-1)
    
    
    ###----------------- WINTER -----------------------
    phyto_winter_data <- .long_file %>%
        filter(season == "Winter", name_of_water_body == "Florida Keys",
               func_type == .phyto
        ) %>%
        dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  func_type, conc, tot_chl_a
        ) %>%
        group_by(hplc_gsfc_id) %>% 
    mutate(ratio = conc/tot_chl_a, .after=conc)
    
    df_phyto_winter <- phyto_winter_data %>% 
        group_by(station) %>% 
        summarize(ratio_avg = mean(ratio), conc_avg = mean(conc), chla_avg = mean(tot_chl_a), across()) %>% 
        distinct(station, .keep_all = T) %>% 
        ungroup()
    ### krigging locations
    df_phyto_winter.loc <- cbind(lon=df_phyto_winter$lon,lat=df_phyto_winter$lat)
    loc.grid_phyto_winter <- list(lon=seq(min(df_phyto_winter$lon, na.rm=T) - adj, max(df_phyto_winter$lon, na.rm=T) + adj, resolution),
                 lat=seq(min(df_phyto_winter$lat, na.rm=T) - adj, max(df_phyto_winter$lat, na.rm=T) + adj, resolution))
    ### limits for plotting
    xlims_phyto_winter <- range(loc.grid_phyto_winter$lon)
    ylims_phyto_winter <- range(loc.grid_phyto_winter$lat)
    
    #phyto krig
    my.krig_phyto_winter <- spatialProcess(df_phyto_winter.loc, df_phyto_winter$ratio_avg)
    ratio_kriged_phyto_winter <- predictSurface(my.krig_phyto_winter, loc.grid_phyto_winter, extrap=F)
    ratio_SE_phyto_winter <- predictSurfaceSE(my.krig_phyto_winter, loc.grid_phyto_winter, extrap=F)
    
    ### ----------------- winter Phyto ratio krig -----------------

    ### corrects the krigged min and max to observed min and max
    if(max(ratio_kriged_phyto_winter$z, na.rm=T) > max(df_phyto_winter$ratio_avg, na.rm=T)){
        ratio_kriged_phyto_winter$z[which(ratio_kriged_phyto_winter$z > max(df_phyto_winter$ratio_avg, na.rm=T))] <- max(df_phyto_winter$ratio_avg, na.rm=T)
    }
    if(min(ratio_kriged_phyto_winter$z, na.rm=T) < min(df_phyto_winter$ratio_avg, na.rm=T)){
        ratio_kriged_phyto_winter$z[which(ratio_kriged_phyto_winter$z < min(df_phyto_winter$ratio_avg, na.rm=T))] <- min(df_phyto_winter$ratio_avg, na.rm=T)
    }
    
    ### color and contour breaks phyto winter
    breaks_phyto_winter <- pretty(df_phyto_winter$ratio_avg, n=10)
    breaks_phyto_winter <- seq(0, 0.6, by=.05) ### use for fixed breaks.
    cols_phyto_winter <- chl_col(length(breaks_phyto_winter)-1)
    
    
    root <- rprojroot::find_rstudio_root_file()
    dir5 <- "//data//plots//"
    path_map <- paste0(root,dir5)
    file <-
    paste0(path_map,'MBON_pigments_',
           .phyto,
           format(Sys.time(), "%y%m%d_%H%M%S") ,
           '.png')
    png(file, height = 11, width = 4, units = 'in', res=300)
    
    par(mfrow=c(4,1),mar=c(4.5,4,2,1),oma=c(4,1,4,1))

    ###SPRING
    imagePlot(ratio_kriged_phyto_spring$x,
              ratio_kriged_phyto_spring$y,
              ratio_kriged_phyto_spring$z,
              col=cols_phyto_spring,breaks=breaks_phyto_spring,asp=1,
              xlab='',ylab='',las=1,
              xlim=xlims_phyto_spring,ylim=ylims_phyto_spring,
              nlevel=length(cols_phyto_spring),legend.width=1,legend.mar=3)
    contour(ratio_kriged_phyto_spring$x,
            ratio_kriged_phyto_spring$y,
            ratio_kriged_phyto_spring$z,
            levels=breaks_phyto_spring,add=T)
    image(ratio_SE_phyto_spring,add=T,breaks=quantile(ratio_SE_phyto_spring$z,c(1,1),na.rm=T),col='white')
    image(topo_lon,topo_lat,topo,breaks=c(-1,100),col='white',add=T)
    plot(world2,col='gray70',add=T)
    contour(topo_lon,topo_lat,topo,add=T,levels=c(-100,-50,-25,-10),col='gray40')
    points(df_phyto_spring$lon,df_phyto_spring$lat,pch=20,col='yellow')
    mtext(expression(paste('Longitude (',degree,'W)')),1,line=3,cex=.75)
    mtext(expression(paste('Latitude (',degree,'N)')),2,line=3,cex=.75)
    mtext(paste(.phyto,':TChla:Spring'),adj=1,cex=.75)
    
    ###SUMMER
    imagePlot(ratio_kriged_phyto_summer$x,
              ratio_kriged_phyto_summer$y,
              ratio_kriged_phyto_summer$z,
              col=cols_phyto_summer,breaks=breaks_phyto_summer,asp=1,
              xlab='',ylab='',las=1,
              xlim=xlims_phyto_summer,ylim=ylims_phyto_summer,
              nlevel=length(cols_phyto_summer),legend.width=1,legend.mar=3)
    contour(ratio_kriged_phyto_summer$x,
            ratio_kriged_phyto_summer$y,
            ratio_kriged_phyto_summer$z,
            levels=breaks_phyto_summer,add=T)
    image(ratio_SE_phyto_summer,add=T,breaks=quantile(ratio_SE_phyto_summer$z,c(1,1),na.rm=T),col='white')
    image(topo_lon,topo_lat,topo,breaks=c(-1,100),col='white',add=T)
    plot(world2,col='gray70',add=T)
    contour(topo_lon,topo_lat,topo,add=T,levels=c(-100,-50,-25,-10),col='gray40')
    points(df_phyto_summer$lon,df_phyto_summer$lat,pch=20,col='yellow')
    mtext(expression(paste('Longitude (',degree,'W)')),1,line=3,cex=.75)
    mtext(expression(paste('Latitude (',degree,'N)')),2,line=3,cex=.75)
    mtext(paste(.phyto,':TChla:Summer'),adj=1,cex=.75)
    
    ###FALL
    imagePlot(ratio_kriged_phyto_autumn$x,
              ratio_kriged_phyto_autumn$y,
              ratio_kriged_phyto_autumn$z,
              col=cols_phyto_autumn,breaks=breaks_phyto_autumn,asp=1,
              xlab='',ylab='',las=1,
              xlim=xlims_phyto_autumn,ylim=ylims_phyto_autumn,
              nlevel=length(cols_phyto_autumn),legend.width=1,legend.mar=3)
    contour(ratio_kriged_phyto_autumn$x,
            ratio_kriged_phyto_autumn$y,
            ratio_kriged_phyto_autumn$z,
            levels=breaks_phyto_autumn,add=T)
    image(ratio_SE_phyto_autumn,add=T,breaks=quantile(ratio_SE_phyto_autumn$z,c(1,1),na.rm=T),col='white')
    image(topo_lon,topo_lat,topo,breaks=c(-1,100),col='white',add=T)
    plot(world2,col='gray70',add=T)
    contour(topo_lon,topo_lat,topo,add=T,levels=c(-100,-50,-25,-10),col='gray40')
    points(df_phyto_autumn$lon,df_phyto_autumn$lat,pch=20,col='yellow')
    mtext(expression(paste('Longitude (',degree,'W)')),1,line=3,cex=.75)
    mtext(expression(paste('Latitude (',degree,'N)')),2,line=3,cex=.75)
    mtext(paste(.phyto,':TChla:Autumn'),adj=1,cex=.75)
    
    ###WINTER
    imagePlot(ratio_kriged_phyto_winter$x,
              ratio_kriged_phyto_winter$y,
              ratio_kriged_phyto_winter$z,
              col=cols_phyto_winter,breaks=breaks_phyto_winter,asp=1,
              xlab='',ylab='',las=1,
              xlim=xlims_phyto_winter,ylim=ylims_phyto_winter,
              nlevel=length(cols_phyto_winter),legend.width=1,legend.mar=3)
    contour(ratio_kriged_phyto_winter$x,
            ratio_kriged_phyto_winter$y,
            ratio_kriged_phyto_winter$z,
            levels=breaks_phyto_winter,add=T)
    image(ratio_SE_phyto_winter,add=T,breaks=quantile(ratio_SE_phyto_winter$z,c(1,1),na.rm=T),col='white')
    image(topo_lon,topo_lat,topo,breaks=c(-1,100),col='white',add=T)
    plot(world2,col='gray70',add=T)
    contour(topo_lon,topo_lat,topo,add=T,levels=c(-100,-50,-25,-10),col='gray40')
    points(df_phyto_winter$lon,df_phyto_winter$lat,pch=20,col='yellow')
    mtext(expression(paste('Longitude (',degree,'W)')),1,line=3,cex=.75)
    mtext(expression(paste('Latitude (',degree,'N)')),2,line=3,cex=.75)
    mtext(paste(.phyto,':TChla:Winter'),adj=1,cex=.75)

    dev.off()

}

map_phyto("chloro", cmtx_longer3)

map_phyto("cyano", cmtx_longer3)

map_phyto("crypto", cmtx_longer3)

map_phyto("diat", cmtx_longer3)

map_phyto("dino", cmtx_longer3)

map_phyto("hapt", cmtx_longer3)

map_phyto("pras", cmtx_longer3)


```
