---
title: "Template for Mapping"
author: "Sebastian DiGeronimo"
date: '2022-06-14'
output: html_document
---

## Creates map of site locations for IMaRS MBON
Uses topography and coastline data from NOAA.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("fields")
library("raster")
library("rerddap")
library("rgdal")
library("ncdf4")
library("magrittr")
library("ggplot2")
library("sf")
library("metR")
library("tidyr")
library("dplyr")
library("cowplot")
library("conflicted")
library("here")
library("stringr")
library("glue")

# prefer select and filter 
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

# fixes issue with reading file
sf::sf_use_s2(FALSE)

# path for map to find/download to, they can be different depending on
# your computer
path_land <- here("data", "map")
path_topo <- here("data", "map")

# Region of Interest Bounding Box                   ======== edit this if needed
# Change the spatial extent
exnt <- c(xmin = -82, # West
          xmax = -80, # East
          ymin = 24.25,# South
          ymax = 25.75   # North 
        )

# inset extent (i.e. smaller map on top)            ======== edit this if needed
exnt2 <-  c(-90, # West
            -75, # East
            20,  # South
            35   # North
            ) 

names(exnt2) <- names(exnt)
```

# Load Combined Pigment Data
```{r load-pigment-data}
pig_dat  <- read.csv(here("data", "processed", "combined_pig_dat.csv")) %>%
    select(-c(pi, 
              bottle, indicate_if_filters_are_replicates, volfilt,
              filter_type:filter_storage_before_shipment_to_gfsc,
              comments:other, collected_with_positive_pressure_or_vaccuum,
              date_extracted_month_day_year,
              cruise, vx_ml))
```


# Download bathymetry and coastline files
**Note:** this will be done automatically if files **do not exists**

Topography, ETOPO1, 0.0166667 degrees, Global (longitude -180 to 180), (Ice Sheet Surface) from https://coastwatch.pfeg.noaa.gov/erddap/griddap/

Global Self-consistent, Hierarchical, High-resolution Geography Database (GSHHG) from https://www.ngdc.noaa.gov/mgg/shorelines/
```{r download-maps}
# source map download function
source(here("scripts", "map_files_dwnlod.R"))

world_download(
    # path_land = path_land,
    path_land = path_land,
    path_topo = path_topo,
    extent = exnt
)
```

# Load coastline and topography data
```{r load-shapefiles}
# select and read coastline file from GSHHS then crop
world2 <- fs::dir_ls(path = path_land,
               recurse = TRUE ,
               regexp = "GSHHS_h_L1.shp") %>%
               st_read(.) %>%
               st_crop(., st_bbox(exnt2))

# read state lines and crop
state <- fs::dir_ls(path = path_land,
               recurse = TRUE ,
               regexp = "WDBII_border_h_L2.shp") %>%
               st_read(.) %>%
               st_crop(., st_bbox(exnt2))


bathy <- nc_open(here(path_topo, "etopo1.nc"))
topo <- ncvar_get(bathy, 'altitude')
topo_lat <- ncvar_get(bathy, 'latitude')
topo_lon <- ncvar_get(bathy, 'longitude')
nc_close(bathy)

# read topography data
bathy <- raster::raster(here(path_topo,"etopo1.nc")) %>%
  as.data.frame(xy = TRUE)

rm(path_topo, path_land)
```

# Plot FL Keys stations on map
```{r plt-fl-keys, fig.height=6, fig.width=10}
lat_lon_dat <- pig_dat %>%
    filter(name_of_water_body == "Florida Keys" & 
           !str_detect(station, "60")) %>%
    select(station, lat, lon) %>% 
    distinct(station, lat, lon)

base_map <-
    ggplot() +
    geom_contour2(data = bathy, 
                  aes(x = x, 
                      y = y, 
                      z = -Altitude), 
                  col = "grey40",
                  breaks = c(100, 50, 25, 10, 0)
                  ) +
    geom_sf(data = world2) +
    
    # TODO: remove ----
    # geom_point(data = lat_lon_dat,
    #            aes(
    #                x = lon,
    #                y = lat
    #                # shape = station,
    #                ),
    #           size = 2,
    #           # alpha = 0.5,
    #           stroke = 1.5,
    #           show.legend = FALSE) +

    # geom_text(data = lat_lon_dat,
    #           aes(x = lon, y = lat, label = station),
    #           vjust = -0.5,
    #           hjust = -0.0001,
    #           size = 5) +

    # scale_shape_manual(values = c(24,25,21)) +
    scale_x_continuous(expand = c(0,0)) + 
    scale_y_continuous(expand = c(0,0), 
                       breaks = seq(floor(exnt[3]), ceiling(exnt[4]), 0.5)
                       ) +
    # coord_sf(xlim = exnt[1:2], ylim = exnt[3:4]) +
    labs(
      x = NULL,
      y = NULL
      ) +
    theme_bw() +
    theme(text = element_text(family = "serif", size = 20),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()
          # legend.position = c(0.85, 0.25),
          # legend.title = element_text(size = 10), 
          # legend.text = element_text(size = 15),
          # legend.key.size = unit(0.6, "cm"),
          # legend.margin=margin(c(0.1,0.1,0.1,0.1), unit='cm')
          )

stn.map <- 
    base_map +
    geom_point(data = lat_lon_dat,
               aes(
                   x = lon,
                   y = lat
                   # shape = station,
                   ),
              size = 2,
              # alpha = 0.5,
              stroke = 1.5,
              show.legend = FALSE) +
    
    labs(
        shape = "Station",
        fill = "Station") +
    coord_sf(xlim = exnt[1:2], ylim = exnt[3:4]) +
     scale_shape_manual(values = c(24,25,21)) 

base_map
stn.map
# plotly::ggplotly(
#     ggplot() +
#   geom_point(data = pig_dat,
#              aes(
#                x = lon,
#                y = lat,
#                fill = station,
#                ##shape = station,
#              ),
#              size = 2,
#              alpha = 0.5,
#              stroke = 1.5,
#              show.legend = FALSE)
# )
```
# Zoomed out map 
```{r plt-large, fig.width=8, fig.height=10}
east.map <- 
      
    ggplot() +
    geom_sf(data = world2) +  
    geom_sf(data = state, color = "grey40", linewidth = 0.2) +
  
    # create red box on map to show sampling locations
    geom_rect(                  
        aes(
          xmin = min(lat_lon_dat$lon) - 0.25, # West
          xmax = max(lat_lon_dat$lon) + 0.25, # East
          ymin = min(lat_lon_dat$lat) - 0.25, # South
          ymax = max(lat_lon_dat$lat) + 0.25  # North
        ),
        color = "red",
        fill = NA
    ) +
    
    scale_x_continuous(expand = c(0,0)) + 
    scale_y_continuous(expand = c(0,0)) +
    coord_sf(xlim = exnt2[1:2], ylim = exnt2[3:4], 
            ) +
    
    theme_bw() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          plot.margin = grid::unit(c(0,0,0,0), "in"),
          panel.border = element_rect(linetype = "solid", 
                                      color = "black", 
                                      fill = NA)
        )

east.map
```

# putting both of the maps together
May need to play with draw_plot(east.map, x, y, width and height)
```{r plt-lg-sm, fig.height=6, fig.width=10}
ggdraw() +
  draw_plot(stn.map) +
  draw_plot(east.map, x = 0.04, y = 0.57, width = 0.1908108, height = 0.4)
```
```{r save-plt-map}
if (FALSE) {
    save_plot(
        filename = here("data", "plots",
                        glue::glue("map",
                                   format(Sys.time(), '_%Y%m%d_%H%M%S'),
                                   ".jpeg")), 
        plot = last_plot(),
        base_height = 6,
        base_width = 10,
        dpi = 600, 
        units = "in", 
        device = 'jpeg')
    
    save_plot(
        filename = here("data", "plots",
                        glue::glue("map",
                                   format(Sys.time(), '_%Y%m%d_%H%M%S'),
                                   ".svg")), 
        plot = last_plot(),
        base_height = 6,
        base_width = 10,
        dpi = 600, 
        units = "in", 
        device = 'svg')
}
```

# ---- Kriging for Fucoxanthin ----
Test to see if it works
```{r filter-fuco-winter}
# filter for pigment and season
conc_name <- "perid"
loc       <- "West Florida Shelf"
szn_sel   <- "Spring"
szn_filt  <- case_when(
    str_detect(szn_sel, "(?i)sp|(?i)su") ~ str_trunc(szn_sel, 2, ellipsis = ""),
    str_detect(szn_sel, "^(?i)a") ~ "F",
    str_detect(szn_sel, "^(?i)w") ~ "W") %>%
    str_to_title()

df_fuco <-
    pig_dat %>%
    pivot_longer(tot_chl_b:dp, names_to = "conc_name", values_to = "conc") %>%
    filter(
        str_detect(season, szn_filt) &
        str_detect(name_of_water_body, loc) &
        str_detect(conc_name, pig)
        ) %>%
    select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
        tot_chl_a, conc_name, conc
        ) %>%
    
    mutate(ratio  = conc/tot_chl_a) %>%
    group_by(station) %>%
    summarise(
        across(c(ratio, conc, tot_chl_a), 
               .fn = list(avg = mean), na.rm = TRUE), 
        across(c(lat, lon), .fn = mean)) %>%
    ungroup()


source("../scripts/krig_fun.R")
map_data(
    df_fuco,
    conc_name = conc_name,
    .title   = szn_sel,
    base_map = base_map
)
```


```{r base-plot-fuco}
if (FALSE) {
png(file, height = 11, width = 4, units = "in", res = 300)
par(
    mfrow = c(3, 1),
    mar = c(4.5, 4, 2, 1),
    oma = c(4, 1, 4, 1)
)

imagePlot(
    ratio_kriged_fuco$x,
    ratio_kriged_fuco$y,
    ratio_kriged_fuco$z,
    col = cols_fuco,
    breaks = breaks_fuco,
    asp = 1,
    xlab = '',
    ylab = '',
    las = 1,
    xlim = xlims_fuco,
    ylim = ylims_fuco,
    nlevel = length(cols_fuco),
    legend.width = 1,
    legend.mar = 3
)

contour(
    ratio_kriged_fuco$x,
    ratio_kriged_fuco$y,
    ratio_kriged_fuco$z,
    levels = breaks_fuco,
    add = TRUE
)

image(
    ratio_SE_fuco,
    add = TRUE,
    breaks = quantile(ratio_SE_fuco$z, c(1, 1), na.rm = TRUE),
    col = 'white'
)

image(
    topo_lon,
    topo_lat,
    topo,
    breaks = c(-1, 100),
    col = 'white',
    add = TRUE
)

# plot(FL,col='gray70',add=TRUE)
plot(world2, col = 'gray70', add = TRUE)
contour(
    topo_lon,
    topo_lat,
    topo,
    add = TRUE,
    levels = c(-100, -50, -25, -10),
    col = 'gray40'
)

points(df_fuco$lon,
       df_fuco$lat,
       pch = 20,
       col = 'yellow')

mtext(expression(paste('Longitude (\u00B0W)')), 1, line = 3, cex = 0.75)

mtext(expression(paste('Latitude (\u00B0N)')), 2, line = 3, cex = 0.75)
mtext('Fucoxanthin:TChla: Winter', adj = 1, cex = 0.75)
}
```

```{r function-for-mapping}
map_phyto <- function(.phyto, .long_file, .season) {
    
    # filter data for season and water body
    phyto_data <- .long_file %>%
        filter(season == .season &
               name_of_water_body == "Florida Keys" &
               func_type == .phyto
        ) %>%
        dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  func_type, conc, tot_chl_a
        ) %>%
        group_by(hplc_gsfc_id) %>% 
    mutate(ratio = conc/tot_chl_a, .after=conc)
    
    # 
    df_phyto <- phyto_data %>% 
        group_by(station) %>% 
        summarize(ratio_avg = mean(ratio), 
                  conc_avg = mean(conc), 
                  chla_avg = mean(tot_chl_a), 
                  across()) %>% 
        distinct(station, .keep_all = TRUE) %>% 
        ungroup()

}


map_phyto("cyano4", cmtx_longer2, "Winter")
map_phyto("fuco", cmtx_longer2, "Summer")
map_phyto("fuco", cmtx_longer2, "Spring")
map_phyto("fuco", cmtx_longer2, "Autumn")

```
```{r}
#Longer chemtax dataframe with combined functional types 
cmtx_longer3 <-
    fs::dir_ls(here(), 
               regexp  = "chemtax_with_metadata", 
               recurse = TRUE) %>% 
    read.csv() %>%
    
    mutate(diat  = diat1 + diat2,
           cyano = cyano2 + cyano4,
           hapt  = hapt6 + hapt8,
           .keep = "unused"
              ) %>% 
    
    pivot_longer(cols = chloro:hapt, 
                 names_to = "func_type", 
                 values_to = "conc") %>%
    
    filter(station != 60) %>%
    select(
       hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  func_type, conc, tot_chl_a 
    )
```
# PROBABLY DONT NEED
```{r all-seasons}
    # ----------------- SPRING -----------------
     phyto_spring_data <- .long_file %>%
        filter(season == "Spring" & name_of_water_body == "Florida Keys" &
               func_type == .phyto
        ) %>%
        dplyr::select(
        hplc_gsfc_id, sample, station, lat, lon, season, year, date_time_utc,
                  func_type, conc, tot_chl_a
        ) %>%
        group_by(hplc_gsfc_id) %>% 
    mutate(ratio = conc/tot_chl_a, .after=conc)
    
    df_phyto_spring <- 
        phyto_spring_data %>% 
        group_by(station) %>% 
        summarize(ratio_avg = mean(ratio), 
                  conc_avg = mean(conc), 
                  chla_avg = mean(tot_chl_a), across()) %>% 
        distinct(station, .keep_all = TRUE) %>% 
        ungroup(

```
