---
title: "New ANOVA Testing"
author: "Sebastian Di Geronimo"
date: 2024-07-12
format: html
editor: source
---

# 1.0 ---- Summary of Document ----



# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  forcats, lubridate, glue, fs, magrittr, here,
  
  # additional
  patchwork,
  vegan
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )
  
source(here("scripts", "misc_functions.R"))
```

```{r}
col_manual <-
  c(
    # "Offshore"  = "blue",
    # "Nearshore" = "red",
    "Offshore"  = "#882255",
    "Nearshore" = "#88CCEE",
    "Winter"    = "#88CCEE",
    "Spring"    = "#44AA99",
    "Summer"    = "#DDCC77",
    "Autumn"    = "#CC6677"
  )
```


## 2.2 Load Dataset

```{r load-data}
# set location to save data and plots
dir_data_save <- here("data", "processed", "new_anova_tests")
dir_plt_save  <- here("data", "plots", "cap_plots")

# ---- set season names and order ---- #
szn <- c("Winter", "Spring", "Summer", "Autumn")

# ---- phytoplankton full names ---- #
phyto_name <-
  here("data", "processed", "04_1_merge_meta_cmtx", "phyto_names.csv") %>%
  read_csv(show_col_types = FALSE) %T>% 
  print()

# ---- load chemtax data in wide format ---- #
cmtx_w_og <-
  here("data", "processed", "04_1_merge_meta_cmtx", "chemtax_w_meta_wide.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  # filter(!str_detect(station, "60|21.5|24|^3$|9B|9.5")) %>%
  filter(!str_detect(station, "60|21.5|24|^3$|9.5")) %>%
  mutate(
    season = fct(season, levels = szn),
    # station = str_replace(station, "9B", "9.5"),
    station = fct_reorder(station, lon)
  ) %>% 
  select(-c(hplc_gsfc_id, cruise:bottle, cluster_code, water_type)) %>%
    
  # add PFT concentrations to get total chlorophyll-a
  rowwise() %>%
  mutate(
    total_chl_a = sum(c_across(chloro:pras)),
    .before = chloro
  ) %>%
  ungroup() %>%
  mutate(
      .after = total_chl_a,
      across(
          total_chl_a:pras,
          ~ if_else(
          str_detect(sample, "WS20278-32"), 
          .x / 10,
            .x)
      )
  ) %>%
  filter(
    .by = c(station, cruise_id),
    depth == min(depth)
    & depth < 20
  ) %>%
  mutate(
    nearoff = 
      if_else(
        str_detect(station, "10|16"), 
        "Nearshore", 
        "Offshore"
        )
    ) %>%
  relocate(nearoff, .after = station) %T>% 
  print()
```

## 2.2.1 Remove Outliers?

```{r remove-outliers}
# remove_out <- menu(c("yes", "no"), title = "Remove Outliers?")
remove_out <-  2
if (remove_out == 1) {
  message("Filtering out outliers!")
  cmtx_w <-
    cmtx_w_og %>%
    filter(
      total_chl_a <= 1
    ) %T>% 
    print()
  
  file_suffix <- "_no_out"
} else {
  message("Keeping outliers!")
  cmtx_w <-
    cmtx_w_og
  file_suffix <- ""
}
```


## 2.3 Convert to Long Format


Pivot longer by taking PFT columns and pivot to two columns of `phyto` for name
and `conc` for concentration.

Next, convert to percentage by `conc` / `total_chl_a`


```{r convert-long-format}
# one column for all PFT and one for concentrations
cmtx_l <-
  cmtx_w %>%
  pivot_longer(
    cols      = chloro:pras,
    names_to  = "phyto",
    values_to = "conc"
  ) %>%
  left_join(phyto_name, by = c("phyto" = "shrt")) %>%
  
  # calc percent
  mutate(percent = conc/total_chl_a) %T>% 
  print()

# extract percent column
cmtx_l_pct <-
  cmtx_l %>%
  select(-conc) %T>% 
  print()

cmtx_w_pct <- 
  cmtx_l_pct %>%
  select(-names) %>%
  pivot_wider(
    data         = .,
    names_from   = c(phyto),
    values_from  = c(percent),
    names_sep    = "_",
    names_repair = janitor::make_clean_names
    ) %>%
  mutate(nearoff = factor(nearoff))
```

# 3.x

Year     = factor
Distance = factor
Season   = factor

```{r}
shelf(emmeans, sjPlot)
  # lmerTest::lmer(
  #   Richness ~ Type * Distance * (1|Year) * (1|Season), 
  #   Reefs_SR
  # )


chla_anova <-
  lmerTest::lmer(
    total_chl_a ~ nearoff * season *  (1 | year),
    mutate(cmtx_w, year = factor(year))
  )

summary(chla_anova)

(chla_anova_means <- 
  emmeans(
    chla_anova,
    ~ nearoff * season
    )
)


contrast(chla_anova_means, "pairwise")

tab_model(chla_anova)
```


```{r}
chla_anova <-
  glm(
    total_chl_a ~  nearoff + season + year,
    data = mutate(cmtx_w, year = factor(year)) # %>%
           # filter(str_detect(station, "10|12|16|18"), total_chl_a < 1)
    )
  

summary(chla_anova)

(chla_anova_means <- 
  emmeans(chla_anova,
      ~  season)
)
pairs(chla_anova_means)

(chla_anova_means <- 
  emmeans(chla_anova,
      ~  season | nearoff)
)

pairs(chla_anova_means)

emmip(ref_grid(chla_anova), nearoff ~ season, style = "factor", CIs = TRUE, type = "response")


plot(chla_anova_means) +
  geom_point(
    data = filter(cmtx_w, str_detect(station, "10|12|16|18"), total_chl_a < 1),
    # data = cmtx_w,
    aes(y = nearoff, x = total_chl_a, group = season),
    shape = 2
  ) +
  coord_flip() +
  facet_grid(~season)
```




# 4.0 Q-Q Plot

when skew is:
- negative = left-skewed
- positive = right-skewed
moments::skewness(<vector>, na.rm = TRUE)

when kurtosis is:
- >3 = more values in tail
- <=3 = more values in not in tail
moments::kurtosis(<vector>, na.rm = TRUE)


## 4.1 Q-Q Plot All

Absolute:
- Shapiro-Wilk p-value < 0.05, therefore NOT normal
- Skewness > 0, right-skewed
- Kurtosis > 3, more value in tail

Log:
- Shapiro-Wilk p-value > 0.05, therefore normal
- Skewness > 0, right-skewed
- Kurtosis < 3, more value NOT in tail

```{r qq-plot-all}
{
# total chlorophyll-a
cmtx_w %$% 
  car::qqPlot(
    total_chl_a, 
    main = "All",
    ylab = "Total Chlorophyll-a"
    )

shapiro.test(cmtx_w$total_chl_a) %>%
    print()

# skewness test
moments::skewness(cmtx_w$total_chl_a, na.rm = TRUE) %>%
  round(., digits = 3) %>%
  paste("Skewness:", .) %>%
  message()

# kurtosis
moments::kurtosis(cmtx_w$total_chl_a, na.rm = TRUE) %>%
  round(., digits = 3) %>%
  paste("Kurtosis:", .) %>%
  message()
}

{
# log total chlorophyll-a
cmtx_w %$% 
  car::qqPlot(
    log(total_chl_a + 0.001),, 
    main = "All",
    ylab = "log(Total Chlorophyll-a)"
    )

shapiro.test(log(cmtx_w$total_chl_a + 0.001)) %>%
    print()

# skewness test
moments::skewness(log(cmtx_w$total_chl_a + 0.001), na.rm = TRUE) %>%
  round(., digits = 3) %>%
  paste("Skewness:", .) %>%
  message()

# kurtosis
moments::kurtosis(log(cmtx_w$total_chl_a + 0.001), na.rm = TRUE) %>%
  round(., digits = 3) %>%
  paste("Kurtosis:", .) %>%
  message()
}
```

## 4.2 Q-Q Plot Seaonal

Winter:
- Shapiro-Wilk p-value < 0.05, therefore NOT normal
- Skewness > 0, right-skewed
- Kurtosis < 3, more value NOT in tail

Spring:
- Shapiro-Wilk p-value < 0.05, therefore NOT normal
- Skewness > 0, right-skewed
- Kurtosis > 3, more value in tail

Summer:
- Shapiro-Wilk p-value < 0.05, therefore NOT normal
- Skewness > 0, right-skewed
- Kurtosis > 3, more value in tail

Autumn:
- Shapiro-Wilk p-value < 0.05, therefore NOT normal
- Skewness > 0, right-skewed
- Kurtosis > 3, more value in tail

```{r qq-plot-seasonal}
walk(
  szn,
  \(.x) {
    message("\n-----\n", .x, "\n-----\n\n")
    cmtx_w %>%
      filter(str_detect(season, .x)) %$%
      car::qqPlot(
        total_chl_a,
        main = .x,
        ylab = "Total Chlorophyll-a"
      ) %>%
      print()

    # shaprio-wilk
    cmtx_w %>%
      filter(str_detect(season, .x)) %$%
      shapiro.test(total_chl_a) %>%
      print()
    
    # skewness test
    cmtx_w %>%
      filter(str_detect(season, .x)) %$%
      moments::skewness(total_chl_a, na.rm = TRUE) %>%
      round(., digits = 3) %>%
      paste("Skewness:", .) %>%
      message()
    
    # kurtosis
    cmtx_w %>%
      filter(str_detect(season, .x)) %$%
      moments::kurtosis(total_chl_a, na.rm = TRUE) %>%
      round(., digits = 3) %>%
      paste("Kurtosis:", .) %>%
      message()
  }
)
```
Log:

Winter:
- Shapiro-Wilk p-value > 0.05, therefore normal
- Skewness > 0, right-skewed
- Kurtosis < 3, more value in tail

Spring:
- Shapiro-Wilk p-value > 0.05, therefore normal
- Skewness > 0, right-skewed
- Kurtosis < 3, more value in tail

Summer:
- Shapiro-Wilk p-value > 0.05, therefore normal
- Skewness > 0, right-skewed
- Kurtosis > 3, more value NOT in tail

Autumn:
- Shapiro-Wilk p-value < 0.05, therefore NOT normal
- Skewness > 0, right-skewed
- Kurtosis < 3, more value in tail

```{r qq-plot-seasonal-log}
walk(
  szn,
  \(.x) {
    message("\n-----\n", .x, "\n-----\n\n")
    cmtx_w %>%
      filter(str_detect(season, .x)) %$%
      car::qqPlot(
        log(total_chl_a + 0.001),
        main = .x,
        ylab = "log(Total Chlorophyll-a)"
      ) %>%
      print()
    
    # shaprio-wilk
    cmtx_w %>%
      filter(str_detect(season, .x)) %$%
      shapiro.test(log(total_chl_a + 0.001)) %>%
      print()
    
    # skewness test
    cmtx_w %>%
      filter(str_detect(season, .x)) %$%
      moments::skewness(log(total_chl_a + 0.001), na.rm = TRUE) %>%
      round(., digits = 3) %>%
      paste("Skewness:", .) %>%
      message()
    
    # kurtosis
    cmtx_w %>%
      filter(str_detect(season, .x)) %$%
      moments::kurtosis(log(total_chl_a + 0.001), na.rm = TRUE) %>%
      round(., digits = 3) %>%
      paste("Kurtosis:", .) %>%
      message()
  }
)



```

# 5.0 CAP

## 5.1 CAP Function
```{r cap-func}
shelf(BiodiversityR)

cap_pctvars <- function(cap_out){
  # A function that takes the output of the CAPdiscrim function and 
  # produces the percent of among-group variability explained by each
  # canonical axis.
  #
  # Inputs: 
  # cap_out = 'CAPDiscrim' function list output
  # 
  # Output: 
  # pct_var = A numerical array with dimensions 1 x p, where p is the
  #           number of canonical axes, corresponding to the percent
  #           of among-group variability explained by each axis.
  #         
  
  # Extract eigenvalues from manova sublist
  eig     <- cap_out[["manova"]][["Eigenvalues"]]
  
  # Convert eigenvalues to analog of canonical variability
  vars    <- eig*cap_out$tot*(cap_out$varm/100)/(eig+1)
  
  # Calculate percent of variability explained by each axis
  pct_var <- 100*vars/sum(vars)
  
  # Trim columns not corresponding to canonical axes
  pct_var <- pct_var[1:ncol(cap_out$x)]
}


cap_extraction <- function(
    .data,
    .data_sub,
    cap_dist,
    scale_fact = 5,
    ld2_range = c(-1.25, 1.25),
    ld2_spc_h = c(-1, 1)) {

  `%!in%` <- Negate(`%in%`)

  # Percent of the total variance explained by the axes used to describe the CAP
  # model (e.g., 0.15 of the TOTAL VARIANCE)
  cap_pct_var <-
    sum(diag(cap_dist$manova$SS$`y[, group]`)) / cap_dist$tot

  # Percent of total variance explained by the CAP model that is explained by EACH
  # CA. (e.g., 0.12 of the 0.15 of the TOTAL VARIANCE from above)
  cap_var_expl <-
    cap_pctvars(cap_dist)
  # CA-II explains NO between group variance w/ 2 groups
  if (length(cap_var_expl) < 3) {
    cap_var_expl[2] <- 0
  }
  
  ## Adds LD2 is not present in the dataframe.
  if ("LD2" %!in% colnames(cap_dist$x)) {
    message(paste("Grouping factor < 3, adding Jittered LD2 Column with range", 
                  ld2_range[1], "to", ld2_range[2], "\n"))
    cap_dist$x <-
      as.data.frame(
        cap_dist$x
      ) %>%
      add_column(
        LD2 = runif(nrow(cap_dist$x), ld2_range[1], ld2_range[2])
      )
    ## Adjust LD2 to jittered values if present in the dataframe.
  } else if ("LD2" %in% colnames(cap_dist$x)) {
    message("Keep LD2")
    cap_dist$x <- 
      as.data.frame(cap_dist$x)
  } else {
    message(paste("Grouping factor < 3, modifying Jittered LD2 Column with range", 
                  ld2_range[1], "to", ld2_range[2], "\n"))
    cap_dist$x <-
      as.data.frame(
        cap_dist$x
      ) %>%
      mutate(
        LD2 = runif(nrow(cap_dist$x), ld2_range[1], ld2_range[2])
      )
  }


  # Add species scores
  cap_dist <-
    add.spec.scores(
      cap_dist,
      .data_sub,
      # Vector scaling factor
      multi = scale_fact
    )


  cap_site <-
    as_tibble(cap_dist$x) %>%
    bind_cols(.data)

  cap_score <-
    as_tibble(
      cap_dist$cproj,
      rownames = "pft"
    ) 
  
  if (length(cap_var_expl) < 3) {
    message(paste("Grouping factor < 3, adding jittered axis from", 
                  ld2_spc_h[1], "to", ld2_spc_h[2], "\n"))
    cap_score <- 
      cap_score %>%
      mutate(
        SpCode = rownames(.),
        LD1_sign = sign(LD1)
      ) %>%
      mutate(
        # LD2 = sample(seq(-1, 1, length = n())),
        LD2 = sample(seq(ld2_spc_h[1], ld2_spc_h[2], length = n())),
        .by = LD1_sign
      )
  }


  return(
    list(
       "full"     = cap_dist,
       "score"    = cap_score,
       "site"     = cap_site,
       "pct_var"  = cap_pct_var,
       "var_expl" = cap_var_expl
    )
  )
  # ---- end of function cap_extraction ---- #
}

```

```{r plot-func}
shelf(shadowtext)

shadowtext::geom_shadowtext()
plot_template <- function(
    cap_ext,
    label_name,
    axes = c(1, 2),
    title,
    .col) {
    
  .cap_site  <- cap_ext$site
  .cap_score <- cap_ext$score
  pct_var    <- cap_ext$pct_var
  var_expl   <- cap_ext$var_expl


  .cap_site <-
    .cap_site %>%
    rename("col_shp" = {{ .col }})

  y_lab <- if_else(
    length(axes) > 1,
    paste0(
      "CA2 (variance explained: ",
      round(var_expl[2], digits = 1),
      "%)"
    ),
    paste0("Jittered axis")
  )


  ggplot() +
    theme_classic() +
    theme(
      axis.title       = element_text(size = 32 / .pt, face = "bold"),
      axis.text        = element_text(size = 22 / .pt, face = "bold", 
                                      family = "serif"),
      legend.text      = element_text(size = 28 / .pt, face = "bold"),
      legend.title     = element_text(size = 32 / .pt, face = "bold"),
      legend.direction = "horizontal",
      # legend.position  = c(0.5, 1.0),
      legend.position  = "top",
      legend.margin    = margin(0, 0, 0, 0, unit = "cm"),
      plot.title       = element_text(size = 34 / .pt)
    ) +
    labs(
      color = glue("{label_name}: "),
      shape = glue("{label_name}: "),
      title =
        paste0(
          title,
          " (",
          format(round(pct_var, 4) * 100, nsmall = 2),
          "%)"
        ),
      x =
        paste0(
          "CA1 (variance explained: ",
          round(var_expl[1], digits = 1),
          "%)"
        ),
      y = y_lab
    ) +
    ggvegan::geom_ordi_axis(
      lty   = "dashed",
      color = "darkgray"
    ) +
    geom_point(
      data = .cap_site,
      aes(
        x     = LD1,
        y     = LD2,
        color = col_shp,
        shape = col_shp
      ),
      size = 10 / .pt,
      alpha = 0.65
    ) +
    stat_ellipse(
      data = .cap_site,
      aes(
        x     = LD1,
        y     = LD2,
        group = col_shp,
        color = col_shp
      ),
      linewidth = 3 / .pt,
      level = 0.90,
      alpha = 0.65,
      show.legend = FALSE
    ) +
    geom_segment(
      data = .cap_score,
      aes(
        x    = 0,
        y    = 0,
        xend = LD1,
        yend = LD2
      ),
      linewidth = 3 / .pt,
      arrow = arrow(length = unit(0.3, "cm")),
      color = "black"
    ) +
    # geom_text(
    geom_shadowtext(
      data = .cap_score,
      aes(
        x = LD1,
        y = LD2,
        label = pft
      ),
      color = "black",
      vjust = "outward",
      hjust = "outward",
      family = "serif",
      bg.color = "white"
    )

  # ---- end of function plot_template ---- #
}
```

## 5.2 CAP on Seasons

LOOCV:
  Winter   Spring   Summer   Autumn 
70.00000 71.69811 67.56757 80.85106 

Overall classification success (m=20) : 72.8813559322034 percent
Winter (n=40) correct: 70 percent
Spring (n=53) correct: 71.6981132075472 percent
Summer (n=37) correct: 67.5675675675676 percent
Autumn (n=47) correct: 80.8510638297872 percent

```{r rel-cap-season}
data_filt <- filter(cmtx_w_pct)

data_sub <- select(data_filt, chloro:pras)

dist_mat <- vegdist(data_sub) 

# CAP
cap_result_szn <-
  CAPdiscrim(
    dist_mat ~ season,
    data         = as.data.frame(data_filt),
    permutations = 0,
    mmax         = 99
  )

# cross validation
cap_result_szn$percent
cap_result_szn$percent.level

cap_extract_szn <- cap_extraction(
  .data        = data_filt,
  .data_sub    = data_sub,
  scale_fact   = 5,
  cap_dist     = cap_result_szn
) 

(cap_rel_szn_plt <-
  plot_template(
    cap_ext    = cap_extract_szn,
    label_name = "Season",
    axes       = c(1, 2),
    title      = "Seasonal",
    .col       = season
  ) +
  scale_color_manual(
    values = col_manual
  )
)

```

LOOCV:
 Winter   Spring   Summer   Autumn 
70.00000 69.81132 67.56757 82.97872 

Overall classification success (m=25) : 72.8813559322034 percent
Winter (n=40) correct: 70 percent
Spring (n=53) correct: 69.811320754717 percent
Summer (n=37) correct: 67.5675675675676 percent
Autumn (n=47) correct: 82.9787234042553 percent

```{r abs-cap-season}
data_filt <- filter(cmtx_w)

data_sub <- select(data_filt, chloro:pras)

dist_mat <- vegdist(data_sub) 

# CAP
cap_result_szn <-
  CAPdiscrim(
    dist_mat ~ season,
    data         = as.data.frame(data_filt),
    permutations = 0,
    mmax         = 99
  )

# cross validation
cap_result_szn$percent
cap_result_szn$percent.level

cap_extract_szn <- cap_extraction(
  .data        = data_filt,
  .data_sub    = data_sub,
  scale_fact   = 5,
  cap_dist     = cap_result_szn
) 

(cap_abs_szn_plt <-
  plot_template(
    cap_ext    = cap_extract_szn,
    label_name = "Season",
    axes       = c(1, 2),
    title      = "Seasonal",
    .col       = season
  ) +
  scale_color_manual(
    values = col_manual
  )
)
```



## 5.3 CAP on Distance From Shore

LOOCV
Nearshore  Offshore 
 73.80952  93.33333
 
Overall classification success (m=2) : 88.7005649717514 percent
Nearshore (n=42) correct: 73.8095238095238 percent
Offshore (n=135) correct: 93.3333333333333 percent


```{r}
data_filt %>%
  select(chloro:pras) %>%
  vegdist(
    .,
    method = "bray"
  ) %>%
  betadisper(
    group = select(data_filt, nearoff)[["nearoff"]]
  ) %T>% print() %>%
     TukeyHSD()
```


```{r cap-dist}
data_filt <- filter(cmtx_w_pct,
                    # str_detect(station, "10|12|16|18")
                    )

data_sub <- select(data_filt, chloro:pras)

dist_mat <- vegdist(data_sub)

# CAP
cap_result_dist <-
  CAPdiscrim(
    dist_mat ~ nearoff,
    data         = as.data.frame(data_filt),
    permutations = 0,
    mmax         = 99
  )

# cross validation
cap_result_dist$percent
cap_result_dist$percent.level

cap_extract_dist <- cap_extraction(
  .data        = data_filt,
  .data_sub    = data_sub,
  scale_fact   = 4,
  cap_dist     = cap_result_dist,
  ld2_range    = c(-1.25, 1.25),
  ld2_spc_h    = c(-1.0, 1.0)
) 


(cap_rel_dist_plt <-
  plot_template(
    cap_ext    = cap_extract_dist,
    label_name = NULL,
    axes       = c(1),
    title      = "Distance from Shore",
    .col       = nearoff
  ) +
  scale_color_manual(
    values = col_manual
  ) +
  coord_cartesian(xlim = c(NA, 4))
)
```

LOOCV
Offshore Nearshore 
 91.11111  78.57143 

Overall classification success (m=2) : 88.135593220339 percent
Offshore (n=135) correct: 91.1111111111111 percent
Nearshore (n=42) correct: 78.5714285714286 percent

```{r abs-cap-dist}
data_filt <- filter(cmtx_w) %>%
    mutate(nearoff = fct(nearoff))

data_sub <- select(data_filt, chloro:pras)

dist_mat <- vegdist(data_sub)

# CAP
cap_result_dist <-
  CAPdiscrim(
    dist_mat ~ nearoff,
    data         = as.data.frame(data_filt),
    permutations = 0,
    mmax         = 99
  )

# cross validation
cap_result_dist$percent
cap_result_dist$percent.level

cap_extract_dist <- cap_extraction(
  .data        = data_filt,
  .data_sub    = data_sub,
  scale_fact   = 4,
  cap_dist     = cap_result_dist,
  ld2_range    = c(-1.25, 1.25),
  ld2_spc_h    = c(-1.0, 1.0)
) 

(cap_abs_dist_plt <-
  plot_template(
    cap_ext    = cap_extract_dist,
    # label_name = "Distance from Shore",
    label_name = NULL,
    axes       = c(1),
    title      = "Distance from Shore",
    .col       = nearoff
  ) +
  scale_color_manual(
    values = col_manual
  )
)

```

## 5.5 Save Plots

```{r}


( 
 cap_relative <- 
  cap_rel_szn_plt/cap_rel_dist_plt +
  plot_annotation(
    # title      = "Composition and abundance",
    tag_levels = "a",
    tag_suffix = ")"
  ) +
  plot_layout(
    # guides = "collect",
    # Adjusts the relative height of each plot such that more vertical space is
    # provided for the trait plot for the spread of the labels
    # heights = (c(1,1.5))
  ) &
  theme(
      plot.background = element_rect(fill = "white", color = NA),
      # plot.tag.position = c(0.1, 0.9),
      plot.margin = margin(b = 1, t = 2, l = 2, r = 20),
    plot.tag.position = c(0.02,0.985),
    plot.tag          = element_text(size = 32/.pt, face = "bold"),
    plot.title        = element_text(size = 32/.pt, face = "bold"),
    # legend.position   = "bottom"
  )
)

( 
cap_absolute <- 
  cap_abs_szn_plt/cap_abs_dist_plt +
  plot_annotation(
    # title      = "Composition and abundance",
    tag_levels = "a",
    tag_suffix = ")"
  ) +
  plot_layout(
    # guides = "collect",
    # Adjusts the relative height of each plot such that more vertical space is
    # provided for the trait plot for the spread of the labels
    # heights = (c(1,1.5))
  ) &
  theme(
      plot.background = element_rect(fill = "white", color = NA),
      # plot.tag.position = c(0.1, 0.9),
      plot.margin = margin(b = 1, t = 2, l = 2, r = 20),
    plot.tag.position = c(0.02,0.985),
    plot.tag          = element_text(size = 32/.pt, face = "bold"),
    plot.title        = element_text(size = 32/.pt, face = "bold"),
    # legend.position   = "bottom"
  )
)

# shelf(camcorder)
# camcorder::gg_record(
#   dir    = here::here("data", "cam_test", "cap_analysis"),
#   device         = "png",
#   height         = 70,
#   width          = 90 * 0.5,
#   dpi            = 600,
#   units          = "mm",
#   bg             = "white"
# )

# shell.exec(here::here("data", "cam_test", "cap_analysis"))
# gg_stop_recording()

# gg_resize_film(width = 170, height = 170,  dpi = 600, units = "mm")

# ---- save season and station ---- #
save_gg(
  plt            = cap_relative,
  save_location  = here(dir_plt_save),
  save_name      = glue("cap_relative", file_suffix),
  verbose        = TRUE,
  # overwrite      = TRUE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "png",
  width          = 170,
  height         = 170,
  dpi            = 600,
  units          = "mm",
  bg             = "white"
)

save_gg(
  plt            = cap_absolute,
  save_location  = here(dir_plt_save),
  save_name      = glue("cap_absolute", file_suffix),
  verbose        = TRUE,
  # overwrite      = TRUE,
  time_stamp_fmt = "%Y%m%d_%H%M%S",
  device         = "png",
  width          = 170,
  height         = 170,
  dpi            = 600,
  units          = "mm",
  bg             = "white"
)
```



```{r}
cmtx_w %>%
    # filter(str_detect(station, "10|12|16|18")) %>%
summarise(
          med = median(total_chl_a, na.rm = TRUE),
          mean = mean(total_chl_a, na.rm = TRUE),
          sd = sd(total_chl_a, na.rm = TRUE),
          max = max(total_chl_a, na.rm = TRUE),
          iqr = quantile(total_chl_a, probs = 0.75, na.rm = TRUE),
          .by = season
          )

cmtx_w %>%
  count(
    station
  )
cmtx_w %>%
  count(
    station,
    season
  )
cmtx_w %>%
    filter(str_detect(station, "10|12|16|18")) %>%
    
  count(
    
    nearoff
  )
  
```

## 5.4 CAP Separate Distance and Season

```{r cap-off-season}
data_filt <- filter(cmtx_w_pct, str_detect(nearoff, "Off"))

data_sub <- select(data_filt, chloro:pras)

dist_mat <- vegdist(data_sub) 

# CAP
cap_dist2 <-
  CAPdiscrim(
    dist_mat ~ season,
    data         = as.data.frame(data_filt),
    permutations = 0,
    mmax         = 99
  )

cap_extract <- cap_extraction(
  .data        = data_filt,
  .data_sub    = data_sub,
  scale_fact   = 5,
  cap_dist = cap_dist2
) 


plot_template(
    cap_ext    = cap_extract,
    label_name = "Season",
    axes       = c(1, 2),
    title      = "Offshore Seasonal",
    .col       = season
)

```


```{r cap-near-season}
data_filt <- filter(cmtx_w_pct, str_detect(nearoff, "Near"))

data_sub <- select(data_filt, chloro:pras)

dist_mat <- vegdist(data_sub) 

# CAP
cap_dist2 <-
  CAPdiscrim(
    dist_mat ~ season,
    data         = as.data.frame(data_filt),
    permutations = 0,
    mmax         = 99
  )

cap_extract <- cap_extraction(
  .data        = data_filt,
  .data_sub    = data_sub,
  scale_fact   = 5,
  cap_dist = cap_dist2
) 


plot_template(
    cap_ext    = cap_extract,
    label_name = "Season",
    axes       = c(1, 2),
    title      = "Nearshore Seasonal",
    .col       = season
)

```

