---
title: "NOAA Wind Data"
author: "Sebastian DiGeronimo"
date: "2023-01-11"
output: html_document
---

----



----- 
```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
    # additional packages, if from GitHub, include username
    rnoaa
)

conflicts_prefer(
    dplyr::filter, 
    dplyr::select
    )

source(here("scripts", "misc_functions.R"))

```

# Hurricanes that Affected the Florida Keys

```{r hurricane-dates}
hurricane_dates <-
  tribble(
    ~date, ~name,
    # "2015-08-30", "Erika", ## 
    
    # "2016-10-12", "Nicole", ##
    
    # "2017-07-31", "Emily", ##
    "2017-09-10", "Irma",
    # "2017-10-29", "Philippe", ##
    
    "2018-09-03", "Gordon", 
    
    # "2019-09-01", "Dorian", ##
    
    # "2020-05-25", "Bertha", ##
    "2020-08-02", "Isaias",
    # "2020-08-26", "Laura", ##
    "2020-09-11", "Sally",
    "2020-11-09", "Eta", 
    
  ) %>%
  mutate(date = ymd(date)) %T>% 
  print()

```

# Add NOAA Key to `.Renvrion`

`rnoaa` package seems promising to gather wind data
<https://docs.ropensci.org/rnoaa/articles/rnoaa.html>

go to <https://www.ncdc.noaa.gov/cdo-web/token> to get token
then paste into `.Renvrion` using below code
copy and paste: NOAAKEY = "<ENTER TOKEN HERE>"

NOTE: `.Renviron` should be ignored in `.gitignore`

```{r wind-speed-data}
# set location to save data and plots
dir_data_save <- here("data", "raw", "noaa_atmosphere")
# path_in <- here("data", "raw")

# wind_url <- "https://meteostat.net/en/station/72211?t=2016-01-01/2021-12-31"


if (str_equal(Sys.getenv("NOAAKEY"), "")) {
  # usethis::edit_r_profile(scope = "project") # <-- old way

  # copy and paste: NOAAKEY = "<ENTER TOKEN HERE>"
  usethis::edit_r_environ(scope = "project")
  message("RESTART R!")
} else {
  message("`NOAAKEY` found in `.Renviron`")
  
  # set noaakey in options
  options(noaakey = Sys.getenv("NOAAKEY"))
}

```
## Test `noaakey` Works with `rnoaa`

```{r test-rnoaa}
stopifnot(
  "Restart R for `.Renvrion` to take effect" = !is.null(getOption("noaakey"))
)


# test that it works
ncdc_locs(locationcategoryid = "CITY", sortfield = "name", sortorder = "desc")

# might want data from https://www1.ncdc.noaa.gov/pub/data/cdo/documentation/gsom-gsoy.pdf
# ncdc_datasets(
#   locationid = "ZIP:33050", 
#   startdate  = "2015-03-01", 
#   enddate    = "2022-05-31"
#   )
```



# NOAA Data


Explain variables:
<https://www1.ncdc.noaa.gov/pub/data/cdo/documentation/gsom-gsoy.pdf>

## Find NOAA Stations in Florida Keys

```{r search-noaa}
noaa_station_search <-
  ncdc_stations(locationid = "ZIP:33050") %$%
  data %>%
  mutate(
    min_date = as_date(mindate),
    max_date = as_date(maxdate),
    name2    = str_remove(name, ", FL US"),
    datasetid = str_remove(id, ":.*")
  ) %>%
  filter(
    !str_detect(name2, "(?i)miami|lauderdale")
    & max_date > as_date("2016-01-01") 
    & min_date < as_date("2020-01-01")
  ) %T>% 
  print() 

# quickly plot
noaa_station_search %>%
  ggplot(aes(x = longitude, y = latitude, color = datasetid)) +
  geom_point() +
  ggrepel::geom_text_repel(aes(label = name2), max.overlaps = 20) +
  coord_cartesian(xlim = c(NA, -80.6), ylim = c(24, 25.5)) +
  labs(color = NULL) +
  theme_bw() +
  theme(legend.position = "bottom")
```


## Extract Time Series from NOAA Stations


Extract from datasets *GHCND*


```{r extract-noaa-data}
noaa_station_data <-
  noaa_station_search %>%
  filter(str_detect(id, "(?i)ghcnd")) %>%
  mutate(
      id2 = str_remove(id, "GHCND:"),
      
      # extract data from NOAA
      data = map(
        id2,
        \(.x) {
          ghcnd(
            stationid = .x,
            date_min  = "2010-01-01",
            add_units = TRUE
          ) %>%
            filter(
              year >= 2016 
              & year <= 2021
            )
        }
      ),
      
      # separate data into individual variables
      vars = map(
        data,
        ~ ghcnd_splitvars(.x) %>%
          map(
            \(.y) {
              .y %>%
                separate_wider_delim(
                  cols  = "date",
                  names = c("year", "month", "day"),
                  delim = "-",
                  cols_remove = FALSE
                )
            }
          )
      )
    ) %>%
    unnest_wider(vars, names_sep = "_") %>%
  select(!contains(c("snow", "snwd")))  
    
noaa_station_data %T>% 
  print() %>%
  select(
    "id_full" = id,     
    "lat" = latitude,
    "lon" = longitude,
    vars_prcp) %>%
  unnest(vars_prcp, names_repair = "unique", keep_empty = TRUE)


names(noaa_station_data)
```


## Extract Precipitation

Units: mL

```{r ts-prcp-stn, fig.width=20}
noaa_station_data %>%
  mutate(
    plt_prcp = map2(
      vars_prcp,
      name2,
      .hurr = hurricane_dates,
      \(.x, .y, .hurr) {
        .x %>%
          ggplot(aes(x = date, y = prcp)) +
          geom_line(na.rm = TRUE) +
          geom_vline(xintercept = .hurr$date, color = "red", linetype = "dashed") +
          geom_text(
            data = .hurr, 
            aes(x = date, y = 2000, label = name,),
            angle = 90,
            vjust = -0.5,
            inherit.aes = FALSE,
            ) +
          # geom_smooth(method = "loess", span = 0.05, se = FALSE) +
          scale_x_date(
            expand = c(0, 0),
            date_breaks = "1 year",
            date_labels = "%Y"
          ) +
          labs(title = .y) +
          theme_bw()
      }
    )
  ) %>%
  pull(plt_prcp)
```





```{r ts-precip, fig.width=25}
prcp_data <-
  noaa_station_data %>%
  select(
    "id_full" = id, 
    "lat" = latitude,
    "lon" = longitude,
    name2, 
    vars_prcp
    ) %>%
  unnest(vars_prcp, names_repair = "unique", keep_empty = TRUE) %>%
  arrange(date) %>%
  mutate(
    name2 = fct(name2, levels = c("MARATHON 6.8 ENE", "DUCK KEY", "CURRY HAMMOCK STATE PARK", 
                                  "MARATHON AIRPORT"))
  )%T>% 
  print()


# average per day for each month
# prcp_monthly_daily <-
#   prcp_data %>%
#   summarize(
#     .by = c(year, month),
#     prcp_daily_avg = mean(prcp, na.rm = TRUE),
#     prcp_daily_sd  = sd(prcp, na.rm = TRUE)
#   ) %>%
#   mutate(
#     date = paste(year, month, sep = "-"),
#     date = ym(date),
#     err_max = prcp_avg + prcp_sd,
#     err_min = prcp_avg - prcp_sd
#   ) %T>% 
#   print()


# ---

# daily average between stations
prcp_daily <-
  prcp_data %>%
  summarize(
    .by = c(date, year, month, day),
    prcp_avg = mean(prcp, na.rm = TRUE),
    prcp_sd  = sd(prcp, na.rm = TRUE),
    prcp_se  = FSA::se(prcp, na.rm = TRUE)
  ) %T>% 
  print()

# for plotting daily average betweeen guages, don't display when daily is low and
# sd is low
prcp_daily_filt <- 
  prcp_daily %>%
  mutate(
    prcp_avg2 = prcp_avg,
    prcp_avg = if_else(prcp_avg < 0.5 & prcp_sd < 0.05, NA, prcp_avg),
    # prcp_avg = if_else(prcp_sd < 100, NA, prcp_avg),
    err_max = prcp_avg + prcp_sd,
    err_min = prcp_avg - prcp_sd
  ) %T>% 
  print()

# sum average daily precipitation
prcp_monthly <- 
  prcp_daily %>%
  summarize(
    .by = c(year, month),
    prcp_sum = sum(prcp_avg, na.rm = TRUE),
  ) %>%
  mutate(
    date = paste(year, month, sep = "-"),
    date = ym(date)
  ) %T>% 
  print()

# ---

# # sum daily precipitation per station per month
# prcp_monthly_station <- 
#   prcp_data %>%
#   summarize(
#     .by = c(id, name2, year, month),
#     prcp_sum = sum(prcp, na.rm = TRUE),
#     # prcp_avg = mean(prcp, na.rm = TRUE),
#     # prcp_sd  = sd(prcp, na.rm = TRUE)
#   ) %>%
#   arrange(id) %>%
#   mutate(
#     .by = id,
#     date = paste(year, month, sep = "-"),
#     date = ym(date),
#     date2 = date + months(1)
#     # err_max = prcp_avg + prcp_sd,
#     # err_min = prcp_avg - prcp_sd
#   ) %>%
#   arrange(date) %T>% 
#   print()
# 
# # average monthly precipitation from the stations
# prcp_monthly3 <-
# 
#   prcp_monthly_station %>%
#   summarize(
#     .by = c(year, month),
#     # prcp_sum = sum(prcp, na.rm = TRUE),
#     prcp_avg = mean(prcp_sum, na.rm = TRUE),
#     prcp_sd  = sd(prcp_sum, na.rm = TRUE),
#     prcp_se  = FSA::se(prcp_sum)
#   ) %>%
#   mutate(
#     date = paste(year, month, sep = "-"),
#     date = ym(date),
#     date2 = date + months(1),
#     date3 = date + 15
#     # err_max = prcp_avg + prcp_sd,
#     # err_min = prcp_avg - prcp_sd
#   ) %T>% 
#   print()

```


```{r ts-precip, fig.width=25}
# precipitation plot each station separately with monthly average
prcp_data %>%
  ggplot(aes(x = date, y = prcp, color = name2)) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_line(
    linewidth = 1.25,
    na.rm = TRUE,
    show.legend = FALSE
    ) +
  geom_vline(xintercept = hurricane_dates$date, color = "red", linetype = "dashed") +
  geom_vline(xintercept = prcp_monthly$date, color = "blue", linetype = "dashed") +
  geom_text(
      data = hurricane_dates, 
      aes(x = date, y = 4000, label = name,),
      angle = 90,
      vjust = -0.5,
      hjust = 1,
      inherit.aes = FALSE,
      ) +
      geom_line(
    data = prcp_monthly,
    aes(x = date + months(1), y = prcp_sum),
    color = "blue",
    inherit.aes = FALSE,
    na.rm = TRUE
  ) +
  geom_point(
    data = prcp_monthly,
    aes(x = date + months(1), y = prcp_sum),
    color = "blue",
    size = 2,
    inherit.aes = FALSE,
    na.rm = TRUE
  ) +

  labs(
    x = NULL
  ) +
  scale_x_date(
    expand = expansion(0, 0),
    date_breaks = "1 year",
    date_labels = "%Y",
    date_minor_breaks = "6 month"
  ) +
  scale_y_continuous(
    expand = expansion(add = c(50, 0)),
    limits = c(0, NA),
    oob    = scales::oob_keep) +
  facet_wrap(~name2, ncol = 1) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.margin = margin(t = -10)
  )


# precipitation plot average daily and monthly average
ggplot() +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_vline(xintercept = hurricane_dates$date, color = "red", linetype = "dashed") +
  geom_text(
      data = hurricane_dates, 
      aes(x = date, y = 2000, label = name,),
      angle = 90,
      vjust = -0.5,
      hjust = 1.5,
      inherit.aes = FALSE,
      ) +
          geom_line(
    data = prcp_monthly,
    aes(x = date + months(1), y = prcp_sum),
    color = "blue",
    inherit.aes = FALSE,
    na.rm = TRUE
  ) +
  geom_point(
    data = prcp_monthly,
    aes(x = date + months(1), y = prcp_sum),
    color = "blue",
    size = 2,
    inherit.aes = FALSE,
    na.rm = TRUE
  ) +
  geom_pointrange(
    data = prcp_daily_filt,
    aes(x = date,  y = prcp_avg, ymin = err_min, ymax = err_max),
    color = "red",
    inherit.aes = FALSE,
    na.rm = TRUE
  ) +

  labs(
    y = expression("Precipitation"~(mL)),
    x = NULL
  ) +
  scale_x_date(
    expand = expansion(0, 0),
    date_breaks = "1 year",
    date_labels = "%Y",
    date_minor_breaks = "6 month"
  ) +
  scale_y_continuous(
    expand = expansion(add = c(50, 0)),
    limits = c(0, NA),
    oob    = scales::oob_keep) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.margin = margin(t = -10)
  )
```

## AWND – Average Wind Speed/Fastest 2-minute

units: m/s

```{r ts-wind-speed, fig.width=25}
awnd_data <-
  noaa_station_data %>%
  select(
    "id_full" = id, 
    "lat" = latitude,
    "lon" = longitude,
    name2, 
    vars_awnd) %>%
  unnest(vars_awnd, names_repair = "unique", keep_empty = FALSE) %>%
  arrange(date) %>%
  relocate(awnd, .after = date) %T>% 
  print()

awnd_monthly <-
  awnd_data %>%
  summarize(
    .by = c(year, month),
    awnd_avg = mean(awnd, na.rm = TRUE),
    awnd_sd  = sd(awnd, na.rm = TRUE)
  ) %>%
  mutate(
    date = paste(year, month, sep = "-"),
    date = ym(date),
    err_max = awnd_avg + awnd_sd,
    err_min = awnd_avg - awnd_sd
  ) %T>% 
  print()
  


# avg wind speed plot daily and monthly average
awnd_data %>%
  ggplot(aes(x = date, y = awnd, color = name2)) +
  geom_line(
    linewidth   = 0.75,
    na.rm       = TRUE,
    show.legend = FALSE,
    color       = "blue"
    ) +
  geom_vline(xintercept = hurricane_dates$date, color = "red", linetype = "dashed") +
  geom_text(
      data = hurricane_dates, 
      aes(x = date, y = 150, label = name,),
      angle = 90,
      vjust = -0.5,
      hjust = 1.5,
      inherit.aes = FALSE,
      ) +
  geom_step(
    data = awnd_monthly,
    aes(x = date, y = awnd_avg),
    color = "red",
    inherit.aes = FALSE,
    na.rm = TRUE
  ) +
  geom_pointrange(
    data = awnd_monthly,
    aes(x = date, y = awnd_avg, ymin = err_min, ymax = err_max),
    color = "red",
    inherit.aes = FALSE,
    na.rm = TRUE
  ) +
  geom_smooth(method = "loess", span = 0.05, show.legend = FALSE) +
  labs(
    x = NULL,
    y = expression("Average Daily Wind Speed"~(m~s^-1)) 
    ) +
  scale_x_date(
    expand = expansion(0, 0),
    date_breaks = "1 year",
    date_labels = "%Y",
    date_minor_breaks = "6 month"
  ) +
  scale_y_continuous(
    expand = expansion(add = c(0, 0)),
    limits = c(0, 150),
    oob    = scales::oob_keep) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.margin = margin(t = -10)
  )
```
## WSF2 – Maximum Wind Speed/Fastest 2-minute

units: m/s

```{r ts-wind-speed, fig.width=25}
wsf2_data <-
  noaa_station_data %>%
  select(
    "id_full" = id,   
    "lat" = latitude,
    "lon" = longitude,
    name2, 
    vars_wsf2) %>%
  unnest(vars_wsf2, names_repair = "unique", keep_empty = FALSE) %>%
  arrange(date) %T>% 
  print()

wsf2_monthly <-
  wsf2_data %>%
  summarize(
    .by = c(year, month),
    wsf2_avg = mean(wsf2, na.rm = TRUE),
    wsf2_sd  = sd(wsf2, na.rm = TRUE)
  ) %>%
  mutate(
    date = paste(year, month, sep = "-"),
    date = ym(date),
    err_max = wsf2_avg + wsf2_sd,
    err_min = wsf2_avg - wsf2_sd
  ) %T>% 
  print()
  

# max 2 min wind plot with monthly average 
wsf2_data %>%
  ggplot(aes(x = date, y = wsf2, color = name2)) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_line(
    linewidth   = 0.75,
    na.rm       = TRUE,
    show.legend = FALSE,
    color       = "blue"
    ) +
  geom_vline(xintercept = hurricane_dates$date, color = "red", linetype = "dashed") +
  geom_text(
      data = hurricane_dates, 
      aes(x = date, y = 250, label = name,),
      angle = 90,
      vjust = -0.5,
      hjust = 1.5,
      inherit.aes = FALSE,
      ) +
  geom_step(
    data = wsf2_monthly,
    aes(x = date, y = wsf2_avg),
    color = "red",
    inherit.aes = FALSE,
    na.rm = TRUE
  ) +
  geom_pointrange(
    data = wsf2_monthly,
    aes(x = date, y = wsf2_avg, ymin = err_min, ymax = err_max),
    color = "red",
    inherit.aes = FALSE,
    na.rm = TRUE
  ) +
  geom_smooth(method = "loess", span = 0.05, show.legend = FALSE) +
  labs(
    x = NULL,
    y = expression("Max 2 Minute Wind Speed"~(m~s^-1)) 
    ) +
  scale_x_date(
    expand = expansion(0, 0),
    date_breaks = "1 year",
    date_labels = "%Y",
    date_minor_breaks = "6 month"
  ) +
  scale_y_continuous(
    expand = expansion(add = c(0, 0)),
    limits = c(0, 250),
    oob    = scales::oob_keep) +
  theme_bw() +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.margin = margin(t = -10)
  )

```

# Join Precipitation, Average Wind Speed and Max 2 Minute Wind Gust

## Daily

```{r join-atm-data}
noaa_atm_merg <-
  full_join(
    awnd_data,
    wsf2_data,
    suffix = c("_awnd", "_wfs2"),
    by = join_by(id_full, lat, lon, name2, id, year, month, day, date),
  ) %>%
  full_join(
    # prcp_daily,
    prcp_data %>%
      select(-c(lat, lon, name2, id_full)) %>%
      pivot_wider(
        names_from   = c(id), 
        values_from  = c(prcp, mflag, qflag, sflag),
        names_sep    = "_",
        names_repair = janitor::make_clean_names
      ),
    by = join_by(date, year, month, day)
  ) %T>%
  print() %>%
  # ---- save ---- #
  save_csv(
    .data          = .,
    save_location  = dir_data_save,
    save_name      = "noaa_atm_merg_daily",
    overwrite      = FALSE,
    verbose        = TRUE,
    time_stamp_fmt = NULL,
    utf_8          = FALSE
  )

```


## Monthly

```{r}
awnd_monthly

prcp_monthly

wsf2_monthly



```

```{r}
wsf2_data 





(wsf2_data %>%
  # slice(1:10) %>%
  mutate(
    # .keep = "used",
    year = year(date),
    date2 = map(
      date,
      \(.x) {
        year(.x) <- 2001
        return(.x)
      }
    )
  ) %>%
  unnest(date2) %>%
  ggplot() +
    geom_line(aes(x = date2, y = wsf2, color = as.factor(year), lab = date)) +
    scale_x_date(
      date_labels = "%b", 
      date_breaks = "1 month"
      ) +
  labs(
    color = NULL,
    x = NULL
  ) +
  theme_bw()) %>%
  plotly::ggplotly()
    

  


```

